### 2019-02-06 sections/appendix.tex
setup correct acm style

% !TEX root=../pldi2019.tex

\section{Appendix}

  \subsection{Evaluation rules}

  % \begin{figure}

    \begin{mathpar}
      \boxed{\RelationE} \\
      \userule{E-Value} \\
      \userule{E-App} \\
      \userule{E-IfTrue} \\
      \userule{E-IfFalse} \\
      \userule{E-Pair} \\
      \userule{E-Ref} \quad
      \userule{E-Deref} \\
      \userule{E-Assign} \\
      \userule{E-Edit} \quad
      \userule{E-Enter} \quad
      \userule{E-Update} \\
      \userule{E-Then} \quad
      \userule{E-Next} \\
      \userule{E-And} \quad
      \userule{E-Or} \\
      \userule{E-Xor} \quad
      \userule{E-Appoint} \quad
      \userule{E-Fail}
    \end{mathpar}

  %   \caption{Evaluation semantics} \label{fig:evaluation-semantics}
  % \end{figure}

  \subsection{Additional typing rules}

  % \begin{figure}[h]

    \begin{mathpar}
      \boxed{\RelationT} \\
      \userule{T-Var} \quad
      \userule{T-Loc} \\
      \userule{T-Abs} \quad
      \userule{T-App} \\
      \userule{T-If} \\
      \userule{T-Pair} \\
      \userule{T-Ref} \quad
      \userule{T-Deref} \\\
      \userule{T-Assign}
    \end{mathpar}

  %   \caption{Additional typing rules}
  % \end{figure}

  \subsection{Proofs}

  \subsubsection{Theorem~\ref{thm:pres-eval}}
\begin{proof}
  We prove Theorem~\ref{thm:pres-eval} by induction on $e$:\\

  \noindent\textbf{Case} $e=\lambda x:\tau.e, e_1 e_2, x, c, l, e_1 \star e_2,
      \If{e_1}{e_2}{e_3},$\\
      $\tuple{e_1, e_2},\unit,\Ref e,!e,e_1 := e_2$ preservation has
      been proven for these cases by Pierce~\cite{books/Pierce02TAPL}.\\

  \noindent\textbf{Case} $\userule{E-Edit}$\\
    \indent Given that $\Gamma,\Sigma\infers\Edit e:\Task \tau$ and $\Sigma\infers s$,
      \refrule{T-Edit} gives us that $\Gamma,\Sigma\infers e:\tau$. The induction
      hypothesis gives us that $e,s\evaluate v,s'$ also preserves, and thus
      $\Gamma,\Sigma\infers v:\tau$ and $\Sigma\infers s'$. Therefore
      $\Gamma,\Sigma\infers\Edit v:\Task\tau$.\\

  \noindent\textbf{Case} $\userule{E-Enter}$\\
    \indent Evaluation does not alter $e$ and $s$, therefore this case holds tivially.\\

  \noindent\textbf{Case} $\userule{E-Update}$\\
    \indent Given that $\Gamma,\Sigma\infers \Edit e:\Task \tau$ and
      $\Sigma\infers s$, \refrule{T-Update} gives us that $\Gamma,\Sigma\infers e:\Ref \tau$.
      The induction hypothesis gives us that $e,s\evaluate l,s'$ also preserves,
      and thus $\Gamma,\Sigma\infers l:\Ref\tau$ and $\Sigma\infers s'$.
      Therefore $\Gamma,\Sigma\infers\Update l:\Task\tau$\\

  \noindent\textbf{Case} $\userule{E-Fail}$\\
    \indent Evaluation does not alter $e$ and $s$, therefore this case holds tivially.\\

  \noindent\textbf{Case} $\userule{E-Then}$\\
    \indent Given that $\Gamma,\Sigma\infers e_1\Then e_2:\Task \tau$ and
      $\Sigma\infers s$, \refrule{T-Then} gives us that
      $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and
      $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$. By the induction
      hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus
      $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Sigma\infers s'$. Therefore
      $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$.\\

  \noindent\textbf{Case} $\userule{E-Next}$\\
    \indent Given that $\Gamma,\Sigma\infers e_1\Next e_2:\Task \tau$ and
      $\Sigma\infers s$, \refrule{T-Next} gives us that
      $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and
      $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$. By the induction
      hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus
      $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Sigma\infers s'$. Therefore
      $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$.\\

  \noindent\textbf{Case} $\userule{E-And}$\\
    \indent  Given that $\Gamma,\Sigma\infers e_1\And e_2:\Task(\tau_1\times\tau_2)$
      and $\Sigma\infers s$, \refrule{T-And} gives us that
      $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and
      $\Gamma,\Sigma\infers e_2:\Task\tau_2$. By the induction hypothesis, we
      know that both $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$
      preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$,
      $\Sigma\infers s'$, $\Gamma,\Sigma\infers t_2:\Task\tau_2$ and
      $\Sigma\infers s''$. Therfore
      $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$\\

  \noindent\textbf{Case} $\userule{E-Or}$\\
    \indent Given that $\Gamma,\Sigma\infers e_1\Or e_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau$ and
      $\Gamma,\Sigma\infers e_2:\Task\tau$. By the induction hypothesis, we have that both
      $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$ preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau$, $\Sigma\infers s'$,
      $\Gamma,\Sigma\infers t_2:\Task\tau$ and $\Sigma\infers s''$. Therefore $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$.\\

  \noindent\textbf{Case} $\userule{E-Xor}$\\
    \indent Evaluation does not alter $e$ and $s$, therefore this case holds tivially.

  % \noindent\textbf{Case} $\userule{E-Appoint}$
  %     Given that $\Gamma,\Sigma\infers u \At e:\Task\tau$ and $\Sigma\infers s$, the induction hypothesis gives us that $e,s\evaluate t,s'$ also preserves, and therefore by \refrule{T-Appoint} $\Gamma,\Sigma\infers u \At t:\Task\tau$.
\end{proof}

\subsubsection{Lemma~\ref{lem:presvalue}}
\begin{proof}
  We prove Lemma~\ref{lem:presvalue} by induction over $e$.\\

  \noindent\textbf{Case} $\Value{(\Edit v,s)}=v$ By T-Edit, if
  $\Gamma,\Sigma\infers \Edit v:\Task\tau$, then $\Gamma,\Sigma\infers v:\tau$.\\

  \noindent\textbf{Case} $\Value{(\Enter \tau,s)}=\bot$ Since this case does not
  lead to a value, the lemma holds trivially.\\

  \noindent\textbf{Case} $\Value{(\Update l,s)}=s(l)$ Given that
  $\Gamma,\Sigma\infers\Update l:\Task \tau$ and $\Sigma\infers s$, we know that
  $\Gamma,\Sigma\infers s(l):\tau$ by definiton.\\

  \noindent\textbf{Case} $\Value{(\Fail,s)}=\bot$ Since this case does not lead
  to a value, the lemma holds trivially.\\

  \noindent\textbf{Case} $\Value{(t_1\Then e_2,s)}=\bot$ Since this case does
  not lead to a value, the lemma holds trivially.\\

  \noindent\textbf{Case} $\Value{(t_2\Next e_2,s)}=\bot$ Since this case does
  not lead to a value, the lemma holds trivially.\\

  \noindent\textbf{Case} $\Value{(t_1\And t_2,s)}=\tuple{v_1, v_2}$ given that
  $\Value{(t_1,s)}=v_1\wedge\Value{(t_2,s)}=v_2$\\ By \refrule{T-And} we have that
  $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and
  $\Gamma,\Sigma\infers t_1:\tau_1$ and $\Gamma,\Sigma\infers t_2:\tau_2$. By the
  induction hypothesis, $ \Value{(t_1,s)}=v_1$ and $\Value{(t_2,s)}=v_2$ preserve,
  and thus $\Gamma,\Sigma\infers v_1:\tau_1$ and $\Gamma,\Sigma\infers v_2:\tau_2$.
  This gives us that $\Gamma,\Sigma\infers \tuple{v_1, v_2}:\Task(\tau_1\times\tau_2)$ \\

  \noindent\textbf{Case} $\Value{(t_1\And t_2,s)}=\bot$ given that
  $\neg(\Value{(t_1,s)}=v_1\wedge\Value{(t_2,s)}=v_2)$\\ Since this case does
  not lead to a value, the lemma holds trivially.\\

  \noindent\textbf{Case} $\Value{(t_1\Or t_2,s)}=v_1$ given that
  $\Value{(t_1,s)}=v_1$\\
  By \refrule{T-Or} we have that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$,
  and $\Gamma,\Sigma\infers t_1:\Task\tau$ and
  $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis, we have
  that $\Gamma,\Sigma\infers v_1:\tau$.\\

  \noindent\textbf{Case} $\Value{(t_1\Or t_2,s)}=v_2$ given that
  $\Value{(t_1,s)}=\bot\wedge\Value{(t_2,s)}=v_2$\\
  By \refrule{T-Or} we have that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$,
  and $\Gamma,\Sigma\infers t_1:\Task\tau$ and
  $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis, we have
  that $\Gamma,\Sigma\infers v_2:\tau$.\\

  \noindent\textbf{Case} $\Value{(t_1\Or t_2,s)}=\bot$ given that
  $\Value{(t_1,s)}=\bot\wedge\Value{(t_2,s)}=\bot$\\ Since this case does not
  lead to a value, the lemma holds trivially.\\

  \noindent\textbf{Case} $\Value{(t_1\Xor t_2,s)}=\bot$ Since this case does not
  lead to a value, the lemma holds trivially.

  % \noindent\textbf{Case} $\Value{(u\At t,s)}=\Value(t,s)$ This case follows
  % directly from the induction hypothesis.\\
\end{proof}

\subsubsection{Theorem~\ref{thm:pres-stride}}
\begin{proof}
  We prove Theorem~\ref{thm:pres-stride} by induction on $e$:\\

  \noindent\textbf{Case} $\userule{S-Fail}$ \\
    \indent  Since this case does not alter the
  expression, the theorem holds trivially.\\

  \noindent\textbf{Case} $\userule{S-Xor}$ \\
    \indent  Since this case does not alter the
  expression, the theorem holds trivially.\\

  \noindent\textbf{Case} $\userule{S-Update}$ \\
    \indent  Since this case does not alter
  the expression, the theorem holds trivially.\\

  \noindent\textbf{Case} $\userule{S-Fill}$ \\
    \indent  Since this case does not alter the
  expression, the theorem holds trivially.\\

  \noindent\textbf{Case} $\userule{S-Edit}$ \\
    \indent  Since this case does not alter the
  expression, the theorem holds trivially.\\

  \noindent\textbf{Case} $\userule{S-And}$ \\
    \indent  Given that
  $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$, by
  \refrule{T-And we} have $\Gamma,\Sigma\infers t_1:\tau_1$ and
  $\Gamma,\Sigma\infers t_2:\tau_2$. By the induction hypothesis, we also have $\Gamma,\Sigma\infers t_1':\tau_1$ and $\Gamma,\Sigma\infers t_2':\tau_2$.
  This gives us that
  $\Gamma,\Sigma\infers t_1'\And t_2':\Task(\tau_1\times\tau_2)$.\\

  \noindent\textbf{Case} $\userule{S-Next}$ \\
    \indent  Given that
  $\Gamma,\Sigma\infers e_1\Next e_2:\Task \tau$, \refrule{T-Then} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
  $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$. By the induction hypothesis,
  we know that $t_1\stride t_1'$ preserves and thus
  $\Gamma,\Sigma\infers t_1':\Task\tau_1$. Therefore
  $\Gamma,\Sigma\infers t_1'\Next e_2:\Task\tau$.\\\\

  \noindent\textbf{Case} $\userule{S-OrLeft}$ \\
    \indent  Given that
  $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$, by \refrule{T-Or} we have
  $\Gamma,\Sigma\infers t_1:\Task\tau$. By the induction hypothesis, we know that
  $t_1\stride t_1'$ preserves and thus $\Gamma,\Sigma\infers t_1':\Task\tau$.\\

  \noindent\textbf{Case}\\
  $\userule{S-OrRight}$ \\
    \indent  Given that
  $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$, by \refrule{T-Or} we have $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis, we know that
  $t_2\stride t_2'$ preserves and thus $\Gamma,\Sigma\infers t_2':\Task\tau$.\\

  \noindent\textbf{Case}\\
  $\userule{S-OrNone}$ \\
    \indent  Given that
  $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$, by \refrule{T-Or} we have
  $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
  By the induction hypothesis, we know that $t_1\stride t_1'$ and
  $t_2\stride t_2'$ preserve, and thus
  $\Gamma,\Sigma\infers t_1'\Or t_2':\Task\tau$.\\

  \noindent\textbf{Case} $\userule{S-ThenStay}$ \\
    \indent  Given that
  $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$,
  by \refrule{T-Then} we have $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
  $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. By the induction
  hypothesis, we know that $t_1\stride t_1'$ preserves, and thus
  $\Gamma,\Sigma\infers t_1'\Then e_2:\Task\tau$.\\

  \noindent\textbf{Case}\\
   $\userule{S-ThenFail}$ \\
    \indent  Given that
  $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$,
  by \refrule{T-Then} we have $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
  $e_2:\tau_1\to\Task\tau$. By the induction
  hypothesis, we know that $t_1\stride t_1'$ preserves, and thus
  $\Gamma,\Sigma\infers t_1'\Then e_2:\Task\tau$.\\

  \noindent\textbf{Case}\\
  $\userule{S-ThenCont}$\\
    \indent  Given that
  $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$, by \refrule{T-Then} we have
  $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
  $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. By the induction hypothesis, we
  know that $t_1\stride t_1'$ preserves. By Lemma~\ref{lem:presvalue}, we know
  that $\Value{(t_1')}=v_1$ preserves. By Theorem~\ref{thm:pres-eval} we know
  that $e_2 v_1\evaluate t_2$ preserves. And finally by the induction hypothesis,
  we know that $t_2\stride t_2'$ preserves. Therefore
  $\Gamma,\Sigma\infers t_2':\Task\tau$.
  %
  % \noindent\textbf{Case} $\userule{S-Appoint}$ Given that
  % $\Gamma,\Sigma\infers u\At t:\Task\tau$, by \refrule{T-Appoint} we have
  % $t:\Task\tau$. By the induction hypothesis, we know that $t,s\stride t',s'$
  % preserves, and thus $\Gamma,\Sigma\infers u\At t':\Task\tau$.\\

\end{proof}

\subsubsection{Theorem~\ref{thm:pres-norm}}

\begin{proof}
  We prove Theorem~\ref{thm:pres-norm} by induction on $e$:\\

  \noindent\textbf{Case} $\userule{N-Done}$ \\
  \indent Given that
  $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that
  $\Gamma,\Sigma\infers t:\Task\tau$ and $\Sigma\infers s'$ by
  Theorem~\ref{thm:pres-eval}. Then by Theorem~\ref{thm:pres-stride}, we have
  $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$.\\

  \noindent\textbf{Case}\\
  $\userule{N-Repeat}$ \\
  \indent Given that
  $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that
  $\Gamma,\Sigma\infers t:\Task\tau$ and $\Sigma\infers s'$ by
  Theorem~\ref{thm:pres-eval}. Then by Theorem~\ref{thm:pres-stride}, we have
  $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$. Then by the
  induction hypothesis, we finally obtain that
  $\Gamma,\Sigma\infers t'':\Task\tau$ and $\Sigma\infers s'''$.\\

\end{proof}

\subsubsection{Theorem~\ref{thm:pres-handle}}

We require the following Lemma for this proof.

\begin{lemma}
  Given that $\Sigma\infers s$, $\Sigma(l)=\tau$ and $\Gamma,\Sigma\infers v:\tau$, it holds that $\Sigma\infers s[l\mapsto v]$
  \label{lemmasigmaconsistent}
\end{lemma}
This lemma follows immediately from definition.

\begin{proof}
  We prove Theorem~\ref{thm:pres-handle} by induction on $e$:

  \noindent\textbf{Case} $\userule{H-Change}$\\
  \indent Given that
  $\Gamma,\Sigma\infers\Edit v:\Task\tau$ and $\Sigma\infers s$, the
  \refrule{H-Change} rule additionally gives us that $v,v':\tau$. Therefore by
  \refrule{T-Edit} we have that $\Gamma,\Sigma\infers\Edit v':\Task\tau$\\

  \noindent\textbf{Case} $\userule{H-Empty}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers\Edit v:\Task\beta$ and $\Sigma\infers s$, the
  \refrule{H-Empty} rule additionally gives us that $v:\tau$. Then by
  \refrule{T-Enter} we have $\Gamma,\Sigma\infers\Enter \tau:\Task\tau$ \\

  \noindent\textbf{Case} $\userule{H-Fill}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers\Enter\tau$ and $\Sigma\infers s$, the \refrule{H-Fill}
  rule additionally gives us that $v':\tau$. Then by \refrule{T-Enter} we have
  $\Gamma,\Sigma\infers \Edit v':\Task\tau$.\\

  \noindent\textbf{Case} $\userule{H-Update}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers\Update l:\Task\tau$ and $\Sigma\infers s$. This gives us
  that $\Sigma(l)=\tau$, and we additionally obtain $s(l),v':\tau$ by
  \refrule{H-Update}. By application of Lemma~\ref{lemmasigmaconsistent} this
  case holds.\\

  \noindent\textbf{Case} $\userule{H-PickLeft}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Sigma\infers s$, then by
  \refrule{T-Xor} we have $\Gamma,\Sigma\infers t_1:\Task \tau$.\\

  \noindent\textbf{Case} $\userule{H-PickRight}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Sigma\infers s$, then by
  \refrule{T-Xor} we have $\Gamma,\Sigma\infers t_2:\Task \tau$. \\

  \noindent\textbf{Case} $\userule{H-Next}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers t_1\Next e_2 :\Task\tau$ and $\Sigma\infers s$. Then by
  \refrule{T-Next}, we have $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
  $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. Then by \refrule{T-Then} we
  obtain $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$.\\

  \noindent\textbf{Case} $\userule{H-PassThen}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$ and $\Sigma\infers s$,
  \refrule{T-Then} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
  $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. By the induction hypothesis, we
  know that $t_1,s\handle{i}t_1',s'$ also preserves and thus
  $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Gamma,Sigma\infers s'$. By
  \refrule{T-Then} we now obtain that
  $\Gamma,\Sigma\infers t_1'\Then e_2:\Task\tau$. \\

  \noindent\textbf{Case} $\userule{H-PassNext}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$ and $\Sigma\infers s$,
  \refrule{T-Next} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
  $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. By the induction hypothesis, we
  know that $t_1,s\handle{i}t_1',s'$ also preserves and thus
  $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Gamma,Sigma\infers s'$. By
  \refrule{T-Next} we now obtain that
  $\Gamma,\Sigma\infers t_1'\Next e_2:\Task\tau$. \\

  \noindent\textbf{Case} $\userule{H-FirstAnd}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and
  $\Sigma\infers s$, \refrule{T-And} gives us that
  $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
  $\Gamma,\Sigma\infers t_2:\Task\tau_2$. By the induction hypothesis, we know
  that $t_1,s\handle{i}t_1',s'$ also preserves and thus
  $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Sigma\infers s'$. Therefore by
  \refrule{T-Next} we obtain
  $\Gamma,\Sigma\infers t_1'\And t_2:\Task(\tau_1\times\tau_2)$.\\

  \noindent\textbf{Case} $\userule{H-SecondAnd}$ \\
  \indent Given that
  $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and
  $\Sigma\infers s$, \refrule{T-And} gives us that
  $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
  $\Gamma,\Sigma\infers t_2:\Task\tau_2$. By the induction hypothesis, we know
  that $t_2,s\handle{i}t_2',s'$ also preserves and thus
  $\Gamma,\Sigma\infers t_2':\Task\tau_2$ and $\Sigma\infers s'$. Therefore by
  \refrule{T-Next} we obtain
  $\Gamma,\Sigma\infers t_1\And t_2':\Task(\tau_1\times\tau_2)$.\\

  \noindent\textbf{Case} $\userule{H-FirstOr}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$,
  \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and
  $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis we know that
  $t_1,s\handle{i}t_1',s'$ also preserves, and therefore
  $\Gamma,\Sigma\infers t_1':\Task\tau$ and $\Sigma\infers s'$. By
  \refrule{T-Or} we now obtain $\Gamma,\Sigma\infers t_1'\Or t_2:\Task\tau$.\\

  \noindent\textbf{Case} $\userule{H-SecondOr}$\\
  \indent  Given that
  $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$,
  \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and
  $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis we know that
  $t_2,s\handle{i}t_2',s'$ also preserves, and therefore
  $\Gamma,\Sigma\infers t_2':\Task\tau$ and $\Sigma\infers s'$. By T-Or we now
  obtain $\Gamma,\Sigma\infers t_1\Or t_2':\Task\tau$.

  % \noindent\textbf{Case} $\userule{H-Appoint}$ Given that
  % $\Gamma,\Sigma\infers u\At t:\Task\tau$, \refrule{T-Assign} gives us that
  % $\Gamma,\Sigma\infers t:\Task\tau$. By the induction hypothesis, we know that
  % $t,s\handle{i}t',s'$ also preserves, and therefore
  % $\Gamma,\Sigma\infers u\At t'$ and $\Sigma\infers s'$.
\end{proof}

% \subsubsection{Theorem~\ref{thmprogressnorm}}
%
% \begin{proof} We prove this Theorem by induction on $e$.\\
%
%   \noindent\textbf{Case} $\userule{S-Edit}$ Given that $\Gamma,\Sigma\infers \Edit v : \Task \tau$, let $e''=\Edit v'$, $s''=s$ and $i=v':\tau$. Then H-Edit gives $\Edit v,s\handle{v'}\Edit v',s$. \\
%
%
%   \noindent\textbf{Case} $\userule{S-Fill}$ Given that $\Gamma,\Sigma\infers \Enter \tau : \Task \tau$, let $e''=\Edit v$, $s''=s$ and $i=v:\tau$. Then H-Fill gives $\Enter \tau,s\handle{v}\Edit v,s$. \\
%
%   \noindent\textbf{Case} $\userule{S-Update}$ Given that $\Gamma,\Sigma\infers \Update l : \Task \tau$, let $e''=\Update l$, $s''=s[l\mapsto v]$ and $i=v:\tau$. Then H-Update gives $\Update l ,s\handle{v}\Update l,s[l\mapsto v]$. \\
%
%   \noindent\textbf{Case} $\userule{S-Fail}$ Since $\Failing(\Fail)=\True$, the theorem holds in this case. \\
%
%   \noindent\textbf{Case} $\userule{S-Xor}$ \todo{Here we have an issue with the defintion of $\Failing$ and the sideconditions of external choice. These definitions should concur, so we can finish the proof, but $\Failing$ is a static property}\\
%
%   \noindent\textbf{Case} $\userule{S-Eval}$ By induction hypothesis we have that there exists an $e'''$, $s'''$ and $i$ such that $e'',s''\handle{i}e''',s'''$.\\
%
%   \noindent\textbf{Case} $\userule{S-ThenStay}$ By the induction hyposthesis we have that there exists an $e''$, $s''$ and $i$ such that $t_1',s'\handle{i}e'',s''$. Then by H-Pass, we know that we can apply the same $i$ to obtain $t_1\Then e_2,s'\handle{i} e''\Then e,s''$.\\
%
%   \noindent\textbf{Case} $\userule{S-ThenFail}$ By the induction hyposthesis we have that there exists an $e''$, $s''$ and $i$ such that $t_1',s'\handle{i}e'',s''$. Then by H-Pass, we know that we can apply the same $i$ to obtain $t_1\Then e_2,s'\handle{i} e''\Then e,s''$.\\
%
%   \noindent\textbf{Case} $\userule{S-ThenCont}$ By the induction hypothesis, we have that there exists an $e''$, $s''''$ and $i$ such that $t_2',s''\handle{i}e'',s''''$.\\
%
%   \noindent\textbf{Case} $\userule{S-OrLeft}$ By the induction hypothesis, we have that there exists an $e''$, $s'$ and $i$ such that $t_1',s\handle{i}e'',s''$.\\
%
%   \noindent\textbf{Case} $\userule{S-OrRight}$ By the induction hypothesis, we have that there exists an $e''$, $s'$ and $i$ such that $t_2',s\handle{i}e'',s''$.\\
%
%   \noindent\textbf{Case} $\userule{S-OrNone}$ \todo{   }\\
%
%   \noindent\textbf{Case} $\userule{S-Next}$ By the induction hypothesis, we have that there exists an $e''$, $s'$ and $i$ such that $t_1',s'\handle{i}e'',s''$. Then by the H-PassNext rule, we know that we can apply the same $i$ to obtain $t_1'\Next e_2,s'\handle{i}e'',s''$.\\
%
%   \noindent\textbf{Case} $\userule{S-And}$ By the induction hypothesis, we have that there exists an $e''$, $s'$ and $i$ such that $t_1',s'\handle{i}e'',s''$. Then by the H-FirstOr rule, we know that we can apply $\First i$ to obtain $t_1'\And t_2',s'\handle{\First i} t_1''\And t_2',s''$
%
% \end{proof}

% \subsubsection{Lemma~\ref{lem:stride-does-not-eval}}
%
% \begin{proof}
%   We prove Lemma~\ref{lem:stride-does-not-eval} by induction on $e$. First we
%   apply evaluation, and we then find ourselves in one of the cases below.\\
%
%   \noindent\textbf{Case} $\userule{S-ThenStay}$\\
%   and $\userule{S-ThenFail}$ The \refrule{E-Then} rule gives us that evaluation
%   has been applied to $t_1$. We therefore can apply the induction hypothesis,
%   from which we obtain that $t_1',s'\evaluate t_1'',s''$ then $t_1'=t_1''$ and
%   $s'=s''$. This, together with the \refrule{E-Then} rule, proves this case.\\
%
%   \noindent\textbf{Case} $\userule{S-ThenCont}$ By application of the induction
%   hypothesis, we obtain that $t_2',s'''\evaluate t_2'',s''''$ then $t_2'=t_2''$
%   and $s'''=s''''$, which immediately proves this case.\\
%
%   \noindent\textbf{Case} $\userule{S-OrLeft}$ The \refrule{E-Or} rule gives us
%   that evaluation has been applied to $t_1$. We therefore can apply the
%   induction hypothesis, from which we obtain that $t_1',s'\evaluate t_1'',s''$
%   then $t_1'=t_1''$ and $s'=s''$, which immediately proves this case.\\
%
%   \noindent\textbf{Case} $\userule{S-OrRight}$The \refrule{E-Or} rule gives us
%   that evaluation has been applied to $t_2$. We therefore can apply the
%   induction hypothesis, from which we obtain that $t_2',s'\evaluate t_2'',s''$
%   then $t_2'=t_2''$ and $s'=s''$, which immediately proves this case.\\
%
%   \noindent\textbf{Case} $\userule{S-OrNone}$ The \refrule{E-Or} rule gives us
%   that evaluation has been applied to both $t_1$ and $t_1$. We assumed that the
%   states do not change, so we know that $s=s'=s''$. We can therefore safely
%   apply the induction hypothesis twice, to obtain that
%   $t_1',s''\evaluate t_1'',s'''$ then $t_1'=t_1''$ and $s''=s'''$, and
%   $t_2',s'''\evaluate t_2'',s''''$ then $t_2'=t_2''$ and $s'''=s''''$. This,
%   together with the \refrule{E-Or} rule proves this case.\\
%
%   \noindent\textbf{Case} $\userule{S-Edit}$ by definiton, $\Edit v$ is a value
%   with respect to evaluation.\\
%
%   \noindent\textbf{Case} $\userule{S-Fill}$ by definition, $\Enter \tau$ is a
%   value with respect to evaluation.\\
%
%   \noindent\textbf{Case} $\userule{S-Update}$ by definiton, $\Update l$ is a
%   value with respect to evaluation.\\
%
%   \noindent\textbf{Case} $\userule{S-Fail}$ by definiton, $\Fail$ is a value
%   with respect to evaluation.\\
%
%   \noindent\textbf{Case} $\userule{S-Xor}$ by definition, $e_1\Xor e_2$ is a
%   value with respect to evaluation.\\
%
%   \noindent\textbf{Case} $\userule{S-Next}$ The \refrule{E-Next} rule gives us
%   that evaluation has been applied to $t_1$. We therefore can apply the
%   induction hypothesis, from which we obtain that $t_1',s'\evaluate t_1'',s''$
%   then $t_1'=t_1''$ and $s'=s''$. This, together with the \refrule{E-Next} rule,
%   proves this case.\\
%
%   \noindent\textbf{Case} $\userule{S-And}$ The \refrule{E-And} rule gives us
%   that evaluation has been applied to both $t_1$ and $t_1$. We assumed that the
%   states do not change, so we know that $s=s'=s''$. We can therefore safely
%   apply the induction hypothesis twice, to obtain that
%   $t_1',s''\evaluate t_1'',s'''$ then $t_1'=t_1''$ and $s''=s'''$, and
%   $t_2',s'''\evaluate t_2'',s''''$ then $t_2'=t_2''$ and $s'''=s''''$. This,
%   together with the \refrule{E-And} rule proves this case.\\
%
%   \noindent\textbf{Case} $\userule{S-Appoint}$ The \refrule{E-Appoint} rule
%   gives us that evaluation has been applied to $t$. We therefore can apply the
%   induction hypothesis, from which we obtain that $t',s'\evaluate t'',s''$ then
%   $t'=t''$ and $s'=s''$. This, together with the \refrule{E-Appoint} rule,
%   proves this case.
%
% \end{proof}

\subsubsection{Theorem~\ref{thm:failing}}

\begin{proof}

  We prove Theorem~\ref{thm:failing} by induction on $e'$.\\

  \noindent\textbf{Case} $e = \Fail$\\
  $\Failing(\Fail,s)=\True$, and there is no handling rule that applies to fail.\\

  \noindent\textbf{Case} $e = \Edit v$\\
   Given that $\Gamma,\Sigma\infers\Edit v:\Task\tau$, $\Failing(\Edit v,s)=\False$, and there exists an input $i$, namely $v':\tau$.\\

  \noindent\textbf{Case} $e = \Enter \tau$\\
  $\Failing(\Enter \tau)=\False$, and there exists an in put i, namely $v:\tau$.\\

  \noindent\textbf{Case} $e = \Update l$\\
  Given that $\Gamma,\Sigma\infers\Update l:\Task\tau$, $\Failing(\Update l,s)=\False$, and there exists an input $i$, namely $v:\tau$.\\

  \noindent\textbf{Case} $e = t_1\Then e_2$\\
   $\Failing(t_1\Then e_2,s)=\Failing(t_1,s)$. If there exitsts an $i$ for $t_1$, then this $i$ also applies to $t_1\Then e_2$. This case therefore holds by the induction hypothesis.\\

  \noindent\textbf{Case} $e = t_1\Next e_2$\\
   $\Failing(t_1\Next e_2,s)=\Failing(t_1,s)$. If there exitsts an $i$ for $t_1$, then this $i$ also applies to $t_1\Next e_2$. This case therefore holds by the induction hypothesis.\\

  \noindent\textbf{Case} $e = e_1\Xor e_2$\\
   We normalise the two expressions first, $e_1,s\stride t_1,s'$, $e_2,s\stride t_2,s'$ and we can then be in two situations. One, we can have that $\Failing(t_1,s')$ and $\Failing(t_2,s')$ are both true. If that is so, then by definiton, we have both $\Failing(e_1\Xor e_2,s)$ and no rule of the hanlding semantics applies, and therefore there exists no input for this case.\\
                                           Or we are in the situation where one or both of the two sub expressions does not fail. In that case, we know that $\Failing(e_1\Xor e_2,s)$ does not hold, and that at least one of the handling rules applies. Therefore, there must be an input $i$, namely $\Left$, $\Right$ or both.\\

  \noindent\textbf{Case} $e = t_1\And t_2$\\
   We can again find ourselfs in one of two situations. In the first case, both sub expressions fail, $\Failing(t_1,s)$ and $\Failing(t_2,s)$. In that case, we know that $\Failing(t_1\And t_2,s)$ also fails by definition. By the induction hypothesis, we know that for both $t_1$ and $t_2$ there is no input that can be handled. Since the only two rules for $\And$ that handle input just pass this input on to one of the two expressions, we know that indeed no $i$ applies.\\
                                           In the case that one or both sub expressions do not fail, then by definition $t_1\And t_2$ not failing under $s$. Again by induction hypothesis, we know that for one or both of the expressions, there exits an $i$ that can be handled. Then by H-FirstAnd and H-SecondAnd, we know that we can pass this $i$, by prefixing it with either $\First$ or $\Second$.\\

  \noindent\textbf{Case} $e = t_1\Or t_2$\\
  We can again find ourselfs in one of two situations. In the first case, both sub expressions fail, $\Failing(t_1,s)$ and $\Failing(t_2,s)$. In that case, we know that $\Failing(t_1\Or t_2,s)$ also fails by definition. By the induction hypothesis, we know that for both $t_1$ and $t_2$ there is no input that can be handled. Since the only two rules for $\Or$ that handle input just pass this input on to one of the two expressions, we know that indeed no $i$ applies.\\
                                           In the case that one or both sub expressions do not fail, then by definition $t_1\Or t_2$ not failing under $s$. Again by induction hypothesis, we know that for one or both of the expressions, there exits an $i$ that can be handled. Then by H-FirstOr and H-SecondOr, we know that we can pass this $i$, by prefixing it with either $\First$ or $\Second$.

  % \noindent\textbf{Case} $e = u \At t$\\
  %  This case follows directly from applying the induction hypothesis, since $\Failing(u \At t,s)=\Failing(t,s)$ and any input that applies to $t$ also applies to $u \At t$.\\
\end{proof}

\subsubsection{Theorem~\ref{thm:safety-i}}
\begin{proof}

  \noindent\textbf{Case} $e=\Edit v:\Task\tau, i= v':\tau$\\
   Given that $\userule{H-Change}$, we have by definition that $\Inputs(\Edit v:\Task\tau,s) = \set{v':\tau,\Empty}$, which includes $v':\tau$.

  \noindent\textbf{Case} $e=\Edit v:\Task\tau, i= \Empty$\\
   Given that $\userule{H-Empty}$, we have by definition that $\Inputs(\Edit v:\Task\tau,s) = \set{v':\tau,\Empty}$, which includes $\Empty$.\\

  \noindent\textbf{Case} $e=\Enter \tau, i= v':\tau$\\
   Given that $\userule{H-Fill}$, we have by definiton that $\Inputs(\Enter \tau,s) = \set{v':\tau }$, which includes $v':\tau$.\\

  \noindent\textbf{Case} $e=\Update l:\Task\tau, i= v':\tau$\\
   Given that $\userule{H-Update}$, we have by definiton that $\Inputs(\Update l:\Task\tau,s)= \set{v':\tau }$, which includes $v':\tau$.\\

  \noindent\textbf{Case} $e=t_1\Xor t_2 , i= \Left$\\
   Given that $\userule{H-PickLeft}$, we have by definition that $\Inputs(t_1\Xor t_2,s) = \set{ \Left,\Right}$, which includes $\Left$.\\

  \noindent\textbf{Case} $e=t_1\Xor t_2 , i= \Right$\\
   Given that $\userule{H-PickRight}$, we have by definition that $\Inputs(t_1\Xor t_2,s) = \set{ \Left,\Right}$, which includes $\Right$.\\

  \noindent\textbf{Case} $e=t_1\Next e_2 , i= \Continue $\\
   Given that $\userule{H-Next}$, we have by definition that $\Inputs(t_1\Next e_2,s) = \Inputs(t_1,s) \cup \set{\Continue\mid \Value{(t_1,s)} = v_1 \wedge \neg\Failing{(e_2 v_1,s\stride)}}$. If the H-Next rule applies, this means that the conditions $\Value{(t_1,s)} = v_1 \wedge \neg\Failing{(e_2 v_1,s\stride)}$ are fulfilled, and therefore $\Continue$ is contained. \\

  \noindent\textbf{Case} $e=t_1\Next e_2 , i\neq\Continue$\\
   Given that $\userule{H-PassNext}$, we have by definition that $\Inputs(t_1\Next e_2,s) = \Inputs(t_1,s) \cup \set{\Continue\mid \Value{(t_1,s)} = v_1 \wedge \neg\Failing{(e_2 v_1,s\stride)}}$. By the induction hypthesis, we have that $i\in \Inputs(t_1,s)$, and by definiton of $\Inputs$, $i$ is therefore also included in this case.\\

  \noindent\textbf{Case} $e=t_1\Then e_2 , i$\\
   Given that $\userule{H-PassThen}$, we have by definition that $\Inputs(t_1\Then e_2,s)= \Inputs(t_1,s)$. By the induction hypthesis, we have that $i\in \Inputs(t_1,s)$, and by definiton of $\Inputs$, $i$ is therefore also included in this case.\\

  \noindent\textbf{Case} $e=t_1\And t_2 , i=\First i$\\
   Given that $\userule{H-FirstAnd}$ we have by definition that $\Inputs(t_1\And t_2,s) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$. By the induction hypothesis, we have that $i\in\Inputs(t_1,s)$, and by definition of $\Inputs$, $\First i$ is therefore als included in this case.\\

  \noindent\textbf{Case} $e=t_1\And t_2 , i=\Second i$\\
   Given that $\userule{H-SecondAnd}$ we have by definition that $\Inputs(t_1\And t_2) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$. By the induction hypothesis, we have that $i\in\Inputs(t_2,s)$, and by definition of $\Inputs$, $\Second i$ is therefore als included in this case.\\

  \noindent\textbf{Case} $e=t_1\Or t_2 , i=\First i$\\
   Given that $\userule{H-FirstOr}$ we have by definition that $\Inputs(t_1\Or t_2,s) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$. By the induction hypothesis, we have that $i\in\Inputs(t_1,s)$, and by definition of $\Inputs$, $\First i$ is therefore als included in this case.\\

  \noindent\textbf{Case} $e=t_1\Or t_2 , i=\Second i$\\
   Given that $\userule{H-FirstOr}$ we have by definition that $\Inputs(t_1\Or t_2,s) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$. By the induction hypothesis, we have that $i\in\Inputs(t_2,s)$, and by definition of $\Inputs$, $\Second i$ is therefore als included in this case.
  %
  % \noindent\textbf{Case} $e=u\At t, i$\\
  %  Given that $\userule{H-Appoint}$, we can directly apply the induction hypothesis to prove this case.\\
\end{proof}

### 2019-02-06 sections/abstract.tex
update magic comments

% !TEX root=../icfp2019.tex

Task-Oriented Programming (\TOP) is a programming paradigm that focusses on modelling real world collaborations between people.
It prescribes a declarative programming style to specify multi-user workflows.
Workflows can be higher-order.
They communicate through typed values on a local or global level.
Such specifications can be turned into interactive applications for different platforms,
supporting collaboration during execution.

In this paper we decompose the rich features of \TOP into elementary language elements,
which makes them suitable for formal treatment.
We use the simply typed lambda calculus, extended with pairs and references, as a base language.
On top of this language, we develop \TOPHAT (TopHat), a calculus for modular interactive workflows.
With \TOPHAT we prepare a way to formally reason about \TOP systems.

We describe \TOPHAT by means of a layered semantics.
These layers consist of multiple big step evaluations on expressions,
and two labelled transition systems, handling user inputs.
We show some interesting properties of this machinery.
This approach allows for comparison with other work in the field.
We place \TOPHAT in perspective with the process calculus \CSP.

### 2019-02-06 sections/comparison.tex
update magic comments

% !TEX root=../icfp2019.tex

\section{Intuition}
\label{sec:comparison}

This section gives an overview of the essential concepts of \TOP and how they are represented in \TOPHAT.
While \TOP is focussed on collaboration, it gives rise to concepts like communication and multiprogramming.
% These concepts are: communication with the environment, communication between subsystems, sequential composition, concurrency, and synchronisation.
These concepts can also be found in process algebras, which raises the question how \TOPHAT and process algebras relate.
While introducing the concepts, we compare them with Hoare's Communicating Sequential Processes (\CSP).
The comparison is based on Hoare's book \cite{books/Hoare85CSP}.
We chose \CSP as an example to stand for the numerous process algebras in existence, which, as far as this section is concerned, are reasonably similar.

We provide comparisons based on two aspects.
First, we compare the scope of the languages, as intended by the authors.
Second, we focus on features common to \TOP and \CSP, and study their realisation.
When we point out differences, we do not argue that the different realisations are incompatible.
As a matter of fact the primitives can certainly be expressed in terms of each other.
Instead, we point out how the systems emphasise certain points of view by choosing different building blocks to express the same concepts.

\paragraph{Scope}

The central objective of \TOP is to coordinate collaboration.
Therefore it has combinators to express ways in which people can work together to reach a common goal.
Tasks can be executed after each other, at the same time, or conditionally.
This motivates the task combinators for steps, parallel combination, and choice.
To employ these patterns of collaboration, users need to communicate.
This motivates the different forms of communication: with the environment, along control flow, and across control flow.
\TOPHAT is a language to specify workflows.

The central objective of \CSP is to model the patterns of behaviour of processes.
These patterns of behaviour manifest themselves in sequences of actions that processes can engage in.
\CSP is a language to specify synchronisation patterns.

\begin{quote}
\noindent In \TOP, communication is a means to an end. \\
In \CSP, communication is the central matter of discourse.
\end{quote}

%%NOTE: Focus on what TOP is good at --TS
% \CSP has a formal semantics that allows various kinds of correctness proofs, including equality of processes, and adherence to a specification.
% This allows applications in program correctness, proofs of deadlock freedom, liveness, or verification of protocols.
%
% \TOP focuses less on formal correctness, and more on practical applicability.
The big difference between \CSP and \TOP is the intended level of abstraction.
While \CSP is used to explicitly model low-level details about communication, \TOPHAT wants to hide details and provide high-level communication mechanisms to the programmer.
It wants to be a calculus with intuitive semantics that facilitates communication between programmers and domain experts.
Such programs are supposed to hide implementation details from domain experts while containing enough information to allow automatic generation of executable applications, including user interfaces.

%%NOTE: Already said in section intro --TS
% Despite the different origins, \TOP and \CSP feature concepts that serve similar purposes.

\subsection{Communication}

Both \TOP and \CSP have two sorts of communication: with the environment, and between subsystems.
In both languages, communication with the \emph{environment} is input and output.
They block evaluation of programs until the environment sends or receives events.
In both languages, communication \emph{between subsystems} does not block.
They evaluate programs as far as possible until further external input is required.

\paragraph{Communication with the environment}

In \TOPHAT, communication with the environment is represented by editors.
A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive any change event with a value of type $\tau$ and additionally the special event $\Empty$ that empties the editor.
The environment can always inspect the current value of an editor.

Editors have no notion of continuation.
Their sole purpose is to interact with the environment, retaining the last value that has been sent to them.
%
They can be used for both input and output.
The interpretation of whether an editor represents input or output is left to the reader.
An empty editor is most commonly interpreted as a prompt to input data, while a filled editor can be seen either as outputting its value, or as an input that comes with a default value.

In \CSP, communication with the environment is represented by prefixing.
The process $(a \to P)$ can engage in action $a$, after which it continues as process $P$.
%%NOTE: Not that important I think, we don't use it in the rest of the text. --TS
% In general, if $B$ is a set of events and $P(x)$ is an expression that evaluates to a process for each $x \in B$, the process $(x:B \to P(x))$ can engage in any one of the actions in $B$, after which it continues as the process determined by $P(x)$.
% Hoare does not specify the language in which $P(x)$ is to be expressed.
% He seems to permit any kind of mathematical formula, or any programming language.
%
Prefixing has no direction of sending and receiving.
Hoare uses the neutral phrase \emph{engaging in an action}.
The interpretation of whether an action stands for input or output is left to the reader.
For example, a vending machine $P = (\textit{coin} \to \textit{choc} \to P)$ is to be interpreted as taking a coin as input and producing chocolate as output.

Sending and receiving of values is modelled by giving additional structure to the names of actions.
For example, an action name could be $\textit{in}.5$, which can be interpreted as inputting the value 5.
%
In \TOP, event values are typed to match the type of the receiving editor.
In \CSP, action names are essentially strings, subject to implicit restrictions.

\begin{example}[Vending machine]

This example demonstrates external communication and choice.
The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins.
\begin{equation*}
  V = \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
\end{equation*}

A vending machine in \TOPHAT that is close in spirit to the one above looks as follows.
\begin{TASK}
  enter Int >>? \n. if n == 1 then edit "small"
    else if n == 2 then edit "large" else fail
\end{TASK}
This machine does not exhibit exactly the same behaviour as the one in \CSP, nor is it intended to.
Instead, it just demonstrates how the same choice
%%NOTE: Problem is already stated --TS
% of buying a small biscuit for one coin or a large biscuit for two coins
can be expressed in \TOP.
The task asks the user to enter an amount of money.
This editor stands for a coin slot in a real machine that freely accepts and returns coins.
There is a continue button that is initially disabled.
When the user has inserted exactly 1 or 2 coins, the continue button becomes enabled.
When the user presses the continue button, the machine dispenses either a large or a small biscuit, depending on the amount of money.

\end{example}

\paragraph{Communication between subsystems}

\TOP has two methods for internal communication, both different from its meth\-od for external communication.
The step combinator represents data flow that follows control flow, and shared data represents data flow across control flow.

The step combinators in $t \Then e$ and $t \Next e$ pass the value of $t$ to the continuation $e$.
% The intuition about sequential composition is to first do one thing, and once it is done, do some other thing.
% In \TOP, tasks never terminate.
% Nonetheless, the notion of sequential composition exists in two forms: the external and the internal step.
They both act like $t$ as long as the step is guarded.
Once it becomes unguarded, the task continues as $e\ v$, where $v$ is the value of $t$.
The task $t \Next e$ requires an input event $\Continue$ in addition to the step being unguarded. % in order to step.
% \CSP's semicolon is like an internal step $t \Then e$ in \TOP where $t$ uses $\Unit$ to signal when it is done, and $e$ never fails.

Shared data sources are assignable references, whose changes in value are immediately visible to all tasks interested in them.
The semantics of \TOPHAT requires all updates to shared data and all enabled internal steps to be processed before any further communication with the environment can take place.
This introduces the possibility of diverging computations, where a cyclic dependency between shared data causes internal steps to be taken continuously without ever being able to engage in external communication.

\CSP uses prefixing for both internal and external communication.
Prefixing becomes internal by using \emph{concealment}.
Processes must take any concealed steps as soon as possible, before any further communication with the environment can take place.
This means concealed actions have priority over exposed ones.
This introduces the possibility of diverging computations, that is processes which continuously take internal steps without ever being able to engage in external communication.

% \paragraph{Sequential composition}

% In \CSP, sequential composition is represented by the semicolon combinator together with the special action $\checkmark$, called \emph{success}.
% We require that if a process engages in $\checkmark$ at all, it must be its single last action.
% When that happens, we say the process terminates successfully.
% The process $P;Q$ behaves like $P$ until $P$ terminates successfully, after which it continues as $Q$.
% If $P$ never terminates successfully, neither does $P;Q$

\subsection{Multiprogramming}

Both \TOP and \CSP are models of \emph{multiprogramming}, as opposed to \emph{multiprocessing}.
This means that one processor evaluates multiple programs in an interleaving fashion.
% instead of multiple processors.
Multiprogramming involves three aspects: concurrency, synchronisation, and nondeterminism.

\paragraph{Concurrency}

Concurrency is the source of parallelism: two things are done at the same time.
%
In \TOP, the parallel combinator $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
All events sent to the combined task are interleaved into a serial stream of inputs.
The tasks are truly independent of each other, all interleavings are permitted.
Events to $t_1$ and $t_2$ must be prefixed by $\First$(irst) and $\Second$(econd) respectively.
This unambiguously renames the events, removing any source of nondeterminism.

In \CSP, there are two different combinators for parallel composition: parallel and interleave.
The parallel process $P \parallel Q$ can take $P$-steps and $Q$-steps in arbitrary interleaving for actions unique to $P$ and $Q$.
Actions in the alphabets of both $P$ and $Q$ must be taken in synchronisation, so if one process wants to take such a step, it must wait until the other is ready to do so.
Synchronised steps are the only occasion where two processes actually do something at the same time.

The interleaving operator $P \mid\mid\mid Q$ does not synchronise any actions of $P$ and $Q$.
All interleavings are permitted.
Actions that occur in both alphabets are nondeterministically taken by either $P$ or $Q$.

\paragraph{Synchronisation}

Synchronisation means that an agent has to pause execution until some condition is met.
The need for synchronisation arises quite naturally in countless variations in situations involving concurrency.
For multiprogramming, synchronisation always means that some of the possible interleavings are forbidden.

For example, the mutual exclusion problem for two parallel processes $a_p \to b_p \to \text{STOP} $ and $a_q \to b_q \to \text{STOP}$ can be stated as:
In no interleaving should there be two adjacent $a$'s.
%
The mutual exclusion problem for two parallel tasks can be stated as follows.
Let \TS{inc} and \TS{dec} be tasks that increment and decrement some shared counter.
Then, in \TS{(inc x >>? dec) <&> (dec x >>? inc)}, the shared counter should at no point be greater than 1.
This means that in no interleaving should there be two adjacent \TS{inc}'s.

The basic method for synchronisation in \TOP is built into the step combinator.
The task $t \Next e$ can only continue execution when two conditions are met:
Task $t$ must have a value $v$, and $e\ v$ must not evaluate to $\Fail$.
Programmers can encode arbitrary conditions in $e\ v$, which are evaluated atomically between interaction steps.
This allows a variety of synchronisation problems to be solved in an intuitive and straight-forward manner.

The basic method for synchronisation in \CSP is synchronised prefixing.
If some action $a$ is in the alphabets of both $P$ and $Q$, then their parallel composition $P \parallel Q$ can take an $a$-step only if both $P$ and $Q$ are ready to take an $a$-step, in which case they both take this step simultaneously.
Both have to wait for the condition that the other process is ready to take the step.
All synchronisation problems in \CSP must be solved by employing synchronised steps in some form.

\begin{example}[Cigarette smokers]

The cigarette smokers problem \cite{books/Downey08LBOS} is a surprisingly tricky synchronisation problem.
We study it here because it demonstrates the capabilities of guarded steps.
% The problem requires waiting for two conditions, waking up only if both conditions are satisfied.
The problem is stated as follows.
In order to smoke a cigarette, three ingredients are required: tobacco, paper, and a match.
There are three smokers, each having one of the ingredients and requiring the other two.
There is an agent that randomly provides two of the ingredients.

Downey models availability of the ingredients using a semaphore for each ingredient.
The agent randomly signals two of the three semaphores.
% The problem asks writing code for the smokers so that the smoker who requires the two signalled semaphores wakes up and proceeds with smoking.
%
% A naïve attempt where each smoker first waits for one and then the other semaphore can lead to deadlock.
% The following pseudo-code is a non-solution.
% Let \emph{paper}, \emph{tobacco} and \emph{match} be semaphores, initialised to 0.
% \begin{TASK}
%   tobaccoSmoker = wait(match); wait(paper); smoke();
%   paperSmoker = wait(tobacco); wait(match); smoke();
%   matchSmoker = wait(tobacco); wait(paper); smoke();
% \end{TASK}
% To see what can go wrong, imagine the three smokers running in parallel.
% Let the agent signal the two semaphores \emph{tobacco} and \emph{paper}.
% One of \emph{paperSmoker} and \emph{matchSmoker} can take their first step, but if \emph{paperSmoker} takes the step, the system deadlocks.
%
The solution proposed by Downey involves an additional mutex, three additional semaphores, three additional threads called \emph{pushers}, and three regular Boolean variables.
The job of the pushers is to record availability of their ingredient in their Boolean variable, and check availability of other resources, waking the correct smoker when appropriate.
The details of the solution are not important here.

What is important is that the implementation of what is essentially waiting for two semaphores requires a substantial amount of additional synchronisation, together with non-trivial conditional statements.
%
A similar effort as in Downey's solution must be made in order to solve the problem using synchronised actions in \CSP.
This is because in \CSP, processes can only synchronise on one action at a time.

\TOP allows a simple solution to this problem, using guarded steps.
Steps can be guarded with arbitrary expressions, and the parallel combinator can be used to watch two shared editors ($\Update$) at the same time.
Let \TS{match}, \TS{paper}, and \TS{tobacco} be references to Booleans.
The smokers are defined as follows.
\begin{TASK}
  let continue = \<<x,y>>. if x /\ y then smoke else fail in
  let tobaccoSmoker = (update match <&> update paper) >>? continue in
  let paperSmoker = (update tobacco <&> update match) >>? continue in
  let matchSmoker = (update tobacco <&> update paper) >>? continue
\end{TASK}
When the agent supplies two of the ingredients by setting the respective shares to \TS{True}, only the step of the smoker that waits for those becomes enabled.

\end{example}

\paragraph{Nondeterminism}

Nondeterminism means that a system can react to the same input in different ways, the choice of which can not be influenced by the environment.

\CSP has an operator to explicitly introduce nondeterminism.
Furthermore, nondeterminism can arise from the combination of some other combinators, for example choice and concealment.
%
Hoare states that nondeterminism is only useful for \emph{specifying} processes, never for \emph{implementing} them \cite{books/Hoare85CSP}.
\CSP can be used for both.
A process with nondeterministic behaviour must always be regarded as a specification, while a deterministic process can be seen as specification or implementation.

\TOP is meant solely for implementation and does not have any form of nondeterminism.
Input events for parallel tasks are disambiguated, internal steps have a well-defined evaluation order, and internal choice is left-biased.

\paragraph{Conclusion}

There are several concepts that exist in both languages, like concurrency, or sending values between concurrent agents.
In \TOP however they are meant to be used on a higher level of abstraction, which allows concise declarative description of patterns of collaboration, without worrying about the implementation details of communication.
By focussing on collaboration, \TOP and \TOPHAT lead to high-level specifications closer to real-world workflows which, at the same time, can be used to generate distributive, multi-user applications to support such workflows.

### 2019-02-06 sections/conclusions.tex
update magic comments

% !TEX root=../icfp2019.tex

\section{Conclusion}

\label{sec:conclusions}

In this paper we have developed a domain-specific calculus for declarative interactive workflows.
We have identified, what we believe to be, the essence of Task-Oriented Programming, and provided corresponding operators in our language.
We have compared our language with \CSP to point out differences and similarities.
Finally, we have proven type safety for our language.

\paragraph{Future work}

There are a couple of ways in which we would like to continue this line of work.

The addition of read-only shared editors would allow watching references without being able to change them.
Some programs could be expressed more faithfully, where we for now require that users do not modify certain shares.

One goal of the formalisation of a language is the ability to reason about programs.
In this paper we reason about the language itself, but it would be nice to prove properties about individual programs.
We would like to prove whether certain programs are equal, for example to show that the monad laws hold for our step combinator.
This requires a notion of equality, which in the presence of side effects most certainly needs some form of coalgebraic input-output conformance.

We have implemented the reduction semantics of our language in Idris, whose type system could aid in a formalisation of those proofs.

Another form of reasoning about programs is static analysis.
Klinik et al.~\cite{conf/ifl/KlinikJP17} have developed a cost analysis for tasks that require resources in order to be executed.
% Put this line back in for final version of paper: In previous work we have developed a cost analysis for tasks that require resources in order to be executed \cite{conf/ifl/KlinikJP17}.
This analysis was developed for a simpler task calculus, and could be brought over to the one developed here.

% Put this line back in for final version of paper: In other previous work, we have looked at building a generic feedback system for rule-based problems \cite{UUCS2017013}.
Naus et al.~\cite{UUCS2017013} have looked at building a generic feedback system for rule-based problems.
A workflow system typically is rule based, as outlined in this work.
It would be interesting to fit the generic feedback system to \TOPHAT in order to support end-users working in applications developed in this calculus.

Additionally, we would like to develop visualisations for \TOPHAT language constructs.
An assistive development environment integrating these visualisations and the presented textual language
would aid domain experts to model workflows in a more accessible manner.

### 2019-02-06 sections/contributions.tex
update magic comments

% !TEX root=../icfp2019.tex

% \paragraph{Contributions}

In this paper, we set out to develop a formal basis for \TOP, called \TOPHAT (TopHat).
Our goal is to specify the basic building blocks of declarative workflows.
We clearly separate them from general purpose programming concepts.
This allows us to prove properties like progress and type preservation.
\TOPHAT should be powerful enough to model complicated real-world situations,
without losing opportunities for formal reasoning.

We present a calculus for modelling declarative workflows, embedded in a simply typed lambda-calculus.
We develop an operational semantics that is driven by user inputs.
The semantics of the task language is clearly separated from the semantics of the underlying host language.
Along the semantics, we present the following semantic observations on tasks: the current value, whether a term is stuck, the current user interface, and the accepted inputs.
%%NOTE: iTasks things are moved to related work
% All these observations exist in iTasks, albeit combined into one big type.
By separating them in the style of a deep embedding \cite{conf/cefp/Gibbons13}, we make explicit where and where not, they depend on each other.

Additionally, we implemented the whole semantic system in the dependently typed programming language Idris \cite{journals/jfp/Brady13},
to gain confidence for our system and check practical applicability.
Furthermore we compare \TOPHAT  with the process algebra \CSP, to show how their concepts, like communication, and sequential and parallel composition, relate.
We show that, where process algebras focus on the \emph{coordination} of processes and the \emph{communication} between them,
\TOP, and therefore \TOPHAT, focus on supporting \emph{tasks} and model ways to \emph{collaborate} on them.

### 2019-02-06 sections/example.tex
update magic comments

% !TEX root=../icfp2019.tex

\section{Example}
\label{sec:example}

In this section we develop a small example program to demonstrate the capabilities of \TOPHAT.
The example is a small flight booking system.
It demonstrates communication on all three levels: with the environment, across control flow, and alongside it.
Also, it shows synchronisation and input validation.

The requirements of the application are as follows.
\begin{enumerate*}
  \item A user has to enter a list of passengers for which to book tickets.
  \item At least one of these passengers has to be an adult.
  \item After a valid list of passengers has been entered, the user has to pick seats.
  \item Only free seats may be picked.
  \item Every passenger must have exactly one seat.
  \item Multiple users should be able to book tickets at the same time.
\end{enumerate*}
For this example we assume that the host language has lists and four functions on them: all, any, intersect, and difference.
%%NOTE: After Peter's comments symbols are not clear --TS
We also make use of string equality ($\equiv$), dereferencing ($!$), reference assignment ($:=$), and expression sequencing ($;$).
For brevity, we omit the type annotations of variable bindings.

\begin{example}[Flight booking]
\label{exm:flight-booking}

We start by defining some type aliases.
A passenger is a pair with name and age.
A seat is a pair with a row number and a seat letter.
\begin{TASK}
  type Passenger = String * Int
  type Seat = Int * String
\end{TASK}

Choosing seats requires reading and updating shared information.
The list of free seats is stored in a reference. % called \TS{freeSeats}.
\begin{TASK}
  let freeSeats = ref [<<1,"A">>, <<1,"B">>, <<1,"C">>, ...]
\end{TASK}

Now we develop our workflow in a top down way.
A flight booking starts with an interactive task \TS{enter (List Passenger)}
where users can enter a list of passengers.
A passenger is valid if the name is not empty and the age is at least 0.
A list of passengers is valid if each passenger is valid, and at least one of the passengers is an adult.
When the user has entered a valid list of passengers, the step after \TS{>>?} becomes enabled,
and the user can proceed to picking seats.
\begin{TASK}
  let valid = \p. not (fst p == "") /\ snd p >= 0 in
  let adult = \p. snd p >= 18 in
  let allValid = \ps. all valid ps /\ any adult ps in
  let bookFlight = enter (List Passenger) >>? \ps.
    if allValid ps then chooseSeats ps else fail
\end{TASK}
A selection of seats is correct if every entered seat is free.
\begin{TASK}
  let correct = \ss. intersect ss !freeSeats == ss in
  let chooseSeats = \ps. enter (List Seat) >>? \ss.
    if correct ss /\ length ps == length ss
      then confirmBooking ps ss else fail
\end{TASK}
The function \TS{confirmBooking} removes the picked seats from the shared list of free seats,
and displays the end result using the \TS{edit}-construct.
\begin{TASK}
  let confirmBooking = \ps. \ss.
    freeSeats := difference !freeSeats ss; edit <<ps, ss>>
\end{TASK}
% It uses a function \TS{difference}, which removes all elements from the second list from the first list.

The main task starts three \TS{bookFlight} tasks,
which could be performed by three different users in parallel
\begin{TASK}
  bookFlight <&> bookFlight <&> bookFlight
\end{TASK}
A screenshot of the running application is shown in \cref{fig:flight-booking}.

All instances of the \TS{bookFlight} task have access to the shared list of free seats.
Rewriting the example in a calculus without side effects would not only be cumbersome,
obfuscating the code with explicit threading of state,
but it would be impossible to model the parallel execution of three \TS{bookFlight} tasks.
It is not known upfront which task will finish first,
and thus it is not possible to thread the free seat list between the parallel tasks.

\end{example}

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/flight-booking.png}
  \caption{
    Running web application of the flight booking example using a translation to \ITASKS.
    It shows three users booking a flight simultaneously.
    The first user entered name and age and continued picking seats.
    The second is entering details of two passengers.
    The ages are not filled in, therefore the \TS{Continue} button is disabled.
    The third user finished a booking.
    The first user can not pick seats \smallcaps{1b} and \smallcaps{1c} any more.
    The message bubble shows that the \TS{age} field only accepts integer values.
  }
  \label{fig:flight-booking}
\end{figure}

### 2019-02-06 sections/introduction.tex
update magic comments

% !TEX root=../icfp2019.tex

\section{Introduction}

Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
\TOP has two aspects.
Firstly, \TOP defines the primitive building blocks that are useful for high-level descriptions of how users collaborate with each other and with applications.
These building blocks are \emph{editors}, \emph{composition}, and \emph{shared data}.
Secondly, \TOP defines a type-driven way to generate applications, including graphical user interfaces, from workflows modelled with said building blocks.

In this paper we study the essence of \TOP.
We claim that at the heart of \TOP lie modular interactive workflows.
Workflows because \TOP programs coordinate collaboration between people.
Interactive because progress of \TOP programs is driven by user input.
Modular because the embedding in a strongly-typed functional language allows powerful abstractions using type-driven techniques and higher-order functions.

Interactive workflows are called \emph{tasks}.
Tasks stand for units of work in the real world, assigned to people.
People can work together in a number of ways, and this is reflected in \TOP by task combinators.
There is sequential composition, parallel composition, and choice.
People need to communicate in order to engage in these forms of collaboration.
This is reflected in \TOP by three kinds of communication mechanisms.
There is data flow \emph{alongside} control flow, where the result of a task is passed onto the next.
There is data flow \emph{across} control flow, where information is shared between multiple tasks.
Finally, there is communication with the \emph{outside} world, where information is entered into the system via input events.
The end points where the outside world interacts with \TOP applications are called editors.
In generated applications, editors can take many forms, like input fields, selection boxes, or map widgets.

iTasks is an implementation of \TOP, in the form of a shallowly embedded domain-specific language in the lazy functional programming language Clean.
It is a library that provides editors and monadic combinators.
iTasks uses the generic programming facilities of Clean to derive web applications from workflow models.
It has been used to model an incident management tool for the Dutch coast guard~\cite{conf/iscram/LijnseJP12}.
Also it has been used numerous times to prototype ideas for Command and Control~\cite{theses/nlda/Kool17, theses/radboud/Stutterheim17}, and in a case study for the Dutch tax authority \cite{conf/sfp/StutterheimAP17}.

### 2019-02-06 sections/language.tex
update magic comments

% !TEX root=../icfp2019.tex

\section{Calculus}
\label{sec:language}

In this section, we present the constructs of the modular interactive workflow
calculus \TOPHAT. We define the host and task language, the types, and
the static semantics. Then we describe the workings of each construct
using examples. These construct are formalised in \cref{sec:semantics}.

\subsection{Expressions}

\label{sub:expressions}
The host language is a simply typed lambda calculus, extended with some basic types and \ML-style references.
The grammar in \cref{fig:language-grammar} defines the syntax of the host language.
It has abstractions, applications, variables, and constants for booleans, integers and strings.
The symbol $\star$ stands for binary operators.
For the result of parallel tasks we need pairs.
Conditionals come in handy for defining guards.
%
References will be used to implement shared editors.
Our treatment of references closely follows the one by \citet{books/Pierce02TAPL}.
Creating a reference using the keyword $\Ref$ yields a location $l$.
Locations are not intended to be directly manipulated by the programmer.
The symbols ! and $:=$ stand for dereferencing and assignment.
The unit value will be used as the result of assignments.

\begin{figure}[h]
  \small
  \usemacro{G-Language-Compact}
  \caption{Language grammar} \label{fig:language-grammar}
\end{figure}

\label{sub:notation}
We use double quotation marks to denote strings.
Integers are denoted by their decimal representation, and booleans are written $\True$ and $\False$.
We freely make use of the logic operators $\Not$, $\land$, and $\lor$, arithmetic operators $+$, $-$, $\times$, and the string append operator $\pp$.
Furthermore, we use standard comparison operations $<$, $\le$, $\equiv$, $\not\equiv$, $\ge$, and $>$.
The symbol $\star$ stands for any of those.
%
\label{sub:abbreviations}
The notation $e_1; e_2$ is an abbreviation for $(\lambda x:\Unit.\ e_2)\ e_1$, where $x$ is a fresh variable.
The notation $\Let x:\tau = e_1 \In e_2$ is an abbreviation for $(\lambda x:\tau.\ e_2)\ e_1$.

\label{sub:pretasks}
The grammar in \cref{fig:task-grammar} specifies the syntactic category of \emph{pretasks}.
Pretasks are tasks that have unevaluated subexpressions.
Each pretask will be discussed in more detail in the following subsections.
We use open symbols ($\Edit, \Enter, \Next, \Xor$) for tasks that require user input, and closed symbols ($\Update, \Then, \Or$) for tasks that can be evaluated without user input.

\begin{figure}[h]
  \small
  \usemacro{G-Pretasks-Compact}
  \caption{Task grammar} \label{fig:task-grammar}
\end{figure}

\paragraph{Typing}
\label{sub:typing}

% Typing of our expressions $e$ is as to be expected.
% and won't be given in this document.
\Cref{fig:type-grammar} shows the grammar of types used by \TOPHAT.
It has functions, pairs, basic types, unit, references, and tasks.

\begin{figure}[h]
  \small
  \usemacro{G-Types-Compact}
  \caption{Type grammar} \label{fig:type-grammar}
\end{figure}

Typing rules are of the form $\RelationT$, which should be read as \enquote{in environment $\Gamma$ and store typing $\Sigma$, expression $e$ has type $\tau$}.
Typing rules for expressions in the host language are presented in the appendix.
The typing rules for pretasks are given in \cref{fig:typing-rules}.

\begin{figure}[h]
  \small
  \begin{mathpar}
    \boxed{\RelationT} \\
    \userule{T-Edit} \quad
    \userule{T-Enter} \quad
    \userule{T-Update} \\
    \userule{T-Then} \quad
    \userule{T-Next} \\
    \userule{T-Fail} \quad
    \userule{T-And} \\
    \userule{T-Or} \quad
    \userule{T-Xor}
  \end{mathpar}
  \caption{Typing rules} \label{fig:typing-rules}
\end{figure}

\subsection{Editors}

Programs in \TOPHAT model interactive workflows.
Interaction means communication with end users.
End users should be able to enter information into the system, change it, clear it, reenter it, and so on.
To do this, we introduce the concept of \emph{editors}.
%
Editors are typed containers that either hold a value or are empty.
Editors that have a value can be \emph{changed} or \emph{emptied}.
Empty editors can be \emph{filled}.
This is depicted as a state diagram in \cref{fig:editor-state} below.

\begin{figure}[h]
  \centering
  \includegraphics[width=\columnwidth,page=3]{figures/drawings-crop.pdf}
  \caption{
    Possible states of an editor and its transitions.
    Shared editors cannot be cleared.
  }
  \label{fig:editor-state}
\end{figure}

Editors stand for various forms of input and output, for example widgets in a \GUI, form fields on a webpage, sensors, or network connections.
%
Consider the editor for a person's age from \cref{exm:flight-booking}.
Users can change the value until they are satisfied with it.
Editors are meant capture this constantly changing nature of user input.
%
The user interface of an editor depends on its type.
This could be an input field for strings, a toggle switch for booleans, or even a map with a pin for locations.
It could also be a parser that tries to parse a line of text to match the type of the editor.

\paragraph{Valued and unvalued editors $(\Edit e, \Enter \tau)$}

Editors that hold an expression $e : \tau$ have type $\Task \tau$.
Empty editors are annotated with a type in order to ensure type safety.
Without that it would be possible to change the type of an editor, by clearing it and filling it with a value of different type.

\paragraph{Shared editors $(\Update e)$}

Shared editors watch references, lifting their value into the task domain.
If $e$ is a reference $\Reference \tau$, then $\Update e$ is of type $\Task \tau$.
Shared editors cannot be cleared, only changed.

Changes to a shared editor are immediately visible to all shared editors watching the same reference.
Imagine two users, Marco and Christopher, both watching shared editors of the same coordinate.
The editors are visualised as a pin on a map.
When Marco moves his pin, he updates the value of the shared editor, thereby changing the value of the reference.
This change is immediately reflected on Christopher's screen: The pin changes its position on his map.
This way Marco and Christopher can work together to edit the same information.

\label{sub:time}
Two other important use cases for shared editors are sensors and time.
Sensors can be represented as external entities that periodically update a shared editor with their current sensor value.
Similarly, the current time can be stored in a shared editor (\TS{update time}) which is periodically updated by a clock.
The actual sensor and the clock are not modelled in \TOPHAT.
We assume that they exist as external users that send update events to the system.
This allows programmers to write tasks that react to sensor values or timeouts.

\subsection{Steps}

Editors represent atomic units of work.
In this section we look at ways to compose smaller tasks into bigger ones.
Composing tasks can be done in two ways, sequential and parallel.
Parallel composition comes in two variants: combining two tasks (\emph{and}-parallel) and choosing between two tasks (\emph{or}-parallel).
We study sequential composition first, and after that combining and choosing.

\paragraph{Internal and external step $(t \Then e, t \Next e)$}
\label{sub:steps}

Sequential composition has a task $t$ on the left and a continuation $e$ on the right.
The accompanying typing rules are \refrule{T-Then} and \refrule{T-Next}.
According to these rules, the left hand side must be a task $t : \Task \tau_1$, and the right hand side $e : \tau_1 \to \Task \tau_2$ must be a function that, given the task value of $t$, calculates the task with which to continue.

Steps are guarded, which means that the step combinators can only proceed when the following conditions are met.
The left hand side must have a value, only then can the right hand side calculate the successor task.
The successor task must not be $\Fail$, introduced below.
The internal step can proceed immediately when these conditions are met.
The external step must additionally receive a continue event $\Continue$.

\begin{example}[Conditional stepping]
\label{exm:conditions}

Consider the following program.
\begin{TASK}
  enter Int >>= \n. if n == 42 then edit "Good" else edit "Bad"
\end{TASK}
Initially the step is guarded because the editor doesn't have a value.
When the user enters an integer, the program continues immediately with either \TS|edit "Good"| or \TS|edit "Bad"|, depending on the input.

\end{example}

\paragraph{Fail $(\Fail)$}
\label{sub:fail}

Fail is a task that never has a value and never accepts input.
The typing rule \refrule{T-Fail} states that it has type $\Task \tau$ for any type $\tau$.
Programmers can use $\Fail$ to tell steps that no sensible successor task can be determined.

\begin{example}[Guarded stepping]

Consider this slight variation on \cref{exm:conditions}.
\begin{TASK}
  enter Int >>= \n. if n == 42 then edit "Good" else fail
\end{TASK}
The user is asked to enter an integer.
As long as the right hand side of $\Then$ evaluates to $\Fail$, the step cannot proceed, and the user can keep editing the integer.
As soon as the value of the left hand side is $42$, the right hand side evaluates to something other than $\Fail$, and the step proceeds to \TS{edit "Good"}.

\end{example}

\begin{example}[Waiting]\label{exm:wait}

\lstset{emph={amount,start,now}}
With the language constructs seen so far it is possible to create a task that waits for a specified amount of time.
To do this,
we make use of a shared editor holding the current time (see \cref{sub:time}),
and a guarded internal step.
\begin{TASK}
  let wait : Int -> Task Unit = \amount : Int.
    update time >>= \start : Int.
    update time >>= \now : Int.
      if now >= start + amount then edit <<>> else fail
\end{TASK}
The first step is immediately taken, resulting in \TS{start} to be the time at the moment \TS{wait} is executed.
The second step is guarded until the current time is greater or equal to the start time plus the requested amount.

\end{example}

\subsection{Parallel}

A common pattern in workflow design is splitting up work into multiple tasks that can be executed simultaneously.
In \TOPHAT, all parallel branches can progress independently, driven by input events.
This requires input events to be tagged in order to reach the intended task.

There are two ways to proceed after a parallel composition.
One way is to wait for all tasks to produce results and combine those,
the other to pick the first available result.
Both ways introduce explicit forks and implicit joins in \TOPHAT.

\paragraph{Combination $(e_1 \And e_2)$}

A combination of two tasks is a parallel \emph{and}.
It has a value only if both branches have a value.
This is reflected in the typing rule \refrule{T-And},
It shows that if the first task has type $\tau_1$,
and the second has type $\tau_2$,
their combination has the pair type $\tau_1 \times \tau_2$.

\begin{example}[Combining]

The task
\begin{TASK}
  enter Int <&> edit " Batman" >>= \<<n, s>>. edit (replicate n "Na" ++ s)
\end{TASK}
can only step when both editors have values.
When it steps, the continuation uses the pair of integer and string to calculate the result.

\end{example}

\paragraph{Internal and external choice $(e_1 \Or e_2, e_1 \Xor e_2)$}

\hfill Internal \\ choice ($\Or$) is a parallel \emph{or}.
It picks the leftmost branch that has a value.
Its typing rule \refrule{T-Or} states that both branches must have the same type $\Task \tau$.
For example $\Enter \Int \Or \Edit 37$ normalizes to $\Edit 37$, because $\Enter \Int$ doesn't have a value.
Users can work on both branches of an internal choice simultaneously.

External choice ($\Xor$) is different in this regard.
An external choice requires users to pick a branch before continuing with it.
This means users cannot work on the branches of an external choice before picking one.

\begin{example}[Delay]

\lstset{emph={proceed}}

We illustrate the use of internal and external choice by means of
an example that asks users to proceed with a given task or to cancel.
If the user does not make a choice within a given time frame, the program proceeds automatically.
The example makes use of the task \TS{wait} from \cref{exm:wait}.
\begin{TASK}
  let cancel : Task Unit = edit <<>> in
  let delay : Int -> Task Unit -> Task Unit =
    \n : Int. \proceed : Task Unit.
    (proceed <?> cancel) <|> (wait n >>= \u : Unit. proceed)
\end{TASK}
Note that \TS{delay} is higher-order.
It is a task which takes another task as parameter.

\end{example}

% \subsection{Appointment}
%
% \fixme{write this subsection}

% It is essential for a workflow system to support multiple collaborators working
% together. Adding this funcitonaltiy to the \TOPHAT calculus is very
% straightforward. We identify users by means of a name $u$ which we can then use
% to label appointed tasks. The sole purpose of this labelling is to be able to
% tell which inputs are expected from what users. Other than that, it has no
% influence on the semantics. For space reasons, we have have excluded multi-user
% support from this paper.
%
% Mss moeten we het schrijven vanuit het oogpunt van labeling: je kunt labels
% toevoegen aan knoppen, als uitleg bij taken en om gebruikers toe te kennen aan
% taken. Vervolgens zou je iets met die gebruikers kunnen doen: alleen taken voor
% hun weergeven in de UI, alleen inputs voor hun genereren met I etc. Past dat?

\subsection{Annotations}

Tasks can be annotated with additional information.
The system can use this information in various ways.
Possible use cases are labels for the user interface, resource consumption information for static resource analysis, or messages for automatic end-user feedback.
Annotations are not covered in this paper.
Our Idris implementation of \TOPHAT supports annotating tasks with user \ID{}s, so that individual tasks in a large workflow can be assigned to different users.
These annotations are used to filter the user interfaces for each user so that they can only see their part of the workflow.

% For now, we only concentrate on appointing tasks to users,
% which is common in workflow modelling.
%
%
%
% \paragraph{Appoint $(u \At e)$}
%
% Users are identified by a name $u$ which is a string.
% They can be appointed tasks by means of the $\At$ combinator.
% The sole purpose of this labelling is to be able to tell users which inputs are expected of them.
% In \cref{sec:handling}, a function is described to calculate which users can take what actions.
% Other than this function, the appointment of tasks has no influence on the semantics.
% All semantic rules in \cref{sec:semantics} simply continue with their operations as if the label was not there.
% \todo{we still have to fix the type of $\Inputs$ and state the type of $u$}

### 2019-02-06 sections/properties.tex
update magic comments

% !TEX root=../icfp2019.tex

\section{Properties}
\label{sec:properties}

% \subsection{Safety}

In order to validate our semantics, we show that our evaluation, normalisation
and handling semantics is type preserving. We additionally prove a progress
theorem for our small-step handling semantics.

Additionally, we show that the normalisation semantics is a big-step semantics.
While at first sight, this might seem obvious, the fact that we are dealing with
state complicates matters.

We show that our failing function $\Failing$ indeed only indicates expressions
that can not be normalised and that allow no further interaction.

Finally, we prove that the function to compute all possible inputs $\Inputs$ is sound and complete.

\subsection{Preservation}
\label{sub:preservation}

We show that the following three preservation Theorems hold.

\begin{theorem}[Preservation under evaluation]
  % For all well typed expressions $e$ and states $s$,
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\tau$ and $\Gamma,\Sigma \infers s$,
  if $e,s \evaluate e',s'$,
  then $\Gamma,\Sigma \infers e':\tau$ and $\Sigma \infers s'$.
  \label{thm:pres-eval}
\end{theorem}

Where $\Sigma \infers s$ means that for all $l\in s$, it holds that
$\Sigma(l)=\Gamma,\Sigma\infers s(l)$.

Theorem~\ref{thm:pres-eval} is shown to be correct by induction over $e$. The full
proof can be found in the appendix.

Moving on, we show that normalisation also preserves, by showing that the
following Theorem holds.

\begin{theorem}[Preservation under normalisation]
  % For all well typed expressions $e$ and states $s$,
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $e,s \normalise e',s'$,
  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers s'$.
  \label{thm:pres-norm}
\end{theorem}

Since this semantics makes use of the value function $\Value$ and the striding semantics $\stride$. We first need to
show that preservation also holds for these two semantics.

\begin{lemma}[Task value preserves]
  % For all well typed expressions $e$ and states $s$,
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $\Value{(e,s)}=v$,
  then $v:\tau$.
  \label{lem:presvalue}
\end{lemma}

\begin{lemma}[Striding preserves]
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $e,s \stride e',s'$,
  then $\Gamma,\Sigma \infers e':\tau$ and $\Sigma \infers s'$.
  \label{thm:pres-stride}
\end{lemma}

Lemma~\ref{lem:presvalue} and Theorem~\ref{thm:pres-stride} are proven in the
appendix by induction over $e$. This subsequently allows us to prove
Theorem~\ref{thm:pres-norm}, again by induction over $e$.

This brings us finally to the type preservation property of the handling semantics.

\begin{theorem}[Preservation under handling]
  % For all well typed expressions $e$, states $s$, and inputs $i$,
  For all expressions $e$, states $s$ and inputs $i$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $ e,s \handle{i} e's'$,
  then $\Gamma,\Sigma\infers e':\Task\tau$ and $\Sigma\infers s'$.
  \label{thm:pres-handle}
\end{theorem}

And again, this is proven by induction over $e$.

From Theorem~\ref{thm:pres-handle} and Theorem~\ref{thm:pres-norm} we directly
obtain that the driving semantics also preserves.

\subsection{Progress}

A well-typed term of a task type is guaranteed to progress after normalisation,
unless it is failing.

We define what we mean with progress in Theorem~\ref{thm:prog-norm}.
\begin{theorem}[Progress under handling]
  For all well typed expressions $e$ and states $s$,
  % For all expressions $e$ and states $s$
  % such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $e,s \normalise e',s'$,
  then either $\Failing(e', s')$
  or there exist $e''$, $s''$, and $i$ such that $e',s'\handle{i} e'',s''$.
  \label{thm:prog-norm}
\end{theorem}

Where a well typed expression $e$ means that $\Gamma,\Sigma\infers e:\tau$ for
some type $\tau$, and a well typed state means that $\Sigma\infers s$.

If an expression $e$ and state $s$ are well-typed, then after normalisation, it
either fails, or there exists some input $i$ that can be handled by it.
In order to prove this Theorem, we require an additional theorem.

% \paragraph{Failing}

We need to show that the failing function $\Failing$ behaves as desired.

\begin{theorem}[Failing means no interaction possible]
  For all well typed expressions $e$ and states $s$,
  % For all expressions $e$ and states $s$
  % such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  and $e,s \normalise e',s'$,
  we have that $\Failing(e',s') = \True$,
  if and only if there is no input $i$
  such that $e',s'\handle{i} e'',s''$ for some $e''$ and $s''$.
  \label{thm:failing}
\end{theorem}

The Theorem above states that an expression $e$ and state $s$ are failing, if,
after normalisation, there exists no input that can be handled by it.
We prove the theorem to be true by induction on $e'$.

% \paragraph{Proof}

We now have the ingredients to prove Theorem~\ref{thm:prog-norm}.

\begin{proof}
  Given $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$ and after
  normalisation $e,s \normalise e',s'$, we find ourselves in either one of the
  following situations:

  There exists an $i$ such that $e',s'\handle{i} e'',s''$.

  There does not exist an $i$ such that $e',s'\handle{i} e'',s''$. In this case, we
  know that $\Failing(e',s')$ must be true, by Theorem~\ref{thm:failing}.
\end{proof}

\subsection{Soundness and Completeness of Inputs}

In order to validate the function that calculates all possible inputs $\Inputs$,
we want to show that the set of possible inputs it produces is both sound and
complete with respect to the handle semantics. By sound we mean that all inputs
in the set of possible inputs can actually be handled by the handle semantics,
and by complete we mend that the set of possible inputs contains all inputs
that can be handled by the handle semantics. Theorem~\ref{thm:safety-i} expresses
exactly this property.

\begin{theorem}[Inputs function is sound and complete]
  For all well typed expressions $e$, states $s$, and inputs $i$,
  % For all expressions $e$, states $s$, and inputs $i$
  % such that $\Gamma,\Sigma \infers e:\tau$ and $\Sigma \infers s$,
  we have that $i \in \Inputs{(e)}$ if and only if $e,s \handle{i} e',s'$.
  \label{thm:safety-i}
\end{theorem}

We prove the above theorem by induction over $e$. The proof is listed in the
appendix.

### 2019-02-06 sections/relatedwork.tex
update magic comments

% !TEX root=../icfp2019.tex

\section{Related work}
\label{sec:relatedwork}

\paragraph{iTasks}

As mentioned earlier, iTasks is an implementation of \TOP. iTasks has many
features, and its basic combinators are versatile and powerful.vSimpler
combinators are implemented by restricting the powerful ones. This is useful for
everyday programming, where having lots of functionality at one's fingertips is
convenient and efficient.

There have been two previous papers that describe semantics of iTasks, by
\citet{conf/ifl/KoopmanPA08} and \citet{conf/ppdp/PlasmeijerLMAK12}.
Both give a different semantics in the form of minimal implementations of a
subset of the interface of iTasks. These semantics however do not make an
explicit distinction between the host language and task language and they do not
provide a formal semantics, which makes formal reasoning difficult.

% Our work differs in three major ways.
% First, we make an explicit distinction between the host language and the task
% language, to emphasise their boundaries.
% Second, we make use of separate semantic functions for querying the current task
% value, determining the continuation after an input event, and producing a user
% interface.
% In iTasks and its aforementioned semantics, these are integrated into the
% \CL{Task} datatype.
% Third, our system does not have the notion of task stability.
% When desired, stability can be introduced by the programmer, but the system
% itself makes no use of it.
% We argue that this does not impact the expressiveness of our language.

% mTasks is another implementation of TOP, designed for coordination of IoT devices. \alert{cite MSc Lubbers}
% mTasks has been used in a demonstrator for home automation. \alert{cite MSc Andrade}

\paragraph{Workflow patterns}

Much research has been done into workflow patterns. They are regarded as special
kind of the design patterns in software engineering. They identify recurring patterns
in workflow systems, much like the combinators defined by \TOPHAT. Work by
Van der Aalst~\cite{journals/dpd/AalstHKB03} defines a comprehensive list of
these pattens, and examines their availability in industry workflow software.
Workflow patterns are usually described in terms of control flow graphs, and no
formal specification is given, which makes comparison and formal reasoning more
difficult.

\paragraph{Petri net}

Petri nets, and more specifically Workflow nets, have also been used to describe
workflows~\cite{journals/jcsc/Aalst98}. They allow for the modelling and
analysis of business processes. A downside of using workflow nets is that the
model is not directly implementable, which is possible with the \TOPHAT calculus.

### 2019-02-06 sections/semantics.tex
update magic comments

% !TEX root=../icfp2019.tex

\section{Semantics}
\label{sec:semantics}

In this section we formalise the semantics of the language constructs described in \cref{sec:language}.
We organise this by following the structure of the language.

Firstly, the task language is embedded in a simply typed lambda calculus.
This requires a specification of the evaluation of terms in the host language, and how it handles the task language.

Secondly, there are two ways to drive evaluation of task expressions, internally by the system itself, and externally by the user.
This is done in two additional semantics, one for the internal normalisation of tasks, and another for the interaction with the end user.

One of our explicit goals is to keep the semantics for evaluation and normalisation separate,
to not mix general purpose programming notions with workflow specific semantics.
This is achieved by letting tasks be values in the host language.

The three main layers of semantics are thus \emph{evaluation}, \emph{normalisation}, and \emph{interaction}.
The semantics, together with \emph{observations}, will be discussed in the following subsections.
\Cref{fig:semantic-functions} shows the relation between the semantics.
It also shows that there are two helper semantics, \emph{handle} and \emph{stride}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\columnwidth,page=5]{figures/drawings-crop.pdf}
  \caption{
    Semantic functions defined in this report and their relation.
  }
  \label{fig:semantic-functions}
\end{figure}

We use the convention that downward arrows are big-step semantics, and rightward arrows are small-step semantics.

\subsection{Evaluating expressions}
\label{sec:evaluation}

The host language evaluates expressions using a big-step semantics.
Evaluating an expression $e$ in state $s$ into a value $v$ in state $s'$ is denoted by $\RelationE$.
To ease reasoning about references, we choose a call-by-value evaluation strategy.

\Cref{fig:value-grammar} shows values with respect to the evaluation semantics.
Tasks are values, and the operands of task constructors are evaluated eagerly.
Exceptions to this are steps and external choice, where some or all of the operands are not evaluated.

\begin{figure}[h]
  \small
  \usemacro{G-Values-Compact}
  \caption{Value grammar} \label{fig:value-grammar}
\end{figure}

The rules to evaluate expressions $e$ do not differ from standard work, except for the task constructs.
The evaluation rules for tasks can be deduced from the value grammar.
They are given in the appendix.

% \begin{figure}[h]
%
%   \begin{mathpar}
%     \userule{E-Edit} \quad
%     \userule{E-Enter} \quad
%     \userule{E-Update} \\
%     \userule{E-Then} \quad
%     \userule{E-Next} \\
%     \userule{E-And} \quad
%     \userule{E-Or} \\
%     \userule{E-Xor} \quad
%     \userule{E-Appoint} \quad
%     \userule{E-Fail}
%   \end{mathpar}
%
%   \caption{Evaluation semantics for pretasks} \label{}
% \end{figure}

\subsection{Task observations}

% \begin{figure*}[b]
%   \small
%   \usemacro{O-Value-Compact} \hfill \usemacro{O-Failing-Compact} \hfill \usemacro{O-Inputs-Compact}
%   \caption{Values} \label{fig:observation-value}
% \end{figure*}

The normalisation and interaction semantics make use of observations on tasks.
Observations are semantic functions on the syntax tree of tasks.
There are four semantic functions: $\Value$ for the current task value, $\Failing$ to determine if a task fails, $\Inputs$ for the currently accepted input events, and a function for generating user interfaces.
The semantics make use of $\Value$ and $\Failing$, while $\Inputs$ is used for proving safety.
The function for user interfaces is not used by the semantics, but by our implementation.
It is only described in passing here.

\paragraph{Observable values $(\Value)$}

Task values are used by steps to calculate the successor task.
Filled editors are tasks which contain values, as are shared editors.
Unvalued editors do not contain values, neither does the fail task.
These facts propagate through all other task constructors.
The partial function $\Value$ associates a value $v$ to tasks $t$ where possible.
Its definition is given in \cref{fig:observation-value}.

\begin{figure}[h]
  \small
  \usemacro{O-Value}
  \caption{Values} \label{fig:observation-value}
\end{figure}

Internal and external steps do not have an observable value, because calculating the value would require evaluation of the continuation.
Parallel composition only has a value when both branches have values, in which case these values are paired.
Internal choice has a value when one of the branches has a value.
When both branches have a value, it takes the value of the left branch.
External choice does not have a value because it waits for user input.

\paragraph{Failing $(\Failing)$}

In \cref{sub:fail} we introduced $\Fail$ to stand for an impossible task.
Combinations of tasks can also be impossible.
Take for example the parallel composition of two fails ($\Fail \And \Fail$).
This expression is equivalent to $\Fail$, because it can not handle input and can not be further normalised.

This intuition is formalised by the function $\Failing$ in \cref{fig:observation-failing}.
It determines whether a task is impossible.
Such tasks are called \emph{failing}.

\begin{figure}[h]
  \small
  \usemacro{O-Failing}
  \caption{Failing} \label{fig:observation-failing}
\end{figure}

Steps whose left hand sides are failing can never proceed because of the lack of an observable value.
Therefore they are itself failing.
The internal choice of two failing tasks is failing.
External choices let the user pick a side and only then evaluate the corresponding subexpression.
To determine if an external choice is failing, it needs to peek into the future to check if both subexpressions are failing.
Annotations are failing only if the inner task is failing.

\paragraph{User interface}

\TOPHAT is designed such that a user interface can be generated from a task's syntax tree.
A possible graphical user interface is shown in \cref{fig:flight-booking}, where tasks are rendered as \HTML pages.
Editors are rendered as input fields,
external choices are represented by two buttons,
and parallel tasks are rendered side by side.
Steps only show the interface of their left hand side.
In case of an external step they are accompanied by a button.
When the guard condition of a step is not fulfilled, the button is disabled.

\subsection{Normalising tasks}
\label{sec:normalise}

The normalisation semantics is responsible for reducing expressions of type $\Task$ until they are ready to handle input.
It is a big-step semantics, and makes use of evaluation of the host language.
We write $\RelationN$ to describe that
an expression $e$ in state $s$ normalises to task $t$ in state $s'$.

Normalisation rules are given in \cref{fig:normalisation-semantics}.
Both rules ensure that expressions are first evaluated by the host language ($\evaluate$), and then by the stride semantics ($\stride$).
These two actions are repeated until the resulting state and task stabilise.

\begin{figure}[h]
  \small

  \boxed{\RelationN}
  \begin{mathpar}
    \userule{N-Done} \\
    \userule{N-Repeat}
  \end{mathpar}
  \caption{Normalisation semantics} \label{fig:normalisation-semantics}

\end{figure}

The split between striding and normalisation is due to mutable references.
Consider the following example, where $s=\set{l\mapsto\False}$.
\begin{TASK}
  (update l >>= \x:Bool. if x then e else fail) <&> (l := True; edit <<>>)
\end{TASK}
\refrule{S-And} reduces this expression in one step to
\begin{TASK}
  (update l >>= \x:Bool. if x then e else fail) <&> (edit <<>>)
\end{TASK}
with $s'=\set{l\mapsto\True}$.
This expression is not normalised, because the left task can take a step.
The issue here lies in the fact that the right task updates $l$.
To overcome this problem, the \refrule{N-Done} and \refrule{N-Repeat} rules ensure that striding is applied until the state $s$ becomes stable and no further normalisation can take place.

The striding semantics is responsible for reducing internal steps and internal choices.
A stride from task $t$ in state $s$ to $t'$ in state $s'$ is denoted by $\RelationS$.
The rules for striding are given in \cref{fig:striding-semantics}.
Tasks like editors, fail and external choice are not further reduced.
For external choice and parallel there are congruence rules.

\begin{figure}[h]
  \small

  \begin{mathpar}
    \boxed{\RelationS}
  \end{mathpar}

  \paragraph{Step}
  \begin{mathpar}
    \userule{S-ThenStay} \\
    \userule{S-ThenFail} \\
    \userule{S-ThenCont}
  \end{mathpar}

  \paragraph{Choose}
  \begin{mathpar}
    \userule{S-OrLeft} \\
    \userule{S-OrRight} \\
    \userule{S-OrNone}
  \end{mathpar}

  \paragraph{Ready}
  \begin{mathpar}
    \userule{S-Edit} \quad \userule{S-Fill} \qquad \userule{S-Update} \\
    \userule{S-Fail} \quad \userule{S-Xor}
  \end{mathpar}

  \paragraph{Congruence}
  \begin{mathpar}
    \userule{S-Next} \quad
    \userule{S-And} \\
    %\userule{S-Appoint}
    % \userule{S-Eval}
  \end{mathpar}

  \caption{Striding semantics} \label{fig:striding-semantics}
\end{figure}

\paragraph{Principles of stepping}
\label{sub:stepping-principles}

Stepping away from task $t_1$ can only be performed when
$t_1$ has a value: $\Value(t_1) = v_1$.
Only then can a new task $t_2$ be calculated from the expression $e$.
On top of that, $t_2$ must not be failing: $\neg\Failing(t_2)$.

These principles lead to the stepping rules in \cref{fig:striding-semantics}.
\refrule{S-ThenStay} does nothing,
because the left side does not have a value.
\refrule{S-ThenFail} covers the case that
the left side has a value but the calculated successor task is failing.
This rule uses the semantics of the host language to evaluate the application $e_2\ v_1$.
When all required conditions are fulfilled, \refrule{S-ThenCont} allows stepping to the successor task.

\paragraph{Principles of choosing}
\label{sub:choosing-principles}

Choosing between two tasks $t_1$ and $t_2$ can only be done when
at least one of them has a value: $\Value(t_1) = v_1 \lor \Value(t_2) = v_2$.
When both have a value, the left task is chosen.
When none has a value, none can be chosen.
These principles lead to the rules \refrule{S-OrLeft}, \refrule{S-OrRight}, and \refrule{S-OrNone},
which encode that the choice operator picks the leftmost task that has a value.

\subsection{Handling user inputs}
\label{sec:handling}

The handling semantics is the outermost layer of the stack of semantics.
It is responsible for performing external steps and choices, and for changing the values of editors.
The rules of the interaction semantics are given in \cref{fig:interaction-semantics}.
The semantics is only applicable to normalised $t$.
Sending an input event $i$ to a task $t$ first handles the event and then prepares the resulting task for the next input by normalising it.

\begin{figure}[h]
  \small
  \begin{mathpar}
    \boxed{\RelationI} \\
    \userule{I-Handle}
  \end{mathpar}
  \caption{Interaction semantics} \label{fig:interaction-semantics}
\end{figure}

Inputs $i$ are formed according to the grammar in \cref{fig:input-grammar}.
Handling input is done by the \emph{handling} semantics shown in \cref{fig:handling-semantics}.
It is a small step semantics with labelled transitions.
It takes a task $t$ in a state $s$ and an input $i$, and yields a new task $t'$ in a new state $s'$.

\begin{figure}[h]
  \small
  \usemacro{G-Inputs-Compact}
  \caption{Input grammar} \label{fig:input-grammar}
\end{figure}

\begin{figure}[h]
  \small

  \begin{mathpar}
    \boxed{\RelationH}
  \end{mathpar}

  \paragraph{Editing}
  \begin{mathpar}
    \userule{H-Change} \quad
    \userule{H-Empty} \\
    \userule{H-Fill} \quad
    \userule{H-Update}
  \end{mathpar}

  \paragraph{Continuing}
  \begin{mathpar}
    \userule{H-Next} \\
    \userule{H-PickLeft} \quad
    \userule{H-PickRight}
  \end{mathpar}

  \paragraph{Passing}
  \begin{mathpar}
    \userule{H-PassThen} \quad \userule{H-PassNext} \\
    \userule{H-FirstAnd} \quad \userule{H-SecondAnd} \\
    \userule{H-FirstOr}  \quad \userule{H-SecondOr}\\
  %  \userule{H-Appoint}
  \end{mathpar}

  \caption{Handling semantics} \label{fig:handling-semantics}
\end{figure}

\smallcaps{H-Change},
\smallcaps{H-Empty},
\smallcaps{H-Fill},
\smallcaps{H-Update}:
Input events $v$ and $\Empty$ are used to change the value of editors and empty them.
Editors only accept values of the correct type.

\smallcaps{H-Next}:
The $\Continue$(ontinue) action triggers an external step.
As with internal stepping, this is only possible if the left side has a value and the continuation is not failing.

\smallcaps{H-PickLeft},
\smallcaps{H-PickRight}:
$\Left$ and $\Right$ are used to pick the left or right option of an external choice.

\smallcaps{H-PassThen},
\smallcaps{H-PassNext}:
The step combinators pass all events other than $\Continue$ to the left side.

\smallcaps{H-FirstAnd},
\smallcaps{H-SecondAnd},
\smallcaps{H-FirstOr},
\smallcaps{H-SecondOr}:
Inputs $\First$(irst) and $\Second$(econd) are used to direct inputs to the correct branch of parallel combinations.

There is a function $\Inputs$ which calculates the possible input events a given task expects.
It takes a normalised task and a state and returns a set of inputs that can be handled.
The definition of this function is listed in \cref{fig:observation-input}.

\begin{figure}[h]
  \small
  \usemacro{O-Inputs}
  \caption{Inputs} \label{fig:observation-input}
\end{figure}

\subsection{Implementation}

The semantics have been implemented in the Idris programming language \cite{journals/jfp/Brady13}.
We use techniques presented by \textcite{journals/entcs/JaskelioffGH11}, \textcite{journals/jfp/Swierstra08}, and \textcite{school/maktoberdorf/PeytonJones01}.
Such an implementation increases the confidence in the correctness and completeness of the semantics.
The source code can be found on GitHub.\footnote{\url{http://github.com/XXX/XXX}}
A proof of concept implementation in Haskell is also available, showing that \TOPHAT can be implemented without dependent types.

A command-line interface is part of this implementation.
It prompts users to type input events, which get parsed and processed by the interaction semantics.

### 2019-02-06 sections/structure.tex
update magic comments

% !TEX root=../icfp2019.tex

% \paragraph{Overview}

In Section~\ref{sec:example} we demonstrate the functionality of \TOPHAT by means of an example,
Section~\ref{sec:comparison} compares \TOP and \TOPHAT to \CSP.
Section~\ref{sec:language} introduces the \TOPHAT calculus syntax
and Section~\ref{sec:semantics} the semantics.
Then in Section~\ref{sec:properties} we show that certain properties hold for the calculus.
We take a look at related work in Section~\ref{sec:relatedwork}
and conclude in Section~\ref{sec:conclusions}.

### 2019-02-06 sections/relatedwork.tex
additions to related work

@@
 model is not directly implementable, which is possible with the \TOPHAT calculus.
+
+% NEW RELATED WORK sections
+
+\paragraph{Business process modeling}
+
+A plethora of languages exists to describe, model, display and execute these
+processes. YAWL by van der Aalst~\cite{DBLP:journals/is/AalstH05} is one of
+these languages. It facilitates modeling and execution of dynamic workflows,
+with support for AND, OR and XOR workflow patterns. YAWL programs consist of
+Nets, and are therefore programmed visually. BPEL\todo{cite} is another popular business process language.
+The standardized language allows for the specification of actions within
+business processes, using an XML format. Processes specified in BEPL are
+executable, just like YAWL.
+
+\paragraph{IFTTT}
+
+If This Then That (IFTTT)\todo{cite} is a web service that allows users to write and
+configure simple conditional statements called applets, that can interact with
+web services and IoT devices. These applets can be seen as tiny workflows, that
+facilitate collaboration between different services and platforms. The
+complexity of the flow is very limited however, and is not geared towards humans
+collaborating.
+
+\paragraph{Google forms \& Wufoo}
+
+Google forms\todo{cite} and Wufoo\todo{cite} are web based survey administration apps that allows
+users to compose forms and gather data from them. They are comparable to TOP and
+iTasks in that they provide an easy way for users to construct forms on the web.
+However, Google forms and Wufoo do not allow users to define what the system
+then should do with this information, and can therefore not be regarded as a
+workflow or business process modeling systems.
+
+\paragraph{Chorus} % (http://www.chorus-home.org/)
+
+Chorus~\cite{chen2017chorus} is a visual programming environment for online
+collaboration apps. The goal of Chorus is to allow users to program their own
+application, though which a collective task can be coordinated. Users work on a
+preset data-type that drives the behavior of the user defined applications.

### 2019-02-06 sections/semantics.tex
scale figures for new format
remove mathpartir and use custom rule macros

@@
   \centering
-  \includegraphics[width=\columnwidth,page=5]{figures/drawings-crop.pdf}
+  \includegraphics[width=0.6\columnwidth,page=5]{figures/drawings-crop.pdf}
   \caption{
@@
 %
-%   \begin{mathpar}
-%     \userule{E-Edit} \quad
-%     \userule{E-Enter} \quad
-%     \userule{E-Update} \\
-%     \userule{E-Then} \quad
-%     \userule{E-Next} \\
-%     \userule{E-And} \quad
-%     \userule{E-Or} \\
-%     \userule{E-Xor} \quad
-%     \userule{E-Appoint} \quad
+%   \begin{gather*}
+%     \userule{E-Edit} \Quad
+%     \userule{E-Enter} \Quad
+%     \userule{E-Update} \Break
+%     \userule{E-Then} \Quad
+%     \userule{E-Next} \Break
+%     \userule{E-And} \Quad
+%     \userule{E-Or} \Break
+%     \userule{E-Xor} \Quad
+%     \userule{E-Appoint} \Quad
 %     \userule{E-Fail}
-%   \end{mathpar}
+%   \end{gather*}
 %
@@
   \boxed{\RelationN}
-  \begin{mathpar}
-    \userule{N-Done} \\
+  \begin{gather*}
+    \userule{N-Done} \Break
     \userule{N-Repeat}
-  \end{mathpar}
+  \end{gather*}
   \caption{Normalisation semantics} \label{fig:normalisation-semantics}
@@
 
-  \begin{mathpar}
+  \begin{gather*}
     \boxed{\RelationS}
-  \end{mathpar}
+  \end{gather*}
 
   \paragraph{Step}
-  \begin{mathpar}
-    \userule{S-ThenStay} \\
-    \userule{S-ThenFail} \\
-    \userule{S-ThenCont}
-  \end{mathpar}
+  \begin{align*}
+    \userule{S-ThenStay} \Quad
+    & \userule{S-ThenFail} \Break
+    & \userule{S-ThenCont}
+  \end{align*}
 
   \paragraph{Choose}
-  \begin{mathpar}
-    \userule{S-OrLeft} \\
-    \userule{S-OrRight} \\
-    \userule{S-OrNone}
-  \end{mathpar}
+  \begin{align*}
+    \userule{S-OrLeft} \Quad
+    & \userule{S-OrRight} \Break
+    & \userule{S-OrNone}
+  \end{align*}
 
   \paragraph{Ready}
-  \begin{mathpar}
-    \userule{S-Edit} \quad \userule{S-Fill} \qquad \userule{S-Update} \\
-    \userule{S-Fail} \quad \userule{S-Xor}
-  \end{mathpar}
+  \begin{gather*}
+    \userule{S-Edit} \Quad \userule{S-Fill} \Quad \userule{S-Update} \Quad
+    \userule{S-Fail} \Quad \userule{S-Xor}
+  \end{gather*}
 
   \paragraph{Congruence}
-  \begin{mathpar}
-    \userule{S-Next} \quad
-    \userule{S-And} \\
+  \begin{gather*}
+    \userule{S-Next} \Quad
+    \userule{S-And}
     %\userule{S-Appoint}
     % \userule{S-Eval}
-  \end{mathpar}
+  \end{gather*}
 
@@
   \small
-  \begin{mathpar}
-    \boxed{\RelationI} \\
+  \begin{gather*}
+    \boxed{\RelationI} \Break
     \userule{I-Handle}
-  \end{mathpar}
+  \end{gather*}
   \caption{Interaction semantics} \label{fig:interaction-semantics}
@@
 
-  \begin{mathpar}
+  \begin{gather*}
     \boxed{\RelationH}
-  \end{mathpar}
+  \end{gather*}
 
   \paragraph{Editing}
-  \begin{mathpar}
-    \userule{H-Change} \quad
-    \userule{H-Empty} \\
-    \userule{H-Fill} \quad
+  \begin{gather*}
+    \userule{H-Change} \Quad
+    \userule{H-Empty} \Break
+    \userule{H-Fill} \Quad
     \userule{H-Update}
-  \end{mathpar}
+  \end{gather*}
 
   \paragraph{Continuing}
-  \begin{mathpar}
-    \userule{H-Next} \\
-    \userule{H-PickLeft} \quad
+  \begin{gather*}
+    \userule{H-Next} \Break
+    \userule{H-PickLeft} \Quad
     \userule{H-PickRight}
-  \end{mathpar}
+  \end{gather*}
 
   \paragraph{Passing}
-  \begin{mathpar}
-    \userule{H-PassThen} \quad \userule{H-PassNext} \\
-    \userule{H-FirstAnd} \quad \userule{H-SecondAnd} \\
-    \userule{H-FirstOr}  \quad \userule{H-SecondOr}\\
+  \begin{gather*}
+    \userule{H-PassThen} \Quad \userule{H-FirstAnd} \Quad \userule{H-FirstOr} \Break
+    \userule{H-PassNext} \Quad \userule{H-SecondAnd} \Quad \userule{H-SecondOr}
   %  \userule{H-Appoint}
-  \end{mathpar}
+  \end{gather*}
 
@@
 
-\smallcaps{H-Change},
-\smallcaps{H-Empty},
-\smallcaps{H-Fill},
-\smallcaps{H-Update}:
+\refrule{H-Change},
+\refrule{H-Empty},
+\refrule{H-Fill},
+\refrule{H-Update}:
 Input events $v$ and $\Empty$ are used to change the value of editors and empty them.
@@
 
-\smallcaps{H-Next}:
+\refrule{H-Next}:
 The $\Continue$(ontinue) action triggers an external step.
@@
 
-\smallcaps{H-PickLeft},
-\smallcaps{H-PickRight}:
+\refrule{H-PickLeft},
+\refrule{H-PickRight}:
 $\Left$ and $\Right$ are used to pick the left or right option of an external choice.
 
-\smallcaps{H-PassThen},
-\smallcaps{H-PassNext}:
+\refrule{H-PassThen},
+\refrule{H-PassNext}:
 The step combinators pass all events other than $\Continue$ to the left side.
 
-\smallcaps{H-FirstAnd},
-\smallcaps{H-SecondAnd},
-\smallcaps{H-FirstOr},
-\smallcaps{H-SecondOr}:
+\refrule{H-FirstAnd},
+\refrule{H-SecondAnd},
+\refrule{H-FirstOr},
+\refrule{H-SecondOr}:
 Inputs $\First$(irst) and $\Second$(econd) are used to direct inputs to the correct branch of parallel combinations.

### 2019-02-06 sections/language.tex
scale figures for new format
remove mathpartir and use custom rule macros
compact code examples

@@
   \small
-  \begin{mathpar}
-    \boxed{\RelationT} \\
-    \userule{T-Edit} \quad
-    \userule{T-Enter} \quad
-    \userule{T-Update} \\
-    \userule{T-Then} \quad
-    \userule{T-Next} \\
-    \userule{T-Fail} \quad
-    \userule{T-And} \\
-    \userule{T-Or} \quad
+  \begin{gather*}
+    \boxed{\RelationT} \Break
+    \userule{T-Edit} \Quad
+    \userule{T-Enter} \Quad
+    \userule{T-Update} \Quad
+    \userule{T-Fail} \Break
+    \userule{T-Then} \Quad
+    \userule{T-Next} \Break
+    \userule{T-And} \Quad
+    \userule{T-Or} \Quad
     \userule{T-Xor}
-  \end{mathpar}
+  \end{gather*}
   \caption{Typing rules} \label{fig:typing-rules}
@@
   \centering
-  \includegraphics[width=\columnwidth,page=3]{figures/drawings-crop.pdf}
+  \includegraphics[width=0.6\columnwidth,page=3]{figures/drawings-crop.pdf}
   \caption{
@@
   let cancel : Task Unit = edit <<>> in
-  let delay : Int -> Task Unit -> Task Unit =
-    \n : Int. \proceed : Task Unit.
+  let delay : Int -> Task Unit -> Task Unit = \n : Int. \proceed : Task Unit.
     (proceed <?> cancel) <|> (wait n >>= \u : Unit. proceed)

### 2019-02-12 sections/relatedwork.tex
Updated related work section with subsections

@@
 
+\subsection{Worfklow modeling}
+
+Much research has been done into workflow modeling. These works focus on
+describing the collaboration between subsystems, rather than the communication
+between them.
+
 \paragraph{Workflow patterns}
 
-Much research has been done into workflow patterns. They are regarded as special
-kind of the design patterns in software engineering. They identify recurring patterns
-in workflow systems, much like the combinators defined by \TOPHAT. Work by
-Van der Aalst~\cite{journals/dpd/AalstHKB03} defines a comprehensive list of
-these pattens, and examines their availability in industry workflow software.
+Workflow patterns are regarded as special kind of the design patterns in
+software engineering. They identify recurring patterns in workflow systems, much
+like the combinators defined by \TOPHAT. Work by Van der
+Aalst~\cite{journals/dpd/AalstHKB03} defines a comprehensive list of these
+pattens, and examines their availability in industry workflow software.
 Workflow patterns are usually described in terms of control flow graphs, and no
@@
 
-\paragraph{Petri net}
+\paragraph{Workflow nets}
 
-Petri nets, and more specifically Workflow nets, have also been used to describe
-workflows~\cite{journals/jcsc/Aalst98}. They allow for the modelling and
-analysis of business processes. A downside of using workflow nets is that the
-model is not directly implementable, which is possible with the \TOPHAT calculus.
+Workflow nets~\cite{journals/jcsc/Aalst98} allow for the modeling and analysis
+of business processes. They are graphical in nature, and clearly display how
+every component is related to each other. A downside of Workflow nets is that
+they do not facilitate higher order constructs and that they are often not
+directly executable.
 
-% NEW RELATED WORK sections
+\paragraph{YAWL}
 
-\paragraph{Business process modeling}
+This is not always the case however. A language based on workflow nets that is
+actually directly executable is YAWL by van der
+Aalst~\cite{DBLP:journals/is/AalstH05}. It facilitates modeling and execution of
+dynamic workflows, with support for AND, OR and XOR workflow patterns. As
+mentioned, YAWL programs consist of Workflow Nets, and are therefore programmed
+visually.
 
-A plethora of languages exists to describe, model, display and execute these
-processes. YAWL by van der Aalst~\cite{DBLP:journals/is/AalstH05} is one of
-these languages. It facilitates modeling and execution of dynamic workflows,
-with support for AND, OR and XOR workflow patterns. YAWL programs consist of
-Nets, and are therefore programmed visually. BPEL\todo{cite} is another popular business process language.
-The standardized language allows for the specification of actions within
-business processes, using an XML format. Processes specified in BEPL are
-executable, just like YAWL.
+\paragraph{BPEL}
+
+BPEL\todo{cite} is another popular business process language. The standardized
+language allows for the specification of actions within business processes,
+using an XML format. Processes specified in BEPL are executable, just like YAWL.
+
+\subsection{Related web services}
 

### 2019-02-12 sections/relatedwork.tex
Added citations to related work

@@
 
-BPEL\todo{cite} is another popular business process language. The standardized
+BPEL~\cite{bpel} is another popular business process language. The standardized
 language allows for the specification of actions within business processes,
@@
 
-If This Then That (IFTTT)\todo{cite} is a web service that allows users to write and
+If This Then That (IFTTT)~\cite{IFTTT} is a web service that allows users to write and
 configure simple conditional statements called applets, that can interact with
@@
 
-Google forms\todo{cite} and Wufoo\todo{cite} are web based survey administration apps that allows
+Google forms~\cite{googleforms} and Wufoo~\cite{wufoo} are web based survey administration apps that allows
 users to compose forms and gather data from them. They are comparable to TOP and

### 2019-02-12 sections/comparison.tex
compact code examples
first stab at integrating processes in intuition section

@@
 
-This section gives an overview of the essential concepts of \TOP and how they are represented in \TOPHAT.
-While \TOP is focussed on collaboration, it gives rise to concepts like communication and multiprogramming.
+%NOTE: Use \emph{process} as a general term for a process in process algebras and a reactive system from reactive programming?
+%NOTE: Main focus should be on \emph{collaboration} between tasks opposed to \emph{communication} between processes.
+
+This section gives an overview of the essential concepts of tasks and how they are represented in \TOPHAT.
+While tasks are focussed on collaboration, they give rise to concepts like communication, concurrency and synchronisation.
 % These concepts are: communication with the environment, communication between subsystems, sequential composition, concurrency, and synchronisation.
-These concepts can also be found in process algebras, which raises the question how \TOPHAT and process algebras relate.
-While introducing the concepts, we compare them with Hoare's Communicating Sequential Processes (\CSP).
-The comparison is based on Hoare's book \cite{books/Hoare85CSP}.
-We chose \CSP as an example to stand for the numerous process algebras in existence, which, as far as this section is concerned, are reasonably similar.
+These concepts can also be found in process algebras like \CSP, \CCS and \PICALC, \fixme{cite}
+and reactive programming languages like \ESTEREL, \HIPHOP, and \fixme{more?}.
+This raises the question how \TOPHAT relates to these algebras and languages.
 
-We provide comparisons based on two aspects.
-First, we compare the scope of the languages, as intended by the authors.
-Second, we focus on features common to \TOP and \CSP, and study their realisation.
+In the next subsections, we give an intuition about tasks, by discussing the concepts collaboration, communication \fixme{and multiprocessing}.
+While introducing these concepts, we compare them with \emph{processes}.
+We define a process as \fixme{definition}.
+% We chose processes as an example to stand for the numerous process algebras in existence, which, as far as this section is concerned, are reasonably similar.
+We focus on features common to tasks and processes, and study their realisation.
 When we point out differences, we do not argue that the different realisations are incompatible.
-As a matter of fact the primitives can certainly be expressed in terms of each other.
+As a matter of fact, primitives can certainly be expressed in terms of each other.
 Instead, we point out how the systems emphasise certain points of view by choosing different building blocks to express the same concepts.
 
-\paragraph{Scope}
+% we provide comparisons based on two aspects.
+% First, we compare the scope of the languages, as intended by the authors.
 
-The central objective of \TOP is to coordinate collaboration.
+\subsection{Collaboration}
+
+The central objective of tasks is to coordinate collaboration.
 Therefore it has combinators to express ways in which people can work together to reach a common goal.
@@
 This motivates the different forms of communication: with the environment, along control flow, and across control flow.
-\TOPHAT is a language to specify workflows.
+Tasks are a way to model workflows.
+% \TOPHAT is a language to specify workflows.
 
-The central objective of \CSP is to model the patterns of behaviour of processes.
+The central objective of processes is to model patterns of their behaviour.
 These patterns of behaviour manifest themselves in sequences of actions that processes can engage in.
-\CSP is a language to specify synchronisation patterns.
+Actions can be events from the environment, or signals send internally between processes.
+Processes are a way to model synchronisation patterns.
 
-\begin{quote}
-\noindent In \TOP, communication is a means to an end. \\
-In \CSP, communication is the central matter of discourse.
-\end{quote}
+% \begin{quote}
+%   \noindent In \TOP, communication is a means to an end. \\
+%   In \CSP, communication is the central matter of discourse.
+% \end{quote}
 
@@
 % \TOP focuses less on formal correctness, and more on practical applicability.
-The big difference between \CSP and \TOP is the intended level of abstraction.
-While \CSP is used to explicitly model low-level details about communication, \TOPHAT wants to hide details and provide high-level communication mechanisms to the programmer.
+The big difference between processes and tasks is the intended level of abstraction.
+While processes are used to explicitly model low-level details about communication, tasks hide these details and provide high-level communication mechanisms to the programmer.
 It wants to be a calculus with intuitive semantics that facilitates communication between programmers and domain experts.
@@
 
-Both \TOP and \CSP have two sorts of communication: with the environment, and between subsystems.
-In both languages, communication with the \emph{environment} is input and output.
+Both tasks and processes have two sorts of communication: with the environment, and between subsystems.
+In both paradigms, communication with the \emph{environment} is input and output.
 They block evaluation of programs until the environment sends or receives events.
-In both languages, communication \emph{between subsystems} does not block.
+In both paradigms, communication \emph{between subsystems} does not block.
 They evaluate programs as far as possible until further external input is required.
@@
 A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive any change event with a value of type $\tau$ and additionally the special event $\Empty$ that empties the editor.
-The environment can always inspect the current value of an editor.
+There is no explicit communication of output,
+However, the environment can always inspect the current value of an editor.
 
@@
 
-In \CSP, communication with the environment is represented by prefixing.
-The process $(a \to P)$ can engage in action $a$, after which it continues as process $P$.
+Processes communicate with the environment by sending or receiving actions
+The \CSP process $(a \to P)$ can engage in action $a$, after which it continues as process $P$.
+% Prefixing has no direction of sending and receiving.
+Hoare uses the neutral phrase \emph{engaging in an action}.
+The interpretation of whether an action stands for input or output is left to the reader.
+For example, a vending machine $P = (\textit{coin} \to \textit{choc} \to P)$ is to be interpreted as taking a coin as input and producing chocolate as output.
 %%NOTE: Not that important I think, we don't use it in the rest of the text. --TS
@@
 % He seems to permit any kind of mathematical formula, or any programming language.
-%
-Prefixing has no direction of sending and receiving.
-Hoare uses the neutral phrase \emph{engaging in an action}.
-The interpretation of whether an action stands for input or output is left to the reader.
-For example, a vending machine $P = (\textit{coin} \to \textit{choc} \to P)$ is to be interpreted as taking a coin as input and producing chocolate as output.
+
+The equivalent process in \ESTEREL is the reactive statement $\text{await} a; P$.
+Here, action $a$ is interpreted as an input.
+To emit an action, one can use $\text{emit} a$.
 
@@
 %
-In \TOP, event values are typed to match the type of the receiving editor.
-In \CSP, action names are essentially strings, subject to implicit restrictions.
+In \TOP, input values are typed to match the type of the receiving editor.
+In \CSP and \ESTEREL, action names are essentially strings, subject to implicit restrictions.
 
@@
 \begin{equation*}
-  V = \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
+  V_\text{\CSP} = \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
+\end{equation*}
+An equivalent reactive system in \ESTEREL could be defined as\fixme{validate}
+\begin{equation*}
+  V_\text{\ESTEREL} = \text{await } \textit{coin}; ( \text{await } \textit{button}; \text{emit } \textit{small} \parallel \text{await } \textit{coin}; \text{await } \textit{button}; \text{emit } \textit{large})
 \end{equation*}
@@
 \begin{TASK}
-  enter Int >>? \n. if n == 1 then edit "small"
-    else if n == 2 then edit "large" else fail
+  enter Int >>? \n. if n == 1 then edit "small" else if n == 2 then edit "large" else fail
 \end{TASK}
-This machine does not exhibit exactly the same behaviour as the one in \CSP, nor is it intended to.
+This machine does not exhibit exactly the same behaviour as the ones in \CSP or \ESTEREL, nor is it intended to.
 Instead, it just demonstrates how the same choice
@@
 % of buying a small biscuit for one coin or a large biscuit for two coins
-can be expressed in \TOP.
+can be expressed in \TOPHAT.
 The task asks the user to enter an amount of money.
@@
 
-\TOP has two methods for internal communication, both different from its meth\-od for external communication.
+\TOPHAT has two methods for internal communication, both different from its method for external communication.
 The step combinator represents data flow that follows control flow, and shared data represents data flow across control flow.
-
+%
 The step combinators in $t \Then e$ and $t \Next e$ pass the value of $t$ to the continuation $e$.
@@
 
+\fixme{Add short intro on communication between non-sequential tasks.}
 Shared data sources are assignable references, whose changes in value are immediately visible to all tasks interested in them.
@@
 
+\fixme{What to do with this paragraph?}
 \CSP uses prefixing for both internal and external communication.
@@
 
-\subsection{Multiprogramming}
+% \subsection{Multiprogramming}
 
-Both \TOP and \CSP are models of \emph{multiprogramming}, as opposed to \emph{multiprocessing}.
+\subsection{Concurrency}
+
+Both tasks and processes are models of \emph{multiprogramming}, as opposed to \emph{multiprocessing}.
 This means that one processor evaluates multiple programs in an interleaving fashion.
@@
 
-\paragraph{Concurrency}
-
 Concurrency is the source of parallelism: two things are done at the same time.
 %
-In \TOP, the parallel combinator $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
+Using tasks $t_1$ and $t_2$, the parallel combination $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
 For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
@@
 
-\paragraph{Synchronisation}
+\subsection{Synchronisation}
 
@@
 
-\paragraph{Nondeterminism}
+\subsection{Nondeterminism}
 
@@
 
-\paragraph{Conclusion}
+\subsection{Recap}
 

### 2019-02-12 sections/relatedwork.tex
typo
workflows are like GOTO

@@
 As mentioned earlier, iTasks is an implementation of \TOP. iTasks has many
-features, and its basic combinators are versatile and powerful.vSimpler
+features, and its basic combinators are versatile and powerful. Simpler
 combinators are implemented by restricting the powerful ones. This is useful for
@@
 between them.
+The systems described here follow a \emph{boxes and arrows} model of specifying workflows.
+Control flow, represented by arrows, usually can go unrestricted from anywhere to anywhere else in a workflow.
+We see \TOP as the functional programming of workflows, as opposed to this GOTO style.
 

### 2019-02-13 sections/comparison.tex
add some more todos

@@
 For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
-All events sent to the combined task are interleaved into a serial stream of inputs.
+All inputs sent to the combined task are interleaved into a serial stream of inputs.
 The tasks are truly independent of each other, all interleavings are permitted.
-Events to $t_1$ and $t_2$ must be prefixed by $\First$(irst) and $\Second$(econd) respectively.
-This unambiguously renames the events, removing any source of nondeterminism.
+Inputs to $t_1$ and $t_2$ must be prefixed by $\First$(irst) and $\Second$(econd) respectively.
+This unambiguously renames the inputs, removing any source of nondeterminism.
 
@@
 Synchronised steps are the only occasion where two processes actually do something at the same time.
-
+%
 The interleaving operator $P \mid\mid\mid Q$ does not synchronise any actions of $P$ and $Q$.
@@
 Actions that occur in both alphabets are nondeterministically taken by either $P$ or $Q$.
+
+\fixme{Add something about reactive programming.}
 
@@
 All synchronisation problems in \CSP must be solved by employing synchronised steps in some form.
+
+\fixme{Add something about reactive programming.}
 
@@
 
+\TOP is meant solely for implementation and does not have any form of nondeterminism.
+Input events for parallel tasks are disambiguated, internal steps have a well-defined evaluation order, and internal choice is left-biased.
+
 \CSP has an operator to explicitly introduce nondeterminism.
@@
 
-\TOP is meant solely for implementation and does not have any form of nondeterminism.
-Input events for parallel tasks are disambiguated, internal steps have a well-defined evaluation order, and internal choice is left-biased.
+\fixme{Add something about reactive programming.}
+\fixme{Add combination function for signals that are emitted at the same instant.}
+
+\subsection{Program generation}
+
+\fixme{Add section about generating executable applications?}
 
@@
 
-There are several concepts that exist in both languages, like concurrency, or sending values between concurrent agents.
-In \TOP however they are meant to be used on a higher level of abstraction, which allows concise declarative description of patterns of collaboration, without worrying about the implementation details of communication.
-By focussing on collaboration, \TOP and \TOPHAT lead to high-level specifications closer to real-world workflows which, at the same time, can be used to generate distributive, multi-user applications to support such workflows.
+There are several concepts that exist for both tasks and processes, like concurrency, or sending values between concurrent agents.
+Using tasks, however, they are meant to be used on a higher level of abstraction, which allows concise declarative description of patterns of collaboration, without worrying about the implementation details of communication.
+By focussing on collaboration, tasks and \TOPHAT lead to high-level specifications closer to real-world workflows which, at the same time, can be used to generate distributive, multi-user applications to support such workflows.

### 2019-02-13 sections/relatedwork.tex

@@
 preset data-type that drives the behavior of the user defined applications.
+
+\subsection{Functional Reactive Programming}
+
+Functional Reactive Programming (FRP) is a paradigm to describe graphs of values, called behaviors, that depend on each other and on external events.
+Behaviors can change over time, or triggered by events.
+When a behavior changes, all other behaviors that depend on it are updated automatically.
+
+There are some concepts that are similar in \TOP and FRP.
+The \emph{stepper} behavior is associated with an event and yields the value of the most recent event.
+This is similar to editors in \TOP.

### 2019-02-13 sections/relatedwork.tex
TOP vs FRP

--- 2019-02-12 sections/relatedwork.tex
@@
 preset data-type that drives the behavior of the user defined applications.
+
+\subsection{Functional Reactive Programming}
+
+Functional Reactive Programming (FRP) is a paradigm to describe dynamic changes of values in a declarative way.
+This is done by specifying networks of values, called behaviors, that can depend on each other and on external events.
+Behaviors can change over time, or triggered by events.
+When a behavior changes, all other behaviors that depend on it are updated automatically.
+The underlying implementation that takes care of the updating usually can tie input devices, like mouse and keyboard, to event streams and behaviors to output facilities, like text fields.
+This allows for declarative specifications of applications with user interfaces.
+
+The idea of FRP was pioneered by \citet{conf/icfp/ElliottH97}, in the meantime there are many variants and implementations, with reactive-banana one of the most well-known \cite{reactive-banana}.
+
+FRP and \TOP are different systems that have different goals in mind.
+Whereas FRP expresses automatically updating data dependencies, \TOP expresses collaboration patterns.
+\TOP has no notion of time.
+Tasks can not change over time, while behaviors can.
+The biggest conceptual difference between a workflow in \TOP and a data network in FRP is that an event to a task only causes updates up until the next step, while an event in FRP propagates through the whole network.
+
+That being said, there are some concepts that are similar in \TOP and FRP.
+The \emph{stepper} behavior for example is associated with an event and yields the value of the most recent event.
+This is similar to editors in \TOP.
+Furthermore, both systems can be used to declaratively program user interfaces.

### 2019-02-13 sections/relatedwork.tex

@@
 This is similar to editors in \TOP.
-Furthermore, both systems can be used to declaratively program user interfaces.
+Furthermore, both systems can be used to declaratively program user interfaces, albeit in FRP the programmer has to construct the GUI elements manually, and connect inputs and outputs to the correct events and behaviors.
+In \TOP graphical user interfaces are automatically derived.

### 2019-02-13 sections/contributions.tex
Added old list-style contributions to current contributions write-up

@@
 
-% \paragraph{Contributions}
+\subsection{Contributions}
 
@@
 \TOP, and therefore \TOPHAT, focus on supporting \emph{tasks} and model ways to \emph{collaborate} on them.
+
+\begin{itemize*}
+  \item A calculus to model workflows, which can be used in a modular way and are driven by user inputs.\\
+  \item Embed this calculus in a simply typed $\lambda$-calculus.\\
+  \item Its operational semantics, presented and defined in a layered way.\\
+  \item Semantic observations on the calculus which can be used to:\\
+  \begin{itemize*}
+    \item calculate stuck configurations,\\
+    \item associate a value with a task,\\
+    \item draw a user interface (\UI) for a task,\\
+    \item calculate input possibilities for the user.\\
+  \end{itemize*}
+  \item A clear separation between \TOP principles and general purpose programming constructs.\\
+\end{itemize*}

### 2019-02-13 sections/relatedwork.tex
related work: session types

@@
 In \TOP graphical user interfaces are automatically derived.
+
+\subsection{Session types}
+
+Session types are a type discipline that can be used to check whether communicating programs conform to a certain protocol.
+Session types are expressions in some process calculus that describe the input/output behavior of such programs.
+Session types are useful for programming languages where modules communicate with each other via messages, like CSP, $\pi$-calculus, or Go, to name a few.
+The only form of messages in \TOP are input events which drive execution, but modules do not communicate using messages.
+Therefore, session types are not applicable to \TOP in the sense used in the literature.
+
+Formal reasoning about \TOP programs is one of our future goals for \TOPHAT.
+The ideas and techniques of session types could be useful for specifying that a list of inputs of a certain form leads to desired task values.
+The details are a topic for future work.

### 2019-02-13 sections/comparison.tex
split comparison into intuition of tasks and move cps to related work

--- 2019-02-06 sections/comparison.tex
@@
 
-This section gives an overview of the essential concepts of \TOP and how they are represented in \TOPHAT.
-While \TOP is focussed on collaboration, it gives rise to concepts like communication and multiprogramming.
-% These concepts are: communication with the environment, communication between subsystems, sequential composition, concurrency, and synchronisation.
-These concepts can also be found in process algebras, which raises the question how \TOPHAT and process algebras relate.
-While introducing the concepts, we compare them with Hoare's Communicating Sequential Processes (\CSP).
-The comparison is based on Hoare's book \cite{books/Hoare85CSP}.
-We chose \CSP as an example to stand for the numerous process algebras in existence, which, as far as this section is concerned, are reasonably similar.
+%NOTE: Use \emph{process} as a general term for a process in process algebras and a reactive system from reactive programming?
+%NOTE: Main focus should be on \emph{collaboration} between tasks opposed to \emph{communication} between processes.
 
-We provide comparisons based on two aspects.
-First, we compare the scope of the languages, as intended by the authors.
-Second, we focus on features common to \TOP and \CSP, and study their realisation.
-When we point out differences, we do not argue that the different realisations are incompatible.
-As a matter of fact the primitives can certainly be expressed in terms of each other.
-Instead, we point out how the systems emphasise certain points of view by choosing different building blocks to express the same concepts.
-
-\paragraph{Scope}
-
-The central objective of \TOP is to coordinate collaboration.
-Therefore it has combinators to express ways in which people can work together to reach a common goal.
-Tasks can be executed after each other, at the same time, or conditionally.
-This motivates the task combinators for steps, parallel combination, and choice.
-To employ these patterns of collaboration, users need to communicate.
-This motivates the different forms of communication: with the environment, along control flow, and across control flow.
-\TOPHAT is a language to specify workflows.
-
-The central objective of \CSP is to model the patterns of behaviour of processes.
-These patterns of behaviour manifest themselves in sequences of actions that processes can engage in.
-\CSP is a language to specify synchronisation patterns.
-
-\begin{quote}
-\noindent In \TOP, communication is a means to an end. \\
-In \CSP, communication is the central matter of discourse.
-\end{quote}
-
-%%NOTE: Focus on what TOP is good at --TS
-% \CSP has a formal semantics that allows various kinds of correctness proofs, including equality of processes, and adherence to a specification.
-% This allows applications in program correctness, proofs of deadlock freedom, liveness, or verification of protocols.
-%
-% \TOP focuses less on formal correctness, and more on practical applicability.
-The big difference between \CSP and \TOP is the intended level of abstraction.
-While \CSP is used to explicitly model low-level details about communication, \TOPHAT wants to hide details and provide high-level communication mechanisms to the programmer.
-It wants to be a calculus with intuitive semantics that facilitates communication between programmers and domain experts.
-Such programs are supposed to hide implementation details from domain experts while containing enough information to allow automatic generation of executable applications, including user interfaces.
-
-%%NOTE: Already said in section intro --TS
-% Despite the different origins, \TOP and \CSP feature concepts that serve similar purposes.
+\subsection{Collaboration}
 
@@
 
-Both \TOP and \CSP have two sorts of communication: with the environment, and between subsystems.
-In both languages, communication with the \emph{environment} is input and output.
-They block evaluation of programs until the environment sends or receives events.
-In both languages, communication \emph{between subsystems} does not block.
-They evaluate programs as far as possible until further external input is required.
-
-\paragraph{Communication with the environment}
-
-In \TOPHAT, communication with the environment is represented by editors.
-A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive any change event with a value of type $\tau$ and additionally the special event $\Empty$ that empties the editor.
-The environment can always inspect the current value of an editor.
-
-Editors have no notion of continuation.
-Their sole purpose is to interact with the environment, retaining the last value that has been sent to them.
-%
-They can be used for both input and output.
-The interpretation of whether an editor represents input or output is left to the reader.
-An empty editor is most commonly interpreted as a prompt to input data, while a filled editor can be seen either as outputting its value, or as an input that comes with a default value.
-
-In \CSP, communication with the environment is represented by prefixing.
-The process $(a \to P)$ can engage in action $a$, after which it continues as process $P$.
-%%NOTE: Not that important I think, we don't use it in the rest of the text. --TS
-% In general, if $B$ is a set of events and $P(x)$ is an expression that evaluates to a process for each $x \in B$, the process $(x:B \to P(x))$ can engage in any one of the actions in $B$, after which it continues as the process determined by $P(x)$.
-% Hoare does not specify the language in which $P(x)$ is to be expressed.
-% He seems to permit any kind of mathematical formula, or any programming language.
-%
-Prefixing has no direction of sending and receiving.
-Hoare uses the neutral phrase \emph{engaging in an action}.
-The interpretation of whether an action stands for input or output is left to the reader.
-For example, a vending machine $P = (\textit{coin} \to \textit{choc} \to P)$ is to be interpreted as taking a coin as input and producing chocolate as output.
-
-Sending and receiving of values is modelled by giving additional structure to the names of actions.
-For example, an action name could be $\textit{in}.5$, which can be interpreted as inputting the value 5.
-%
-In \TOP, event values are typed to match the type of the receiving editor.
-In \CSP, action names are essentially strings, subject to implicit restrictions.
-
-\begin{example}[Vending machine]
-
-This example demonstrates external communication and choice.
-The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins.
-\begin{equation*}
-  V = \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
-\end{equation*}
-
-A vending machine in \TOPHAT that is close in spirit to the one above looks as follows.
-\begin{TASK}
-  enter Int >>? \n. if n == 1 then edit "small"
-    else if n == 2 then edit "large" else fail
-\end{TASK}
-This machine does not exhibit exactly the same behaviour as the one in \CSP, nor is it intended to.
-Instead, it just demonstrates how the same choice
-%%NOTE: Problem is already stated --TS
-% of buying a small biscuit for one coin or a large biscuit for two coins
-can be expressed in \TOP.
-The task asks the user to enter an amount of money.
-This editor stands for a coin slot in a real machine that freely accepts and returns coins.
-There is a continue button that is initially disabled.
-When the user has inserted exactly 1 or 2 coins, the continue button becomes enabled.
-When the user presses the continue button, the machine dispenses either a large or a small biscuit, depending on the amount of money.
-
-\end{example}
+The equivalent process in \ESTEREL is the reactive statement $\text{await} a; P$.
+Here, action $a$ is interpreted as an input.
+To emit an action, one can use $\text{emit} a$.
 
@@
 
-\TOP has two methods for internal communication, both different from its meth\-od for external communication.
-The step combinator represents data flow that follows control flow, and shared data represents data flow across control flow.
+% \subsection{Multiprogramming}
 
-The step combinators in $t \Then e$ and $t \Next e$ pass the value of $t$ to the continuation $e$.
-% The intuition about sequential composition is to first do one thing, and once it is done, do some other thing.
-% In \TOP, tasks never terminate.
-% Nonetheless, the notion of sequential composition exists in two forms: the external and the internal step.
-They both act like $t$ as long as the step is guarded.
-Once it becomes unguarded, the task continues as $e\ v$, where $v$ is the value of $t$.
-The task $t \Next e$ requires an input event $\Continue$ in addition to the step being unguarded. % in order to step.
-% \CSP's semicolon is like an internal step $t \Then e$ in \TOP where $t$ uses $\Unit$ to signal when it is done, and $e$ never fails.
-
-Shared data sources are assignable references, whose changes in value are immediately visible to all tasks interested in them.
-The semantics of \TOPHAT requires all updates to shared data and all enabled internal steps to be processed before any further communication with the environment can take place.
-This introduces the possibility of diverging computations, where a cyclic dependency between shared data causes internal steps to be taken continuously without ever being able to engage in external communication.
-
-\CSP uses prefixing for both internal and external communication.
-Prefixing becomes internal by using \emph{concealment}.
-Processes must take any concealed steps as soon as possible, before any further communication with the environment can take place.
-This means concealed actions have priority over exposed ones.
-This introduces the possibility of diverging computations, that is processes which continuously take internal steps without ever being able to engage in external communication.
-
-% \paragraph{Sequential composition}
-
-% In \CSP, sequential composition is represented by the semicolon combinator together with the special action $\checkmark$, called \emph{success}.
-% We require that if a process engages in $\checkmark$ at all, it must be its single last action.
-% When that happens, we say the process terminates successfully.
-% The process $P;Q$ behaves like $P$ until $P$ terminates successfully, after which it continues as $Q$.
-% If $P$ never terminates successfully, neither does $P;Q$
-
-\subsection{Multiprogramming}
-
-Both \TOP and \CSP are models of \emph{multiprogramming}, as opposed to \emph{multiprocessing}.
-This means that one processor evaluates multiple programs in an interleaving fashion.
-% instead of multiple processors.
-Multiprogramming involves three aspects: concurrency, synchronisation, and nondeterminism.
-
-\paragraph{Concurrency}
-
-Concurrency is the source of parallelism: two things are done at the same time.
-%
-In \TOP, the parallel combinator $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
-For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
-All events sent to the combined task are interleaved into a serial stream of inputs.
-The tasks are truly independent of each other, all interleavings are permitted.
-Events to $t_1$ and $t_2$ must be prefixed by $\First$(irst) and $\Second$(econd) respectively.
-This unambiguously renames the events, removing any source of nondeterminism.
-
-In \CSP, there are two different combinators for parallel composition: parallel and interleave.
-The parallel process $P \parallel Q$ can take $P$-steps and $Q$-steps in arbitrary interleaving for actions unique to $P$ and $Q$.
-Actions in the alphabets of both $P$ and $Q$ must be taken in synchronisation, so if one process wants to take such a step, it must wait until the other is ready to do so.
-Synchronised steps are the only occasion where two processes actually do something at the same time.
-
-The interleaving operator $P \mid\mid\mid Q$ does not synchronise any actions of $P$ and $Q$.
-All interleavings are permitted.
-Actions that occur in both alphabets are nondeterministically taken by either $P$ or $Q$.
-
-\paragraph{Synchronisation}
-
-Synchronisation means that an agent has to pause execution until some condition is met.
-The need for synchronisation arises quite naturally in countless variations in situations involving concurrency.
-For multiprogramming, synchronisation always means that some of the possible interleavings are forbidden.
-
-For example, the mutual exclusion problem for two parallel processes $a_p \to b_p \to \text{STOP} $ and $a_q \to b_q \to \text{STOP}$ can be stated as:
-In no interleaving should there be two adjacent $a$'s.
-%
-The mutual exclusion problem for two parallel tasks can be stated as follows.
-Let \TS{inc} and \TS{dec} be tasks that increment and decrement some shared counter.
-Then, in \TS{(inc x >>? dec) <&> (dec x >>? inc)}, the shared counter should at no point be greater than 1.
-This means that in no interleaving should there be two adjacent \TS{inc}'s.
-
-The basic method for synchronisation in \TOP is built into the step combinator.
-The task $t \Next e$ can only continue execution when two conditions are met:
-Task $t$ must have a value $v$, and $e\ v$ must not evaluate to $\Fail$.
-Programmers can encode arbitrary conditions in $e\ v$, which are evaluated atomically between interaction steps.
-This allows a variety of synchronisation problems to be solved in an intuitive and straight-forward manner.
-
-The basic method for synchronisation in \CSP is synchronised prefixing.
-If some action $a$ is in the alphabets of both $P$ and $Q$, then their parallel composition $P \parallel Q$ can take an $a$-step only if both $P$ and $Q$ are ready to take an $a$-step, in which case they both take this step simultaneously.
-Both have to wait for the condition that the other process is ready to take the step.
-All synchronisation problems in \CSP must be solved by employing synchronised steps in some form.
-
-\begin{example}[Cigarette smokers]
-
-The cigarette smokers problem \cite{books/Downey08LBOS} is a surprisingly tricky synchronisation problem.
-We study it here because it demonstrates the capabilities of guarded steps.
-% The problem requires waiting for two conditions, waking up only if both conditions are satisfied.
-The problem is stated as follows.
-In order to smoke a cigarette, three ingredients are required: tobacco, paper, and a match.
-There are three smokers, each having one of the ingredients and requiring the other two.
-There is an agent that randomly provides two of the ingredients.
-
-Downey models availability of the ingredients using a semaphore for each ingredient.
-The agent randomly signals two of the three semaphores.
-% The problem asks writing code for the smokers so that the smoker who requires the two signalled semaphores wakes up and proceeds with smoking.
-%
-% A naïve attempt where each smoker first waits for one and then the other semaphore can lead to deadlock.
-% The following pseudo-code is a non-solution.
-% Let \emph{paper}, \emph{tobacco} and \emph{match} be semaphores, initialised to 0.
-% \begin{TASK}
-%   tobaccoSmoker = wait(match); wait(paper); smoke();
-%   paperSmoker = wait(tobacco); wait(match); smoke();
-%   matchSmoker = wait(tobacco); wait(paper); smoke();
-% \end{TASK}
-% To see what can go wrong, imagine the three smokers running in parallel.
-% Let the agent signal the two semaphores \emph{tobacco} and \emph{paper}.
-% One of \emph{paperSmoker} and \emph{matchSmoker} can take their first step, but if \emph{paperSmoker} takes the step, the system deadlocks.
-%
-The solution proposed by Downey involves an additional mutex, three additional semaphores, three additional threads called \emph{pushers}, and three regular Boolean variables.
-The job of the pushers is to record availability of their ingredient in their Boolean variable, and check availability of other resources, waking the correct smoker when appropriate.
-The details of the solution are not important here.
-
-What is important is that the implementation of what is essentially waiting for two semaphores requires a substantial amount of additional synchronisation, together with non-trivial conditional statements.
-%
-A similar effort as in Downey's solution must be made in order to solve the problem using synchronised actions in \CSP.
-This is because in \CSP, processes can only synchronise on one action at a time.
-
-\TOP allows a simple solution to this problem, using guarded steps.
-Steps can be guarded with arbitrary expressions, and the parallel combinator can be used to watch two shared editors ($\Update$) at the same time.
-Let \TS{match}, \TS{paper}, and \TS{tobacco} be references to Booleans.
-The smokers are defined as follows.
-\begin{TASK}
-  let continue = \<<x,y>>. if x /\ y then smoke else fail in
-  let tobaccoSmoker = (update match <&> update paper) >>? continue in
-  let paperSmoker = (update tobacco <&> update match) >>? continue in
-  let matchSmoker = (update tobacco <&> update paper) >>? continue
-\end{TASK}
-When the agent supplies two of the ingredients by setting the respective shares to \TS{True}, only the step of the smoker that waits for those becomes enabled.
-
-\end{example}
-
-\paragraph{Nondeterminism}
-
-Nondeterminism means that a system can react to the same input in different ways, the choice of which can not be influenced by the environment.
-
-\CSP has an operator to explicitly introduce nondeterminism.
-Furthermore, nondeterminism can arise from the combination of some other combinators, for example choice and concealment.
-%
-Hoare states that nondeterminism is only useful for \emph{specifying} processes, never for \emph{implementing} them \cite{books/Hoare85CSP}.
-\CSP can be used for both.
-A process with nondeterministic behaviour must always be regarded as a specification, while a deterministic process can be seen as specification or implementation.
-
-\TOP is meant solely for implementation and does not have any form of nondeterminism.
-Input events for parallel tasks are disambiguated, internal steps have a well-defined evaluation order, and internal choice is left-biased.
-
-\paragraph{Conclusion}
-
-There are several concepts that exist in both languages, like concurrency, or sending values between concurrent agents.
-In \TOP however they are meant to be used on a higher level of abstraction, which allows concise declarative description of patterns of collaboration, without worrying about the implementation details of communication.
-By focussing on collaboration, \TOP and \TOPHAT lead to high-level specifications closer to real-world workflows which, at the same time, can be used to generate distributive, multi-user applications to support such workflows.
+\fixme{Add something about reactive programming.}
+\fixme{Add combination function for signals that are emitted at the same instant.}

### 2019-02-13 sections/intuition.tex
split comparison into intuition of tasks and move cps to related work

% !TEX root=../icfp2019.tex

\section{Intuition}
\label{sec:intuition}

This section gives an overview of the core concepts of tasks in task oriented programming (\TOP).

\subsection{Tasks model real world collaborations between people and machines}

The central objective of tasks is to \emph{coordinate collaboration}.
Therefore it has combinators to express ways in which people can work together to reach a common goal.
Tasks can be executed after each other, at the same time, or conditionally.
This motivates the task combinators for steps, parallel combination, and choice.
One can say that tasks are a way to model \emph{workflows}.
% \TOPHAT is a language to specify workflows.

\subsection{Tasks are reusable}

\fixme{Extend}
Tasks are modular and composable,
can be abstracted over zero or more arguments,
arguments could themselves be tasks,
thus we have higher order workflows.

\subsection{Tasks are driven by user input}

\fixme{Extend}
Tasks process input events from users.
Sensors are also regarded as users.
Gives rise to communication with the environment.

In \TOPHAT, communication with the environment is represented by editors.
A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive any change event with a value of type $\tau$ and additionally the special event $\Empty$ that empties the editor.
There is no explicit communication of output.
% However, the environment can always inspect the current value of an editor.

Editors have no notion of continuation.
Their sole purpose is to interact with the environment, retaining the last value that has been sent to them.
%
They can be used for both input and output.
The interpretation of whether an editor represents input or output is left to the reader.
An empty editor is most commonly interpreted as a prompt to input data, while a filled editor can be seen either as outputting its value, or as an input that comes with a default value.

\begin{example}[Vending machine]

This example demonstrates external communication and choice.
The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins.
\begin{equation*}
  \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
\end{equation*}
% An equivalent reactive system in \ESTEREL could be defined as\fixme{validate}
% \begin{equation*}
  % V_\text{\ESTEREL} = \text{await } \textit{coin}; ( \text{await } \textit{button}; \text{emit } \textit{small} \parallel \text{await } \textit{coin}; \text{await } \textit{button}; \text{emit } \textit{large})
% \end{equation*}

A vending machine in \TOPHAT that is close in spirit to the one above looks as follows.
\begin{TASK}
  enter Int >>? \n. if n == 1 then edit "small" else if n == 2 then edit "large" else fail
\end{TASK}
This machine does not exhibit exactly the same behaviour as the ones in \CSP, nor is it intended to.
Instead, it just demonstrates how the same choice
%%NOTE: Problem is already stated --TS
% of buying a small biscuit for one coin or a large biscuit for two coins
can be expressed in \TOPHAT.
The task asks the user to enter an amount of money.
This editor stands for a coin slot in a real machine that freely accepts and returns coins.
There is a continue button that is initially disabled.
When the user has inserted exactly 1 or 2 coins, the continue button becomes enabled.
When the user presses the continue button, the machine dispenses either a large or a small biscuit, depending on the amount of money.

\end{example}

\subsection{Tasks can be observed for their current status at all time}

\fixme{Extend}
The value of a task can be present or not.

Because every task has an observable value of a specified type,
it is always possible to generate a (graphical) user interface from a \TOPHAT specification.
%NOTE: depends on: editors
The structure of the specification prescribes the structure of the \UI and possible input events.
Editors form the leaves of such trees.

\subsection{Tasks are never done}

%NOTE: depends on: editors
As users can continuously update editors, tasks never terminate.

The decision to move on to the next task is not performed by the task itself.
It is based on a calculation to determine the next tasks.
The step combinators in $t \Then e$ and $t \Next e$ pass the value of $t$ to the continuation $e$.
They both act like $t$ as long as the step is guarded.
Once it becomes unguarded, the task continues as $e\ v$, where $v$ is the observed value of $t$.
% The continuation $e$ is a function which calculates the next task or $\Fail$, signalling no continuation is present.
The task $t \Next e$ requires an input event $\Continue$ in addition to the step being unguarded. % in order to step.

The step combinators give rise to a method of internal communication.
They represent data flow that \emph{follows} control flow.

% The intuition about sequential composition is to first do one thing, and once it is done, do some other thing.
% The notion of sequential composition exists in two forms: the external and the internal step.

\subsection{Tasks can share information}

Second method of internal communication between tasks is shared data.
Shared data represents data flow \emph{across} control flow.
Shared data sources are assignable references, whose changes in value are immediately visible to all tasks interested in them.
The semantics of \TOPHAT requires all updates to shared data and all enabled internal steps to be processed before any further communication with the environment can take place.

\fixme{?}
This introduces the possibility of diverging computations, where a cyclic dependency between shared data causes internal steps to be taken continuously without ever being able to engage in external communication.

\begin{example}[Cigarette smokers]

The cigarette smokers problem \cite{books/Downey08LBOS} is a surprisingly tricky synchronisation problem.
We study it here because it demonstrates the capabilities of guarded steps.
% The problem requires waiting for two conditions, waking up only if both conditions are satisfied.
The problem is stated as follows.
In order to smoke a cigarette, three ingredients are required: tobacco, paper, and a match.
There are three smokers, each having one of the ingredients and requiring the other two.
There is an agent that randomly provides two of the ingredients.

Downey models availability of the ingredients using a semaphore for each ingredient.
The agent randomly signals two of the three semaphores.
% The problem asks writing code for the smokers so that the smoker who requires the two signalled semaphores wakes up and proceeds with smoking.
%
% A naïve attempt where each smoker first waits for one and then the other semaphore can lead to deadlock.
% The following pseudo-code is a non-solution.
% Let \emph{paper}, \emph{tobacco} and \emph{match} be semaphores, initialised to 0.
% \begin{TASK}
%   tobaccoSmoker = wait(match); wait(paper); smoke();
%   paperSmoker = wait(tobacco); wait(match); smoke();
%   matchSmoker = wait(tobacco); wait(paper); smoke();
% \end{TASK}
% To see what can go wrong, imagine the three smokers running in parallel.
% Let the agent signal the two semaphores \emph{tobacco} and \emph{paper}.
% One of \emph{paperSmoker} and \emph{matchSmoker} can take their first step, but if \emph{paperSmoker} takes the step, the system deadlocks.
%
The solution proposed by Downey involves an additional mutex, three additional semaphores, three additional threads called \emph{pushers}, and three regular Boolean variables.
The job of the pushers is to record availability of their ingredient in their Boolean variable, and check availability of other resources, waking the correct smoker when appropriate.

The details of the solution are not important here.
What is important is that the implementation of what is essentially waiting for two semaphores requires a substantial amount of additional synchronisation, together with non-trivial conditional statements.

% A similar effort as in Downey's solution must be made in order to solve the problem using synchronised actions in \CSP.
% This is because in \CSP, processes can only synchronise on one action at a time.

\TOPHAT allows a simple solution to this problem, using guarded steps.
Steps can be guarded with arbitrary expressions, and the parallel combinator can be used to watch two shared editors ($\Update$) at the same time.
Let \TS{match}, \TS{paper}, and \TS{tobacco} be references to Booleans.
The smokers are defined as follows.
\begin{TASK}
  let continue = \<<x,y>>. if x /\ y then smoke else fail in
  let tobaccoSmoker = (update match <&> update paper) >>? continue in
  let paperSmoker = (update tobacco <&> update match) >>? continue in
  let matchSmoker = (update tobacco <&> update paper) >>? continue
\end{TASK}
When the agent supplies two of the ingredients by setting the respective shares to \TS{True}, only the step of the smoker that waits for those becomes enabled.

\end{example}

\subsection{Tasks are predictable}

Concurrency is the source of parallelism: two things are done at the same time.
%
Using tasks $t_1$ and $t_2$, the parallel combination $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
All inputs sent to the combined task are interleaved into a serial stream of inputs.
The tasks are truly independent of each other, all interleavings are permitted.
Inputs to $t_1$ and $t_2$ must be prefixed by $\First$(irst) and $\Second$(econd) respectively.
This unambiguously renames the inputs, removing any source of nondeterminism.

The basic method for synchronisation in \TOP is built into the step combinator.
The task $t \Next e$ can only continue execution when two conditions are met:
Task $t$ must have a value $v$, and $e\ v$ must not evaluate to $\Fail$.
Programmers can encode arbitrary conditions in $e\ v$, which are evaluated atomically between interaction steps.
This allows a variety of synchronisation problems to be solved in an intuitive and straight-forward manner.

\TOPHAT is meant solely for implementation and does not have any form of nondeterminism.
Input events for parallel tasks are disambiguated, internal steps have a well-defined evaluation order, and internal choice is left-biased.

\subsection{Recap}

While tasks are focussed on collaboration, they give rise to concepts like communication, concurrency and synchronisation.
Using tasks, however, they are meant to be used on a higher level of abstraction, which allows concise declarative description of patterns of collaboration, without worrying about the implementation details of communication.
By focussing on collaboration, tasks and \TOPHAT lead to high-level specifications closer to real-world workflows which, at the same time, can be used to generate distributive, multi-user applications to support such workflows.

To employ these patterns of collaboration, users need to communicate.
This motivates different forms of communication: with the environment, along control flow, and across control flow.

### 2019-02-13 sections/relatedwork.tex
split comparison into intuition of tasks and move cps to related work

--- 2019-02-13 sections/relatedwork.tex
@@
 
+\subsection{Process algebras}
+
+The central objective of processes is to model patterns of their behaviour.
+These patterns of behaviour manifest themselves in sequences of actions that processes can engage in.
+Actions can be events from the environment, or signals send internally between processes.
+Processes are a way to model synchronisation patterns.
+
+% \begin{quote}
+%   \noindent In \TOP, communication is a means to an end. \\
+%   In \CSP, communication is the central matter of discourse.
+% \end{quote}
+
+%%NOTE: Already said in section intro --TS
+% Despite the different origins, \TOP and \CSP feature concepts that serve similar purposes.
+
+%%NOTE: Focus on what TOP is good at --TS
+% \CSP has a formal semantics that allows various kinds of correctness proofs, including equality of processes, and adherence to a specification.
+% This allows applications in program correctness, proofs of deadlock freedom, liveness, or verification of protocols.
+%
+% \TOP focuses less on formal correctness, and more on practical applicability.
+The big difference between processes and tasks is the intended level of abstraction.
+While processes are used to explicitly model low-level details about communication, tasks hide these details and provide high-level communication mechanisms to the programmer.
+It wants to be a calculus with intuitive semantics that facilitates communication between programmers and domain experts.
+Such programs are supposed to hide implementation details from domain experts while containing enough information to allow automatic generation of executable applications, including user interfaces.
+
+\paragraph{Communication}
+
+Both tasks and processes have two sorts of communication: with the environment, and between subsystems.
+In both paradigms, communication with the \emph{environment} is input and output.
+They block evaluation of programs until the environment sends or receives events.
+In both paradigms, communication \emph{between subsystems} does not block.
+They evaluate programs as far as possible until further external input is required.
+
+Processes communicate with the environment by sending or receiving actions
+The \CSP process $(a \to P)$ can engage in action $a$, after which it continues as process $P$.
+% Prefixing has no direction of sending and receiving.
+Hoare uses the neutral phrase \emph{engaging in an action}.
+The interpretation of whether an action stands for input or output is left to the reader.
+For example, a vending machine $P = (\textit{coin} \to \textit{choc} \to P)$ is to be interpreted as taking a coin as input and producing chocolate as output.
+%%NOTE: Not that important I think, we don't use it in the rest of the text. --TS
+% In general, if $B$ is a set of events and $P(x)$ is an expression that evaluates to a process for each $x \in B$, the process $(x:B \to P(x))$ can engage in any one of the actions in $B$, after which it continues as the process determined by $P(x)$.
+% Hoare does not specify the language in which $P(x)$ is to be expressed.
+% He seems to permit any kind of mathematical formula, or any programming language.
+
+Sending and receiving of values is modelled by giving additional structure to the names of actions.
+For example, an action name could be $\textit{in}.5$, which can be interpreted as inputting the value 5.
+%
+In \TOP, input values are typed to match the type of the receiving editor.
+In \CSP action names are essentially strings, subject to implicit restrictions.
+
+\paragraph{Communication between subsystems}
+
+%
+% \CSP's semicolon is like an internal step $t \Then e$ in \TOP where $t$ uses $\Unit$ to signal when it is done, and $e$ never fails.
+\CSP uses prefixing for both internal and external communication.
+Prefixing becomes internal by using \emph{concealment}.
+Processes must take any concealed steps as soon as possible, before any further communication with the environment can take place.
+This means concealed actions have priority over exposed ones.
+This introduces the possibility of diverging computations, that is processes which continuously take internal steps without ever being able to engage in external communication.
+
+% \paragraph{Sequential composition}
+
+% In \CSP, sequential composition is represented by the semicolon combinator together with the special action $\checkmark$, called \emph{success}.
+% We require that if a process engages in $\checkmark$ at all, it must be its single last action.
+% When that happens, we say the process terminates successfully.
+% The process $P;Q$ behaves like $P$ until $P$ terminates successfully, after which it continues as $Q$.
+% If $P$ never terminates successfully, neither does $P;Q$
+
+\paragraph{Concurrency}
+
+Both tasks and processes are models of \emph{multiprogramming}, as opposed to \emph{multiprocessing}.
+This means that one processor evaluates multiple programs in an interleaving fashion.
+% instead of multiple processors.
+Multiprogramming involves three aspects: concurrency, synchronisation, and nondeterminism.
+
+In \CSP, there are two different combinators for parallel composition: parallel and interleave.
+The parallel process $P \parallel Q$ can take $P$-steps and $Q$-steps in arbitrary interleaving for actions unique to $P$ and $Q$.
+Actions in the alphabets of both $P$ and $Q$ must be taken in synchronisation, so if one process wants to take such a step, it must wait until the other is ready to do so.
+Synchronised steps are the only occasion where two processes actually do something at the same time.
+%
+The interleaving operator $P \mid\mid\mid Q$ does not synchronise any actions of $P$ and $Q$.
+All interleavings are permitted.
+Actions that occur in both alphabets are nondeterministically taken by either $P$ or $Q$.
+
+\paragraph{Synchronisation}
+
+Synchronisation means that an agent has to pause execution until some condition is met.
+The need for synchronisation arises quite naturally in countless variations in situations involving concurrency.
+For multiprogramming, synchronisation always means that some of the possible interleavings are forbidden.
+
+For example, the mutual exclusion problem for two parallel processes $a_p \to b_p \to \text{STOP} $ and $a_q \to b_q \to \text{STOP}$ can be stated as:
+In no interleaving should there be two adjacent $a$'s.
+%
+The mutual exclusion problem for two parallel tasks can be stated as follows.
+Let \TS{inc} and \TS{dec} be tasks that increment and decrement some shared counter.
+Then, in \TS{(inc x >>? dec) <&> (dec x >>? inc)}, the shared counter should at no point be greater than 1.
+This means that in no interleaving should there be two adjacent \TS{inc}'s.
+
+The basic method for synchronisation in \CSP is synchronised prefixing.
+If some action $a$ is in the alphabets of both $P$ and $Q$, then their parallel composition $P \parallel Q$ can take an $a$-step only if both $P$ and $Q$ are ready to take an $a$-step, in which case they both take this step simultaneously.
+Both have to wait for the condition that the other process is ready to take the step.
+All synchronisation problems in \CSP must be solved by employing synchronised steps in some form.
+
+\paragraph{Nondeterminism}
+
+Nondeterminism means that a system can react to the same input in different ways, the choice of which can not be influenced by the environment.
+
+\CSP has an operator to explicitly introduce nondeterminism.
+Furthermore, nondeterminism can arise from the combination of some other combinators, for example choice and concealment.
+%
+Hoare states that nondeterminism is only useful for \emph{specifying} processes, never for \emph{implementing} them \cite{books/Hoare85CSP}.
+\CSP can be used for both.
+A process with nondeterministic behaviour must always be regarded as a specification, while a deterministic process can be seen as specification or implementation.
+
 \subsection{Related web services}

### 2019-02-13 sections/relatedwork.tex
use brittisch spelling of behaviour

--- 2019-02-13 sections/relatedwork.tex
@@
 Functional Reactive Programming (FRP) is a paradigm to describe dynamic changes of values in a declarative way.
-This is done by specifying networks of values, called behaviors, that can depend on each other and on external events.
-Behaviors can change over time, or triggered by events.
-When a behavior changes, all other behaviors that depend on it are updated automatically.
-The underlying implementation that takes care of the updating usually can tie input devices, like mouse and keyboard, to event streams and behaviors to output facilities, like text fields.
+This is done by specifying networks of values, called behaviours, that can depend on each other and on external events.
+Behaviours can change over time, or triggered by events.
+When a behaviour changes, all other behaviours that depend on it are updated automatically.
+The underlying implementation that takes care of the updating usually can tie input devices, like mouse and keyboard, to event streams and behaviours to output facilities, like text fields.
 This allows for declarative specifications of applications with user interfaces.
@@
 \TOP has no notion of time.
-Tasks can not change over time, while behaviors can.
+Tasks can not change over time, while behaviours can.
 The biggest conceptual difference between a workflow in \TOP and a data network in FRP is that an event to a task only causes updates up until the next step, while an event in FRP propagates through the whole network.
@@
 That being said, there are some concepts that are similar in \TOP and FRP.
-The \emph{stepper} behavior for example is associated with an event and yields the value of the most recent event.
+The \emph{stepper} behaviour, for example, is associated with an event and yields the value of the most recent event.
 This is similar to editors in \TOP.
-Furthermore, both systems can be used to declaratively program user interfaces, albeit in FRP the programmer has to construct the GUI elements manually, and connect inputs and outputs to the correct events and behaviors.
+Furthermore, both systems can be used to declaratively program user interfaces, albeit in FRP the programmer has to construct the GUI elements manually, and connect inputs and outputs to the correct events and behaviours.
 In \TOP graphical user interfaces are automatically derived.
@@
 Session types are a type discipline that can be used to check whether communicating programs conform to a certain protocol.
-Session types are expressions in some process calculus that describe the input/output behavior of such programs.
+Session types are expressions in some process calculus that describe the input/output behaviour of such programs.
 Session types are useful for programming languages where modules communicate with each other via messages, like CSP, $\pi$-calculus, or Go, to name a few.

### 2019-02-13 sections/relatedwork.tex
make clear tasks can change over time, but only after events
add more abbreviations and fix spelling in related work

@@
 
-\subsection{Worfklow modeling}
+\subsection{Worfklow modelling}
 
-Much research has been done into workflow modeling. These works focus on
+Much research has been done into workflow modelling. These works focus on
 describing the collaboration between subsystems, rather than the communication
@@
 Control flow, represented by arrows, usually can go unrestricted from anywhere to anywhere else in a workflow.
-We see \TOP as the functional programming of workflows, as opposed to this GOTO style.
+We see \TOP as the functional programming of workflows, as opposed to this \emph{goto}-style.
 
@@
 software engineering. They identify recurring patterns in workflow systems, much
-like the combinators defined by \TOPHAT. Work by Van der
-Aalst~\cite{journals/dpd/AalstHKB03} defines a comprehensive list of these
+like the combinators defined by \TOPHAT. Work by \citet{journals/dpd/AalstHKB03}
+defines a comprehensive list of these
 pattens, and examines their availability in industry workflow software.
@@
 
-\paragraph{Workflow nets}
+\paragraph{Workflow Nets}
 
-Workflow nets~\cite{journals/jcsc/Aalst98} allow for the modeling and analysis
+Workflow Nets~(\WFN)\cite{journals/jcsc/Aalst98} allow for the modelling and analysis
 of business processes. They are graphical in nature, and clearly display how
-every component is related to each other. A downside of Workflow nets is that
+every component is related to each other. A downside of \WFN is that
 they do not facilitate higher order constructs and that they are often not
@@
 
-This is not always the case however. A language based on workflow nets that is
-actually directly executable is YAWL by van der
-Aalst~\cite{DBLP:journals/is/AalstH05}. It facilitates modeling and execution of
-dynamic workflows, with support for AND, OR and XOR workflow patterns. As
-mentioned, YAWL programs consist of Workflow Nets, and are therefore programmed
+This is not always the case however. A language based on \WFN that is
+actually directly executable is \YAWL by \citet{DBLP:journals/is/AalstH05}.
+It facilitates modelling and execution of
+dynamic workflows, with support for \emph{and}, \emph{or} and \emph{xor} workflow patterns. As
+mentioned, \YAWL programs consist of \WFN, and are therefore programmed
 visually.
@@
 
-BPEL~\cite{bpel} is another popular business process language. The standardized
+\BPEL~\cite{bpel} is another popular business process language. The standardised
 language allows for the specification of actions within business processes,
-using an XML format. Processes specified in BEPL are executable, just like YAWL.
+using an \XML format. Processes specified in \BPEL are executable, just like \YAWL.
 
@@
 
-If This Then That (IFTTT)~\cite{IFTTT} is a web service that allows users to write and
+If This Then That (\IFTTT)~\cite{IFTTT} is a web service that allows users to write and
 configure simple conditional statements called applets, that can interact with
@@
 
-Google forms~\cite{googleforms} and Wufoo~\cite{wufoo} are web based survey administration apps that allows
-users to compose forms and gather data from them. They are comparable to TOP and
+Google Forms~\cite{googleforms} and Wufoo~\cite{wufoo} are web based survey administration apps that allows
+users to compose forms and gather data from them. They are comparable to \TOP and
 iTasks in that they provide an easy way for users to construct forms on the web.
-However, Google forms and Wufoo do not allow users to define what the system
+However, Google Forms and Wufoo do not allow users to define what the system
 then should do with this information, and can therefore not be regarded as a
-workflow or business process modeling systems.
+workflow or business process modelling systems.
 
@@
 application, though which a collective task can be coordinated. Users work on a
-preset data-type that drives the behavior of the user defined applications.
+preset data-type that drives the behaviour of the user defined applications.
 
@@
 
-Functional Reactive Programming (FRP) is a paradigm to describe dynamic changes of values in a declarative way.
+Functional Reactive Programming (\FRP) is a paradigm to describe dynamic changes of values in a declarative way.
 This is done by specifying networks of values, called behaviours, that can depend on each other and on external events.
@@
 
-The idea of FRP was pioneered by \citet{conf/icfp/ElliottH97}, in the meantime there are many variants and implementations, with reactive-banana one of the most well-known \cite{reactive-banana}.
+The idea of \FRP was pioneered by \citet{conf/icfp/ElliottH97}, in the meantime there are many variants and implementations, with reactive-banana one of the most well-known \cite{reactive-banana}.
 
-FRP and \TOP are different systems that have different goals in mind.
-Whereas FRP expresses automatically updating data dependencies, \TOP expresses collaboration patterns.
+\FRP and \TOP are different systems that have different goals in mind.
+Whereas \FRP expresses automatically updating data dependencies, \TOP expresses collaboration patterns.
 \TOP has no notion of time.
-Tasks can not change over time, while behaviours can.
-The biggest conceptual difference between a workflow in \TOP and a data network in FRP is that an event to a task only causes updates up until the next step, while an event in FRP propagates through the whole network.
+Tasks can not change spontaneous over time, while behaviours can.
+Only input events can change task values.
+The biggest conceptual difference between a workflow in \TOP and a data network in \FRP is that an event to a task only causes updates up until the next step, while an event in \FRP propagates through the whole network.
 
-That being said, there are some concepts that are similar in \TOP and FRP.
+That being said, there are some concepts that are similar in \TOP and \FRP.
 The \emph{stepper} behaviour, for example, is associated with an event and yields the value of the most recent event.
 This is similar to editors in \TOP.
-Furthermore, both systems can be used to declaratively program user interfaces, albeit in FRP the programmer has to construct the GUI elements manually, and connect inputs and outputs to the correct events and behaviours.
+Furthermore, both systems can be used to declaratively program user interfaces, albeit in \FRP the programmer has to construct the \GUI elements manually, and connect inputs and outputs to the correct events and behaviours.
 In \TOP graphical user interfaces are automatically derived.
@@
 Session types are expressions in some process calculus that describe the input/output behaviour of such programs.
-Session types are useful for programming languages where modules communicate with each other via messages, like CSP, $\pi$-calculus, or Go, to name a few.
+Session types are useful for programming languages where modules communicate with each other via messages, like \CSP, \PICALC, or Go, to name a few.
 The only form of messages in \TOP are input events which drive execution, but modules do not communicate using messages.

### 2019-02-13 sections/relatedwork.tex

--- 2019-02-13 sections/relatedwork.tex
@@
 
-Functional Reactive Programming (FRP) is a paradigm to describe dynamic changes of values in a declarative way.
-This is done by specifying networks of values, called behaviors, that can depend on each other and on external events.
-Behaviors can change over time, or triggered by events.
-When a behavior changes, all other behaviors that depend on it are updated automatically.
-The underlying implementation that takes care of the updating usually can tie input devices, like mouse and keyboard, to event streams and behaviors to output facilities, like text fields.
+Functional Reactive Programming (\FRP) is a paradigm to describe dynamic changes of values in a declarative way.
+This is done by specifying networks of values, called behaviours, that can depend on each other and on external events.
+Behaviours can change over time, or triggered by events.
+When a behaviour changes, all other behaviours that depend on it are updated automatically.
+The underlying implementation that takes care of the updating usually can tie input devices, like mouse and keyboard, to event streams and behaviours to output facilities, like text fields.
 This allows for declarative specifications of applications with user interfaces.
 
-The idea of FRP was pioneered by \citet{conf/icfp/ElliottH97}, in the meantime there are many variants and implementations, with reactive-banana one of the most well-known \cite{reactive-banana}.
+The idea of \FRP was pioneered by \citet{conf/icfp/ElliottH97}, in the meantime there are many variants and implementations, with reactive-banana one of the most well-known \cite{reactive-banana}.
 
-FRP and \TOP are different systems that have different goals in mind.
-Whereas FRP expresses automatically updating data dependencies, \TOP expresses collaboration patterns.
+\FRP and \TOP are different systems that have different goals in mind.
+Whereas \FRP expresses automatically updating data dependencies, \TOP expresses collaboration patterns.
 \TOP has no notion of time.
-Tasks can not change over time, while behaviors can.
-The biggest conceptual difference between a workflow in \TOP and a data network in FRP is that an event to a task only causes updates up until the next step, while an event in FRP propagates through the whole network.
+Tasks can not change spontaneous over time, while behaviours can.
+Only input events can change task values.
+The biggest conceptual difference between a workflow in \TOP and a data network in \FRP is that an event to a task only causes updates up until the next step, while an event in \FRP propagates through the whole network.
 
-That being said, there are some concepts that are similar in \TOP and FRP.
-The \emph{stepper} behavior for example is associated with an event and yields the value of the most recent event.
+That being said, there are some concepts that are similar in \TOP and \FRP.
+The \emph{stepper} behaviour, for example, is associated with an event and yields the value of the most recent event.
 This is similar to editors in \TOP.
-Furthermore, both systems can be used to declaratively program user interfaces, albeit in FRP the programmer has to construct the GUI elements manually, and connect inputs and outputs to the correct events and behaviors.
+Furthermore, both systems can be used to declaratively program user interfaces, albeit in \FRP the programmer has to construct the \GUI elements manually, and connect inputs and outputs to the correct events and behaviours.
 In \TOP graphical user interfaces are automatically derived.
@@
 Session types are a type discipline that can be used to check whether communicating programs conform to a certain protocol.
-Session types are expressions in some process calculus that describe the input/output behavior of such programs.
-Session types are useful for programming languages where modules communicate with each other via messages, like CSP, $\pi$-calculus, or Go, to name a few.
+Session types are expressions in some process calculus that describe the input/output behaviour of such programs.
+Session types are useful for programming languages where modules communicate with each other via messages, like \CSP, \PICALC, or Go, to name a few.
 The only form of messages in \TOP are input events which drive execution, but modules do not communicate using messages.

### 2019-02-13 sections/intuition.tex
add todo
some stuff in intuition

@@
 The central objective of tasks is to \emph{coordinate collaboration}.
-Therefore it has combinators to express ways in which people can work together to reach a common goal.
+Therefore \TOPHAT has combinators to express ways in which people can work together to reach a common goal.
 Tasks can be executed after each other, at the same time, or conditionally.
@@
 
+\fixme{Add example about higher order tasks}
+
 \subsection{Tasks are driven by user input}
@@
 A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive any change event with a value of type $\tau$ and additionally the special event $\Empty$ that empties the editor.
-There is no explicit communication of output.
+There are no output events, instead the environment can always inspect a task's value.
 % However, the environment can always inspect the current value of an editor.
@@
 This example demonstrates external communication and choice.
-The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins.
+The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins. \todo{remove CSP?}
 \begin{equation*}
@@
 The structure of the specification prescribes the structure of the \UI and possible input events.
-Editors form the leaves of such trees.
+Editors form the leaves of such trees \todo{which trees?}.
 
@@
 
-The decision to move on to the next task is not performed by the task itself.
+The decision to move on to the next task is not taken by the task itself.
 It is based on a calculation to determine the next tasks.
@@
 The semantics of \TOPHAT requires all updates to shared data and all enabled internal steps to be processed before any further communication with the environment can take place.
-
-\fixme{?}
-This introduces the possibility of diverging computations, where a cyclic dependency between shared data causes internal steps to be taken continuously without ever being able to engage in external communication.
 
@@
 
-While tasks are focussed on collaboration, they give rise to concepts like communication, concurrency and synchronisation.
+The focus on collaboration gives rise to concepts like communication, concurrency and synchronisation.
 Using tasks, however, they are meant to be used on a higher level of abstraction, which allows concise declarative description of patterns of collaboration, without worrying about the implementation details of communication.

### 2019-02-13 sections/intuition.tex
add higher order example: twice

@@
 
-\fixme{Add example about higher order tasks}
+\begin{example}[Twice]
+
+A simple higher-order task which performs a task twice can be specified as follows.
+\begin{TASK}
+  let twice : Task $\tau$ -> Task $\tau$ = \t : Task $\tau$.
+    t >>? \x : $\tau$. t
+\end{TASK}
+By using the step combinator $\When$ we specify that the given task \TS{t} should be performed a second time.
+The result \TS{x} from the first run is ignored.
+
+\end{example}
 

### 2019-02-13 sections/intuition.tex
remove mentioning processes in vending machine example

@@
 
+% The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins. \todo{remove CSP?}
+% \begin{equation*}
+%   \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
+% \end{equation*}
+% % An equivalent reactive system in \ESTEREL could be defined as\fixme{validate}
+% % \begin{equation*}
+%   % V_\text{\ESTEREL} = \text{await } \textit{coin}; ( \text{await } \textit{button}; \text{emit } \textit{small} \parallel \text{await } \textit{coin}; \text{await } \textit{button}; \text{emit } \textit{large})
+% % \end{equation*}
+% A vending machine in \TOPHAT that is close in spirit to the one above looks as follows.
+
 This example demonstrates external communication and choice.
-The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins. \todo{remove CSP?}
-\begin{equation*}
-  \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
-\end{equation*}
-% An equivalent reactive system in \ESTEREL could be defined as\fixme{validate}
-% \begin{equation*}
-  % V_\text{\ESTEREL} = \text{await } \textit{coin}; ( \text{await } \textit{button}; \text{emit } \textit{small} \parallel \text{await } \textit{coin}; \text{await } \textit{button}; \text{emit } \textit{large})
-% \end{equation*}
-
-A vending machine in \TOPHAT that is close in spirit to the one above looks as follows.
+A vending machine in \TOPHAT that dispenses a small biscuit for one coin and a large biscuit for two coins can be specified as follows
 \begin{TASK}
-  enter Int >>? \n. if n == 1 then edit "small" else if n == 2 then edit "large" else fail
+  let vend : Task String = enter Int >>? \n. if n == 1 then edit "small" else if n == 2 then edit "large" else fail
 \end{TASK}
-This machine does not exhibit exactly the same behaviour as the ones in \CSP, nor is it intended to.
-Instead, it just demonstrates how the same choice
-%%NOTE: Problem is already stated --TS
-% of buying a small biscuit for one coin or a large biscuit for two coins
-can be expressed in \TOPHAT.
 The task asks the user to enter an amount of money.
@@
 
+% This machine does not exhibit exactly the same behaviour as the ones in \CSP, nor is it intended to.
+% Instead, it just demonstrates how the same choice
+%%NOTE: Problem is already stated --TS
+% of buying a small biscuit for one coin or a large biscuit for two coins
+% can be expressed in \TOPHAT.
 \end{example}

### 2019-02-14 sections/intuition.tex
fix wrong macro name
Changes to section 3.1 and 3.2

@@
 
-This section gives an overview of the core concepts of tasks in task oriented programming (\TOP).
+This section gives an overview of the core concepts of tasks in task oriented
+programming (\TOP).
 
@@
 
-The central objective of tasks is to \emph{coordinate collaboration}.
-Therefore \TOPHAT has combinators to express ways in which people can work together to reach a common goal.
-Tasks can be executed after each other, at the same time, or conditionally.
-This motivates the task combinators for steps, parallel combination, and choice.
-One can say that tasks are a way to model \emph{workflows}.
+The central objective of \TOP is to \emph{coordinate collaboration}.
+To this end, \TOPHAT has basic tasks called editors, whith which users can
+interact, and combinators to express ways in which people can work together on
+these tasks. Tasks can be executed after each other, at the same time, or
+conditionally. This motivates the task combinators for steps, parallel
+combination, and choice.
+%One can say that tasks are a way to model \emph{workflows}.
 % \TOPHAT is a language to specify workflows.
@@
 
-\fixme{Extend}
-Tasks are modular and composable,
-can be abstracted over zero or more arguments,
-arguments could themselves be tasks,
-thus we have higher order workflows.
+Tasks are extremely flexible. They are modular; \fixme{why exactly?}. Tasks are
+composable, using the aforementioned combinators. And tasks are higher order;
+they can be abstracted over zero or more arguments, and those arguments could
+be tasks themselves.
 
@@
 
-A simple higher-order task which performs a task twice can be specified as follows.
+A simple higher-order task which performs a task twice can be specified as
+follows.
 \begin{TASK}
-  let twice : Task $\tau$ -> Task $\tau$ = \t : Task $\tau$.
-    t >>? \x : $\tau$. t
+  let twice : Task $\tau$ -> Task $\tau$ = \t : Task $\tau$. t >>? \x : $\tau$. t
 \end{TASK}
-By using the step combinator $\When$ we specify that the given task \TS{t} should be performed a second time.
-The result \TS{x} from the first run is ignored.
+By using the step combinator $\Next$ we specify that the given task \TS{t}
+should be performed a second time. The result \TS{x} from the first run is
+ignored.
 
 \end{example}
+
+\fixme{conclude something about why this is convenient}
 

### 2019-02-14 sections/intuition.tex
Changes to text section 3.3

@@
 
-\fixme{Extend}
-Tasks process input events from users.
-Sensors are also regarded as users.
-Gives rise to communication with the environment.
+The evaluation of tasks is driven by input events from users. When the system
+receives a valid event, it is applied to the current task, which results in a
+new task. This gives rise to communication with the users of the system.
 
-In \TOPHAT, communication with the environment is represented by editors.
-A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive any change event with a value of type $\tau$ and additionally the special event $\Empty$ that empties the editor.
-There are no output events, instead the environment can always inspect a task's value.
+In \TOPHAT, this communication with the environment is represented by editors,
+which are denoted by a box symbol.
+A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive a change
+event with a value of type $\tau$ and additionally the special event $\Empty$
+that empties the editor.
+There are no external output events, instead the new task can be observed, as
+described in the next section.
 % However, the environment can always inspect the current value of an editor.
 
-Editors have no notion of continuation.
-Their sole purpose is to interact with the environment, retaining the last value that has been sent to them.
+Editors have no notion of continuation. Their sole purpose is to facilitate
+interact with the user, retaining the last value that has been sent to them.
 %
-They can be used for both input and output.
-The interpretation of whether an editor represents input or output is left to the reader.
-An empty editor is most commonly interpreted as a prompt to input data, while a filled editor can be seen either as outputting its value, or as an input that comes with a default value.
+Since editors can be observed in for example a user interface, they can be used
+for both input and output. An empty editor is most commonly interpreted as a
+prompt to input data, while a filled editor can be seen either as outputting its
+value, or as an input that comes with a default or current value.
 
 \begin{example}[Vending machine]
-
+\label{Vending machine}
 % The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins. \todo{remove CSP?}
@@
 \begin{TASK}
-  let vend : Task String = enter Int >>? \n. if n == 1 then edit "small" else if n == 2 then edit "large" else fail
+  let vend : Task String = enter Int >>? \n. if n == 1 then edit "small"
+                                                       else if n == 2 then edit "large"
+                                                                      else fail
 \end{TASK}
@@
 This editor stands for a coin slot in a real machine that freely accepts and returns coins.
-There is a continue button that is initially disabled.
+There is a continue button that is initially disabled, due to the fact that the right hand side of the step combinator results in $\Fail$.
 When the user has inserted exactly 1 or 2 coins, the continue button becomes enabled.

### 2019-02-14 sections/intuition.tex
Changes to section 3.4

@@
 
-\subsection{Tasks can be observed for their current status at all time}
+\subsection{Tasks can be observed}
 
-\fixme{Extend}
-The value of a task can be present or not.
+Several oberservations can be made of a task. The most obvious and basic
+observation is determining the value of a task. Not all tasks have a value
+though, for example, the empty editor $\Enter \tau$ does not have a value.
 
-Because every task has an observable value of a specified type,
-it is always possible to generate a (graphical) user interface from a \TOPHAT specification.
-%NOTE: depends on: editors
-The structure of the specification prescribes the structure of the \UI and possible input events.
-Editors form the leaves of such trees \todo{which trees?}.
+We can also observe what potential events a task can take. This is essential in
+generating a (graphical) user interface for tasks. For example, the task
+$\Edit 5$ can take a new value as input, or it can be cleared, resulting in an
+empty editor of type Int.
+
+In order to complete the user interface, we also need to be able to render a
+task. This is done in a modular and recursive fashion. Every combinator renders
+their sub-tasks in a certain way. For example, when rendering a task that is
+composed of a step combinator, we simply render the left side of it.
+
+When we are able to render tasks, and we are able to list all possible inputs,
+we can obviously construct a user interface, that displays the current state of
+the system, together with buttons that show the potential actions a user can
+apply.
+
+Another useful observation is to determine whether a task results in $\Fail$.
+This is used by the step combinator, as demonstrated in
+Example~\ref{Vending machine}, and the choice combinator, to prevent users from
+picking a task that fails.
 

### 2019-02-14 sections/intuition.tex
changes to section 3.5

@@
 %NOTE: depends on: editors
-As users can continuously update editors, tasks never terminate.
+Users can continuously interact with tasks, they will never terminate. The event
+send by a user leads to a new task that can always receive a new event.
+
+The new task can either be the same task, but with a new value for the changed
+editor, or the system can decide to move on to the next task.
 

### 2019-02-14 sections/intuition.tex
Changes to intro of section 3.6

@@
 
-Second method of internal communication between tasks is shared data.
+We have already seen communication between tasks by means of the step
+combinator. The resulting value from the left task is passed on to the task on
+the right. The second method of internal communication between tasks is shared
+data.
 Shared data represents data flow \emph{across} control flow.

### 2019-02-15 sections/intuition.tex
spelling

@@
 The central objective of \TOP is to \emph{coordinate collaboration}.
-To this end, \TOPHAT has basic tasks called editors, whith which users can
+To this end, \TOPHAT has basic tasks called editors, with which users can
 interact, and combinators to express ways in which people can work together on
@@
 
-Several oberservations can be made of a task. The most obvious and basic
+Several observations can be made of a task. The most obvious and basic
 observation is determining the value of a task. Not all tasks have a value

### 2019-02-15 sections/conclusions.tex
Reworked conclusion and future work

@@
 
-In this paper we have developed a domain-specific calculus for declarative interactive workflows.
-We have identified, what we believe to be, the essence of Task-Oriented Programming, and provided corresponding operators in our language.
-We have compared our language with \CSP to point out differences and similarities.
-Finally, we have proven type safety for our language.
+In this paper we have developed a domain-specific calculus for declarative interactive workflows, named \TOPHAT.
+We have shown that our calculus is modular, and clearly seperated from the host language it is embedded in.
+By developing this calculus, we have identified the essence of Task-Oriented Programming.
+In addition to the formal semantics, we have also demonstrated how to build a user interface from a task description.
+We have compared \TOPHAT with workflow moddeling languages, process algebras, functional reactive programming and session types to point out differences and similarities.
+Finally, we have proven type safety and progress for our language.
 
@@
 
-The addition of read-only shared editors would allow watching references without being able to change them.
-Some programs could be expressed more faithfully, where we for now require that users do not modify certain shares.
+One of the main motivations to formalize Task Oriented Programmig is to be able to reason about programs.
+In this paper we reason about the language itself, but it would be nice to prove properties about individual programs.
+To this end, we are very interested to see if it is possible to develop an axiomatic semantics for \TOPHAT that allows us to do so.
+There are certain properties of our calculus that make this particularly complex; we have to deal with parallelism, user interacton and references.
 
-One goal of the formalisation of a language is the ability to reason about programs.
-In this paper we reason about the language itself, but it would be nice to prove properties about individual programs.
-We would like to prove whether certain programs are equal, for example to show that the monad laws hold for our step combinator.
+We would also like to prove whether certain programs are equivalent, for example to show that the monad laws hold for our step combinator.
 This requires a notion of equality, which in the presence of side effects most certainly needs some form of coalgebraic input-output conformance.
@@
 Naus et al.~\cite{UUCS2017013} have looked at building a generic feedback system for rule-based problems.
-A workflow system typically is rule based, as outlined in this work.
+A workflow system typically is rule based, as outlined in their work.
 It would be interesting to fit the generic feedback system to \TOPHAT in order to support end-users working in applications developed in this calculus.
@@
 would aid domain experts to model workflows in a more accessible manner.
+A system that visualizes iTask programs has been developed in the past~\cite{DBLP:conf/sfp/StutterheimPA14}.

### 2019-02-15 sections/intuition.tex
Fixed minor nit 10 from reviewer #3
add something about higher-order tasks

@@
 be tasks themselves.
+
+From a programmer's point of view,
+tasks should not just model their real world equivalents.
+They should be reusable units in a programming language.
+Similar tasks should be specified once,
+and reused in new context.
+Taking this further,
+by specifying tasks that take other tasks as arguments,
+programmers can have a lot of freedom to create their own reusable collaboration patterns.
+In other words,
+programmers would like abstraction and higher-order tasks.
 
@@
 % The continuation $e$ is a function which calculates the next task or $\Fail$, signalling no continuation is present.
-The task $t \Next e$ requires an input event $\Continue$ in addition to the step being unguarded. % in order to step.
+The task $t \Next e$ requires an input event $\Continue$ (Continue) in addition to the step being unguarded. % in order to step.
 

### 2019-02-15 sections/introduction.tex
add new introduction

@@
 
-Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
-\TOP has two aspects.
-Firstly, \TOP defines the primitive building blocks that are useful for high-level descriptions of how users collaborate with each other and with applications.
-These building blocks are \emph{editors}, \emph{composition}, and \emph{shared data}.
-Secondly, \TOP defines a type-driven way to generate applications, including graphical user interfaces, from workflows modelled with said building blocks.
+\subsection{Task-oriented programming}
 
-In this paper we study the essence of \TOP.
-We claim that at the heart of \TOP lie modular interactive workflows.
-Workflows because \TOP programs coordinate collaboration between people.
-Interactive because progress of \TOP programs is driven by user input.
-Modular because the embedding in a strongly-typed functional language allows powerful abstractions using type-driven techniques and higher-order functions.
+Many applications these days are developed to support workflows in institutions and businesses.
+Take for example expenses declarations, order processing, and emergency management.
+Not all of these workflows happen inside organisations.
+Some occur on the boundary with customers, like flight bookings or tax returns.
+What they all have in common is,
+that they need to interact with different people (end users, tax officers, customers, etc.)
+and they use information from multiple sources (input forms, databases, sensors, etc.).
 
-Interactive workflows are called \emph{tasks}.
-Tasks stand for units of work in the real world, assigned to people.
-People can work together in a number of ways, and this is reflected in \TOP by task combinators.
-There is sequential composition, parallel composition, and choice.
-People need to communicate in order to engage in these forms of collaboration.
-This is reflected in \TOP by three kinds of communication mechanisms.
-There is data flow \emph{alongside} control flow, where the result of a task is passed onto the next.
-There is data flow \emph{across} control flow, where information is shared between multiple tasks.
-Finally, there is communication with the \emph{outside} world, where information is entered into the system via input events.
-The end points where the outside world interacts with \TOP applications are called editors.
-In generated applications, editors can take many forms, like input fields, selection boxes, or map widgets.
+We call such interactive units of work based on information sources \emph{tasks}.
+Tasks model collaboration between users and are driven by work users do.
+Users could be people out in the field or sitting behind their desks,
+as well as machines doing calculations or fetching data.
 
-iTasks is an implementation of \TOP, in the form of a shallowly embedded domain-specific language in the lazy functional programming language Clean.
-It is a library that provides editors and monadic combinators.
-iTasks uses the generic programming facilities of Clean to derive web applications from workflow models.
-It has been used to model an incident management tool for the Dutch coast guard~\cite{conf/iscram/LijnseJP12}.
-Also it has been used numerous times to prototype ideas for Command and Control~\cite{theses/nlda/Kool17, theses/radboud/Stutterheim17}, and in a case study for the Dutch tax authority \cite{conf/sfp/StutterheimAP17}.
+Task-oriented programming (\TOP) is a programming paradigm which searches for the sweet spot between faithful modelling workflows
+and rapid prototyping of multi-user web applications supporting this workflow.
+% Task-oriented programming (\TOP) is a programming paradigm to support these ways of working.
+The paradigm takes collaboration patterns as the base of modelling
+and anticipates on user's need to interact and share information.
+A language that supports \TOP should allow to specify tasks from real world scenarios,
+and be able to generate multi-user web applications to support these scenarios.
+By choosing the right level of abstraction,
+\TOP focusses on creating workflow specifications,
+while automatically providing common development duties like designing \GUI\ s, connecting to databases, and communicating between servers and client.
+
+Currently, we know of two frameworks using \TOP: \ITASKS and \MTASKS.
+\ITASKS is a full fledged \TOP framework able to generate rich client and server applications from a single source in the functional programming language Clean.
+\MTASKS is a subset of \ITASKS,
+focussing on \IOT devices and deployment on micro controllers.
+Both have been used to model real world scenarios like \fixme{add}.
+
+In this paper we tend to identify the essence of \TOP in formal as well as an informal way.
+These core principles lead to a formal system for \TOP called \TOPHAT.
+\TOPHAT paves the way to formal reasoning about applications written using the \TOP paradigm,
+like those in \ITASKS and \MTASKS.
+
+\TOP applications give rise to concepts like concurrency, synchronisation, \GUI-programming and program generation.
+In this paper we identify common concepts with already existing languages and frameworks,
+and describe their similarities and differences.
+
+\subsection{Contributions}
+
+Our contributions to workflow modelling, functional programming language design, and rapid application development are as follows.
+
+\begin{enumerate}
+
+  \item
+    We informally describe the essential concepts of task-oriented programming (\TOP).
+    We have a strong focus on modelling using collaboration
+    while keeping a constant desire to rapidly generate executable applications into account.
+
+  \item
+    We present a formal calculus for \TOP, arising naturally from above essential concepts.
+    Hereby we provide ground work to apply formal reasoning about \TOP specifications in future work.
+
+  \item
+    Using both informal and formal descriptions, we compare \TOP and \TOPHAT with multiple related work in the area,
+    ranging from business process modelling languages, to process algebras and reactive programming frameworks.
+
+\end{enumerate}
+
+\subsection{Structure}
+
+\input{sections/structure}
+
+\subsection{Old}
+
+\input{sections/introduction-old}

### 2019-02-15 sections/introduction.tex
improve intro

@@
 
-\subsection{Task-oriented programming}
+\subsection{Tasks}
 
 Many applications these days are developed to support workflows in institutions and businesses.
-Take for example expenses declarations, order processing, and emergency management.
-Not all of these workflows happen inside organisations.
-Some occur on the boundary with customers, like flight bookings or tax returns.
+Take for example expense declarations, order processing, and emergency management.
+Some of these workflows occur on the boundary between organisations and customers,
+like flight bookings or tax returns.
 What they all have in common is,
@@
 
+\subsection{Task-oriented programming}
+
 Task-oriented programming (\TOP) is a programming paradigm which searches for the sweet spot between faithful modelling workflows
-and rapid prototyping of multi-user web applications supporting this workflow.
+and rapid prototyping of multi-user web applications supporting these workflows.
 % Task-oriented programming (\TOP) is a programming paradigm to support these ways of working.
-The paradigm takes collaboration patterns as the base of modelling
-and anticipates on user's need to interact and share information.
-A language that supports \TOP should allow to specify tasks from real world scenarios,
-and be able to generate multi-user web applications to support these scenarios.
-By choosing the right level of abstraction,
-\TOP focusses on creating workflow specifications,
-while automatically providing common development duties like designing \GUI\ s, connecting to databases, and communicating between servers and client.
+\TOP focusses on modelling collaboration patterns.
+This gives rise to user's need to interact and share information.
+Next to that, \TOP automatically provides solutions to common development jobs like designing \GUI\ s, connecting to databases, and communicating between servers and clients.
+Therefore,
+a language that supports \TOP should choose the right level of abstraction to support two things.
+First, it should allow to specify tasks from real world scenarios.
+Second, it should be able to generate multi-user web applications to support these scenarios.
+
+\subsection{Utilisation}
 
@@
 Both have been used to model real world scenarios like \fixme{add}.
+
+\subsection{Contributions}
 
@@
 and describe their similarities and differences.
-
-\subsection{Contributions}
 

### 2019-02-15 sections/introduction.tex
merge old and new intros

@@
 We call such interactive units of work based on information sources \emph{tasks}.
-Tasks model collaboration between users and are driven by work users do.
+% Tasks stand for units of work in the real world, assigned to people.
+Tasks model real world collaboration between users,
+are driven by work they do,
+and are assigned to some user.
 Users could be people out in the field or sitting behind their desks,
@@
 
+% Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
 Task-oriented programming (\TOP) is a programming paradigm which searches for the sweet spot between faithful modelling workflows
@@
 Next to that, \TOP automatically provides solutions to common development jobs like designing \GUI\ s, connecting to databases, and communicating between servers and clients.
+
 Therefore,
 a language that supports \TOP should choose the right level of abstraction to support two things.
-First, it should allow to specify tasks from real world scenarios.
-Second, it should be able to generate multi-user web applications to support these scenarios.
+% \TOP has two aspects.
+% First, it should allow to specify tasks from real world scenarios.
+Firstly, it should provide primitive building blocks that are useful for high-level descriptions of how users collaborate with each other and with machines.
+These building blocks are \emph{editors}, \emph{composition}, and \emph{shared data}.
+% Second, it should be able to generate multi-user web applications to support these scenarios.
+Secondly, it should be able to generate applications, including graphical user interfaces, from workflows modelled with said building blocks.
+
+Users can work together in a number of ways, and this is reflected in \TOP by task compositions.
+There is sequential composition, parallel composition, and choice.
+Users need to communicate in order to engage in these forms of collaboration.
+This is reflected in \TOP by three kinds of communication mechanisms.
+There is data flow \emph{alongside} control flow, where the result of a task is passed onto the next.
+There is data flow \emph{across} control flow, where information is shared between multiple tasks.
+Finally, there is communication with the \emph{outside} world, where information is entered into the system via input events.
+The end points where the outside world interacts with \TOP applications are called editors.
+In generated applications, editors can take many forms, like input fields, selection boxes, or map widgets.
+
+\fixme{Change if title changes.}
+In this paper we study the essence of \TOP.
+We claim that at the heart of \TOP lie modular interactive workflows.
+Workflows because \TOP programs coordinate collaboration between people.
+Interactive because progress of \TOP programs is driven by user input.
+Modular because the embedding in a strongly-typed functional language allows powerful abstractions using type-driven techniques and higher-order functions.
 
@@
 Currently, we know of two frameworks using \TOP: \ITASKS and \MTASKS.
-\ITASKS is a full fledged \TOP framework able to generate rich client and server applications from a single source in the functional programming language Clean.
+
+\ITASKS is an implementation of \TOP, in the form of a shallowly embedded domain-specific language in the lazy functional programming language Clean.
+It is a library that provides editors, monadic combinators, and shared data sources.
+\ITASKS uses the generic programming facilities of Clean to derive rich client and server applications from a single source.
+It has been used to model an incident management tool for the Dutch coast guard~\cite{conf/iscram/LijnseJP12}.
+Also it has been used numerous times to prototype ideas for Command and Control~\cite{theses/nlda/Kool17, theses/radboud/Stutterheim17}, and in a case study for the Dutch tax authority \cite{conf/sfp/StutterheimAP17}.
+
 \MTASKS is a subset of \ITASKS,
 focussing on \IOT devices and deployment on micro controllers.
-Both have been used to model real world scenarios like \fixme{add}.
+\fixme{Add examples.}
+
+\fixme{Say what lacks in these approaches: no formal reasoning about properties.}
 
 \subsection{Contributions}
+
+\input{sections/contributions}
 
@@
 
-\input{sections/structure}
-
-\subsection{Old}
-
-\input{sections/introduction-old}
+In Section~\ref{sec:example} we demonstrate the functionality of \TOPHAT by means of an example,
+Section~\ref{sec:intuition} gives an overview of the essential concepts of \TOP and \TOPHAT.
+Section~\ref{sec:language} introduces the \TOPHAT calculus syntax
+and Section~\ref{sec:semantics} the semantics.
+Then in Section~\ref{sec:properties} we show that certain properties hold for the calculus.
+We take a look at related work in Section~\ref{sec:relatedwork}
+and conclude in Section~\ref{sec:conclusions}.

### 2019-02-16 sections/intuition.tex
intuition: tasks are reusable

--- 2019-02-15 sections/intuition.tex
@@
 To this end, \TOPHAT has basic tasks called editors, with which users can
-interact, and combinators to express ways in which people can work together on
-these tasks. Tasks can be executed after each other, at the same time, or
+interact, and combinators to express ways in which people can work together.
+Tasks can be executed after each other, at the same time, or
 conditionally. This motivates the task combinators for steps, parallel
@@
 
-Tasks are extremely flexible. They are modular; \fixme{why exactly?}. Tasks are
-composable, using the aforementioned combinators. And tasks are higher order;
-they can be abstracted over zero or more arguments, and those arguments could
-be tasks themselves.
+Tasks are modular in the following ways.
+First, larger tasks are composed from smaller ones.
+Second, tasks are first-class, they can be arguments and results of functions.
+Third, tasks can be values of other tasks.
+These aspects make it possible for programmers to model custom collaboration patterns.
 
@@
 
-A simple higher-order task which performs a task twice can be specified as
-follows.
+A higher-order function which executes a given task twice.
+The task value of the first execution is ignored.
+
+\fixme{Is that our official syntax?}
 \begin{TASK}
-  let twice : Task $\tau$ -> Task $\tau$ = \t : Task $\tau$. t >>? \x : $\tau$. t
+  let twice : Task $\tau$ -> Task $\tau$ = \t : Task $\tau$. t >>? \x : $\tau$
 \end{TASK}
-By using the step combinator $\Next$ we specify that the given task \TS{t}
-should be performed a second time. The result \TS{x} from the first run is
-ignored.
+\qed
 
 \end{example}
-
-\fixme{conclude something about why this is convenient}
 

### 2019-02-18 sections/intuition.tex
simplify sentence
tasks are driven by user input

@@
 
-This section gives an overview of the core concepts of tasks in task oriented
-programming (\TOP).
+This section gives an overview of the core concepts of task oriented programming (\TOP).
 
@@
 
-The evaluation of tasks is driven by input events from users. When the system
-receives a valid event, it is applied to the current task, which results in a
-new task. This gives rise to communication with the users of the system.
+Input events drive evaluation of tasks.
+When the system receives a valid event, it applies the event to the current task, which results in a new task.
+In this way the system communicates with the environment.
 
-In \TOPHAT, this communication with the environment is represented by editors,
-which are denoted by a box symbol.
-A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive a change
-event with a value of type $\tau$ and additionally the special event $\Empty$
-that empties the editor.
-There are no external output events, instead the new task can be observed, as
-described in the next section.
-% However, the environment can always inspect the current value of an editor.
+In \TOPHAT, editors are the basic method of communication with the environment.
+There are different editors, denoted by different box symbols.
+An editor of the form $\Edit v$, where $v$ is of type $\tau$, reacts to two kinds of events:
+First, change events $v$ with a value of type $\tau$ and second, clear events $\Empty$ which empty the editor.
 
-Editors have no notion of continuation. Their sole purpose is to facilitate
-interact with the user, retaining the last value that has been sent to them.
-%
-Since editors can be observed in for example a user interface, they can be used
-for both input and output. An empty editor is most commonly interpreted as a
-prompt to input data, while a filled editor can be seen either as outputting its
-value, or as an input that comes with a default or current value.
+The sole purpose of editors is to interact with the user by retaining the last value that has been sent to them.
+There are no output events.
+As values of editors can be observed, for example by a user interface, editors serve as facilities for both input and output.
+An empty editor $\Enter \tau$ stands for a prompt to input data, while a filled editor can be seen either as outputting its value, or as an input that comes with a default value.
 

### 2019-02-18 sections/intuition.tex
remove dead code
vending machine

@@
 \label{Vending machine}
-% The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins. \todo{remove CSP?}
-% \begin{equation*}
-%   \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
-% \end{equation*}
-% % An equivalent reactive system in \ESTEREL could be defined as\fixme{validate}
-% % \begin{equation*}
-%   % V_\text{\ESTEREL} = \text{await } \textit{coin}; ( \text{await } \textit{button}; \text{emit } \textit{small} \parallel \text{await } \textit{coin}; \text{await } \textit{button}; \text{emit } \textit{large})
-% % \end{equation*}
-% A vending machine in \TOPHAT that is close in spirit to the one above looks as follows.
 
 This example demonstrates external communication and choice.
-A vending machine in \TOPHAT that dispenses a small biscuit for one coin and a large biscuit for two coins can be specified as follows
+It is a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins.
 \begin{TASK}
@@
 \end{TASK}
-The task asks the user to enter an amount of money.
+The editor $\Enter \Int$ asks the user to enter an amount of money.
 This editor stands for a coin slot in a real machine that freely accepts and returns coins.
@@
 When the user has inserted exactly 1 or 2 coins, the continue button becomes enabled.
-When the user presses the continue button, the machine dispenses either a large or a small biscuit, depending on the amount of money.
+When the user presses the continue button, the machine dispenses either a large or a small biscuit, depending on the amount of money. \qed
 
-% This machine does not exhibit exactly the same behaviour as the ones in \CSP, nor is it intended to.
-% Instead, it just demonstrates how the same choice
-%%NOTE: Problem is already stated --TS
-% of buying a small biscuit for one coin or a large biscuit for two coins
-% can be expressed in \TOPHAT.
 \end{example}

### 2019-02-18 sections/intuition.tex
tasks can be observed

@@
 
-Several observations can be made of a task. The most obvious and basic
-observation is determining the value of a task. Not all tasks have a value
-though, for example, the empty editor $\Enter \tau$ does not have a value.
+Several observations can be made of tasks.
+One of those is determining the value of a task $\Value(t)$.
+Not all tasks have a value, for example the empty editor $\Enter \tau$, which makes $\Value$ a partial function.
 
-We can also observe what potential events a task can take. This is essential in
-generating a (graphical) user interface for tasks. For example, the task
-$\Edit 5$ can take a new value as input, or it can be cleared, resulting in an
-empty editor of type Int.
+Another observation is the set of events a task can react to.
+This is important for generating user interfaces.
+For example, the task $\Edit 5$ can react to value events $v$ and empty events $\Empty$.
 
-In order to complete the user interface, we also need to be able to render a
-task. This is done in a modular and recursive fashion. Every combinator renders
-their sub-tasks in a certain way. For example, when rendering a task that is
-composed of a step combinator, we simply render the left side of it.
+In order to render a user interface, the system needs to observe a task's user interface.
+This is done compositionally.
+User interfaces of combined tasks are composed of the user interfaces of the components.
+For example of two tasks combined with a step combinator, only the left hand side is rendered.
+Two parallel tasks are rendered next to each other.
 
-When we are able to render tasks, and we are able to list all possible inputs,
-we can obviously construct a user interface, that displays the current state of
-the system, together with buttons that show the potential actions a user can
-apply.
+In order for the system to actually render a task, it needs to observe the task's value, possible inputs, and user interface.
+With this information, the system can then display the current state of the task, together with buttons that show the actions a user can engage in.
 
-Another useful observation is to determine whether a task results in $\Fail$.
-This is used by the step combinator, as demonstrated in
-Example~\ref{Vending machine}, and the choice combinator, to prevent users from
-picking a task that fails.
+The final observation is to determine whether a task results in $\Fail$.
+The step combinator and the choice combinator use this to prevent users from picking a failing task.
 

### 2019-02-18 sections/intuition.tex
Tasks are never done

@@
 
-%NOTE: depends on: editors
-Users can continuously interact with tasks, they will never terminate. The event
-send by a user leads to a new task that can always receive a new event.
+Tasks never terminate, they always keep reacting to events.
+Editors can always be changed or cleared, and step combinators move on to new tasks.
 
-The new task can either be the same task, but with a new value for the changed
-editor, or the system can decide to move on to the next task.
+In a step $t \Next e$, the decision to move on to e is taken by $\Next$, not by t.
+The decision is based on a speculative evaluation of $e$.
+The step combinators in $t \Then e$ and $t \Next e$ pass the value $v$ of $t$ to the continuation $e$.
+Both steps act like $t$ as long as the step is guarded.
+A step is guarded as long as the speculative evaluation $e\ v$ yields $\Fail$.
+Once it becomes unguarded, the task continues as $e\ v$.
+The task $t \Next e$ requires a continue event $\Continue$ in addition to the step being unguarded.
+We designed speculative evaluation so that it does not have side effects.
 
-The decision to move on to the next task is not taken by the task itself.
-It is based on a calculation to determine the next tasks.
-The step combinators in $t \Then e$ and $t \Next e$ pass the value of $t$ to the continuation $e$.
-They both act like $t$ as long as the step is guarded.
-Once it becomes unguarded, the task continues as $e\ v$, where $v$ is the observed value of $t$.
-% The continuation $e$ is a function which calculates the next task or $\Fail$, signalling no continuation is present.
-The task $t \Next e$ requires an input event $\Continue$ (Continue) in addition to the step being unguarded. % in order to step.
-
-The step combinators give rise to a method of internal communication.
+The step combinators give rise to a method of internal communication:
 They represent data flow that \emph{follows} control flow.
-
-% The intuition about sequential composition is to first do one thing, and once it is done, do some other thing.
-% The notion of sequential composition exists in two forms: the external and the internal step.
 

### 2019-02-18 sections/intuition.tex
wording
Tasks can share information

@@
 The task $t \Next e$ requires a continue event $\Continue$ in addition to the step being unguarded.
-We designed speculative evaluation so that it does not have side effects.
+Speculative evaluation is designed so that it does not have side effects.
 
@@
 
-We have already seen communication between tasks by means of the step
-combinator. The resulting value from the left task is passed on to the task on
-the right. The second method of internal communication between tasks is shared
-data.
-Shared data represents data flow \emph{across} control flow.
-Shared data sources are assignable references, whose changes in value are immediately visible to all tasks interested in them.
+The step combinator is one form of internal communication, where task values are passed to continuations.
+Another form of internal communication is shared data.
+Shared data enables data flow \emph{across} control flow, in particular between parallel tasks.
+Shared data sources are assignable references whose changes are immediately visible to all tasks interested in them.
 The semantics of \TOPHAT requires all updates to shared data and all enabled internal steps to be processed before any further communication with the environment can take place.

### 2019-02-18 sections/intuition.tex
cigarette smokers

@@
 Shared data sources are assignable references whose changes are immediately visible to all tasks interested in them.
+If $r$ is a reference of type $\tau$, then $\Update r$ is an editor whose value is that of $r$.
+
 The semantics of \TOPHAT requires all updates to shared data and all enabled internal steps to be processed before any further communication with the environment can take place.
@@
 There is an agent that randomly provides two of the ingredients.
+The difficulty lies in the requirement that only the smoker may proceed whose missing ingredients are present.
 
-Downey models availability of the ingredients using a semaphore for each ingredient.
+Downey models availability of the ingredients with a semaphore for each ingredient.
 The agent randomly signals two of the three semaphores.
-% The problem asks writing code for the smokers so that the smoker who requires the two signalled semaphores wakes up and proceeds with smoking.
-%
-% A naïve attempt where each smoker first waits for one and then the other semaphore can lead to deadlock.
-% The following pseudo-code is a non-solution.
-% Let \emph{paper}, \emph{tobacco} and \emph{match} be semaphores, initialised to 0.
-% \begin{TASK}
-%   tobaccoSmoker = wait(match); wait(paper); smoke();
-%   paperSmoker = wait(tobacco); wait(match); smoke();
-%   matchSmoker = wait(tobacco); wait(paper); smoke();
-% \end{TASK}
-% To see what can go wrong, imagine the three smokers running in parallel.
-% Let the agent signal the two semaphores \emph{tobacco} and \emph{paper}.
-% One of \emph{paperSmoker} and \emph{matchSmoker} can take their first step, but if \emph{paperSmoker} takes the step, the system deadlocks.
-%
 The solution proposed by Downey involves an additional mutex, three additional semaphores, three additional threads called \emph{pushers}, and three regular Boolean variables.
@@
 The details of the solution are not important here.
-What is important is that the implementation of what is essentially waiting for two semaphores requires a substantial amount of additional synchronisation, together with non-trivial conditional statements.
-
-% A similar effort as in Downey's solution must be made in order to solve the problem using synchronised actions in \CSP.
-% This is because in \CSP, processes can only synchronise on one action at a time.
+What is important is that the implementation of what is essentially deadlock-free waiting for two semaphores requires a substantial amount of additional synchronisation, together with non-trivial conditional statements.
 
 \TOPHAT allows a simple solution to this problem, using guarded steps.
-Steps can be guarded with arbitrary expressions, and the parallel combinator can be used to watch two shared editors ($\Update$) at the same time.
+Steps can be guarded with arbitrary expressions, and the parallel combinator can be used to watch two shared editors at the same time.
 Let \TS{match}, \TS{paper}, and \TS{tobacco} be references to Booleans.
@@
 \end{TASK}
-When the agent supplies two of the ingredients by setting the respective shares to \TS{True}, only the step of the smoker that waits for those becomes enabled.
+When the agent supplies two of the ingredients by setting the respective shares to \TS{True}, only the step of the smoker that waits for those becomes enabled. \qed
 

### 2019-02-18 sections/intuition.tex
tasks are predictable

@@
 
-Concurrency is the source of parallelism: two things are done at the same time.
-%
-Using tasks $t_1$ and $t_2$, the parallel combination $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
+The parallel combinator $\And$ introduces concurrency: Two things are done at the same time.
+Let $t_1$ and $t_2$ be tasks.
+The parallel combination $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
 For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
-All inputs sent to the combined task are interleaved into a serial stream of inputs.
+The inputs sent to the component tasks are interleaved into a serial stream, which is sent to the parallel combinator.
+We assume that such a serialization is always possible.
 The tasks are truly independent of each other, all interleavings are permitted.
-Inputs to $t_1$ and $t_2$ must be prefixed by $\First$(irst) and $\Second$(econd) respectively.
+The environment must prefix events to $t_1$ and $t_2$ by $\First$ (first) and $\Second$ (second) respectively.
 This unambiguously renames the inputs, removing any source of nondeterminism.
 
-The basic method for synchronisation in \TOP is built into the step combinator.
+With concurrency comes the need for synchronisation, in situations where only some but not all interleavings are desired.
+The basic method for synchronisation in \TOPHAT is built into the step combinator.
 The task $t \Next e$ can only continue execution when two conditions are met:
@@
 
+\Citet{books/Hoare85CSP} states that nondeterminism is only ever useful for \emph{specifying} systems, never for implementing them.
 \TOPHAT is meant solely for implementation and does not have any form of nondeterminism.

### 2019-02-18 sections/intuition.tex
recap

@@
 
-The focus on collaboration gives rise to concepts like communication, concurrency and synchronisation.
-Using tasks, however, they are meant to be used on a higher level of abstraction, which allows concise declarative description of patterns of collaboration, without worrying about the implementation details of communication.
-By focussing on collaboration, tasks and \TOPHAT lead to high-level specifications closer to real-world workflows which, at the same time, can be used to generate distributive, multi-user applications to support such workflows.
+To model collaboration, one needs communication, concurrency, and synchronisation.
+In \TOP, collaboration is specified on a high level of abstraction, hiding the details of communication.
+For example, the cigarette smokers communicate with each other, but the programs do not explicitly mention sending or receiving events.
 
-To employ these patterns of collaboration, users need to communicate.
-This motivates different forms of communication: with the environment, along control flow, and across control flow.
+By focussing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate distributed, multi-user applications to support such workflows.

### 2019-02-18 sections/intuition.tex
basic buliding blocks

@@
 
-This section gives an overview of the core concepts of task oriented programming (\TOP).
+This section gives an overview over the core concepts of task-oriented programming.
 
@@
 The central objective of \TOP is to \emph{coordinate collaboration}.
-To this end, \TOPHAT has basic tasks called editors, with which users can
-interact, and combinators to express ways in which people can work together.
-Tasks can be executed after each other, at the same time, or
-conditionally. This motivates the task combinators for steps, parallel
-combination, and choice.
-%One can say that tasks are a way to model \emph{workflows}.
-% \TOPHAT is a language to specify workflows.
+The basic building blocks of \TOPHAT are editors and combinators.
+Editors are tasks with which users interact directly.
+Combinators express ways in which people can work together.
+Tasks can be executed after each other, at the same time, or conditionally.
+This motivates the combinators step, parallel, and choice.
 

### 2019-02-18 sections/intuition.tex
fix twice example
wording

@@
 A higher-order function which executes a given task twice.
+\begin{TASK}
+  let twice : Task $\tau$ -> Task $\tau$ = \t : Task $\tau$. t >>? \x . t
+\end{TASK}
 The task value of the first execution is ignored.
-
-\fixme{Is that our official syntax?}
-\begin{TASK}
-  let twice : Task $\tau$ -> Task $\tau$ = \t : Task $\tau$. t >>? \x : $\tau$
-\end{TASK}
 \qed
@@
 An editor of the form $\Edit v$, where $v$ is of type $\tau$, reacts to two kinds of events:
-First, change events $v$ with a value of type $\tau$ and second, clear events $\Empty$ which empty the editor.
+Change events $v$ with a value of type $\tau$ and clear events $\Empty$ which empty the editor.
 
-The sole purpose of editors is to interact with the user by retaining the last value that has been sent to them.
+The sole purpose of editors is to interact with users by remembering the last value that has been sent to them.
 There are no output events.
 As values of editors can be observed, for example by a user interface, editors serve as facilities for both input and output.
-An empty editor $\Enter \tau$ stands for a prompt to input data, while a filled editor can be seen either as outputting its value, or as an input that comes with a default value.
+An empty editor $\Enter \tau$ stands for a prompt to input data, while a filled editor can be seen either as outputting a value, or as an input that comes with a default value.
 
@@
 This editor stands for a coin slot in a real machine that freely accepts and returns coins.
-There is a continue button that is initially disabled, due to the fact that the right hand side of the step combinator results in $\Fail$.
+There is a continue button that is initially disabled, due to the fact that the left hand side of the step combinator has no value.
 When the user has inserted exactly 1 or 2 coins, the continue button becomes enabled.

### 2019-02-18 sections/intuition.tex
wording

@@
 
-Several observations can be made of tasks.
+Several observations can be made about tasks.
 One of those is determining the value of a task $\Value(t)$.
@@
 
-In a step $t \Next e$, the decision to move on to e is taken by $\Next$, not by t.
+In a step $t \Next e$, the decision to move on is taken by $\Next$, not by $t$.
 The decision is based on a speculative evaluation of $e$.
@@
 Both steps act like $t$ as long as the step is guarded.
-A step is guarded as long as the speculative evaluation $e\ v$ yields $\Fail$.
-Once it becomes unguarded, the task continues as $e\ v$.
-The task $t \Next e$ requires a continue event $\Continue$ in addition to the step being unguarded.
+A step is guarded if either the left task has no value, or the speculative evaluation $e\ v$ yields $\Fail$.
+Once it becomes unguarded, the step continues as $e\ v$.
+The task $t \Next e$ additionally requires a continue event $\Continue$ to proceed.
 Speculative evaluation is designed so that it does not have side effects.
 
-The step combinators give rise to a method of internal communication:
+The step combinators give rise to a form of internal communication:
 They represent data flow that \emph{follows} control flow.

### 2019-02-18 sections/introduction.tex
Fixes in 1.1

@@
 \section{Introduction}
+
+\fixme{Common intro??}
 
@@
 like flight bookings or tax returns.
-What they all have in common is,
+What they all have in common is
 that they need to interact with different people (end users, tax officers, customers, etc.)
@@
 
-We call such interactive units of work based on information sources \emph{tasks}.
+We call interactive units of work based on information sources \emph{tasks}.
 % Tasks stand for units of work in the real world, assigned to people.
 Tasks model real world collaboration between users,
-are driven by work they do,
+are driven by work users do,
 and are assigned to some user.

### 2019-02-18 sections/introduction.tex
Fix in 1.2

@@
 % Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
-Task-oriented programming (\TOP) is a programming paradigm which searches for the sweet spot between faithful modelling workflows
+Task-oriented programming (\TOP) is a programming paradigm which aims to be the sweet spot between faithful modelling workflows
 and rapid prototyping of multi-user web applications supporting these workflows.

### 2019-02-18 sections/contributions.tex
merge old and new intros
fixes in 1.4

@@
 % !TEX root=../icfp2019.tex
-
-\subsection{Contributions}
 
@@
 The semantics of the task language is clearly separated from the semantics of the underlying host language.
-Along the semantics, we present the following semantic observations on tasks: the current value, whether a term is stuck, the current user interface, and the accepted inputs.
+Along with the semantics, we present the following semantic observations on tasks: the current value, whether a term is stuck, the current user interface, and the accepted inputs.
+\fixme{we do not provide observation for user interface!}
 %%NOTE: iTasks things are moved to related work

### 2019-02-18 sections/introduction.tex
fixes in 1.4

@@
 
-In this paper we tend to identify the essence of \TOP in formal as well as an informal way.
+In this paper we aim to identify the essence of \TOP in formal as well as an informal way.
 These core principles lead to a formal system for \TOP called \TOPHAT.
@@
 \TOP applications give rise to concepts like concurrency, synchronisation, \GUI-programming and program generation.
-In this paper we identify common concepts with already existing languages and frameworks,
-and describe their similarities and differences.
+In this paper we identify common concepts from already existing languages and frameworks,
+and describe their similarities and differences.\fixme{What do you mean with this sentence?}
 
@@
     We informally describe the essential concepts of task-oriented programming (\TOP).
-    We have a strong focus on modelling using collaboration
+    \TOP has a strong focus on modelling using collaboration
     while keeping a constant desire to rapidly generate executable applications into account.
@@
   \item
-    We present a formal calculus for \TOP, arising naturally from above essential concepts.
-    Hereby we provide ground work to apply formal reasoning about \TOP specifications in future work.
+    We present a formal calculus for \TOP, arising naturally from the aforementioned essential concepts.
+    In doing so, we lay the ground work to allow formal reasoning about \TOP specifications in future work.
 
   \item
-    Using both informal and formal descriptions, we compare \TOP and \TOPHAT with multiple related work in the area,
+    Using both informal and formal descriptions, we compare \TOP and \TOPHAT with multiple related works in the areas,
     ranging from business process modelling languages, to process algebras and reactive programming frameworks.

### 2019-02-18 sections/relatedwork.tex
some stuff in intuition
merge all communication paragraphs
process algebras wip

@@
 
-The central objective of processes is to model patterns of their behaviour.
-These patterns of behaviour manifest themselves in sequences of actions that processes can engage in.
-Actions can be events from the environment, or signals send internally between processes.
-Processes are a way to model synchronisation patterns.
+There are two main differences between \TOP and process algebras.
+
+The first is a difference in scope.
+Process algebras focus on modelling the input/output behaviour of processes, by explicitly stating which actions are sent and received at certain points in the program.
+The goal of process algebras is formal reasoning about the interaction between subprocesses.
+Typically, one wishes to prove properties such as deadlock-freedom, liveness, or adherence to a protocol specification.
+
+The focus of \TOP is to model collaboration patterns, with the explicit goal of not having to specify how exactly subtasks communicate.
+The declarative specification of data dependencies between subtasks enables our system to hide such details.
+
+TODO: continue here
 
@@
 % \TOP focuses less on formal correctness, and more on practical applicability.
-The big difference between processes and tasks is the intended level of abstraction.
+The big difference between processes and tasks is the level of detail in which programmers need to specify communication between modules.
 While processes are used to explicitly model low-level details about communication, tasks hide these details and provide high-level communication mechanisms to the programmer.
@@
 They evaluate programs as far as possible until further external input is required.
+
+% \paragraph{Communication with the environment}
 
@@
 
-\paragraph{Communication between subsystems}
+% \paragraph{Communication between subsystems}
 
-%
 % \CSP's semicolon is like an internal step $t \Then e$ in \TOP where $t$ uses $\Unit$ to signal when it is done, and $e$ never fails.

### 2019-02-19 sections/relatedwork.tex
the difference between TOP and process algebras

@@
 There are two main differences between \TOP and process algebras.
-
 The first is a difference in scope.
 Process algebras focus on modelling the input/output behaviour of processes, by explicitly stating which actions are sent and received at certain points in the program.
-The goal of process algebras is formal reasoning about the interaction between subprocesses.
+The goal of process algebras is formal reasoning about the interaction between processes.
 Typically, one wishes to prove properties such as deadlock-freedom, liveness, or adherence to a protocol specification.
 
-The focus of \TOP is to model collaboration patterns, with the explicit goal of not having to specify how exactly subtasks communicate.
-The declarative specification of data dependencies between subtasks enables our system to hide such details.
+The focus of \TOP on the other hand is to model collaboration patterns, with the explicit goal of not having to specify how exactly subtasks communicate.
+The declarative specification of data dependencies between subtasks enables \TOP to hide such details.
 
-TODO: continue here
-
-% \begin{quote}
-%   \noindent In \TOP, communication is a means to an end. \\
-%   In \CSP, communication is the central matter of discourse.
-% \end{quote}
-
-%%NOTE: Already said in section intro --TS
-% Despite the different origins, \TOP and \CSP feature concepts that serve similar purposes.
-
-%%NOTE: Focus on what TOP is good at --TS
-% \CSP has a formal semantics that allows various kinds of correctness proofs, including equality of processes, and adherence to a specification.
-% This allows applications in program correctness, proofs of deadlock freedom, liveness, or verification of protocols.
-%
-% \TOP focuses less on formal correctness, and more on practical applicability.
-The big difference between processes and tasks is the level of detail in which programmers need to specify communication between modules.
-While processes are used to explicitly model low-level details about communication, tasks hide these details and provide high-level communication mechanisms to the programmer.
-It wants to be a calculus with intuitive semantics that facilitates communication between programmers and domain experts.
-Such programs are supposed to hide implementation details from domain experts while containing enough information to allow automatic generation of executable applications, including user interfaces.
+The second difference is in the way tasks communicate.
+There are two forms of communication between tasks: Passing values to continuations and sharing data.
+This is different from communication in process algebras, which is based on message-passing.
 

### 2019-02-19 sections/relatedwork.tex
related work: process algebras

@@
 
-The second difference is in the way tasks communicate.
+The second difference concerns internal communication.
 There are two forms of communication between tasks: Passing values to continuations and sharing data.
@@
 
-\paragraph{Communication}
+\paragraph{Similarities}
+There are some aspects that are similar in \TOPHAT and process algebras.
+Internal communication in Hoare's CSP \cite{books/Hoare85CSP} is introduced with the concealment operator.
+The semantics of CSP requires that all concealed actions are handled to exhaustion before any action with the environment can take place.
+This is somewhat similar to \TOPHAT, where all enabled internal steps must be taken until the system can react to input events again.
+Contrast this with Milner's CCS \cite{books/Milner89CAC}, where concealed actions are visible to the outside as $\tau$-actions, and can be interleaved with external communication.
 
-Both tasks and processes have two sorts of communication: with the environment, and between subsystems.
-In both paradigms, communication with the \emph{environment} is input and output.
-They block evaluation of programs until the environment sends or receives events.
-In both paradigms, communication \emph{between subsystems} does not block.
-They evaluate programs as far as possible until further external input is required.
-
-% \paragraph{Communication with the environment}
-
-Processes communicate with the environment by sending or receiving actions
-The \CSP process $(a \to P)$ can engage in action $a$, after which it continues as process $P$.
-% Prefixing has no direction of sending and receiving.
-Hoare uses the neutral phrase \emph{engaging in an action}.
-The interpretation of whether an action stands for input or output is left to the reader.
-For example, a vending machine $P = (\textit{coin} \to \textit{choc} \to P)$ is to be interpreted as taking a coin as input and producing chocolate as output.
-%%NOTE: Not that important I think, we don't use it in the rest of the text. --TS
-% In general, if $B$ is a set of events and $P(x)$ is an expression that evaluates to a process for each $x \in B$, the process $(x:B \to P(x))$ can engage in any one of the actions in $B$, after which it continues as the process determined by $P(x)$.
-% Hoare does not specify the language in which $P(x)$ is to be expressed.
-% He seems to permit any kind of mathematical formula, or any programming language.
-
-Sending and receiving of values is modelled by giving additional structure to the names of actions.
-For example, an action name could be $\textit{in}.5$, which can be interpreted as inputting the value 5.
-%
-In \TOP, input values are typed to match the type of the receiving editor.
-In \CSP action names are essentially strings, subject to implicit restrictions.
-
-% \paragraph{Communication between subsystems}
-
-% \CSP's semicolon is like an internal step $t \Then e$ in \TOP where $t$ uses $\Unit$ to signal when it is done, and $e$ never fails.
-\CSP uses prefixing for both internal and external communication.
-Prefixing becomes internal by using \emph{concealment}.
-Processes must take any concealed steps as soon as possible, before any further communication with the environment can take place.
-This means concealed actions have priority over exposed ones.
-This introduces the possibility of diverging computations, that is processes which continuously take internal steps without ever being able to engage in external communication.
-
-% \paragraph{Sequential composition}
-
-% In \CSP, sequential composition is represented by the semicolon combinator together with the special action $\checkmark$, called \emph{success}.
-% We require that if a process engages in $\checkmark$ at all, it must be its single last action.
-% When that happens, we say the process terminates successfully.
-% The process $P;Q$ behaves like $P$ until $P$ terminates successfully, after which it continues as $Q$.
-% If $P$ never terminates successfully, neither does $P;Q$
-
-\paragraph{Concurrency}
-
-Both tasks and processes are models of \emph{multiprogramming}, as opposed to \emph{multiprocessing}.
-This means that one processor evaluates multiple programs in an interleaving fashion.
-% instead of multiple processors.
-Multiprogramming involves three aspects: concurrency, synchronisation, and nondeterminism.
-
-In \CSP, there are two different combinators for parallel composition: parallel and interleave.
-The parallel process $P \parallel Q$ can take $P$-steps and $Q$-steps in arbitrary interleaving for actions unique to $P$ and $Q$.
-Actions in the alphabets of both $P$ and $Q$ must be taken in synchronisation, so if one process wants to take such a step, it must wait until the other is ready to do so.
-Synchronised steps are the only occasion where two processes actually do something at the same time.
-%
-The interleaving operator $P \mid\mid\mid Q$ does not synchronise any actions of $P$ and $Q$.
-All interleavings are permitted.
-Actions that occur in both alphabets are nondeterministically taken by either $P$ or $Q$.
-
-\paragraph{Synchronisation}
-
-Synchronisation means that an agent has to pause execution until some condition is met.
-The need for synchronisation arises quite naturally in countless variations in situations involving concurrency.
-For multiprogramming, synchronisation always means that some of the possible interleavings are forbidden.
-
-For example, the mutual exclusion problem for two parallel processes $a_p \to b_p \to \text{STOP} $ and $a_q \to b_q \to \text{STOP}$ can be stated as:
-In no interleaving should there be two adjacent $a$'s.
-%
-The mutual exclusion problem for two parallel tasks can be stated as follows.
-Let \TS{inc} and \TS{dec} be tasks that increment and decrement some shared counter.
-Then, in \TS{(inc x >>? dec) <&> (dec x >>? inc)}, the shared counter should at no point be greater than 1.
-This means that in no interleaving should there be two adjacent \TS{inc}'s.
-
-The basic method for synchronisation in \CSP is synchronised prefixing.
-If some action $a$ is in the alphabets of both $P$ and $Q$, then their parallel composition $P \parallel Q$ can take an $a$-step only if both $P$ and $Q$ are ready to take an $a$-step, in which case they both take this step simultaneously.
-Both have to wait for the condition that the other process is ready to take the step.
-All synchronisation problems in \CSP must be solved by employing synchronised steps in some form.
-
-\paragraph{Nondeterminism}
-
-Nondeterminism means that a system can react to the same input in different ways, the choice of which can not be influenced by the environment.
-
-\CSP has an operator to explicitly introduce nondeterminism.
-Furthermore, nondeterminism can arise from the combination of some other combinators, for example choice and concealment.
-%
-Hoare states that nondeterminism is only useful for \emph{specifying} processes, never for \emph{implementing} them \cite{books/Hoare85CSP}.
-\CSP can be used for both.
-A process with nondeterministic behaviour must always be regarded as a specification, while a deterministic process can be seen as specification or implementation.
+Another similarity between \TOPHAT and process algebras, or any system with concurrency for that matter, is the need for synchronisation.
+Broadly speaking, concurrency means that different parts of a program can interact with the environment independently, in an interleaved manner.
+Synchronisation means that only some, but not all, of the possible interleavings are desirable.
+The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise TODO: say that synchronization in TOPHAT is the best.
 

### 2019-02-19 sections/conclusions.tex
Take out future work

@@
 Finally, we have proven type safety and progress for our language.
-
-\paragraph{Future work}
-
-There are a couple of ways in which we would like to continue this line of work.
-
-One of the main motivations to formalize Task Oriented Programmig is to be able to reason about programs.
-In this paper we reason about the language itself, but it would be nice to prove properties about individual programs.
-To this end, we are very interested to see if it is possible to develop an axiomatic semantics for \TOPHAT that allows us to do so.
-There are certain properties of our calculus that make this particularly complex; we have to deal with parallelism, user interacton and references.
-
-We would also like to prove whether certain programs are equivalent, for example to show that the monad laws hold for our step combinator.
-This requires a notion of equality, which in the presence of side effects most certainly needs some form of coalgebraic input-output conformance.
-
-We have implemented the reduction semantics of our language in Idris, whose type system could aid in a formalisation of those proofs.
-
-Another form of reasoning about programs is static analysis.
-Klinik et al.~\cite{conf/ifl/KlinikJP17} have developed a cost analysis for tasks that require resources in order to be executed.
-% Put this line back in for final version of paper: In previous work we have developed a cost analysis for tasks that require resources in order to be executed \cite{conf/ifl/KlinikJP17}.
-This analysis was developed for a simpler task calculus, and could be brought over to the one developed here.
-
-% Put this line back in for final version of paper: In other previous work, we have looked at building a generic feedback system for rule-based problems \cite{UUCS2017013}.
-Naus et al.~\cite{UUCS2017013} have looked at building a generic feedback system for rule-based problems.
-A workflow system typically is rule based, as outlined in their work.
-It would be interesting to fit the generic feedback system to \TOPHAT in order to support end-users working in applications developed in this calculus.
-
-Additionally, we would like to develop visualisations for \TOPHAT language constructs.
-An assistive development environment integrating these visualisations and the presented textual language
-would aid domain experts to model workflows in a more accessible manner.
-A system that visualizes iTask programs has been developed in the past~\cite{DBLP:conf/sfp/StutterheimPA14}.

### 2019-02-19 sections/futurework.tex
Take out future work

% !TEX root=../icfp2019.tex

\section{Future work}

There are a couple of ways in which we would like to continue this line of work.

One of the main motivations to formalize Task Oriented Programmig is to be able to reason about programs.
In this paper we reason about the language itself, but it would be nice to prove properties about individual programs.
To this end, we are very interested to see if it is possible to develop an axiomatic semantics for \TOPHAT that allows us to do so.
There are certain properties of our calculus that make this particularly complex; we have to deal with parallelism, user interacton and references.

We would also like to prove whether certain programs are equivalent, for example to show that the monad laws hold for our step combinator.
This requires a notion of equality, which in the presence of side effects most certainly needs some form of coalgebraic input-output conformance.

We have implemented the reduction semantics of our language in Idris, whose type system could aid in a formalisation of those proofs.

Another form of reasoning about programs is static analysis.
Klinik et al.~\cite{conf/ifl/KlinikJP17} have developed a cost analysis for tasks that require resources in order to be executed.
% Put this line back in for final version of paper: In previous work we have developed a cost analysis for tasks that require resources in order to be executed \cite{conf/ifl/KlinikJP17}.
This analysis was developed for a simpler task calculus, and could be brought over to the one developed here.

% Put this line back in for final version of paper: In other previous work, we have looked at building a generic feedback system for rule-based problems \cite{UUCS2017013}.
Naus et al.~\cite{UUCS2017013} have looked at building a generic feedback system for rule-based problems.
A workflow system typically is rule based, as outlined in their work.
It would be interesting to fit the generic feedback system to \TOPHAT in order to support end-users working in applications developed in this calculus.

Additionally, we would like to develop visualizations for \TOPHAT language constructs.
An assistive development environment integrating these visualizations and the presented textual language
would aid domain experts to model workflows in a more accessible manner.
A system that visualizes iTask programs has been developed in the past~\cite{DBLP:conf/sfp/StutterheimPA14}.

### 2019-02-19 sections/relatedwork.tex
minor fixes to related work

@@
 \label{sec:relatedwork}
+
+\subsection{Task-oriented programming implementations}
 
@@
 
+\paragraph{mTasks}
+\fixme{Hier iets gezelligs over mTasks}
 % Our work differs in three major ways.
@@
 Synchronisation means that only some, but not all, of the possible interleavings are desirable.
-The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise TODO: say that synchronization in TOPHAT is the best.
+The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise TODO: say that synchronisation in TOPHAT is the best.
 
-\subsection{Related web services}
+\subsection{Reactive programming}
 
-\paragraph{IFTTT}
+\paragraph{Esterel \& HipHop}
 
-If This Then That (\IFTTT)~\cite{IFTTT} is a web service that allows users to write and
-configure simple conditional statements called applets, that can interact with
-web services and IoT devices. These applets can be seen as tiny workflows, that
-facilitate collaboration between different services and platforms. The
-complexity of the flow is very limited however, and is not geared towards humans
-collaborating.
-
-\paragraph{Google forms \& Wufoo}
-
-Google Forms~\cite{googleforms} and Wufoo~\cite{wufoo} are web based survey administration apps that allows
-users to compose forms and gather data from them. They are comparable to \TOP and
-iTasks in that they provide an easy way for users to construct forms on the web.
-However, Google Forms and Wufoo do not allow users to define what the system
-then should do with this information, and can therefore not be regarded as a
-workflow or business process modelling systems.
-
-\paragraph{Chorus} % (http://www.chorus-home.org/)
-
-Chorus~\cite{chen2017chorus} is a visual programming environment for online
-collaboration apps. The goal of Chorus is to allow users to program their own
-application, though which a collective task can be coordinated. Users work on a
-preset data-type that drives the behaviour of the user defined applications.
-
-\subsection{Functional Reactive Programming}
+\paragraph{Functional reactive programming}
 
@@
 The details are a topic for future work.
+
+\subsection{Related web services}
+
+\paragraph{IFTTT}
+
+If This Then That (\IFTTT)~\cite{IFTTT} is a web service that allows users to write and
+configure simple conditional statements called applets, that can interact with
+web services and IoT devices. These applets can be seen as tiny workflows, that
+facilitate collaboration between different services and platforms. The
+complexity of the flow is very limited however, and is not geared towards humans
+collaborating.
+
+\paragraph{Google forms \& Wufoo}
+
+Google Forms~\cite{googleforms} and SurveyMonkey~\cite{wufoo} (and their various other tools like Wufoo, Precise Polling and Zoomerang) are web based survey administration apps that allows
+users to compose forms and gather data from them. They are comparable to \TOP and
+iTasks in that they provide an easy way for users to construct forms on the web.
+However, Google Forms and Wufoo do not allow users to define what the system
+then should do with this information, and can therefore not be regarded as a
+workflow or business process modelling systems.
+
+\paragraph{Chorus} % (http://www.chorus-home.org/)
+
+Chorus~\cite{chen2017chorus} is a visual programming environment for online
+collaboration apps. The goal of Chorus is to allow users to program their own
+application, though which a collective task can be coordinated. Users work on a
+preset data-type that drives the behaviour of the user defined applications.

### 2019-02-19 sections/conclusions.tex
returned future work, and named section "outlook"

--- 2019-02-15 sections/conclusions.tex
@@
 
-\paragraph{Future work}
+\subsection{Future work}
 
@@
 
-Additionally, we would like to develop visualisations for \TOPHAT language constructs.
-An assistive development environment integrating these visualisations and the presented textual language
+Additionally, we would like to develop visualizations for \TOPHAT language constructs.
+An assistive development environment integrating these visualizations and the presented textual language
 would aid domain experts to model workflows in a more accessible manner.

### 2019-02-19 sections/relatedwork.tex
add Cooper and Krishnasmurti
added fixme's
add Meyerovich et al
cite FrTime and Flapjax

@@
 \label{sec:relatedwork}
+
+\fixme{Intro related work}
 
@@
 
-The idea of \FRP was pioneered by \citet{conf/icfp/ElliottH97}, in the meantime there are many variants and implementations, with reactive-banana one of the most well-known \cite{reactive-banana}.
+The idea of \FRP was pioneered by \citet{conf/icfp/ElliottH97}.
+In the meantime there are many variants and implementations, where reactive-banana \cite{reactive-banana}, FrTime \cite{CooperK04}, and Flapjax \cite{conf/oopsla/MeyerovichGBCGBK09} belong to the most well-known.
 

### 2019-02-19 sections/outlook.tex
rename future work to outlook

% !TEX root=../icfp2019.tex

\section{Outlook}

\fixme{write this section}

### 2019-02-19 sections/introduction.tex
Fixed contributions in introduction

@@
 
-\subsection{Contributions}
+\subsection{Scope}
 
-\input{sections/contributions}
+In this paper, we set out to develop a formal basis for \TOP, called \TOPHAT (TopHat).
+Our goal is to specify the basic building blocks of declarative workflows.
+We clearly separate them from general purpose programming concepts.
+This allows us to prove properties like progress and type preservation.
+\TOPHAT should be powerful enough to model complicated real-world situations,
+without losing opportunities for formal reasoning.
+
+By separating them in the style of a deep embedding \cite{conf/cefp/Gibbons13}, we make explicit where and where not, they depend on each other. \fixme{what to do with this sentence}
+
+We show that, where process algebras focus on the \emph{coordination} of processes and the \emph{communication} between them,
+\TOP, and therefore \TOPHAT, focus on supporting \emph{tasks} and model ways to \emph{collaborate} on them.
 
@@
 
+\subsection{Contributions}
+
 Our contributions to workflow modelling, functional programming language design, and rapid application development are as follows.
 
-\begin{enumerate}
+\begin{itemize}
+  \item
+    We informally describe the essential concepts of task-oriented programming (\TOP).
 
   \item
-    We informally describe the essential concepts of task-oriented programming (\TOP).
-    \TOP has a strong focus on modelling using collaboration
-    while keeping a constant desire to rapidly generate executable applications into account.
+    We present a formal calculus for modelling declarative workflows, embedded in a simply typed lambda-calculus.
+    This calculus is based on the aforementioned essential \TOP concepts.
 
   \item
-    We present a formal calculus for \TOP, arising naturally from the aforementioned essential concepts.
-    In doing so, we lay the ground work to allow formal reasoning about \TOP specifications in future work.
+    We develop an operational semantics that is driven by user input.
+    The semantics of the task language is clearly separated from the semantics of the underlying host language.
 
   \item
-    Using both informal and formal descriptions, we compare \TOP and \TOPHAT with multiple related works in the areas,
+    Along with the semantics, we present the following semantic observations on tasks:
+    the current value, whether a term is stuck, the current user interface, and the accepted inputs. \fixme{we do not provide observation for user interface!}
+
+  \item
+    We implemented the whole semantic system in the dependently typed programming language Idris \cite{journals/jfp/Brady13}.
+
+  \item
+    We have implemented \TOPHAT in iTasks to demonstrate that the first is a subset of the second.
+
+  \item
+    Using both informal and formal descriptions, we compare \TOP with multiple related works in areas,
     ranging from business process modelling languages, to process algebras and reactive programming frameworks.
 
-\end{enumerate}
+    \item
+      We prove progress and type preservation for our calculus.
+
+\end{itemize}
 

### 2019-02-19 sections/introduction.tex
contributions

@@
   \item
-    We informally describe the essential concepts of task-oriented programming (\TOP).
+    We describe the essential concepts of task-oriented programming.
 
@@
   \item
+    We prove progress and type preservation for our calculus.
+
+  \item
+    Using both the essential concepts and the formal calculus, we compare \TOP with multiple related works, in areas ranging from business process modelling to process algebras and reactive programming.
+
+  \item
     We implemented the whole semantic system in the dependently typed programming language Idris \cite{journals/jfp/Brady13}.
@@
   \item
-    We have implemented \TOPHAT in iTasks to demonstrate that the first is a subset of the second.
+    %To create executable applications, we implemented \TOPHAT in iTasks.
 
-  \item
-    Using both informal and formal descriptions, we compare \TOP with multiple related works in areas,
-    ranging from business process modelling languages, to process algebras and reactive programming frameworks.
-
-    \item
-      We prove progress and type preservation for our calculus.
+    To create executable applications, we implemented the task layer of \TOPHAT in iTasks.
+    This also demonstrates that the former is a subset of the latter.
 

### 2019-02-19 sections/conclusions.tex
fixme
use citet

@@
 
+\fixme{moet passen bij de contributions}
 In this paper we have developed a domain-specific calculus for declarative interactive workflows, named \TOPHAT.
@@
 Another form of reasoning about programs is static analysis.
-Klinik et al.~\cite{conf/ifl/KlinikJP17} have developed a cost analysis for tasks that require resources in order to be executed.
+\citet{conf/ifl/KlinikJP17} have developed a cost analysis for tasks that require resources in order to be executed.
 % Put this line back in for final version of paper: In previous work we have developed a cost analysis for tasks that require resources in order to be executed \cite{conf/ifl/KlinikJP17}.
@@
 % Put this line back in for final version of paper: In other previous work, we have looked at building a generic feedback system for rule-based problems \cite{UUCS2017013}.
-Naus et al.~\cite{UUCS2017013} have looked at building a generic feedback system for rule-based problems.
+\citet{UUCS2017013} have looked at building a generic feedback system for rule-based problems.
 A workflow system typically is rule based, as outlined in their work.

### 2019-02-19 sections/introduction.tex
small fixes in intro

@@
 Finally, there is communication with the \emph{outside} world, where information is entered into the system via input events.
+\fixme{Should this be mentioned earlier?}
 The end points where the outside world interacts with \TOP applications are called editors.
 In generated applications, editors can take many forms, like input fields, selection boxes, or map widgets.
-
-\fixme{Change if title changes.}
-In this paper we study the essence of \TOP.
-We claim that at the heart of \TOP lie modular interactive workflows.
-Workflows because \TOP programs coordinate collaboration between people.
-Interactive because progress of \TOP programs is driven by user input.
-Modular because the embedding in a strongly-typed functional language allows powerful abstractions using type-driven techniques and higher-order functions.
 
@@
 
+\fixme{Rewrite}
+
 In this paper, we set out to develop a formal basis for \TOP, called \TOPHAT (TopHat).
@@
 In this paper we aim to identify the essence of \TOP in formal as well as an informal way.
-These core principles lead to a formal system for \TOP called \TOPHAT.
-\TOPHAT paves the way to formal reasoning about applications written using the \TOP paradigm,
-like those in \ITASKS and \MTASKS.
+These core principles lead to a formal system for \TOP, called \TOPHAT,
+which paves the way to formal reasoning.
+Applications written using the \TOP paradigm, like those in \ITASKS and \MTASKS, currently lack such a feature.
 
@@
 and describe their similarities and differences.\fixme{What do you mean with this sentence?}
+
+\fixme{Change or remove.}
+In this paper we study the essence of \TOP.
+We claim that at the heart of \TOP lie modular interactive workflows.
+Workflows because \TOP programs coordinate collaboration between people.
+Interactive because progress of \TOP programs is driven by user input.
+Modular because the embedding in a strongly-typed functional language allows powerful abstractions using type-driven techniques and higher-order functions.
 

### 2019-02-19 sections/intuition.tex
clarify difference between shared data and shared editor
removed three qed's
fixes in intuition

@@
 Tasks can be executed after each other, at the same time, or conditionally.
-This motivates the combinators step, parallel, and choice.
+This motivates the combinators step ($\Then$, $\Next$), parallel ($\And$), and choice ($\Or$, $\Xor$).
+
+\begin{example}[Breakfast]
+\label{exm:breakfast}
+
+The following program shows the different collaboration operators in action.
+Users have a choice whether they want tea or coffee.
+They always get an egg.
+The drink and the food are prepared in parallel.
+When drink and food are prepared, users can step to eating the result.
+
+\begin{TASK}
+  let makeBreakfast : Task Drink -> Task Food -> Task Unit =
+    \makeDrink . \makeFood . makeDrink <&> makeFood >>? \<<d, f>>. eat d f
+  let main : Task Unit =
+    makeBreakfast (makeTea <?> makeCoffee) makeEgg
+\end{TASK}
+
+\end{example}
 
@@
 
-\begin{example}[Twice]
-
-A higher-order function which executes a given task twice.
-\begin{TASK}
-  let twice : Task $\tau$ -> Task $\tau$ = \t : Task $\tau$. t >>? \x . t
-\end{TASK}
-The task value of the first execution is ignored.
-\qed
-
-\end{example}
+\Cref{exm:breakfast} demonstrates how tasks can be parameterized by other tasks: makeBreakfast is a collaboration pattern that always works the same way, regardless of which food and drink are being prepared.
 
@@
 In this way the system communicates with the environment.
+Input events are synchronous.
+We assume systems react instantaneous to each such event.
 
@@
 Change events $v$ with a value of type $\tau$ and clear events $\Empty$ which empty the editor.
+This results in a typed empty editor $\Enter \tau$, which is annotated with the same type.
 
@@
 This example demonstrates external communication and choice.
-It is a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins.
+It is a vending machine that dispenses a biscuit for one coin and a chocolate bar for two coins.
 \begin{TASK}
-  let vend : Task String = enter Int >>? \n. if n == 1 then edit "small"
-    else if n == 2 then edit "large"
+  let vend : Task Snack = enter Int >>? \n. if n == 1 then edit Biscuit
+    else if n == 2 then edit ChocolateBar
     else fail
@@
 When the user has inserted exactly 1 or 2 coins, the continue button becomes enabled.
-When the user presses the continue button, the machine dispenses either a large or a small biscuit, depending on the amount of money. \qed
+When the user presses the continue button, the machine dispenses either a biscuit or a chocolate bar, depending on the amount of money.
+Snacks are modelled using a custom data type.
 
@@
 
-The final observation is to determine whether a task results in $\Fail$.
+The final observation is to determine whether a task results in $\Fail$ (failure).
 The step combinator and the choice combinator use this to prevent users from picking a failing task.
@@
 The task $t \Next e$ additionally requires a continue event $\Continue$ to proceed.
-Speculative evaluation is designed so that it does not have side effects.
+Speculative evaluation is designed so that possible side effects are undone.
 
@@
 Shared data sources are assignable references whose changes are immediately visible to all tasks interested in them.
+Users can not directly interact with shared data, a shared editor is required for that.
 If $r$ is a reference of type $\tau$, then $\Update r$ is an editor whose value is that of $r$.
@@
 \end{TASK}
-When the agent supplies two of the ingredients by setting the respective shares to \TS{True}, only the step of the smoker that waits for those becomes enabled. \qed
+When the agent supplies two of the ingredients by setting the respective shares to \TS{True}, only the step of the smoker that waits for those becomes enabled.
 
@@
 
-The parallel combinator $\And$ introduces concurrency: Two things are done at the same time.
 Let $t_1$ and $t_2$ be tasks.
 The parallel combination $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
+Therefore $\And$ introduces concurrency.
 For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.

### 2019-02-19 sections/intuition.tex
breakfast example

--- 2019-02-18 sections/intuition.tex
@@
 
+\begin{example}[Breakfast]
+
+The following program shows the different collaboration operators in action.
+The user has a choice whether they want tea or coffee.
+They always get an egg.
+The drink and the food are prepared in parallel.
+When drink and food are prepared, the user can step to eating the result.
+
+\begin{TASK}
+  let makeBreakfast : Task Drink -> Task Food -> Task Unit =
+    \makeDrink . \makeFood . makeDrink <&> makeFood >>? \<<d, f>>. eat d f
+  let main : Task Unit =
+    makeBreakfast (makeTea <?> makeCoffee) makeEgg
+\end{TASK}
+
+\end{example}
+
 \subsection{Tasks are reusable}
@@
 The task value of the first execution is ignored.
-\qed
 
@@
 When the user has inserted exactly 1 or 2 coins, the continue button becomes enabled.
-When the user presses the continue button, the machine dispenses either a large or a small biscuit, depending on the amount of money. \qed
+When the user presses the continue button, the machine dispenses either a large or a small biscuit, depending on the amount of money.
 

### 2019-02-19 sections/intuition.tex
use breakfast as reusable example

@@
 \begin{example}[Breakfast]
+\label{exm:breakfast}
 
 The following program shows the different collaboration operators in action.
-The user has a choice whether they want tea or coffee.
+Users have a choice whether they want tea or coffee.
 They always get an egg.
 The drink and the food are prepared in parallel.
-When drink and food are prepared, the user can step to eating the result.
+When drink and food are prepared, users can step to eating the result.
 
@@
 
-\begin{example}[Twice]
-
-A higher-order function which executes a given task twice.
-\begin{TASK}
-  let twice : Task $\tau$ -> Task $\tau$ = \t : Task $\tau$. t >>? \x . t
-\end{TASK}
-The task value of the first execution is ignored.
-
-\end{example}
+\Cref{exm:breakfast} demonstrates how tasks can be parameterized by other tasks: makeBreakfast is a collaboration pattern that always works the same way, regardless of which food and drink are being prepared.
 

### 2019-02-19 sections/relatedwork.tex
added fixme to related work
small fixes and more fixmes

@@
 \fixme{Hier iets gezelligs over mTasks}
+
 % Our work differs in three major ways.
@@
 
-\paragraph{Workflow Nets}
+\paragraph{Workflow Nets \& YAWL}
 
-Workflow Nets~(\WFN)\cite{journals/jcsc/Aalst98} allow for the modelling and analysis
-of business processes. They are graphical in nature, and clearly display how
-every component is related to each other. A downside of \WFN is that
-they do not facilitate higher order constructs and that they are often not
-directly executable.
+Workflow Nets~(\WFN) \cite{journals/jcsc/Aalst98} allow for the modelling and analysis of business processes.
+They are graphical in nature, and clearly display how every component is related to each other.
+A downside of \WFN is that they do not facilitate higher order constructs.
+Also, they are often not directly executable.
 
-\paragraph{YAWL}
-
-This is not always the case however. A language based on \WFN that is
-actually directly executable is \YAWL by \citet{DBLP:journals/is/AalstH05}.
-It facilitates modelling and execution of
-dynamic workflows, with support for \emph{and}, \emph{or} and \emph{xor} workflow patterns. As
-mentioned, \YAWL programs consist of \WFN, and are therefore programmed
-visually.
+A language based on \WFN that is actually directly executable is \YAWL by \citet{DBLP:journals/is/AalstH05}.
+It facilitates modelling and execution of dynamic workflows, with support for \emph{and}, \emph{or} and \emph{xor} workflow patterns.
+As mentioned, \YAWL programs consist of \WFN, and are therefore programmed visually.
 
@@
 using an \XML format. Processes specified in \BPEL are executable, just like \YAWL.
+
+\fixme{Is this higher order? What combinators can be used? Other limitations?}
 
@@
 \paragraph{Esterel \& HipHop}
+
+\fixme{Write}
 

### 2019-02-19 sections/intuition.tex
intuition: collaboration patterns

--- 2019-02-19 sections/intuition.tex
@@
 
-\subsection{Tasks model real world collaborations between people and machines}
-
 The central objective of \TOP is to \emph{coordinate collaboration}.
-The basic building blocks of \TOPHAT are editors and combinators.
-Editors are tasks with which users interact directly.
-Combinators express ways in which people can work together.
+The basic building blocks of \TOPHAT for this are task combinators.
+They express ways in which people can work together.
 Tasks can be executed after each other, at the same time, or conditionally.
@@
 The following program shows the different collaboration operators in action.
-Users have a choice whether they want tea or coffee.
+Users have a choice ($\Xor$) whether they want tea or coffee.
 They always get an egg.
-The drink and the food are prepared in parallel.
-When drink and food are prepared, users can step to eating the result.
+The drink and the food are prepared in parallel ($\And$).
+When drink and food are prepared, users can step ($\Next$) to eating the results.
 

### 2019-02-19 sections/introduction.tex
scope

@@
 
-\fixme{Rewrite}
+In this paper, we set out to develop a formal basis for \TOP, called \TOPHAT (TopHat).
+Our goal is to pave the way for formal treatment of \TOP programs.
 
-In this paper, we set out to develop a formal basis for \TOP, called \TOPHAT (TopHat).
-Our goal is to specify the basic building blocks of declarative workflows.
-We clearly separate them from general purpose programming concepts.
-This allows us to prove properties like progress and type preservation.
-\TOPHAT should be powerful enough to model complicated real-world situations,
-without losing opportunities for formal reasoning.
+\ITASKS is the de-facto reference implementation of \TOP.
+It has been designed for developing real-world applications, and is constantly being extended and improved with this goal in mind.
+All the bells and whistles of \ITASKS, the different variations of task combinators, and the details that come with real-world requirements, make it hard to see what the essence of \TOP is.
+It also makes formal reasoning about \TOP programs impossible.
 
-By separating them in the style of a deep embedding \cite{conf/cefp/Gibbons13}, we make explicit where and where not, they depend on each other. \fixme{what to do with this sentence}
+In this paper, we want to take a step back and look at the essence of \TOP.
+We do this both formally and informally.
+Informally in the sense that we give an intuitive description of the features that define task-oriented programming.
+Formally in the sense that we develop a calculus which formalizes these features as language constructs, and we give them semantics in the style that is common in programming language research.
 
-We show that, where process algebras focus on the \emph{coordination} of processes and the \emph{communication} between them,
-\TOP, and therefore \TOPHAT, focus on supporting \emph{tasks} and model ways to \emph{collaborate} on them.
+One of the defining aspects of \ITASKS is its embedding in a functional programming language.
+This creates a synergy where \ITASKS profits from the expressivity of Clean, while enhancing the language with functionality for user interaction.
+The tight integration of task DSL and host language makes it sometimes difficult to see where the boundaries are.
 
-In this paper we aim to identify the essence of \TOP in formal as well as an informal way.
-These core principles lead to a formal system for \TOP, called \TOPHAT,
-which paves the way to formal reasoning.
-Applications written using the \TOP paradigm, like those in \ITASKS and \MTASKS, currently lack such a feature.
-
-\TOP applications give rise to concepts like concurrency, synchronisation, \GUI-programming and program generation.
-In this paper we identify common concepts from already existing languages and frameworks,
-and describe their similarities and differences.\fixme{What do you mean with this sentence?}
-
-\fixme{Change or remove.}
-In this paper we study the essence of \TOP.
-We claim that at the heart of \TOP lie modular interactive workflows.
-Workflows because \TOP programs coordinate collaboration between people.
-Interactive because progress of \TOP programs is driven by user input.
-Modular because the embedding in a strongly-typed functional language allows powerful abstractions using type-driven techniques and higher-order functions.
+In \TOPHAT, we separate the task layer and the underlying host language, both syntactically and semantically.
+In this way we make explicit which of the synergetic effects of \TOP come from the task layer, and which come from functional programming.
 

### 2019-02-19 sections/conclusions.tex
match conclusion with contributions

@@
 
-\fixme{moet passen bij de contributions}
-In this paper we have developed a domain-specific calculus for declarative interactive workflows, named \TOPHAT.
-We have shown that our calculus is modular, and clearly seperated from the host language it is embedded in.
-By developing this calculus, we have identified the essence of Task-Oriented Programming.
-In addition to the formal semantics, we have also demonstrated how to build a user interface from a task description.
+In this paper we have identified and intuitively described the essence of task-oriented programming.
+We then formalized this essence by developing a domain-specific calculus for declarative interactive workflows, called \TOPHAT.
+The task calculus and the host language are clearly separated, to make explicit where the boundaries are.
+The semantics of the task layer is driven by user input.
 We have compared \TOPHAT with workflow moddeling languages, process algebras, functional reactive programming and session types to point out differences and similarities.

### 2019-02-19 sections/conclusions.tex
future work

@@
 
-One of the main motivations to formalize Task Oriented Programmig is to be able to reason about programs.
+One of the main motivations to formalize task-oriented programming is to be able to reason about programs.
 In this paper we reason about the language itself, but it would be nice to prove properties about individual programs.
 To this end, we are very interested to see if it is possible to develop an axiomatic semantics for \TOPHAT that allows us to do so.
-There are certain properties of our calculus that make this particularly complex; we have to deal with parallelism, user interacton and references.
+There are certain properties of our calculus that make this particularly complex:
+We have to deal with parallelism, user interaction, and references.
 
@@
 This requires a notion of equality, which in the presence of side effects most certainly needs some form of coalgebraic input-output conformance.
-
-We have implemented the reduction semantics of our language in Idris, whose type system could aid in a formalisation of those proofs.
+We have implemented the reduction semantics of our language in Idris, whose type system could aid in the formalisation of such proofs.
 
@@
 \citet{conf/ifl/KlinikJP17} have developed a cost analysis for tasks that require resources in order to be executed.
-% Put this line back in for final version of paper: In previous work we have developed a cost analysis for tasks that require resources in order to be executed \cite{conf/ifl/KlinikJP17}.
 This analysis was developed for a simpler task calculus, and could be brought over to the one developed here.
 
-% Put this line back in for final version of paper: In other previous work, we have looked at building a generic feedback system for rule-based problems \cite{UUCS2017013}.
 \citet{UUCS2017013} have looked at building a generic feedback system for rule-based problems.

### 2019-02-20 sections/relatedwork.tex
added more about BPEL

@@
 language allows for the specification of actions within business processes,
-using an \XML format. Processes specified in \BPEL are executable, just like \YAWL.
-
-\fixme{Is this higher order? What combinators can be used? Other limitations?}
+using an \XML format.
+The language is mainly used for coordinating web services.
+Two workflow patterns are supported; execution of services can be done sequential or in parallel.
+On top of that, processes can be guarded by conditionals.
+There is no support for higher order processes however.
+Processes described in \BPEL can be regarded as activity graphs, and they can also be rendered as such.
+The specified processes in \BPEL are directly executable, just like \YAWL.
 

### 2019-02-20 appendices/proofs.tex
restructure appendices
Overhaul of proof appendix

--- 2019-02-06 sections/appendix.tex
@@
-% !TEX root=../pldi2019.tex
+% !TEX root=../icfp2019.tex
 
-\section{Appendix}
+\section{Proofs}
 
-  \subsection{Evaluation rules}
+\subsection{Theorem~\ref{thm:pres-eval}}
 
-  % \begin{figure}
-
-    \begin{mathpar}
-      \boxed{\RelationE} \\
-      \userule{E-Value} \\
-      \userule{E-App} \\
-      \userule{E-IfTrue} \\
-      \userule{E-IfFalse} \\
-      \userule{E-Pair} \\
-      \userule{E-Ref} \quad
-      \userule{E-Deref} \\
-      \userule{E-Assign} \\
-      \userule{E-Edit} \quad
-      \userule{E-Enter} \quad
-      \userule{E-Update} \\
-      \userule{E-Then} \quad
-      \userule{E-Next} \\
-      \userule{E-And} \quad
-      \userule{E-Or} \\
-      \userule{E-Xor} \quad
-      \userule{E-Appoint} \quad
-      \userule{E-Fail}
-    \end{mathpar}
-
-  %   \caption{Evaluation semantics} \label{fig:evaluation-semantics}
-  % \end{figure}
-
-  \subsection{Additional typing rules}
-
-  % \begin{figure}[h]
-
-    \begin{mathpar}
-      \boxed{\RelationT} \\
-      \userule{T-Var} \quad
-      \userule{T-Loc} \\
-      \userule{T-Abs} \quad
-      \userule{T-App} \\
-      \userule{T-If} \\
-      \userule{T-Pair} \\
-      \userule{T-Ref} \quad
-      \userule{T-Deref} \\\
-      \userule{T-Assign}
-    \end{mathpar}
-
-  %   \caption{Additional typing rules}
-  % \end{figure}
-
-  \subsection{Proofs}
-
-  \subsubsection{Theorem~\ref{thm:pres-eval}}
 \begin{proof}
@@
 
-  \noindent\textbf{Case} $e=\lambda x:\tau.e, e_1 e_2, x, c, l, e_1 \star e_2,
-      \If{e_1}{e_2}{e_3},$\\
-      $\tuple{e_1, e_2},\unit,\Ref e,!e,e_1 := e_2$ preservation has
-      been proven for these cases by Pierce~\cite{books/Pierce02TAPL}.\\
+  \case
+    {$e=\lambda x:\tau.e, e_1 e_2, x, c, l, e_1 \star e_2,\If{e_1}{e_2}{e_3},\tuple{e_1, e_2},\unit,\Ref e,!e,e_1 := e_2$}
+    {Preservation has been proven for these cases by Pierce~\cite{books/Pierce02TAPL}.}
 
-  \noindent\textbf{Case} $\userule{E-Edit}$\\
-    \indent Given that $\Gamma,\Sigma\infers\Edit e:\Task \tau$ and $\Sigma\infers s$,
-      \refrule{T-Edit} gives us that $\Gamma,\Sigma\infers e:\tau$. The induction
-      hypothesis gives us that $e,s\evaluate v,s'$ also preserves, and thus
-      $\Gamma,\Sigma\infers v:\tau$ and $\Sigma\infers s'$. Therefore
-      $\Gamma,\Sigma\infers\Edit v:\Task\tau$.\\
+  \case
+    {$\userule{E-Edit}$}
+    {Given that $\Gamma,\Sigma\infers\Edit e:\Task \tau$ and $\Sigma\infers s$,\refrule{T-Edit} gives us that $\Gamma,\Sigma\infers e:\tau$.
+    The induction hypothesis gives us that $e,s\evaluate v,s'$ also preserves, and thus $\Gamma,\Sigma\infers v:\tau$ and $\Sigma\infers s'$.
+    Therefore $\Gamma,\Sigma\infers\Edit v:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{E-Enter}$\\
-    \indent Evaluation does not alter $e$ and $s$, therefore this case holds tivially.\\
+  \case
+    {$\userule{E-Enter}$}
+    {Evaluation does not alter $e$ and $s$, therefore this case holds trivially.}
 
-  \noindent\textbf{Case} $\userule{E-Update}$\\
-    \indent Given that $\Gamma,\Sigma\infers \Edit e:\Task \tau$ and
-      $\Sigma\infers s$, \refrule{T-Update} gives us that $\Gamma,\Sigma\infers e:\Ref \tau$.
-      The induction hypothesis gives us that $e,s\evaluate l,s'$ also preserves,
-      and thus $\Gamma,\Sigma\infers l:\Ref\tau$ and $\Sigma\infers s'$.
-      Therefore $\Gamma,\Sigma\infers\Update l:\Task\tau$\\
+  \case
+    {$\userule{E-Update}$}
+    {Given that $\Gamma,\Sigma\infers \Edit e:\Task \tau$ and $\Sigma\infers s$, \refrule{T-Update} gives us that $\Gamma,\Sigma\infers e:\Ref \tau$.
+    The induction hypothesis gives us that $e,s\evaluate l,s'$ also preserves, and thus $\Gamma,\Sigma\infers l:\Ref\tau$ and $\Sigma\infers s'$.
+    Therefore $\Gamma,\Sigma\infers\Update l:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{E-Fail}$\\
-    \indent Evaluation does not alter $e$ and $s$, therefore this case holds tivially.\\
+  \case
+    {$\userule{E-Fail}$}
+    {Evaluation does not alter $e$ and $s$, therefore this case holds trivially.}
 
-  \noindent\textbf{Case} $\userule{E-Then}$\\
-    \indent Given that $\Gamma,\Sigma\infers e_1\Then e_2:\Task \tau$ and
-      $\Sigma\infers s$, \refrule{T-Then} gives us that
-      $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and
-      $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$. By the induction
-      hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus
-      $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Sigma\infers s'$. Therefore
-      $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$.\\
+  \case
+    {$\userule{E-Then}$}
+    {Given that $\Gamma,\Sigma\infers e_1\Then e_2:\Task \tau$ and $\Sigma\infers s$, \refrule{T-Then} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$.
+    By the induction hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Sigma\infers s'$.
+    Therefore $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{E-Next}$\\
-    \indent Given that $\Gamma,\Sigma\infers e_1\Next e_2:\Task \tau$ and
-      $\Sigma\infers s$, \refrule{T-Next} gives us that
-      $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and
-      $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$. By the induction
-      hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus
-      $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Sigma\infers s'$. Therefore
-      $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$.\\
+  \case
+    {$\userule{E-Next}$}
+    {Given that $\Gamma,\Sigma\infers e_1\Next e_2:\Task \tau$ and $\Sigma\infers s$, \refrule{T-Next} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$.
+    By the induction hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Sigma\infers s'$.
+    Therefore $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{E-And}$\\
-    \indent  Given that $\Gamma,\Sigma\infers e_1\And e_2:\Task(\tau_1\times\tau_2)$
-      and $\Sigma\infers s$, \refrule{T-And} gives us that
-      $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and
-      $\Gamma,\Sigma\infers e_2:\Task\tau_2$. By the induction hypothesis, we
-      know that both $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$
-      preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$,
-      $\Sigma\infers s'$, $\Gamma,\Sigma\infers t_2:\Task\tau_2$ and
-      $\Sigma\infers s''$. Therfore
-      $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$\\
+  \case
+    {$\userule{E-And}$}
+    {Given that $\Gamma,\Sigma\infers e_1\And e_2:\Task(\tau_1\times\tau_2)$ and $\Sigma\infers s$, \refrule{T-And} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\Task\tau_2$.
+    By the induction hypothesis, we know that both $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$ preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$, $\Sigma\infers s'$, $\Gamma,\Sigma\infers t_2:\Task\tau_2$ and $\Sigma\infers s''$.
+    Therefore $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$.}
 
-  \noindent\textbf{Case} $\userule{E-Or}$\\
-    \indent Given that $\Gamma,\Sigma\infers e_1\Or e_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau$ and
-      $\Gamma,\Sigma\infers e_2:\Task\tau$. By the induction hypothesis, we have that both
-      $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$ preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau$, $\Sigma\infers s'$,
-      $\Gamma,\Sigma\infers t_2:\Task\tau$ and $\Sigma\infers s''$. Therefore $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$.\\
+  \case
+    {$\userule{E-Or}$}
+    {Given that $\Gamma,\Sigma\infers e_1\Or e_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau$ and $\Gamma,\Sigma\infers e_2:\Task\tau$.
+    By the induction hypothesis, we have that both $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$ preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau$, $\Sigma\infers s'$, $\Gamma,\Sigma\infers t_2:\Task\tau$ and $\Sigma\infers s''$.
+    Therefore $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{E-Xor}$\\
-    \indent Evaluation does not alter $e$ and $s$, therefore this case holds tivially.
+  \case
+    {$\userule{E-Xor}$}
+    {Evaluation does not alter $e$ and $s$, therefore this case holds trivially.}
 
@@
 
-\subsubsection{Lemma~\ref{lem:presvalue}}
+\subsection{Lemma~\ref{lem:presvalue}}
+
 \begin{proof}
@@
 
-  \noindent\textbf{Case} $\Value{(\Edit v,s)}=v$ By T-Edit, if
-  $\Gamma,\Sigma\infers \Edit v:\Task\tau$, then $\Gamma,\Sigma\infers v:\tau$.\\
+\case {$\Value{(\Edit v,s)}=v$}
+      {By T-Edit, if $\Gamma,\Sigma\infers \Edit v:\Task\tau$, then $\Gamma,\Sigma\infers v:\tau$.}
 
-  \noindent\textbf{Case} $\Value{(\Enter \tau,s)}=\bot$ Since this case does not
-  lead to a value, the lemma holds trivially.\\
+\case {$\Value{(\Enter \tau,s)}=\bot$ }
+      {Since this case does not lead to a value, the lemma holds trivially.}
 
-  \noindent\textbf{Case} $\Value{(\Update l,s)}=s(l)$ Given that
-  $\Gamma,\Sigma\infers\Update l:\Task \tau$ and $\Sigma\infers s$, we know that
-  $\Gamma,\Sigma\infers s(l):\tau$ by definiton.\\
+\case {$\Value{(\Update l,s)}=s(l)$ }
+      {Given that $\Gamma,\Sigma\infers\Update l:\Task \tau$ and $\Sigma\infers s$,
+      we know that $\Gamma,\Sigma\infers s(l):\tau$ by definiton.}
 
-  \noindent\textbf{Case} $\Value{(\Fail,s)}=\bot$ Since this case does not lead
-  to a value, the lemma holds trivially.\\
+\case {$\Value{(\Fail,s)}=\bot$ }
+      {Since this case does not lead to a value, the lemma holds trivially.}
 
-  \noindent\textbf{Case} $\Value{(t_1\Then e_2,s)}=\bot$ Since this case does
-  not lead to a value, the lemma holds trivially.\\
+\case {$\Value{(t_1\Then e_2,s)}=\bot$}
+      { Since this case does not lead to a value, the lemma holds trivially.}
 
-  \noindent\textbf{Case} $\Value{(t_2\Next e_2,s)}=\bot$ Since this case does
-  not lead to a value, the lemma holds trivially.\\
+\case {$\Value{(t_2\Next e_2,s)}=\bot$ }
+      {Since this case does not lead to a value, the lemma holds trivially.}
 
-  \noindent\textbf{Case} $\Value{(t_1\And t_2,s)}=\tuple{v_1, v_2}$ given that
-  $\Value{(t_1,s)}=v_1\wedge\Value{(t_2,s)}=v_2$\\ By \refrule{T-And} we have that
-  $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and
-  $\Gamma,\Sigma\infers t_1:\tau_1$ and $\Gamma,\Sigma\infers t_2:\tau_2$. By the
-  induction hypothesis, $ \Value{(t_1,s)}=v_1$ and $\Value{(t_2,s)}=v_2$ preserve,
-  and thus $\Gamma,\Sigma\infers v_1:\tau_1$ and $\Gamma,\Sigma\infers v_2:\tau_2$.
-  This gives us that $\Gamma,\Sigma\infers \tuple{v_1, v_2}:\Task(\tau_1\times\tau_2)$ \\
+\case {$\Value{(t_1\And t_2,s)}=\tuple{v_1, v_2}$ given that $\Value{(t_1,s)}=v_1\wedge\Value{(t_2,s)}=v_2$}
+      {By \refrule{T-And} we have that $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and $\Gamma,\Sigma\infers t_1:\tau_1$ and $\Gamma,\Sigma\infers t_2:\tau_2$.
+      By the induction hypothesis, $ \Value{(t_1,s)}=v_1$ and $\Value{(t_2,s)}=v_2$ preserve, and thus $\Gamma,\Sigma\infers v_1:\tau_1$ and $\Gamma,\Sigma\infers v_2:\tau_2$.
+      This gives us that $\Gamma,\Sigma\infers \tuple{v_1, v_2}:\Task(\tau_1\times\tau_2)$.}
 
-  \noindent\textbf{Case} $\Value{(t_1\And t_2,s)}=\bot$ given that
-  $\neg(\Value{(t_1,s)}=v_1\wedge\Value{(t_2,s)}=v_2)$\\ Since this case does
-  not lead to a value, the lemma holds trivially.\\
+\case {$\Value{(t_1\And t_2,s)}=\bot$ given that $\neg(\Value{(t_1,s)}=v_1\wedge\Value{(t_2,s)}=v_2)$}
+      { Since this case does not lead to a value, the lemma holds trivially.}
 
-  \noindent\textbf{Case} $\Value{(t_1\Or t_2,s)}=v_1$ given that
-  $\Value{(t_1,s)}=v_1$\\
+\case
+{$\Value{(t_1\Or t_2,s)}=v_1$ given that
+  $\Value{(t_1,s)}=v_1$}{
   By \refrule{T-Or} we have that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$,
@@
   $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis, we have
-  that $\Gamma,\Sigma\infers v_1:\tau$.\\
+  that $\Gamma,\Sigma\infers v_1:\tau$.}
 
-  \noindent\textbf{Case} $\Value{(t_1\Or t_2,s)}=v_2$ given that
-  $\Value{(t_1,s)}=\bot\wedge\Value{(t_2,s)}=v_2$\\
+\case
+{$\Value{(t_1\Or t_2,s)}=v_2$ given that
+  $\Value{(t_1,s)}=\bot\wedge\Value{(t_2,s)}=v_2$}{
   By \refrule{T-Or} we have that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$,
@@
   $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis, we have
-  that $\Gamma,\Sigma\infers v_2:\tau$.\\
+  that $\Gamma,\Sigma\infers v_2:\tau$.}
 
-  \noindent\textbf{Case} $\Value{(t_1\Or t_2,s)}=\bot$ given that
-  $\Value{(t_1,s)}=\bot\wedge\Value{(t_2,s)}=\bot$\\ Since this case does not
-  lead to a value, the lemma holds trivially.\\
+\case
+{$\Value{(t_1\Or t_2,s)}=\bot$ given that
+  $\Value{(t_1,s)}=\bot\wedge\Value{(t_2,s)}=\bot$}{ Since this case does not
+  lead to a value, the lemma holds trivially.}
 
-  \noindent\textbf{Case} $\Value{(t_1\Xor t_2,s)}=\bot$ Since this case does not
-  lead to a value, the lemma holds trivially.
+\case{$\Value{(t_1\Xor t_2,s)}=\bot$ }{Since this case does not
+  lead to a value, the lemma holds trivially.}
 
@@
 
-\subsubsection{Theorem~\ref{thm:pres-stride}}
+\subsection{Theorem~\ref{thm:pres-stride}}
+
 \begin{proof}
@@
 
-  \noindent\textbf{Case} $\userule{S-Fail}$ \\
-    \indent  Since this case does not alter the
-  expression, the theorem holds trivially.\\
+\case{$\userule{S-Fail}$}
+     {Since this case does not alter the expression, the theorem holds trivially.}
 
-  \noindent\textbf{Case} $\userule{S-Xor}$ \\
-    \indent  Since this case does not alter the
-  expression, the theorem holds trivially.\\
+\case{$\userule{S-Xor}$}
+     {Since this case does not alter the expression, the theorem holds trivially.}
 
-  \noindent\textbf{Case} $\userule{S-Update}$ \\
-    \indent  Since this case does not alter
-  the expression, the theorem holds trivially.\\
+\case{$\userule{S-Update}$}
+     {Since this case does not alter the expression, the theorem holds trivially.}
 
-  \noindent\textbf{Case} $\userule{S-Fill}$ \\
-    \indent  Since this case does not alter the
-  expression, the theorem holds trivially.\\
+\case{$\userule{S-Fill}$}
+     {Since this case does not alter the expression, the theorem holds trivially.}
 
-  \noindent\textbf{Case} $\userule{S-Edit}$ \\
-    \indent  Since this case does not alter the
-  expression, the theorem holds trivially.\\
+\case{$\userule{S-Edit}$}
+     {Since this case does not alter the expression, the theorem holds trivially.}
 
-  \noindent\textbf{Case} $\userule{S-And}$ \\
-    \indent  Given that
-  $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$, by
-  \refrule{T-And we} have $\Gamma,\Sigma\infers t_1:\tau_1$ and
-  $\Gamma,\Sigma\infers t_2:\tau_2$. By the induction hypothesis, we also have $\Gamma,\Sigma\infers t_1':\tau_1$ and $\Gamma,\Sigma\infers t_2':\tau_2$.
-  This gives us that
-  $\Gamma,\Sigma\infers t_1'\And t_2':\Task(\tau_1\times\tau_2)$.\\
+\case{$\userule{S-And}$}
+     {Given that $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$, by \refrule{T-And we} have $\Gamma,\Sigma\infers t_1:\tau_1$ and $\Gamma,\Sigma\infers t_2:\tau_2$.
+     By the induction hypothesis, we also have $\Gamma,\Sigma\infers t_1':\tau_1$ and $\Gamma,\Sigma\infers t_2':\tau_2$.
+     This gives us that $\Gamma,\Sigma\infers t_1'\And t_2':\Task(\tau_1\times\tau_2)$.}
 
-  \noindent\textbf{Case} $\userule{S-Next}$ \\
-    \indent  Given that
+\case{$\userule{S-Next}$}
+       {Given that
   $\Gamma,\Sigma\infers e_1\Next e_2:\Task \tau$, \refrule{T-Then} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
@@
   $\Gamma,\Sigma\infers t_1':\Task\tau_1$. Therefore
-  $\Gamma,\Sigma\infers t_1'\Next e_2:\Task\tau$.\\\\
+  $\Gamma,\Sigma\infers t_1'\Next e_2:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{S-OrLeft}$ \\
-    \indent  Given that
+\case{$\userule{S-OrLeft}$}
+       {Given that
   $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$, by \refrule{T-Or} we have
   $\Gamma,\Sigma\infers t_1:\Task\tau$. By the induction hypothesis, we know that
-  $t_1\stride t_1'$ preserves and thus $\Gamma,\Sigma\infers t_1':\Task\tau$.\\
+  $t_1\stride t_1'$ preserves and thus $\Gamma,\Sigma\infers t_1':\Task\tau$.}
 
-  \noindent\textbf{Case}\\
-  $\userule{S-OrRight}$ \\
-    \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$, by \refrule{T-Or} we have $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis, we know that
-  $t_2\stride t_2'$ preserves and thus $\Gamma,\Sigma\infers t_2':\Task\tau$.\\
+\case{$\userule{S-OrRight}$ }
+     {Given that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$, by \refrule{T-Or} we have $\Gamma,\Sigma\infers t_2:\Task\tau$.
+     By the induction hypothesis, we know that $t_2\stride t_2'$ preserves and thus $\Gamma,\Sigma\infers t_2':\Task\tau$.}
 
-  \noindent\textbf{Case}\\
-  $\userule{S-OrNone}$ \\
-    \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$, by \refrule{T-Or} we have
-  $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
-  By the induction hypothesis, we know that $t_1\stride t_1'$ and
-  $t_2\stride t_2'$ preserve, and thus
-  $\Gamma,\Sigma\infers t_1'\Or t_2':\Task\tau$.\\
+  \case{$\userule{S-OrNone}$}
+  {Given that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$, by \refrule{T-Or} we have $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
+  By the induction hypothesis, we know that $t_1\stride t_1'$ and $t_2\stride t_2'$ preserve, and thus $\Gamma,\Sigma\infers t_1'\Or t_2':\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{S-ThenStay}$ \\
-    \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$,
-  by \refrule{T-Then} we have $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
-  $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. By the induction
-  hypothesis, we know that $t_1\stride t_1'$ preserves, and thus
-  $\Gamma,\Sigma\infers t_1'\Then e_2:\Task\tau$.\\
+  \case{$\userule{S-ThenStay}$ }
+  {Given that $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$, by \refrule{T-Then} we have $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$.
+  By the induction hypothesis, we know that $t_1\stride t_1'$ preserves, and thus $\Gamma,\Sigma\infers t_1'\Then e_2:\Task\tau$.}
 
-  \noindent\textbf{Case}\\
-   $\userule{S-ThenFail}$ \\
-    \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$,
+  \case{$\userule{S-ThenFail}$}
+  {Given that $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$,
   by \refrule{T-Then} we have $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
@@
   hypothesis, we know that $t_1\stride t_1'$ preserves, and thus
-  $\Gamma,\Sigma\infers t_1'\Then e_2:\Task\tau$.\\
+  $\Gamma,\Sigma\infers t_1'\Then e_2:\Task\tau$.}
 
-  \noindent\textbf{Case}\\
-  $\userule{S-ThenCont}$\\
-    \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$, by \refrule{T-Then} we have
+  \case{$\userule{S-ThenCont}$}
+  {Given that $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$, by \refrule{T-Then} we have
   $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
@@
   we know that $t_2\stride t_2'$ preserves. Therefore
-  $\Gamma,\Sigma\infers t_2':\Task\tau$.
+  $\Gamma,\Sigma\infers t_2':\Task\tau$.}
   %
@@
 
-\subsubsection{Theorem~\ref{thm:pres-norm}}
+\subsection{Theorem~\ref{thm:pres-norm}}
 
@@
 
-\subsubsection{Theorem~\ref{thm:pres-handle}}
+\subsection{Theorem~\ref{thm:pres-handle}}
 
@@
 
-  \noindent\textbf{Case} $\userule{H-Change}$\\
-  \indent Given that
-  $\Gamma,\Sigma\infers\Edit v:\Task\tau$ and $\Sigma\infers s$, the
-  \refrule{H-Change} rule additionally gives us that $v,v':\tau$. Therefore by
-  \refrule{T-Edit} we have that $\Gamma,\Sigma\infers\Edit v':\Task\tau$\\
+  \case
+    {$\userule{H-Change}$}
+    {Given that $\Gamma,\Sigma\infers\Edit v:\Task\tau$ and $\Sigma\infers s$, the \refrule{H-Change} rule additionally gives us that $v,v':\tau$.
+Therefore by \refrule{T-Edit} we have that $\Gamma,\Sigma\infers\Edit v':\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{H-Empty}$\\
-  \indent  Given that
-  $\Gamma,\Sigma\infers\Edit v:\Task\beta$ and $\Sigma\infers s$, the
-  \refrule{H-Empty} rule additionally gives us that $v:\tau$. Then by
-  \refrule{T-Enter} we have $\Gamma,\Sigma\infers\Enter \tau:\Task\tau$ \\
+  \case
+    {$\userule{H-Empty}$}
+    {Given that $\Gamma,\Sigma\infers\Edit v:\Task\beta$ and $\Sigma\infers s$, the \refrule{H-Empty} rule additionally gives us that $v:\tau$.
+    Then by \refrule{T-Enter} we have $\Gamma,\Sigma\infers\Enter \tau:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{H-Fill}$\\
-  \indent  Given that
-  $\Gamma,\Sigma\infers\Enter\tau$ and $\Sigma\infers s$, the \refrule{H-Fill}
-  rule additionally gives us that $v':\tau$. Then by \refrule{T-Enter} we have
-  $\Gamma,\Sigma\infers \Edit v':\Task\tau$.\\
+  \case
+    {$\userule{H-Fill}$}
+    {Given that $\Gamma,\Sigma\infers\Enter\tau$ and $\Sigma\infers s$, the \refrule{H-Fill} rule additionally gives us that $v':\tau$.
+    Then by \refrule{T-Enter} we have $\Gamma,\Sigma\infers \Edit v':\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{H-Update}$\\
-  \indent  Given that
-  $\Gamma,\Sigma\infers\Update l:\Task\tau$ and $\Sigma\infers s$. This gives us
-  that $\Sigma(l)=\tau$, and we additionally obtain $s(l),v':\tau$ by
-  \refrule{H-Update}. By application of Lemma~\ref{lemmasigmaconsistent} this
-  case holds.\\
+  \case
+    {$\userule{H-Update}$}
+    {Given that $\Gamma,\Sigma\infers\Update l:\Task\tau$ and $\Sigma\infers s$.
+    This gives us that $\Sigma(l)=\tau$, and we additionally obtain $s(l),v':\tau$ by \refrule{H-Update}.
+    By application of Lemma~\ref{lemmasigmaconsistent} this case holds.}
 
-  \noindent\textbf{Case} $\userule{H-PickLeft}$\\
-  \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Sigma\infers s$, then by
-  \refrule{T-Xor} we have $\Gamma,\Sigma\infers t_1:\Task \tau$.\\
+  \case
+    {$\userule{H-PickLeft}$}
+    {Given that $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Sigma\infers s$,
+    then by \refrule{T-Xor} we have $\Gamma,\Sigma\infers t_1:\Task \tau$.}
 
-  \noindent\textbf{Case} $\userule{H-PickRight}$\\
-  \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Sigma\infers s$, then by
-  \refrule{T-Xor} we have $\Gamma,\Sigma\infers t_2:\Task \tau$. \\
+  \case
+    {$\userule{H-PickRight}$}
+    {Given that $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Sigma\infers s$,
+    then by \refrule{T-Xor} we have $\Gamma,\Sigma\infers t_2:\Task \tau$.}
 
-  \noindent\textbf{Case} $\userule{H-Next}$\\
-  \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Next e_2 :\Task\tau$ and $\Sigma\infers s$. Then by
-  \refrule{T-Next}, we have $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
-  $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. Then by \refrule{T-Then} we
-  obtain $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$.\\
+  \case
+    {$\userule{H-Next}$}
+    {Given that $\Gamma,\Sigma\infers t_1\Next e_2 :\Task\tau$ and $\Sigma\infers s$.
+    Then by \refrule{T-Next}, we have $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$.
+    Then by \refrule{T-Then} we obtain $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{H-PassThen}$\\
-  \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$ and $\Sigma\infers s$,
-  \refrule{T-Then} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
-  $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. By the induction hypothesis, we
-  know that $t_1,s\handle{i}t_1',s'$ also preserves and thus
-  $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Gamma,Sigma\infers s'$. By
-  \refrule{T-Then} we now obtain that
-  $\Gamma,\Sigma\infers t_1'\Then e_2:\Task\tau$. \\
+  \case
+    {$\userule{H-PassThen}$}
+    {Given that $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$ and $\Sigma\infers s$,
+    \refrule{T-Then} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$.
+    By the induction hypothesis, we know that $t_1,s\handle{i}t_1',s'$ also preserves and thus $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Gamma,Sigma\infers s'$.
+    By \refrule{T-Then} we now obtain that $\Gamma,\Sigma\infers t_1'\Then e_2:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{H-PassNext}$\\
-  \indent  Given that
+  \case
+    {$\userule{H-PassNext}$}
+    {Given that
   $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$ and $\Sigma\infers s$,
@@
   \refrule{T-Next} we now obtain that
-  $\Gamma,\Sigma\infers t_1'\Next e_2:\Task\tau$. \\
+  $\Gamma,\Sigma\infers t_1'\Next e_2:\Task\tau$. }
 
-  \noindent\textbf{Case} $\userule{H-FirstAnd}$\\
-  \indent  Given that
+  \case
+    {$\userule{H-FirstAnd}$}
+    {Given that
   $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and
@@
   \refrule{T-Next} we obtain
-  $\Gamma,\Sigma\infers t_1'\And t_2:\Task(\tau_1\times\tau_2)$.\\
+  $\Gamma,\Sigma\infers t_1'\And t_2:\Task(\tau_1\times\tau_2)$.}
 
-  \noindent\textbf{Case} $\userule{H-SecondAnd}$ \\
-  \indent Given that
+  \case
+    {$\userule{H-SecondAnd}$}
+    {Given that
   $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and
@@
   \refrule{T-Next} we obtain
-  $\Gamma,\Sigma\infers t_1\And t_2':\Task(\tau_1\times\tau_2)$.\\
+  $\Gamma,\Sigma\infers t_1\And t_2':\Task(\tau_1\times\tau_2)$.}
 
-  \noindent\textbf{Case} $\userule{H-FirstOr}$\\
-  \indent  Given that
+  \case
+    {$\userule{H-FirstOr}$}
+    {Given that
   $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$,
@@
   $\Gamma,\Sigma\infers t_1':\Task\tau$ and $\Sigma\infers s'$. By
-  \refrule{T-Or} we now obtain $\Gamma,\Sigma\infers t_1'\Or t_2:\Task\tau$.\\
+  \refrule{T-Or} we now obtain $\Gamma,\Sigma\infers t_1'\Or t_2:\Task\tau$.}
 
-  \noindent\textbf{Case} $\userule{H-SecondOr}$\\
-  \indent  Given that
-  $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$,
-  \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and
-  $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis we know that
-  $t_2,s\handle{i}t_2',s'$ also preserves, and therefore
-  $\Gamma,\Sigma\infers t_2':\Task\tau$ and $\Sigma\infers s'$. By T-Or we now
-  obtain $\Gamma,\Sigma\infers t_1\Or t_2':\Task\tau$.
+  \case
+    {$\userule{H-SecondOr}$}
+    {Given that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
+    By the induction hypothesis we know that $t_2,s\handle{i}t_2',s'$ also preserves, and therefore $\Gamma,\Sigma\infers t_2':\Task\tau$ and $\Sigma\infers s'$.
+    By T-Or we now obtain $\Gamma,\Sigma\infers t_1\Or t_2':\Task\tau$.}
 
@@
 
-% \subsubsection{Theorem~\ref{thmprogressnorm}}
+% \subsection{Theorem~\ref{thmprogressnorm}}
 %
@@
 
-% \subsubsection{Lemma~\ref{lem:stride-does-not-eval}}
+% \subsection{Lemma~\ref{lem:stride-does-not-eval}}
 %
@@
 
-\subsubsection{Theorem~\ref{thm:failing}}
+\subsection{Theorem~\ref{thm:failing}}
 
@@
 
-  \noindent\textbf{Case} $e = \Fail$\\
-  $\Failing(\Fail,s)=\True$, and there is no handling rule that applies to fail.\\
+\case{$e = \Fail$}
+     {$\Failing(\Fail,s)=\True$, and there is no handling rule that applies to fail.}
 
-  \noindent\textbf{Case} $e = \Edit v$\\
-   Given that $\Gamma,\Sigma\infers\Edit v:\Task\tau$, $\Failing(\Edit v,s)=\False$, and there exists an input $i$, namely $v':\tau$.\\
+\case{$e = \Edit v$}
+       {$\Gamma,\Sigma\infers\Edit v:\Task\tau$, $\Failing(\Edit v,s)=\False$, and there exists an input $i$, namely $v':\tau$.}
 
-  \noindent\textbf{Case} $e = \Enter \tau$\\
-  $\Failing(\Enter \tau)=\False$, and there exists an in put i, namely $v:\tau$.\\
+\case{$e = \Enter \tau$}
+       {$\Failing(\Enter \tau)=\False$, and there exists an in put i, namely $v:\tau$.}
 
-  \noindent\textbf{Case} $e = \Update l$\\
-  Given that $\Gamma,\Sigma\infers\Update l:\Task\tau$, $\Failing(\Update l,s)=\False$, and there exists an input $i$, namely $v:\tau$.\\
+\case{$e = \Update l$}
+       {Given that $\Gamma,\Sigma\infers\Update l:\Task\tau$, $\Failing(\Update l,s)=\False$, and there exists an input $i$, namely $v:\tau$.}
 
-  \noindent\textbf{Case} $e = t_1\Then e_2$\\
-   $\Failing(t_1\Then e_2,s)=\Failing(t_1,s)$. If there exitsts an $i$ for $t_1$, then this $i$ also applies to $t_1\Then e_2$. This case therefore holds by the induction hypothesis.\\
+\case{$e = t_1\Then e_2$}
+       {$\Failing(t_1\Then e_2,s)=\Failing(t_1,s)$. If there exists an $i$ for $t_1$, then this $i$ also applies to $t_1\Then e_2$. This case therefore holds by the induction hypothesis.}
 
-  \noindent\textbf{Case} $e = t_1\Next e_2$\\
-   $\Failing(t_1\Next e_2,s)=\Failing(t_1,s)$. If there exitsts an $i$ for $t_1$, then this $i$ also applies to $t_1\Next e_2$. This case therefore holds by the induction hypothesis.\\
+\case{$e = t_1\Next e_2$}
+       {$\Failing(t_1\Next e_2,s)=\Failing(t_1,s)$. If there exists an $i$ for $t_1$, then this $i$ also applies to $t_1\Next e_2$. This case therefore holds by the induction hypothesis.}
 
-  \noindent\textbf{Case} $e = e_1\Xor e_2$\\
-   We normalise the two expressions first, $e_1,s\stride t_1,s'$, $e_2,s\stride t_2,s'$ and we can then be in two situations.
+\case{$e = e_1\Xor e_2$}
+       {We normalise the two expressions first, $e_1,s\stride t_1,s'$, $e_2,s\stride t_2,s'$ and we can then be in two situations.
 One, we can have that $\Failing(t_1,s')$ and $\Failing(t_2,s')$ are both true.
-If that is so, then by definiton, we have both $\Failing(e_1\Xor e_2,s)$ and no rule of the hanlding semantics applies, and therefore there exists no input for this case.\\
+If that is so, then by definition, we have both $\Failing(e_1\Xor e_2,s)$ and no rule of the hanlding semantics applies, and therefore there exists no input for this case.\\
                                            Or we are in the situation where one or both of the two sub expressions does not fail.
 In that case, we know that $\Failing(e_1\Xor e_2,s)$ does not hold, and that at least one of the handling rules applies.
-Therefore, there must be an input $i$, namely $\Left$, $\Right$ or both.\\
+Therefore, there must be an input $i$, namely $\Left$, $\Right$ or both.}
 
-  \noindent\textbf{Case} $e = t_1\And t_2$\\
-   We can again find ourselfs in one of two situations.
+\case{$e = t_1\And t_2$}
+       {We can again find ourselves in one of two situations.
 In the first case, both sub expressions fail, $\Failing(t_1,s)$ and $\Failing(t_2,s)$.
@@
 Again by induction hypothesis, we know that for one or both of the expressions, there exits an $i$ that can be handled.
-Then by H-FirstAnd and H-SecondAnd, we know that we can pass this $i$, by prefixing it with either $\First$ or $\Second$.\\
+Then by H-FirstAnd and H-SecondAnd, we know that we can pass this $i$, by prefixing it with either $\First$ or $\Second$.}
 
-  \noindent\textbf{Case} $e = t_1\Or t_2$\\
-  We can again find ourselfs in one of two situations.
+\case{$e = t_1\Or t_2$}
+       {We can again find ourselves in one of two situations.
 In the first case, both sub expressions fail, $\Failing(t_1,s)$ and $\Failing(t_2,s)$.
@@
 Again by induction hypothesis, we know that for one or both of the expressions, there exits an $i$ that can be handled.
-Then by H-FirstOr and H-SecondOr, we know that we can pass this $i$, by prefixing it with either $\First$ or $\Second$.
+Then by H-FirstOr and H-SecondOr, we know that we can pass this $i$, by prefixing it with either $\First$ or $\Second$.}
 
@@
 
-\subsubsection{Theorem~\ref{thm:safety-i}}
+\subsection{Theorem~\ref{thm:safety-i}}
+
 \begin{proof}
 
-  \noindent\textbf{Case} $e=\Edit v:\Task\tau, i= v':\tau$\\
-   Given that $\userule{H-Change}$, we have by definition that $\Inputs(\Edit v:\Task\tau,s) = \set{v':\tau,\Empty}$, which includes $v':\tau$.
+  \case{$e=\Edit v:\Task\tau, i= v':\tau$}
+       {Given that $\userule{H-Change}$, we have by definition that $\Inputs(\Edit v:\Task\tau,s) = \set{v':\tau,\Empty}$, which includes $v':\tau$.}
 
-  \noindent\textbf{Case} $e=\Edit v:\Task\tau, i= \Empty$\\
-   Given that $\userule{H-Empty}$, we have by definition that $\Inputs(\Edit v:\Task\tau,s) = \set{v':\tau,\Empty}$, which includes $\Empty$.\\
+  \case{$e=\Edit v:\Task\tau, i= \Empty$}
+       {Given that $\userule{H-Empty}$, we have by definition that $\Inputs(\Edit v:\Task\tau,s) = \set{v':\tau,\Empty}$, which includes $\Empty$.}
 
-  \noindent\textbf{Case} $e=\Enter \tau, i= v':\tau$\\
-   Given that $\userule{H-Fill}$, we have by definiton that $\Inputs(\Enter \tau,s) = \set{v':\tau }$, which includes $v':\tau$.\\
+  \case{$e=\Enter \tau, i= v':\tau$}
+       {Given that $\userule{H-Fill}$, we have by definition that $\Inputs(\Enter \tau,s) = \set{v':\tau }$, which includes $v':\tau$.}
 
-  \noindent\textbf{Case} $e=\Update l:\Task\tau, i= v':\tau$\\
-   Given that $\userule{H-Update}$, we have by definiton that $\Inputs(\Update l:\Task\tau,s)= \set{v':\tau }$, which includes $v':\tau$.\\
+  \case{$e=\Update l:\Task\tau, i= v':\tau$}
+       {Given that $\userule{H-Update}$, we have by definition that $\Inputs(\Update l:\Task\tau,s)= \set{v':\tau }$, which includes $v':\tau$.}
 
-  \noindent\textbf{Case} $e=t_1\Xor t_2 , i= \Left$\\
-   Given that $\userule{H-PickLeft}$, we have by definition that $\Inputs(t_1\Xor t_2,s) = \set{ \Left,\Right}$, which includes $\Left$.\\
+  \case{$e=t_1\Xor t_2 , i= \Left$}
+       {Given that $\userule{H-PickLeft}$, we have by definition that $\Inputs(t_1\Xor t_2,s) = \set{ \Left,\Right}$, which includes $\Left$.}
 
-  \noindent\textbf{Case} $e=t_1\Xor t_2 , i= \Right$\\
-   Given that $\userule{H-PickRight}$, we have by definition that $\Inputs(t_1\Xor t_2,s) = \set{ \Left,\Right}$, which includes $\Right$.\\
+  \case{$e=t_1\Xor t_2 , i= \Right$}
+       {Given that $\userule{H-PickRight}$, we have by definition that $\Inputs(t_1\Xor t_2,s) = \set{ \Left,\Right}$, which includes $\Right$.}
 
-  \noindent\textbf{Case} $e=t_1\Next e_2 , i= \Continue $\\
-   Given that $\userule{H-Next}$, we have by definition that $\Inputs(t_1\Next e_2,s) = \Inputs(t_1,s) \cup \set{\Continue\mid \Value{(t_1,s)} = v_1 \wedge \neg\Failing{(e_2 v_1,s\stride)}}$.
-If the H-Next rule applies, this means that the conditions $\Value{(t_1,s)} = v_1 \wedge \neg\Failing{(e_2 v_1,s\stride)}$ are fulfilled, and therefore $\Continue$ is contained.
-\\
+  \case{$e=t_1\Next e_2 , i= \Continue $}
+       {Given that $\userule{H-Next}$, we have by definition that $\Inputs(t_1\Next e_2,s) = \Inputs(t_1,s) \cup \set{\Continue\mid \Value{(t_1,s)} = v_1 \wedge \neg\Failing{(e_2 v_1,s\stride)}}$.
+       If the H-Next rule applies, this means that the conditions $\Value{(t_1,s)} = v_1 \wedge \neg\Failing{(e_2 v_1,s\stride)}$ are fulfilled, and therefore $\Continue$ is contained.}
 
-  \noindent\textbf{Case} $e=t_1\Next e_2 , i\neq\Continue$\\
-   Given that $\userule{H-PassNext}$, we have by definition that $\Inputs(t_1\Next e_2,s) = \Inputs(t_1,s) \cup \set{\Continue\mid \Value{(t_1,s)} = v_1 \wedge \neg\Failing{(e_2 v_1,s\stride)}}$.
-By the induction hypthesis, we have that $i\in \Inputs(t_1,s)$, and by definiton of $\Inputs$, $i$ is therefore also included in this case.\\
+  \case{$e=t_1\Next e_2 , i\neq\Continue$}
+       {Given that $\userule{H-PassNext}$, we have by definition that $\Inputs(t_1\Next e_2,s) = \Inputs(t_1,s) \cup \set{\Continue\mid \Value{(t_1,s)} = v_1 \wedge \neg\Failing{(e_2 v_1,s\stride)}}$.
+       By the induction hypothesis, we have that $i\in \Inputs(t_1,s)$, and by definition of $\Inputs$, $i$ is therefore also included in this case.}
 
-  \noindent\textbf{Case} $e=t_1\Then e_2 , i$\\
-   Given that $\userule{H-PassThen}$, we have by definition that $\Inputs(t_1\Then e_2,s)= \Inputs(t_1,s)$.
-By the induction hypthesis, we have that $i\in \Inputs(t_1,s)$, and by definiton of $\Inputs$, $i$ is therefore also included in this case.\\
+  \case{$e=t_1\Then e_2 , i$}
+       {Given that $\userule{H-PassThen}$, we have by definition that $\Inputs(t_1\Then e_2,s)= \Inputs(t_1,s)$.
+       By the induction hypothesis, we have that $i\in \Inputs(t_1,s)$, and by definition of $\Inputs$, $i$ is therefore also included in this case.}
 
-  \noindent\textbf{Case} $e=t_1\And t_2 , i=\First i$\\
-   Given that $\userule{H-FirstAnd}$ we have by definition that $\Inputs(t_1\And t_2,s) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$.
-By the induction hypothesis, we have that $i\in\Inputs(t_1,s)$, and by definition of $\Inputs$, $\First i$ is therefore als included in this case.\\
+  \case{$e=t_1\And t_2 , i=\First i$}
+       {Given that $\userule{H-FirstAnd}$ we have by definition that $\Inputs(t_1\And t_2,s) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$.
+       By the induction hypothesis, we have that $i\in\Inputs(t_1,s)$, and by definition of $\Inputs$, $\First i$ is therefore als included in this case.}
 
-  \noindent\textbf{Case} $e=t_1\And t_2 , i=\Second i$\\
-   Given that $\userule{H-SecondAnd}$ we have by definition that $\Inputs(t_1\And t_2) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$.
-By the induction hypothesis, we have that $i\in\Inputs(t_2,s)$, and by definition of $\Inputs$, $\Second i$ is therefore als included in this case.\\
+  \case{$e=t_1\And t_2 , i=\Second i$}
+       {Given that $\userule{H-SecondAnd}$ we have by definition that $\Inputs(t_1\And t_2) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$.
+       By the induction hypothesis, we have that $i\in\Inputs(t_2,s)$, and by definition of $\Inputs$, $\Second i$ is therefore als included in this case.}
 
-  \noindent\textbf{Case} $e=t_1\Or t_2 , i=\First i$\\
-   Given that $\userule{H-FirstOr}$ we have by definition that $\Inputs(t_1\Or t_2,s) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$.
-By the induction hypothesis, we have that $i\in\Inputs(t_1,s)$, and by definition of $\Inputs$, $\First i$ is therefore als included in this case.\\
+  \case{$e=t_1\Or t_2 , i=\First i$}
+       {Given that $\userule{H-FirstOr}$ we have by definition that $\Inputs(t_1\Or t_2,s) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$.
+       By the induction hypothesis, we have that $i\in\Inputs(t_1,s)$, and by definition of $\Inputs$, $\First i$ is therefore als included in this case.}
 
-  \noindent\textbf{Case} $e=t_1\Or t_2 , i=\Second i$\\
-   Given that $\userule{H-FirstOr}$ we have by definition that $\Inputs(t_1\Or t_2,s) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$.
-By the induction hypothesis, we have that $i\in\Inputs(t_2,s)$, and by definition of $\Inputs$, $\Second i$ is therefore als included in this case.
+  \case{$e=t_1\Or t_2 , i=\Second i$}
+       {Given that $\userule{H-FirstOr}$ we have by definition that $\Inputs(t_1\Or t_2,s) = \set{\First\ i \mid i \in \Inputs(t_1,s)} \cup \set{\Second\ i \mid i \in \Inputs(t_2,s)}$.
+       By the induction hypothesis, we have that $i\in\Inputs(t_2,s)$, and by definition of $\Inputs$, $\Second i$ is therefore als included in this case.}
   %

### 2019-02-20 sections/outlook.tex
Initial setup outlook section

@@
 
-\fixme{write this section}
+The work presented in this paper describes a formal foundation for task oriented programming.
+For the first time, this programming paradigm has been formalised.
+The formalisation allows us to reason about \TOP and programs written in \TOPHAT.
+So far, we have proven type safety and progress, but this is merely the beginning.
+In future work, we plan on utilising the formalisation to do much more.
+
+First of all, we would like to show that certain properties hold for individual programs.
+We would also like to be able to define what it means for programs to be equal.
+
+A more in depth description of future work can be found in Section~\ref{sec:conclusions}.

### 2019-02-20 sections/abstract.tex
added todo in abstract
tweaks to abstract

@@
 
-Task-Oriented Programming (\TOP) is a programming paradigm that focusses on modelling real world collaborations between people.
+Software that models how people work is omnipresent in todays society.
+Current languages and frameworks are limited in flexibility and lack the proper level of abstraction.
+Task-Oriented Programming (\TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
 It prescribes a declarative programming style to specify multi-user workflows.
@@
 
-In this paper we decompose the rich features of \TOP into elementary language elements,
+In this paper we give a description of the \TOP paradigm and then decompose the rich features of \TOP into elementary language elements,
 which makes them suitable for formal treatment.
@@
 We place \TOPHAT in perspective with the process calculus \CSP.
+
+\fixme{Why?}
+
+% Context: What is the broad context of the work? What is the importance of the general research area?
+% Inquiry: What problem or question does the paper address? How has this problem or question been addressed by others (if at all)?
+% Approach: What was done that unveiled new knowledge?
+% Knowledge: What new facts were uncovered? If the research was not results oriented, what new capabilities are enabled by the work?
+% Grounding: What argument, feasibility proof, artifacts, or results and evaluation support this work?
+% Importance: Why does this work matter?

### 2019-02-20 sections/intuition.tex
intuition collaboration: remove editors
Fix in wording

@@
 The central objective of \TOP is to \emph{coordinate collaboration}.
-The basic building blocks of \TOPHAT for this are task combinators.
+The basic building blocks of \TOPHAT for expressing collaboration are task combinators.
 They express ways in which people can work together.
 Tasks can be executed after each other, at the same time, or conditionally.
-This motivates the combinators step ($\Then$, $\Next$), parallel ($\And$), and choice ($\Or$, $\Xor$).
+This motivates the combinators step, parallel, and choice.
 
@@
 They represent data flow that \emph{follows} control flow.
+\fixme{example with paper deadline}
 

### 2019-02-20 sections/intuition.tex
Changed breakfast example

@@
 
+% \begin{TASK}
+%   let makeBreakfast : Task Drink -> Task Food -> Task Unit =
+%     \makeDrink . \makeFood . makeDrink <&> makeFood >>? \<<d, f>>. eat d f
+%   let main : Task Unit =
+%     makeBreakfast (makeTea <?> makeCoffee) makeEgg
+% \end{TASK}
+
 \begin{TASK}
-  let makeBreakfast : Task Drink -> Task Food -> Task Unit =
-    \makeDrink . \makeFood . makeDrink <&> makeFood >>? \<<d, f>>. eat d f
+  let makeBreakfast : Task Drink -> Task Food -> Task <<Drink,Food>> =
+    \makeDrink . \makeFood . makeDrink <&> makeFood
   let main : Task Unit =
-    makeBreakfast (makeTea <?> makeCoffee) makeEgg
+    (makeBreakfast (makeTea <?> makeCoffee) makeEgg) >>? \<<d, f>>. eat d f
 \end{TASK}
+
+% \begin{TASK}
+%   let makeBreakfast : Task Drink -> Task Food -> Task Breakfast =
+%     \makeDrink . \makeFood . makeDrink <&> makeFood
+%   let main : Task Unit =
+%     (makeBreakfast (makeTea <?> makeCoffee) makeEgg) >>? \breakfast. eat breakfast
+% \end{TASK}
 

### 2019-02-20 sections/outlook.tex
updated outlook

--- 2019-02-19 sections/outlook.tex
@@
 
-\section{Outlook}
+\subsection{Outlook}
 
-\fixme{write this section}
+At this point we have specified a formal language for task-oriented programming,
+given its semantics, and proved its safety.
+The main motive to formalise this paradigm, is to be able to reason about tasks.
+In future work, we plan on utilising the formalisation to do so.
+
+Firstly, we would like to express properties of tasks and prove them.
+For example, one would like to prove that, no matter what, in Example~\ref{exm:abstract} a paper is always submitted.
+Secondly, we would like to point out what it means for two tasks to be equal.
+One could have noticed that some operators have a monadic or applicative feeling.
+The combination of $\And$ and $\Edit$ could form a (lax) monoidal functor,
+$\Or$ is similar to applicative choice,
+and $\Then$ looks like a bind operation.
+We need a correct understand of equivalency of tasks,
+takin the interactive setting into account,
+to prove this.
+Thirdly, we do not know yet if the more complex combinators of \ITASKS are expressible in the basic combinators of \TOPHAT.
+We implemented \TOPHAT on top of \ITASKS, so we know it is a subset,
+but we also know \ITASKS can do more.
+
+% The work presented in this paper describes a formal foundation for task oriented programming.
+% For the first time, this programming paradigm has been formalised.
+% The formalisation allows us to reason about \TOP and programs written in \TOPHAT.
+% So far, we have proven type safety and progress, but this is merely the beginning.
+% In future work, we plan on utilising the formalisation to do much more.
+
+% First of all, we would like to show that certain properties hold for individual programs.
+% We would also like to be able to define what it means for programs to be equal.
+
+A more in depth description of future work can be found in Section~\ref{sec:conclusions}.

### 2019-02-20 sections/intuition.tex
remove parens in breakfast example
make parameters italic
use `haveBreakfast`
add stub for abstract example

@@
 
+\lstset{emph={makeDrink,makeFood}}
+
 % \begin{TASK}
 %   let makeBreakfast : Task Drink -> Task Food -> Task Unit =
-%     \makeDrink . \makeFood . makeDrink <&> makeFood >>? \<<d, f>>. eat d f
+%     \makeDrink. \makeFood. makeDrink <&> makeFood >>? \<<d, f>>. eat d f
 %   let main : Task Unit =
@@
   let makeBreakfast : Task Drink -> Task Food -> Task <<Drink,Food>> =
-    \makeDrink . \makeFood . makeDrink <&> makeFood
+    \makeDrink. \makeFood. makeDrink <&> makeFood
   let main : Task Unit =
-    (makeBreakfast (makeTea <?> makeCoffee) makeEgg) >>? \<<d, f>>. eat d f
+    makeBreakfast (makeTea <?> makeCoffee) makeEgg >>? haveBreakfast
 \end{TASK}
@@
 %   let makeBreakfast : Task Drink -> Task Food -> Task Breakfast =
-%     \makeDrink . \makeFood . makeDrink <&> makeFood
+%     \makeDrink. \makeFood. makeDrink <&> makeFood
 %   let main : Task Unit =
-%     (makeBreakfast (makeTea <?> makeCoffee) makeEgg) >>? \breakfast. eat breakfast
+%     makeBreakfast (makeTea <?> makeCoffee) makeEgg >>? \breakfast. eat breakfast
 % \end{TASK}
@@
 They represent data flow that \emph{follows} control flow.
-\fixme{example with paper deadline}
 
@@
 
+\lstset{emph={text}}
+\begin{example}[Abstract submission]
+\label{exm:abstract}
+
+The task of writing an abstract for a conference is never done.
+\fixme{Extend}
+
+% \begin{TASK}
+%   let submit : String -> Task Unit =
+%     \text. delay 30 (sendTo "icfp19@sigplan.org" text)
+%   let write : Task Unit =
+%     edit "Abstract: ..." >>= \abstract. if 500 <= words abstract <= 700 then submit abstract else fail
+% \end{TASK}
+
+\begin{TASK}
+  let submit : Ref String -> Task String = \text. sendTo "icfp19@sigplan.org" text
+  let write : Ref String -> Task String = \text. update text
+  let main : Task String = write abstract <|> delay 30 (submit abstract)
+\end{TASK}
+
+The \TS{delay} task is specified in Example~\ref{exm:delay} in Section~\ref{sec:language}.
+
+\end{example}
+
 \subsection{Recap}

### 2019-02-21 sections/outlook.tex
add stub for abstract example
two minor changes in outlook

@@
 Firstly, we would like to express properties of tasks and prove them.
-For example, one would like to prove that, no matter what, in Example~\ref{exm:abstract} a paper is always submitted.
-Secondly, we would like to point out what it means for two tasks to be equal.
+For example, one would like to prove that, no matter what, in Example~\ref{exm:abstract} an abstract is always submitted.
+Secondly, we would like explore what it means for two tasks to be equal.
 One could have noticed that some operators have a monadic or applicative feeling.
@@
 We need a correct understand of equivalency of tasks,
-takin the interactive setting into account,
+taking the interactive setting into account,
 to prove this.

### 2019-02-21 sections/relatedwork.tex
fix heading from wofoo to surveymonkey
add hiphop to related work

@@
 
-\paragraph{Esterel \& HipHop}
+\paragraph{HipHop}
 
-\fixme{Write}
+HipHop \fixme{cite} is a programming language to develop synchronous reactive web systems.
+From a single source, both server and client applications can be generated.
+Programs are written in the Hop language, a Scheme dialect.
+Communication is based on a reactive layer embedded in Hop.
+Each reactive component starts by specifying possible input and output events.
+The component then proceeds as a state machine.
+
+Input events are sent to such a machine programatically using Hop, or are explicitly wired to events from the client.
+They are optionally associated with a Hop value.
+As Hop is a dynamic language, and HipHop uses strings to identify events, events and their possible associated values are not statically checked.
+Events are aggregated till the moment the machine as asked to react.
+% then are perceived as conceptually simultaneous.
+The machine is executed and reacts by building a multi-set of output events.
+The execution of a HipHop machine is atomic.
+The set of inputs is not influenced by the current computation.
+
+As with \TOPHAT, HipHop is a \DSL embedded in a general purpose programming language.
+% the same way \TOPHAT is embedded into the simply typed $\lambda$-calculus.
+Another similarity is that both specifications lead to executable server and client applications from a single source.
+However, HipHop is more low level regarding its specification.
+Where \TOPHAT takes tasks and collaboration as a starting point,
+HipHop focusses on synchronous communication and atom execution of reactive machines.
+
+This comes forward in the way both systems define events.
+In HipHop programmers can define and use their own events.
+Inputs in \TOPHAT are not extensible and not visible to the developer.
+They are a completely separate entity living on the semantic level.
+
+Another important difference is the way both systems handle events.
+In HipHop the programmer has control over the moment a machine should processes its events.
+This could be just one event, or a multi-set of events that are processed simultaneously.
+\TOPHAT always processes an input at the moment it occurs and only handles a single event in one instance.
+
+\paragraph{Esterel}
+
+The set of HipHop reactive statements is based on those of the Esterel language.
+Esterel is a programming language
 
@@
 
-\paragraph{Google forms \& Wufoo}
+\paragraph{Google forms \& SurveyMonkey}
 

### 2019-02-21 sections/relatedwork.tex
add paragraph in process algebras

@@
 
+\paragraph{Differences}
 There are two main differences between \TOP and process algebras.
@@
 Synchronisation means that only some, but not all, of the possible interleavings are desirable.
-The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise TODO: say that synchronisation in TOPHAT is the best.
+The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise \fixme{Say that synchronisation in TOPHAT is the best.}
 

### 2019-02-21 appendices/proofs.tex
fixed typo and some formatting stuff

@@
 
-  \case
-    {$\userule{H-PassNext}$}
-    {Given that
-  $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$ and $\Sigma\infers s$,
-  \refrule{T-Next} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
-  $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. By the induction hypothesis, we
-  know that $t_1,s\handle{i}t_1',s'$ also preserves and thus
-  $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Gamma,Sigma\infers s'$. By
-  \refrule{T-Next} we now obtain that
-  $\Gamma,\Sigma\infers t_1'\Next e_2:\Task\tau$. }
+\case{$\userule{H-PassNext}$}
+     {Given that $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Next} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$.
+     By the induction hypothesis, we know that $t_1,s\handle{i}t_1',s'$ also preserves and thus $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Gamma,Sigma\infers s'$.
+     By \refrule{T-Next} we now obtain that $\Gamma,\Sigma\infers t_1'\Next e_2:\Task\tau$. }
 
-  \case
-    {$\userule{H-FirstAnd}$}
-    {Given that
-  $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and
-  $\Sigma\infers s$, \refrule{T-And} gives us that
-  $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
-  $\Gamma,\Sigma\infers t_2:\Task\tau_2$. By the induction hypothesis, we know
-  that $t_1,s\handle{i}t_1',s'$ also preserves and thus
-  $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Sigma\infers s'$. Therefore by
-  \refrule{T-Next} we obtain
-  $\Gamma,\Sigma\infers t_1'\And t_2:\Task(\tau_1\times\tau_2)$.}
+\case{$\userule{H-FirstAnd}$}
+     {Given that $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and $\Sigma\infers s$, \refrule{T-And} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers t_2:\Task\tau_2$.
+     By the induction hypothesis, we know that $t_1,s\handle{i}t_1',s'$ also preserves and thus $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Sigma\infers s'$.
+     Therefore by \refrule{T-Next} we obtain $\Gamma,\Sigma\infers t_1'\And t_2:\Task(\tau_1\times\tau_2)$.}
 
-  \case
-    {$\userule{H-SecondAnd}$}
-    {Given that
-  $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and
-  $\Sigma\infers s$, \refrule{T-And} gives us that
-  $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and
-  $\Gamma,\Sigma\infers t_2:\Task\tau_2$. By the induction hypothesis, we know
-  that $t_2,s\handle{i}t_2',s'$ also preserves and thus
-  $\Gamma,\Sigma\infers t_2':\Task\tau_2$ and $\Sigma\infers s'$. Therefore by
-  \refrule{T-Next} we obtain
-  $\Gamma,\Sigma\infers t_1\And t_2':\Task(\tau_1\times\tau_2)$.}
+\case{$\userule{H-SecondAnd}$}
+     {Given that $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and $\Sigma\infers s$, \refrule{T-And} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers t_2:\Task\tau_2$.
+     By the induction hypothesis, we know that $t_2,s\handle{i}t_2',s'$ also preserves and thus $\Gamma,\Sigma\infers t_2':\Task\tau_2$ and $\Sigma\infers s'$.
+     Therefore by \refrule{T-Next} we obtain $\Gamma,\Sigma\infers t_1\And t_2':\Task(\tau_1\times\tau_2)$.}
 
-  \case
-    {$\userule{H-FirstOr}$}
-    {Given that
-  $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$,
-  \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and
-  $\Gamma,\Sigma\infers t_2:\Task\tau$. By the induction hypothesis we know that
-  $t_1,s\handle{i}t_1',s'$ also preserves, and therefore
-  $\Gamma,\Sigma\infers t_1':\Task\tau$ and $\Sigma\infers s'$. By
-  \refrule{T-Or} we now obtain $\Gamma,\Sigma\infers t_1'\Or t_2:\Task\tau$.}
+\case{$\userule{H-FirstOr}$}
+     {Given that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
+     By the induction hypothesis we know that $t_1,s\handle{i}t_1',s'$ also preserves, and therefore $\Gamma,\Sigma\infers t_1':\Task\tau$ and $\Sigma\infers s'$.
+     By \refrule{T-Or} we now obtain $\Gamma,\Sigma\infers t_1'\Or t_2:\Task\tau$.}
 
-  \case
-    {$\userule{H-SecondOr}$}
+\case{$\userule{H-SecondOr}$}
      {Given that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
@@
 \case{$e = t_1\Then e_2$}
-       {$\Failing(t_1\Then e_2,s)=\Failing(t_1,s)$. If there exists an $i$ for $t_1$, then this $i$ also applies to $t_1\Then e_2$. This case therefore holds by the induction hypothesis.}
+     {$\Failing(t_1\Then e_2,s)=\Failing(t_1,s)$.
+     If there exists an $i$ for $t_1$, then this $i$ also applies to $t_1\Then e_2$.
+     This case therefore holds by the induction hypothesis.}
 
 \case{$e = t_1\Next e_2$}
-       {$\Failing(t_1\Next e_2,s)=\Failing(t_1,s)$. If there exists an $i$ for $t_1$, then this $i$ also applies to $t_1\Next e_2$. This case therefore holds by the induction hypothesis.}
+     {$\Failing(t_1\Next e_2,s)=\Failing(t_1,s)$.
+     If there exists an $i$ for $t_1$, then this $i$ also applies to $t_1\Next e_2$.
+     This case therefore holds by the induction hypothesis.}
 
@@
      One, we can have that $\Failing(t_1,s')$ and $\Failing(t_2,s')$ are both true.
-If that is so, then by definition, we have both $\Failing(e_1\Xor e_2,s)$ and no rule of the hanlding semantics applies, and therefore there exists no input for this case.\\
+     If that is so, then by definition, we have both $\Failing(e_1\Xor e_2,s)$ and no rule of the handling semantics applies, and therefore there exists no input for this case.
+
      Or we are in the situation where one or both of the two sub expressions does not fail.
@@
      By the induction hypothesis, we know that for both $t_1$ and $t_2$ there is no input that can be handled.
-Since the only two rules for $\And$ that handle input just pass this input on to one of the two expressions, we know that indeed no $i$ applies.\\
+     Since the only two rules for $\And$ that handle input just pass this input on to one of the two expressions, we know that indeed no $i$ applies.
+
      In the case that one or both sub expressions do not fail, then by definition $t_1\And t_2$ not failing under $s$.
@@
      By the induction hypothesis, we know that for both $t_1$ and $t_2$ there is no input that can be handled.
-Since the only two rules for $\Or$ that handle input just pass this input on to one of the two expressions, we know that indeed no $i$ applies.\\
+     Since the only two rules for $\Or$ that handle input just pass this input on to one of the two expressions, we know that indeed no $i$ applies.
+
      In the case that one or both sub expressions do not fail, then by definition $t_1\Or t_2$ not failing under $s$.

### 2019-02-21 sections/relatedwork.tex
Minor fixes to HipHop

@@
 
-HipHop \fixme{cite} is a programming language to develop synchronous reactive web systems.
+HipHop \fixme{cite} is a programming language tailored to the development synchronous reactive web systems.
 From a single source, both server and client applications can be generated.
@@
 As Hop is a dynamic language, and HipHop uses strings to identify events, events and their possible associated values are not statically checked.
-Events are aggregated till the moment the machine as asked to react.
+Events are aggregated until the moment the machine as asked to react.
 % then are perceived as conceptually simultaneous.
@@
 The execution of a HipHop machine is atomic.
-The set of inputs is not influenced by the current computation.
+The set of inputs is not influenced by the current computations.
 
@@
 
-This comes forward in the way both systems define events.
+This difference in focus shows in the way both systems define events.
 In HipHop programmers can define and use their own events.
@@
 
-Another important difference is the way both systems handle events.
-In HipHop the programmer has control over the moment a machine should processes its events.
+Another important difference is the way in which both systems handle events.
+In HipHop the programmer decides when a machine should processes its events.
 This could be just one event, or a multi-set of events that are processed simultaneously.
-\TOPHAT always processes an input at the moment it occurs and only handles a single event in one instance.
+\TOPHAT always processes an input the moment it occurs and only handles a single event in one instance.
 

### 2019-02-21 sections/relatedwork.tex
add citations on HipHop and Esterel

@@
 
-\paragraph{HipHop}
+\paragraph{HipHop \& Esterel}
 
-HipHop \fixme{cite} is a programming language tailored to the development synchronous reactive web systems.
+HipHop \cite{berry2011hiphop} is a programming language tailored to the development synchronous reactive web systems.
 From a single source, both server and client applications can be generated.
@@
 Communication is based on a reactive layer embedded in Hop.
+The set of HipHop reactive statements is based on those of the Esterel language \cite{boussinot1991esterel}.
 Each reactive component starts by specifying possible input and output events.
@@
 Another similarity is that both specifications lead to executable server and client applications from a single source.
-However, HipHop is more low level regarding its specification.
+However, both HipHop and Esterel are more low level regarding their specification.
 Where \TOPHAT takes tasks and collaboration as a starting point,
@@
 \TOPHAT always processes an input the moment it occurs and only handles a single event in one instance.
-
-\paragraph{Esterel}
-
-The set of HipHop reactive statements is based on those of the Esterel language.
-Esterel is a programming language
 

### 2019-02-21 appendices/proofs.tex
remove superfluous newlines in proofs

@@
 \begin{proof}
-  We prove Theorem~\ref{thm:pres-eval} by induction on $e$:\\
+  We prove Theorem~\ref{thm:pres-eval} by induction on $e$:
 
@@
 \begin{proof}
-  We prove Lemma~\ref{lem:presvalue} by induction over $e$.\\
+  We prove Lemma~\ref{lem:presvalue} by induction over $e$.
 
@@
   % \noindent\textbf{Case} $\Value{(u\At t,s)}=\Value(t,s)$ This case follows
-  % directly from the induction hypothesis.\\
+  % directly from the induction hypothesis.
 \end{proof}
@@
 \begin{proof}
-  We prove Theorem~\ref{thm:pres-stride} by induction on $e$:\\
+  We prove Theorem~\ref{thm:pres-stride} by induction on $e$:
 
@@
   % $t:\Task\tau$. By the induction hypothesis, we know that $t,s\stride t',s'$
-  % preserves, and thus $\Gamma,\Sigma\infers u\At t':\Task\tau$.\\
+  % preserves, and thus $\Gamma,\Sigma\infers u\At t':\Task\tau$.
 
@@
 \begin{proof}
-  We prove Theorem~\ref{thm:pres-norm} by induction on $e$:\\
+  We prove Theorem~\ref{thm:pres-norm} by induction on $e$:
 
-  \noindent\textbf{Case} $\userule{N-Done}$ \\
-  \indent Given that
+  \case{$\userule{N-Done}$}
+    {Given that
     $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that
@@
     Theorem~\ref{thm:pres-eval}. Then by Theorem~\ref{thm:pres-stride}, we have
-  $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$.\\
+    $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$.}
 
-  \noindent\textbf{Case}\\
-  $\userule{N-Repeat}$ \\
-  \indent Given that
+  \case
+    {$\userule{N-Repeat}$}
+    { Given that
     $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that
@@
     induction hypothesis, we finally obtain that
-  $\Gamma,\Sigma\infers t'':\Task\tau$ and $\Sigma\infers s'''$.\\
+    $\Gamma,\Sigma\infers t'':\Task\tau$ and $\Sigma\infers s'''$.}
 
@@
 
-  We prove Theorem~\ref{thm:failing} by induction on $e'$.\\
+  We prove Theorem~\ref{thm:failing} by induction on $e'$.
 
@@
 
-  % \noindent\textbf{Case} $e = u \At t$\\
-  %  This case follows directly from applying the induction hypothesis, since $\Failing(u \At t,s)=\Failing(t,s)$ and any input that applies to $t$ also applies to $u \At t$.\\
+  % \noindent\textbf{Case} $e = u \At t$
+  %  This case follows directly from applying the induction hypothesis, since $\Failing(u \At t,s)=\Failing(t,s)$ and any input that applies to $t$ also applies to $u \At t$.
 \end{proof}
@@
   %
-  % \noindent\textbf{Case} $e=u\At t, i$\\
-  %  Given that $\userule{H-Appoint}$, we can directly apply the induction hypothesis to prove this case.\\
+  % \noindent\textbf{Case} $e=u\At t, i$
+  %  Given that $\userule{H-Appoint}$, we can directly apply the induction hypothesis to prove this case.
 \end{proof}

### 2019-02-21 sections/introduction.tex
small bit on mtasks

@@
 It has been used to model an incident management tool for the Dutch coast guard~\cite{conf/iscram/LijnseJP12}.
-Also it has been used numerous times to prototype ideas for Command and Control~\cite{theses/nlda/Kool17, theses/radboud/Stutterheim17}, and in a case study for the Dutch tax authority \cite{conf/sfp/StutterheimAP17}.
+Also it has been used numerous times to prototype ideas for Command and Control~\cite{theses/nlda/Kool17, theses/radboud/Stutterheim17}, and in a case study for the Dutch tax authority~\cite{conf/sfp/StutterheimAP17}.
 
@@
 focussing on \IOT devices and deployment on micro controllers.
-\fixme{Add examples.}
+It has been used to control home thermostats and other home automation applications \fixme{cite}.
 
-\fixme{Say what lacks in these approaches: no formal reasoning about properties.}
+Both implementations currently lack formal semantics which can be used to prove properties about tasks.
 
@@
     Along with the semantics, we present the following semantic observations on tasks:
-    the current value, whether a term is stuck, the current user interface, and the accepted inputs. \fixme{we do not provide observation for user interface!}
+    the current value, whether a term is stuck, the current user interface, and the accepted inputs.
 

### 2019-02-21 sections/intuition.tex
fix breakfast, remove abstract examples

--- 2019-02-20 sections/intuition.tex
@@
   let makeBreakfast : Task Drink -> Task Food -> Task <<Drink,Food>> =
-    \makeDrink . \makeFood . makeDrink <&> makeFood
-  let main : Task Unit =
-    (makeBreakfast (makeTea <?> makeCoffee) makeEgg) >>? \<<d, f>>. eat d f
+    \makeDrink. \makeFood. makeDrink <&> makeFood in
+  makeBreakfast (makeTea <?> makeCoffee) makeEgg >>? haveBreakfast
 \end{TASK}
@@
 
-\Cref{exm:breakfast} demonstrates how tasks can be parameterized by other tasks: makeBreakfast is a collaboration pattern that always works the same way, regardless of which food and drink are being prepared.
+\Cref{exm:breakfast} demonstrates how tasks can be parameterised by other tasks: \TS{makeBreakfast} is a collaboration pattern that always works the same way, regardless of which food and drink are being prepared.
 
@@
 
+% \lstset{emph={text}}
+% \begin{example}[Abstract submission]
+% \label{exm:abstract}
+%
+% The task of writing an abstract for a conference is never done.
+% Let \TS{abstract} be a reference to a string.
+%
+% % \begin{TASK}
+% %   let submit : String -> Task Unit =
+% %     \text. delay 30 (sendTo "icfp19@sigplan.org" text)
+% %   let write : Task Unit =
+% %     edit "Abstract: ..." >>= \abstract. if 500 <= words abstract <= 700 then submit abstract else fail
+% % \end{TASK}
+%
+% \begin{TASK}
+%   let submit : Ref String -> Task String = \text. sendTo "icfp19@sigplan.org" text in
+%   update abstract <|> delay 30 (submit abstract)
+% \end{TASK}
+%
+% The \TS{delay} task is specified in Example~\ref{exm:delay} in Section~\ref{sec:language}.
+%
+% \end{example}
+
 \subsection{Recap}

### 2019-02-21 sections/semantics.tex
small fixes and more fixmes
shuffle around pars for readability

@@
 
+\fixme{Add $\UserInterface$ to appendix and refer.}
 \TOPHAT is designed such that a user interface can be generated from a task's syntax tree.
@@
 Inputs $i$ are formed according to the grammar in \cref{fig:input-grammar}.
-Handling input is done by the \emph{handling} semantics shown in \cref{fig:handling-semantics}.
-It is a small step semantics with labelled transitions.
-It takes a task $t$ in a state $s$ and an input $i$, and yields a new task $t'$ in a new state $s'$.
+%
+There is a function $\Inputs$ which calculates the possible input events a given task expects.
+It takes a normalised task and a state and returns a set of inputs that can be handled.
+The definition of this function is listed in \cref{fig:observation-input}.
 
@@
 \end{figure}
+
+\begin{figure}[h]
+  \small
+  \usemacro{O-Inputs}
+  \caption{Inputs} \label{fig:observation-input}
+\end{figure}
+
+Handling input is done by the \emph{handling} semantics shown in \cref{fig:handling-semantics}.
+It is a small step semantics with labelled transitions.
+It takes a task $t$ in a state $s$ and an input $i$, and yields a new task $t'$ in a new state $s'$.
 
@@
 
+%FIXME: Revert rule names in front to normal sentences?
 \refrule{H-Change},
@@
 
-There is a function $\Inputs$ which calculates the possible input events a given task expects.
-It takes a normalised task and a state and returns a set of inputs that can be handled.
-The definition of this function is listed in \cref{fig:observation-input}.
-
-\begin{figure}[h]
-  \small
-  \usemacro{O-Inputs}
-  \caption{Inputs} \label{fig:observation-input}
-\end{figure}
-
 \subsection{Implementation}
@@
 We use techniques presented by \textcite{journals/entcs/JaskelioffGH11}, \textcite{journals/jfp/Swierstra08}, and \textcite{school/maktoberdorf/PeytonJones01}.
-Such an implementation increases the confidence in the correctness and completeness of the semantics.
+% Such an implementation increases the confidence in the correctness and completeness of the semantics.
 The source code can be found on GitHub.\footnote{\url{http://github.com/XXX/XXX}}
-A proof of concept implementation in Haskell is also available, showing that \TOPHAT can be implemented without dependent types.
-
+% A proof of concept implementation in Haskell is also available, showing that \TOPHAT can be implemented without dependent types.
 A command-line interface is part of this implementation.
 It prompts users to type input events, which get parsed and processed by the interaction semantics.
+
+Also, we mad an implementation of \TOPHAT combinators into \ITASKS,
+so that \TOPHAT specifications can be compiled to runnable applications.
+This shows that \TOPHAT is a subset of \ITASKS.

### 2019-02-21 sections/relatedwork.tex
more references

@@
 
-HipHop \cite{berry2011hiphop} is a programming language tailored to the development synchronous reactive web systems.
+HipHop \cite{berry2011hiphop,journals/corr/BerryS13} is a programming language tailored to the development synchronous reactive web systems.
 From a single source, both server and client applications can be generated.
@@
 Communication is based on a reactive layer embedded in Hop.
-The set of HipHop reactive statements is based on those of the Esterel language \cite{boussinot1991esterel}.
+The set of HipHop reactive statements is based on those of the Esterel language \cite{boussinot1991esterel,journals/scp/BerryG92}.
 Each reactive component starts by specifying possible input and output events.

### 2019-02-21 sections/relatedwork.tex
add something about mtasks to related work

@@
 
-As mentioned earlier, iTasks is an implementation of \TOP. iTasks has many
-features, and its basic combinators are versatile and powerful. Simpler
-combinators are implemented by restricting the powerful ones. This is useful for
-everyday programming, where having lots of functionality at one's fingertips is
-convenient and efficient.
+As mentioned earlier, iTasks is an implementation of \TOP.
+iTasks has many features, and its basic combinators are versatile and powerful.
+Simpler combinators are implemented by restricting the powerful ones.
+This is useful for everyday programming, where having lots of functionality at one's fingertips is convenient and efficient.
 
-There have been two previous papers that describe semantics of iTasks, by
-\citet{conf/ifl/KoopmanPA08} and \citet{conf/ppdp/PlasmeijerLMAK12}.
-Both give a different semantics in the form of minimal implementations of a
-subset of the interface of iTasks. These semantics however do not make an
-explicit distinction between the host language and task language and they do not
-provide a formal semantics, which makes formal reasoning difficult.
-
-\paragraph{mTasks}
-\fixme{Hier iets gezelligs over mTasks}
+There have been two previous papers that describe semantics of iTasks, by \citet{conf/ifl/KoopmanPA08} and \citet{conf/ppdp/PlasmeijerLMAK12}.
+Both give a different semantics in the form of minimal implementations of a subset of the interface of iTasks.
+These semantics however do not make an explicit distinction between the host language and task language and they do not provide a formal semantics, which makes formal reasoning difficult.
 
@@
 
-% mTasks is another implementation of TOP, designed for coordination of IoT devices. \alert{cite MSc Lubbers}
-% mTasks has been used in a demonstrator for home automation. \alert{cite MSc Andrade}
+\paragraph{mTasks}
+
+The \MTASKS framework is an implementation of \TOP geared towards \IOT devices.
+As \TOPHAT its basic combinators are a subset of \ITASKS.
+They are similar to those of \TOPHAT.
+However, on \IOT devices it is useful to continue running tasks endlessly,
+which is done in \MTASKS using a \CL{forever} combinator.
+This is currently not possible in \TOPHAT.
+
+As for \ITASKS, there is currently no formal semantics for \MTASKS.
 

### 2019-02-21 sections/abstract.tex
Removed fixme
add todo in abstract

@@
 Current languages and frameworks are limited in flexibility and lack the proper level of abstraction.
-Task-Oriented Programming (\TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
+Task-oriented programming (\TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
 It prescribes a declarative programming style to specify multi-user workflows.
@@
 
-\fixme{Why?}
+\fixme{Add Knowledge, Grounding and Importance}
 

### 2019-02-22 sections/introduction.tex
moved first part of 1.1 to intro

@@
 \section{Introduction}
-
-\fixme{Common intro??}
-
-\subsection{Tasks}
 
@@
 and they use information from multiple sources (input forms, databases, sensors, etc.).
+
+\subsection{Tasks}
 
@@
 Informally in the sense that we give an intuitive description of the features that define task-oriented programming.
-Formally in the sense that we develop a calculus which formalizes these features as language constructs, and we give them semantics in the style that is common in programming language research.
+Formally in the sense that we develop a calculus which formalises these features as language constructs, and we give them semantics in the style that is common in programming language research.
 

### 2019-02-22 sections/relatedwork.tex
Added related work common intro
Added mtasks cite to related work

@@
 
-\fixme{Intro related work}
+The work presented in this paper lies on the boundary of many areas of study.
+People have looked at the problem of how to model and coordinate collaboration from many different perspectives.
+The following subsections give an overview of related work from the many different areas.
 
@@
 
-The \MTASKS framework is an implementation of \TOP geared towards \IOT devices.
+The \MTASKS framework \cite{koopman2018task} is an implementation of \TOP geared towards \IOT devices.
 As \TOPHAT its basic combinators are a subset of \ITASKS.

### 2019-02-22 sections/intuition.tex
add todo
extended intuition of breakfast example

@@
     \makeDrink. \makeFood. makeDrink <&> makeFood in
-  makeBreakfast (makeTea <?> makeCoffee) makeEgg >>? haveBreakfast
+  makeBreakfast (makeTea <?> makeCoffee) makeEgg >>? enjoyBreakfast
 \end{TASK}
@@
 
+The way the combinators are defined matches real life closely.
+When we want to have breakfast, we have to complete several other tasks first before we can do so.
+We decide what we want to have and then prepare it.
+We can prepare the different items we have for breakfast in parallel, but not at the same time.
+For example, it is impossible to scramble eggs, and put on the kettle for tea simultaneously.
+Instead, what is meant by parallel is that the order in which we do tasks and the smaller tasks that they are composed of, does not matter.
+Then finally, only when every item we want to have for breakfast is ready, can we sit down and enjoy it.
 \end{example}

### 2019-02-22 sections/introduction.tex
Added cite to mTasks temperature paper
Hide fixme from intro

@@
 Finally, there is communication with the \emph{outside} world, where information is entered into the system via input events.
-\fixme{Should this be mentioned earlier?}
+%\fixme{Should this be mentioned earlier?}
 The end points where the outside world interacts with \TOP applications are called editors.
@@
 focussing on \IOT devices and deployment on micro controllers.
-It has been used to control home thermostats and other home automation applications \fixme{cite}.
+It has been used to control home thermostats and other home automation applications \cite{koopman2018task}.
 

### 2019-02-22 sections/relatedwork.tex
fixed fix me from related work

@@
 Synchronisation means that only some, but not all, of the possible interleavings are desirable.
-The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise \fixme{Say that synchronisation in TOPHAT is the best.}
+The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise and intuitive syntonisation code.
 

### 2019-02-22 sections/abstract.tex
make abstract pure ascii

--- 2019-02-20 sections/abstract.tex
@@
 
-Software that models how people work is omnipresent in todays society.
-Current languages and frameworks are limited in flexibility and lack the proper level of abstraction.
-Task-Oriented Programming (\TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
+Software that models how people work is omnipresent in today's society.
+Current languages and frameworks are limited in flexibility and lack a good level of abstraction.
+Task-oriented programming (TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
 It prescribes a declarative programming style to specify multi-user workflows.
@@
 
-In this paper we give a description of the \TOP paradigm and then decompose the rich features of \TOP into elementary language elements,
+In this paper we give a description of the TOP paradigm and then decompose the rich features of TOP into elementary language elements,
 which makes them suitable for formal treatment.
 We use the simply typed lambda calculus, extended with pairs and references, as a base language.
-On top of this language, we develop \TOPHAT (TopHat), a calculus for modular interactive workflows.
-With \TOPHAT we prepare a way to formally reason about \TOP systems.
+On top of this language, we develop TopHat, a calculus for modular interactive workflows.
+With TopHat we prepare a way to formally reason about TOP systems.
 
-We describe \TOPHAT by means of a layered semantics.
+We describe TopHat by means of a layered semantics.
 These layers consist of multiple big step evaluations on expressions,
@@
 This approach allows for comparison with other work in the field.
-We place \TOPHAT in perspective with the process calculus \CSP.
-
-\fixme{Why?}
+We place TopHat in perspective with the process calculus CSP.
 

### 2019-02-22 sections/abstract.tex
structure abstract following CIAKGI
Context Inquiry Approach Knowledge Grounding Importance

@@
 
+% Context
 Software that models how people work is omnipresent in today's society.
@@
 Workflows can be higher-order.
-They communicate through typed values on a local or global level.
-Such specifications can be turned into interactive applications for different platforms,
-supporting collaboration during execution.
+They communicate through typed values on a local and global level.
+Such specifications can be turned into interactive applications for different platforms, supporting collaboration during execution.
 
-In this paper we give a description of the TOP paradigm and then decompose the rich features of TOP into elementary language elements,
-which makes them suitable for formal treatment.
+% Inquiry
+TOP has been around for more than a decade, in the form of iTasks, which is tailored for real-world usability.
+So far, it has not been given a formalization in the style common in programming language research.
+
+% Approach
+In this paper we give a description of the TOP paradigm and then decompose its rich features into elementary language elements, which makes them suitable for formal treatment.
 We use the simply typed lambda calculus, extended with pairs and references, as a base language.
 On top of this language, we develop TopHat, a calculus for modular interactive workflows.
-With TopHat we prepare a way to formally reason about TOP systems.
+We describe TopHat by means of a layered semantics.
+These layers consist of multiple big step evaluations on expressions, and two labelled transition systems, handling user inputs.
 
-We describe TopHat by means of a layered semantics.
-These layers consist of multiple big step evaluations on expressions,
-and two labelled transition systems, handling user inputs.
-We show some interesting properties of this machinery.
+% Knowledge
+With TopHat we prepare a way to formally reason about TOP languages and programs.
+We show some interesting properties of the language.
 This approach allows for comparison with other work in the field.
-We place TopHat in perspective with the process calculus CSP.
+
+% Grounding
+%  ??? De taal zelf is eigenlijk het artifact
+
+% Importance
+TOP has been employed in projects with the Dutch coast guard, tax office, and navy.
+Our work matters because formal program verification is important for mission-critical software, especially for systems with concurrency.
 

### 2019-02-25 sections/abstract.tex
abstract: Grounding
examples -> case studies
All changes based on johan's comments

@@
 % Grounding
-%  ??? De taal zelf is eigenlijk het artifact
+We have implemented the semantic rules of TopHat in Idris, and the task layer on top of iTasks.
+This shows that our approach is feasible, and lets us demonstrate the concepts by means of illustrative case studies.
 
@@
 TOP has been employed in projects with the Dutch coast guard, tax office, and navy.
-Our work matters because formal program verification is important for mission-critical software, especially for systems with concurrency.
+Our work matters because Formal program verification is important for mission-critical software, especially for systems with concurrency.
 

### 2019-02-25 sections/introduction.tex
All changes based on johan's comments

@@
   \item
-    Using both the essential concepts and the formal calculus, we compare \TOP with multiple related works, in areas ranging from business process modelling to process algebras and reactive programming.
+    Using both the essential concepts and the formal calculus, we compare \TOP with related work, in areas ranging from business process modelling to process algebras and reactive programming.
 

### 2019-02-25 sections/language.tex
remove old hfill
Fixed minor nit 11 by reviewer #3
small fixes and more fixmes
add stub for abstract example
fixed intro section 4
All changes based on johan's comments

@@
 
-In this section, we present the constructs of the modular interactive workflow
-calculus \TOPHAT. We define the host and task language, the types, and
-the static semantics. Then we describe the workings of each construct
-using examples. These construct are formalised in \cref{sec:semantics}.
+In this section, we present the constructs of \TOPHAT, our modular interactive workflow calculus.
+We define the host and task language, the types, and the static semantics.
+Then we describe the workings of each construct using examples.
+These constructs are formalised in \cref{sec:semantics}.
 
@@
 Users can change the value until they are satisfied with it.
-Editors are meant capture this constantly changing nature of user input.
+Editors are meant to capture this constantly changing nature of user input.
 %
@@
 The successor task must not be $\Fail$, introduced below.
+This is enforced on the semantic level, as described in the next section.
 The internal step can proceed immediately when these conditions are met.
@@
 
-\hfill Internal \\ choice ($\Or$) is a parallel \emph{or}.
+Internal choice ($\Or$) is a parallel \emph{or}.
 It picks the leftmost branch that has a value.
@@
 \begin{example}[Delay]
+\label{exm:delay}
 

### 2019-02-25 sections/outlook.tex
change reference to abstract example to breakfast
All changes based on johan's comments

@@
 Firstly, we would like to express properties of tasks and prove them.
-For example, one would like to prove that, no matter what, in Example~\ref{exm:abstract} an abstract is always submitted.
-Secondly, we would like explore what it means for two tasks to be equal.
+For example, one would like to prove that, no matter what, in Example~\ref{exm:breakfast} breakfast is always being served.
+Secondly, we would like to explore what it means for two tasks to be equal.
 One could have noticed that some operators have a monadic or applicative feeling.
@@
 and $\Then$ looks like a bind operation.
-We need a correct understand of equivalency of tasks,
+We need a correct understand of equivalence of tasks,
 taking the interactive setting into account,

### 2019-02-25 sections/properties.tex
All changes based on johan's comments

@@
 
-Moving on, we show that normalisation also preserves, by showing that the
-following Theorem holds.
+Moving on, we show that normalisation also preserves types, by showing that the following Theorem holds.
 
@@
 
-Since this semantics makes use of the value function $\Value$ and the striding semantics $\stride$. We first need to
-show that preservation also holds for these two semantics.
+Since this semantics makes use of the value function $\Value$ and the striding semantics $\stride$,
+we first need to show that preservation also holds for these two semantics.
 
@@
   if $e,s \stride e',s'$,
-  then $\Gamma,\Sigma \infers e':\tau$ and $\Sigma \infers s'$.
+  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers s'$.
   \label{thm:pres-stride}
@@
 
-Lemma~\ref{lem:presvalue} and Theorem~\ref{thm:pres-stride} are proven in the
+Lemma~\ref{lem:presvalue} and Lemma~\ref{thm:pres-stride} are proven in the
 appendix by induction over $e$. This subsequently allows us to prove
@@
   such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
-  if $ e,s \handle{i} e's'$,
+  if $ e,s \handle{i} e',s'$,
   then $\Gamma,\Sigma\infers e':\Task\tau$ and $\Sigma\infers s'$.
@@
 From Theorem~\ref{thm:pres-handle} and Theorem~\ref{thm:pres-norm} we directly
-obtain that the driving semantics also preserves.
+obtain that the driving semantics also preserves types.
 
@@
 
-If an expression $e$ and state $s$ are well-typed, then after normalisation, it
-either fails, or there exists some input $i$ that can be handled by it.
+If an expression $e$ and state $s$ are well-typed, then after normalisation, the pair $e',s'$
+either fails, or there exists some input $i$ that can be handled by it under the handling semantics.
 In order to prove this Theorem, we require an additional theorem.
@@
 In order to validate the function that calculates all possible inputs $\Inputs$,
-we want to show that the set of possible inputs it produces is both sound and
-complete with respect to the handle semantics. By sound we mean that all inputs
-in the set of possible inputs can actually be handled by the handle semantics,
-and by complete we mend that the set of possible inputs contains all inputs
-that can be handled by the handle semantics. Theorem~\ref{thm:safety-i} expresses
-exactly this property.
+we want to show that the set of possible inputs it produces is both sound and complete with respect to the handle semantics.
+By sound we mean that all inputs in the set of possible inputs can actually be handled by the handle semantics,
+and by complete we mean that the set of possible inputs contains all inputs that can be handled by the handle semantics.
+Theorem~\ref{thm:safety-i} expresses exactly this property.
 
@@
   % such that $\Gamma,\Sigma \infers e:\tau$ and $\Sigma \infers s$,
-  we have that $i \in \Inputs{(e)}$ if and only if $e,s \handle{i} e',s'$.
+  we have that $i \in \Inputs{(e,s)}$ if and only if $e,s \handle{i} e',s'$.
   \label{thm:safety-i}

### 2019-02-25 sections/relatedwork.tex
All changes based on johan's comments

@@
 
-Much research has been done into workflow modelling. These works focus on
-describing the collaboration between subsystems, rather than the communication
-between them.
+Much research has been done into workflow modelling.
+This work focusses on describing the collaboration between subsystems,
+rather than the communication between them.
 The systems described here follow a \emph{boxes and arrows} model of specifying workflows.
@@
 
-HipHop \cite{berry2011hiphop,journals/corr/BerryS13} is a programming language tailored to the development synchronous reactive web systems.
+HipHop \cite{berry2011hiphop,journals/corr/BerryS13} is a programming language tailored to the development of synchronous reactive web systems.
 From a single source, both server and client applications can be generated.
@@
 As Hop is a dynamic language, and HipHop uses strings to identify events, events and their possible associated values are not statically checked.
-Events are aggregated until the moment the machine as asked to react.
+Events are aggregated until the moment the machine is asked to react.
 % then are perceived as conceptually simultaneous.
@@
 Where \TOPHAT takes tasks and collaboration as a starting point,
-HipHop focusses on synchronous communication and atom execution of reactive machines.
+HipHop focusses on synchronous communication and atomic execution of reactive machines.
 
@@
 Another important difference is the way in which both systems handle events.
-In HipHop the programmer decides when a machine should processes its events.
+In HipHop the programmer decides when a machine should process its events.
 This could be just one event, or a multi-set of events that are processed simultaneously.
@@
 \TOP has no notion of time.
-Tasks can not change spontaneous over time, while behaviours can.
+Tasks cannot change spontaneous over time, while behaviours can.
 Only input events can change task values.
@@
 preset data-type that drives the behaviour of the user defined applications.
+\fixme{compare with tophat}

### 2019-02-26 appendices/proofs.tex
Revert changes to types

@@
     {$\userule{H-Empty}$}
-    {Given that $\Gamma,\Sigma\infers\Edit v:\Task\beta$ and $\Sigma\infers s$, the \refrule{H-Empty} rule additionally gives us that $v:\tau$.
+    {Given that $\Gamma,\Sigma\infers\Edit v:\Task\tau$ and $\Sigma\infers s$, the \refrule{H-Empty} rule additionally gives us that $v:\tau$.
     Then by \refrule{T-Enter} we have $\Gamma,\Sigma\infers\Enter \tau:\Task\tau$.}

### 2019-02-26 sections/abstract.tex
Added loads of fixme's based on Rinus and Alejandro's comments

@@
 Current languages and frameworks are limited in flexibility and lack a good level of abstraction.
+\fixme{Rinus: het is vast waar, maar ik zou het iets zachter formuleren: focus op een concreet toepassingsgebied, maar minder op ...}
 Task-oriented programming (TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
@@
 TOP has been around for more than a decade, in the form of iTasks, which is tailored for real-world usability.
-So far, it has not been given a formalization in the style common in programming language research.
+So far, it has not been given a formalization in the style common in programming language research. \fixme{Rinus: not been given a formalization which is very suitable for formal reasoning.}
 

### 2019-02-26 sections/example.tex
scale figures for new format
compact code examples
make figure bigger
move figure in example
Added loads of fixme's based on Rinus and Alejandro's comments

@@
 \end{enumerate*}
-For this example we assume that the host language has lists and four functions on them: all, any, intersect, and difference.
+For this example we assume that the host language has lists and four functions on them: all, any, intersect, and difference. \fixme{Alejandro: small description may be needed} %nico: of the functions
 %%NOTE: After Peter's comments symbols are not clear --TS
@@
 For brevity, we omit the type annotations of variable bindings.
+
+\begin{figure}[b]
+  \includegraphics[width=0.8\columnwidth]{figures/flight-booking.png}
+  \caption{
+    Running web application of the flight booking example using a translation to \ITASKS.
+    It shows three users booking a flight simultaneously.
+    The first user entered name and age and continued picking seats.
+    The second is entering details of two passengers.
+    The ages are not filled in, therefore the \TS{Continue} button is disabled.
+    The third user finished a booking.
+    The first user can not pick seats \smallcaps{1b} and \smallcaps{1c} any more.
+    The message bubble shows that the \TS{age} field only accepts integer values.
+  }
+  \label{fig:flight-booking}
+\end{figure}
 
@@
 \end{TASK}
+\fixme{Alejandro: At this point I wonder, what is $\Enter$}
 A selection of seats is correct if every entered seat is free.
@@
   let chooseSeats = \ps. enter (List Seat) >>? \ss.
-    if correct ss /\ length ps == length ss
-      then confirmBooking ps ss else fail
+    if correct ss /\ length ps == length ss then confirmBooking ps ss else fail
 \end{TASK}
+
 The function \TS{confirmBooking} removes the picked seats from the shared list of free seats,
-and displays the end result using the \TS{edit}-construct.
+and displays the end result using the \TS{edit}-construct.\fixme{Alejandro: Give \TS{edit} a proper name?}
 \begin{TASK}
@@
 \end{example}
-
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{figures/flight-booking.png}
-  \caption{
-    Running web application of the flight booking example using a translation to \ITASKS.
-    It shows three users booking a flight simultaneously.
-    The first user entered name and age and continued picking seats.
-    The second is entering details of two passengers.
-    The ages are not filled in, therefore the \TS{Continue} button is disabled.
-    The third user finished a booking.
-    The first user can not pick seats \smallcaps{1b} and \smallcaps{1c} any more.
-    The message bubble shows that the \TS{age} field only accepts integer values.
-  }
-  \label{fig:flight-booking}
-\end{figure}

### 2019-02-26 sections/introduction.tex
Added loads of fixme's based on Rinus and Alejandro's comments

@@
 \section{Introduction}
+
+\fixme{Alejandro: The introduction seems to be more about why Task-Oriented Programming
+is nice and useful. Personally, I would like it to tell me another
+story: given that some people find TOP useful (as shown by the many
+examples), why should be care about formalizing it and which challenges
+does it pose? I think this is what people from ICFP would care about more.}
 
@@
 \subsection{Task-oriented programming}
-
+\fixme{Alejandro: Not entirely convincing}%Nico: hij heeft het hier volgens mij over de titel van de sectie en het woord aims in de eerste zin. Misschien moeten we een hardere claim doen hier?
 % Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
 Task-oriented programming (\TOP) is a programming paradigm which aims to be the sweet spot between faithful modelling workflows
-and rapid prototyping of multi-user web applications supporting these workflows.
+and rapid prototyping of multi-user web applications supporting these workflows.\fixme{Rinus: Hier zou ik een referentie naar ons ppdp2012 paper verwachten.}
 % Task-oriented programming (\TOP) is a programming paradigm to support these ways of working.
@@
 Firstly, it should provide primitive building blocks that are useful for high-level descriptions of how users collaborate with each other and with machines.
-These building blocks are \emph{editors}, \emph{composition}, and \emph{shared data}.
+These building blocks are: \emph{editors}, \emph{composition}, and \emph{shared data}.
 % Second, it should be able to generate multi-user web applications to support these scenarios.
@@
 
-Currently, we know of two frameworks using \TOP: \ITASKS and \MTASKS.
+Currently, we know of two frameworks implementing \TOP: \ITASKS and \MTASKS.
 
@@
 
-Both implementations currently lack formal semantics which can be used to prove properties about tasks.
+Both implementations currently lack formal semantics which can be used to prove properties about tasks.\fixme{Rinus: can be used.. -> is suited}
 
@@
 This creates a synergy where \ITASKS profits from the expressivity of Clean, while enhancing the language with functionality for user interaction.
-The tight integration of task DSL and host language makes it sometimes difficult to see where the boundaries are.
+The tight integration of task DSL and a host language makes it sometimes difficult to see where the boundaries are.
 

### 2019-02-26 sections/intuition.tex
Added loads of fixme's based on Rinus and Alejandro's comments

@@
 Instead, what is meant by parallel is that the order in which we do tasks and the smaller tasks that they are composed of, does not matter.
+\fixme{Alejandro has placed a line and exclamation under this sentence. I'm not sure what to do here.}
 Then finally, only when every item we want to have for breakfast is ready, can we sit down and enjoy it.
@@
 This results in a typed empty editor $\Enter \tau$, which is annotated with the same type.
+\fixme{Alejandro: unclear} %Nico: He has marked the above paragraph as being unclear.
 
@@
 
+\fixme{Alejandro: What is a value? What is a type?}
+
 \subsection{Tasks can be observed}
@@
 Several observations can be made about tasks.
-One of those is determining the value of a task $\Value(t)$.
+One of those is determining the value of a task $\Value(t)$. \fixme{Should we change the type of Value to Maybe v? Both johan and alejandro mention that it is not completely clear to them.}
 Not all tasks have a value, for example the empty editor $\Enter \tau$, which makes $\Value$ a partial function.
@@
 The final observation is to determine whether a task results in $\Fail$ (failure).
-The step combinator and the choice combinator use this to prevent users from picking a failing task.
+The step combinator $\Next$ and the choice combinator $\Xor$ use this to prevent users from picking a failing task.
 
@@
 The parallel combination $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
-Therefore $\And$ introduces concurrency.
+Therefore $\And$ introduces concurrency. \fixme{Alejandro: But before you said it was interleaving!}%nico: en later doen we dit ook.
 For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
@@
 
-To model collaboration, one needs communication, concurrency, and synchronisation.
+To model collaboration, one needs communication, concurrency, and synchronisation. \fixme{Alejandro: Is this a basic calculus to why iTasks is elaborated?} %nico: not sure what he means here...
 In \TOP, collaboration is specified on a high level of abstraction, hiding the details of communication.

### 2019-02-26 sections/language.tex
added fixme about beta
Added loads of fixme's based on Rinus and Alejandro's comments

@@
 \label{sub:expressions}
-The host language is a simply typed lambda calculus, extended with some basic types and \ML-style references.
+The host language is a simply typed lambda calculus, extended with some basic types and \ML-style references.\fixme{Alejandro: why references?}
 The grammar in \cref{fig:language-grammar} defines the syntax of the host language.
@@
 \end{figure}
-
+\fixme{Alejandro about l: different syntax from vars?}
 \label{sub:notation}
@@
 The grammar in \cref{fig:task-grammar} specifies the syntactic category of \emph{pretasks}.
-Pretasks are tasks that have unevaluated subexpressions.
+Pretasks are tasks that have unevaluated subexpressions. \fixme{Alejandro: How do you define unevaluated?}
 Each pretask will be discussed in more detail in the following subsections.
@@
 The typing rules for pretasks are given in \cref{fig:typing-rules}.
-
+\fixme{Say someting about why we need beta, and use beta in the acutal typing rule}
+\fixme{Alejandro: Is there anything surprising? Say what}
 \begin{figure}[h]
@@
 
-\paragraph{Internal and external step $(t \Then e, t \Next e)$}
+\paragraph{Internal and external step $(t \Then e, t \Next e)$}\fixme{Alejandro: The naming is confusing to me}
 \label{sub:steps}
@@
 An external choice requires users to pick a branch before continuing with it.
-This means users cannot work on the branches of an external choice before picking one.
+This means users cannot work on the branches of an external choice before picking one. \fixme{Alejandro: Is this actually implied?}
 

### 2019-02-26 sections/properties.tex
Added loads of fixme's based on Rinus and Alejandro's comments

@@
 
-Additionally, we show that the normalisation semantics is a big-step semantics.
+Additionally, we show that the normalisation semantics is a big-step semantics. \fixme{Alejandro: ?}
 While at first sight, this might seem obvious, the fact that we are dealing with
@@
 
-\subsection{Preservation}
+\subsection{Type preservation}
 \label{sub:preservation}
@@
 
-\begin{theorem}[Preservation under evaluation]
+\begin{theorem}[Type preservation under evaluation]
   % For all well typed expressions $e$ and states $s$,
@@
 
-\begin{theorem}[Preservation under normalisation]
+\begin{theorem}[Type preservation under normalisation]
   % For all well typed expressions $e$ and states $s$,
@@
 
-\begin{lemma}[Task value preserves]
+\begin{lemma}[Task value preserves types]
   % For all well typed expressions $e$ and states $s$,
@@
 
-\begin{lemma}[Striding preserves]
+\begin{lemma}[Striding preserves types]
   For all expressions $e$ and states $s$
@@
 
-\begin{theorem}[Preservation under handling]
+\begin{theorem}[Type preservation under handling]
   % For all well typed expressions $e$, states $s$, and inputs $i$,

### 2019-02-26 sections/relatedwork.tex
Added loads of fixme's based on Rinus and Alejandro's comments

@@
 \paragraph{iTasks}
-
+\fixme{Rinus: Waarom geen korte opsomming van iTask features die niet zijn meegenomen in TopHat?}
 As mentioned earlier, iTasks is an implementation of \TOP.
@@
 These semantics however do not make an explicit distinction between the host language and task language and they do not provide a formal semantics, which makes formal reasoning difficult.
+
+\fixme{Rinus: Volgens mij is er wel degelijk een formele semantiek, omdat heel iTask wordt uitgelegd in termen van standaard Clean functies,
+en Clean heeft een semantiek (graaf herschrijf systeem).
+Dat de beschrijving onhandig (wellicht te low level is) is voor het formeel redeneren over iTask programma's is een beter argument.}
+\fixme{Alejandro: What is missing from \TOPHAT to describe i or m Tasks}
 

### 2019-02-26 sections/semantics.tex
changed UI fixme to comment
All changes based on johan's comments
Added loads of fixme's based on Rinus and Alejandro's comments

@@
 \label{sec:semantics}
+
+\fisme{Alejandro: In the formalization you show quite some figures. I think that for
+each figure you need to be more explicit in the text about what is
+unsurprising and what is new or should get the attention of the reader.}
 
@@
 \end{figure}
-
+\fixme{Alejandro over tasks: I was not expecting these to be values...}
 The rules to evaluate expressions $e$ do not differ from standard work, except for the task constructs.
 The evaluation rules for tasks can be deduced from the value grammar.
-They are given in the appendix.
+They are given in the appendix. \fixme{Alejandro: What is interesting about them?}
 
@@
 To determine if an external choice is failing, it needs to peek into the future to check if both subexpressions are failing.
-Annotations are failing only if the inner task is failing.
 
@@
 
-\fixme{Add $\UserInterface$ to appendix and refer.}
+%\fixme{Add $\UserInterface$ to appendix and refer.}
 \TOPHAT is designed such that a user interface can be generated from a task's syntax tree.
@@
 To overcome this problem, the \refrule{N-Done} and \refrule{N-Repeat} rules ensure that striding is applied until the state $s$ becomes stable and no further normalisation can take place.
+\fixme{The example is not 100\% clear to me}
 
@@
 \end{figure}
-
+\fixme{Fi, Si are sort of a path to the task}
 \begin{figure}[h]

### 2019-02-26 sections/introduction.tex
some fixmes

@@
 Task-oriented programming (\TOP) is a programming paradigm which aims to be the sweet spot between faithful modelling workflows
-and rapid prototyping of multi-user web applications supporting these workflows.\fixme{Rinus: Hier zou ik een referentie naar ons ppdp2012 paper verwachten.}
+and rapid prototyping of multi-user web applications supporting these workflows \cite{conf/ppdp/PlasmeijerLMAK12}.
 % Task-oriented programming (\TOP) is a programming paradigm to support these ways of working.

### 2019-02-26 sections/intuition.tex
some fixmes

@@
 
-To model collaboration, one needs communication, concurrency, and synchronisation. \fixme{Alejandro: Is this a basic calculus to why iTasks is elaborated?} %nico: not sure what he means here...
-In \TOP, collaboration is specified on a high level of abstraction, hiding the details of communication.
+Collaboration in the real world consists of three aspects: communication, concurrency, and synchronisation.
+These aspects are reflected in \TOP on a high level of abstraction, hiding the details of communication.
 For example, the cigarette smokers communicate with each other, but the programs do not explicitly mention sending or receiving events.

### 2019-02-26 sections/language.tex
unevaluated means not in normal form

@@
 The grammar in \cref{fig:task-grammar} specifies the syntactic category of \emph{pretasks}.
-Pretasks are tasks that have unevaluated subexpressions. \fixme{Alejandro: How do you define unevaluated?}
+Pretasks are tasks that have unevaluated subexpressions.
+This means they are considered to be in normal form for the host language, even if their subexpressions are not in normal form.
 Each pretask will be discussed in more detail in the following subsections.

### 2019-02-26 sections/relatedwork.tex
examples of stuff that we didn't include in TopHat

@@
 \paragraph{iTasks}
-\fixme{Rinus: Waarom geen korte opsomming van iTask features die niet zijn meegenomen in TopHat?}
 As mentioned earlier, iTasks is an implementation of \TOP.
@@
 This is useful for everyday programming, where having lots of functionality at one's fingertips is convenient and efficient.
+\TOPHAT on the other hand does not include the many different variations of the step- and parallel combinators of \ITASKS.
+To name two examples, the combinators \verb+(>>|)+ and \verb+(||-)+ are variations of step and parallel that ignore the value of the left task.
 
@@
 Dat de beschrijving onhandig (wellicht te low level is) is voor het formeel redeneren over iTask programma's is een beter argument.}
-\fixme{Alejandro: What is missing from \TOPHAT to describe i or m Tasks}
 

### 2019-02-26 sections/relatedwork.tex
distinction between explicit and implicit formal semantics

@@
 Both give a different semantics in the form of minimal implementations of a subset of the interface of iTasks.
-These semantics however do not make an explicit distinction between the host language and task language and they do not provide a formal semantics, which makes formal reasoning difficult.
-
-\fixme{Rinus: Volgens mij is er wel degelijk een formele semantiek, omdat heel iTask wordt uitgelegd in termen van standaard Clean functies,
-en Clean heeft een semantiek (graaf herschrijf systeem).
-Dat de beschrijving onhandig (wellicht te low level is) is voor het formeel redeneren over iTask programma's is een beter argument.}
+These semantics however do not make an explicit distinction between the host language and task language and they do not provide an explicit formal semantics.
+\ITASKS implicitly has a formal semantics, inherited from the graph-rewriting semantics of Clean.
+This semantics however does not lend itself well for formal reasoning.
 

### 2019-02-26 sections/abstract.tex
some fixmes
focus on non-programmers leads to inflexibility

--- 2019-02-25 sections/abstract.tex
@@
 Software that models how people work is omnipresent in today's society.
-Current languages and frameworks are limited in flexibility and lack a good level of abstraction.
+Current languages and frameworks often focus on usability by non-programmers, sacrificing flexibility and high level of abstraction.
 Task-oriented programming (TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
@@
 TOP has been around for more than a decade, in the form of iTasks, which is tailored for real-world usability.
-So far, it has not been given a formalization in the style common in programming language research.
+So far, it has not been given a formalization which is very suitable for formal reasoning.
 

### 2019-02-26 sections/example.tex
describe all, any, intersect, difference
briefly explain Enter

@@
 \end{enumerate*}
-For this example we assume that the host language has lists and four functions on them: all, any, intersect, and difference. \fixme{Alejandro: small description may be needed} %nico: of the functions
+For this example we assume that the host language has lists and four functions on them: all, any, intersect, and difference.
+The functions all and any check if all or any elements in a list satisfy a given predicate.
+The functions intersect and difference compute the set-intersection and set-difference of two lists.
 %%NOTE: After Peter's comments symbols are not clear --TS
@@
 and the user can proceed to picking seats.
+The task $\Enter$ is an empty editor that asks for a value of the given type.
 \begin{TASK}
@@
 \end{TASK}
-\fixme{Alejandro: At this point I wonder, what is $\Enter$}
 A selection of seats is correct if every entered seat is free.

### 2019-02-26 sections/example.tex
call editor an editor

@@
 The function \TS{confirmBooking} removes the picked seats from the shared list of free seats,
-and displays the end result using the \TS{edit}-construct.\fixme{Alejandro: Give \TS{edit} a proper name?}
+and displays the end result using an editor \TS{edit}.
 \begin{TASK}

### 2019-02-26 sections/intuition.tex
make it clearer what partial function means
by giving an example with bottom

@@
 Several observations can be made about tasks.
-One of those is determining the value of a task $\Value(t)$. \fixme{Should we change the type of Value to Maybe v? Both johan and alejandro mention that it is not completely clear to them.}
+One of those is determining the value of a task $\Value(t)$.
+% \fixme{Should we change the type of Value to Maybe v? Both johan and alejandro mention that it is not completely clear to them.}
+% I don't think that we should use Maybe. We should use the term partial function and bottom. -- mkl
 Not all tasks have a value, for example the empty editor $\Enter \tau$, which makes $\Value$ a partial function.
+For example, $\Value(\Edit 5) = 5$, $\Value(\Enter \Int) = \bot$.
 

### 2019-02-26 sections/language.tex
references because shared data

@@
 \label{sub:expressions}
-The host language is a simply typed lambda calculus, extended with some basic types and \ML-style references.\fixme{Alejandro: why references?}
+The host language is a simply typed lambda calculus, extended with some basic types and \ML-style references.
+We use references to represent shared data sources.
 The grammar in \cref{fig:language-grammar} defines the syntax of the host language.

### 2019-02-26 sections/language.tex
explain difference between x and l
briefly explain internal vs. external

@@
 Creating a reference using the keyword $\Ref$ yields a location $l$.
+While $x$ denotes program variables, $l$ denotes store locations.
 Locations are not intended to be directly manipulated by the programmer.
@@
 \end{figure}
-\fixme{Alejandro about l: different syntax from vars?}
 \label{sub:notation}
@@
 
-\paragraph{Internal and external step $(t \Then e, t \Next e)$}\fixme{Alejandro: The naming is confusing to me}
+\paragraph{Internal and external step $(t \Then e, t \Next e)$}
 \label{sub:steps}
@@
 Sequential composition has a task $t$ on the left and a continuation $e$ on the right.
+External steps ($\Next$) must be triggered by the user, while internal steps ($\Then$) are taken automatically.
 The accompanying typing rules are \refrule{T-Then} and \refrule{T-Next}.

### 2019-02-27 sections/abstract.tex
soften abstract a bit

@@
 TOP has been around for more than a decade, in the form of iTasks, which is tailored for real-world usability.
-So far, it has not been given a formalization which is very suitable for formal reasoning.
+So far, it has not been given a formalisation which is suitable for formal reasoning.
 
@@
 TOP has been employed in projects with the Dutch coast guard, tax office, and navy.
-Our work matters because Formal program verification is important for mission-critical software, especially for systems with concurrency.
+Our work matters because formal program verification is important for mission-critical software, especially for systems with concurrency.
 

### 2019-02-27 sections/example.tex
new paragraph after itemize in example

@@
 \end{enumerate*}
-For this example we assume that the host language has lists and four functions on them: all, any, intersect, and difference.
-The functions all and any check if all or any elements in a list satisfy a given predicate.
-The functions intersect and difference compute the set-intersection and set-difference of two lists.
+
+For this example we assume that the host language has lists and four functions on them: \TS{all}, \TS{any}, \TS{intersect}, and \TS{difference}.
+The functions \TS{all} and \TS{any} check if all or any elements in a list satisfy a given predicate.
+The functions \TS{intersect} and \TS{difference} compute the set-intersection and set-difference of two lists.
 %%NOTE: After Peter's comments symbols are not clear --TS

### 2019-02-27 sections/example.tex
move comment on symbols up in example

@@
 Now we develop our workflow in a top down way.
-A flight booking starts with an interactive task \TS{enter (List Passenger)}
+Our flight booking starts with an interactive task \TS{enter (List Passenger)}.
+A task $\Enter \tau$ is an empty editor that asks for a value of the given type $\tau$.
 where users can enter a list of passengers.
@@
 and the user can proceed to picking seats.
-The task $\Enter$ is an empty editor that asks for a value of the given type.
 \begin{TASK}

### 2019-02-27 sections/language.tex
reference where evaluation will be discusses in pretask section

@@
 The grammar in \cref{fig:task-grammar} specifies the syntactic category of \emph{pretasks}.
-Pretasks are tasks that have unevaluated subexpressions.
-This means they are considered to be in normal form for the host language, even if their subexpressions are not in normal form.
+Pretasks are tasks that have unevaluated subexpressions with respect to the host language.
+Who expressions are evaluated will be discussed in Section~\ref{sec:evaluation}.
 Each pretask will be discussed in more detail in the following subsections.
@@
 The typing rules for pretasks are given in \cref{fig:typing-rules}.
+
 \fixme{Say someting about why we need beta, and use beta in the acutal typing rule}

### 2019-02-27 appendices/proofs.tex
use cleveref consistently

--- 2019-02-21 appendices/proofs.tex
@@
 
-\subsection{Theorem~\ref{thm:pres-eval}}
+\subsection{\cref{thm:pres-eval}}
 
 \begin{proof}
-  We prove Theorem~\ref{thm:pres-eval} by induction on $e$:
+  We prove \cref{thm:pres-eval} by induction on $e$:
 
@@
 
-\subsection{Lemma~\ref{lem:presvalue}}
+\subsection{\cref{lem:presvalue}}
 
 \begin{proof}
-  We prove Lemma~\ref{lem:presvalue} by induction over $e$.
+  We prove \cref{lem:presvalue} by induction over $e$.
 
@@
 
-\subsection{Theorem~\ref{thm:pres-stride}}
+\subsection{\cref{thm:pres-stride}}
 
 \begin{proof}
-  We prove Theorem~\ref{thm:pres-stride} by induction on $e$:
+  We prove \cref{thm:pres-stride} by induction on $e$:
 
@@
   $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$. By the induction hypothesis, we
-  know that $t_1\stride t_1'$ preserves. By Lemma~\ref{lem:presvalue}, we know
-  that $\Value{(t_1')}=v_1$ preserves. By Theorem~\ref{thm:pres-eval} we know
+  know that $t_1\stride t_1'$ preserves. By \cref{lem:presvalue}, we know
+  that $\Value{(t_1')}=v_1$ preserves. By \cref{thm:pres-eval} we know
   that $e_2 v_1\evaluate t_2$ preserves. And finally by the induction hypothesis,
@@
 
-\subsection{Theorem~\ref{thm:pres-norm}}
+\subsection{\Cref{thm:pres-norm}}
 
 \begin{proof}
-  We prove Theorem~\ref{thm:pres-norm} by induction on $e$:
+  We prove \cref{thm:pres-norm} by induction on $e$:
 
@@
     {Given that
-    $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that
-    $\Gamma,\Sigma\infers t:\Task\tau$ and $\Sigma\infers s'$ by
-    Theorem~\ref{thm:pres-eval}. Then by Theorem~\ref{thm:pres-stride}, we have
-    $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$.}
+    $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that $\Gamma,\Sigma\infers t:\Task\tau$ and $\Sigma\infers s'$ by
+    \Cref{thm:pres-eval}.
+    Then by \cref{thm:pres-stride}, we have $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$.}
 
@@
     {$\userule{N-Repeat}$}
-    { Given that
-    $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that
-    $\Gamma,\Sigma\infers t:\Task\tau$ and $\Sigma\infers s'$ by
-    Theorem~\ref{thm:pres-eval}. Then by Theorem~\ref{thm:pres-stride}, we have
-    $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$. Then by the
-    induction hypothesis, we finally obtain that
-    $\Gamma,\Sigma\infers t'':\Task\tau$ and $\Sigma\infers s'''$.}
+    { Given that $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that $\Gamma,\Sigma\infers t:\Task\tau$ and $\Sigma\infers s'$ by \cref{thm:pres-eval}.
+    Then by \cref{thm:pres-stride}, we have $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$.
+    Then by the induction hypothesis, we finally obtain that $\Gamma,\Sigma\infers t'':\Task\tau$ and $\Sigma\infers s'''$.}
 
@@
 
-\subsection{Theorem~\ref{thm:pres-handle}}
+\subsection{\cref{thm:pres-handle}}
 
@@
 \begin{proof}
-  We prove Theorem~\ref{thm:pres-handle} by induction on $e$:
+  We prove \cref{thm:pres-handle} by induction on $e$:
 
@@
     This gives us that $\Sigma(l)=\tau$, and we additionally obtain $s(l),v':\tau$ by \refrule{H-Update}.
-    By application of Lemma~\ref{lemmasigmaconsistent} this case holds.}
+    By application of \cref{lemmasigmaconsistent} this case holds.}
 
@@
 
-% \subsection{Theorem~\ref{thmprogressnorm}}
+% \subsection{\cref{thmprogressnorm}}
 %
@@
 
-% \subsection{Lemma~\ref{lem:stride-does-not-eval}}
+% \subsection{\cref{lem:stride-does-not-eval}}
 %
 % \begin{proof}
-%   We prove Lemma~\ref{lem:stride-does-not-eval} by induction on $e$. First we
+%   We prove \cref{lem:stride-does-not-eval} by induction on $e$. First we
 %   apply evaluation, and we then find ourselves in one of the cases below.\\
@@
 
-\subsection{Theorem~\ref{thm:failing}}
+\subsection{\cref{thm:failing}}
 
@@
 
-  We prove Theorem~\ref{thm:failing} by induction on $e'$.
+  We prove \cref{thm:failing} by induction on $e'$.
 
@@
 
-\subsection{Theorem~\ref{thm:safety-i}}
+\subsection{\cref{thm:safety-i}}
 

### 2019-02-27 sections/introduction.tex
some fixmes
use cleveref consistently

@@
 
-Both implementations currently lack formal semantics which can be used to prove properties about tasks.\fixme{Rinus: can be used.. -> is suited}
+Both implementations currently lack formal semantics which are suited to prove properties about tasks.
 
@@
 
-In Section~\ref{sec:example} we demonstrate the functionality of \TOPHAT by means of an example,
-Section~\ref{sec:intuition} gives an overview of the essential concepts of \TOP and \TOPHAT.
-Section~\ref{sec:language} introduces the \TOPHAT calculus syntax
-and Section~\ref{sec:semantics} the semantics.
-Then in Section~\ref{sec:properties} we show that certain properties hold for the calculus.
-We take a look at related work in Section~\ref{sec:relatedwork}
-and conclude in Section~\ref{sec:conclusions}.
+In \cref{sec:example} we demonstrate the functionality of \TOPHAT by means of an example,
+\Cref{sec:intuition} gives an overview of the essential concepts of \TOP and \TOPHAT.
+\Cref{sec:language} introduces the \TOPHAT calculus syntax
+and \cref{sec:semantics} the semantics.
+Then in \cref{sec:properties} we show that certain properties hold for the calculus.
+We take a look at related work in \cref{sec:relatedwork}
+and conclude in \cref{sec:conclusions}.

### 2019-02-27 sections/outlook.tex
use cleveref consistently

@@
 Firstly, we would like to express properties of tasks and prove them.
-For example, one would like to prove that, no matter what, in Example~\ref{exm:breakfast} breakfast is always being served.
+For example, one would like to prove that, no matter what, in \cref{exm:breakfast} breakfast is always being served.
 Secondly, we would like to explore what it means for two tasks to be equal.
@@
 
-A more in depth description of future work can be found in Section~\ref{sec:conclusions}.
+A more in depth description of future work can be found in \cref{sec:conclusions}.

### 2019-02-27 sections/properties.tex
use cleveref consistently

@@
 
-Theorem~\ref{thm:pres-eval} is shown to be correct by induction over $e$. The full
+\Cref{thm:pres-eval} is shown to be correct by induction over $e$. The full
 proof can be found in the appendix.
@@
 
-Lemma~\ref{lem:presvalue} and Lemma~\ref{thm:pres-stride} are proven in the
+\Cref{lem:presvalue} and \cref{thm:pres-stride} are proven in the
 appendix by induction over $e$. This subsequently allows us to prove
-Theorem~\ref{thm:pres-norm}, again by induction over $e$.
+\Cref{thm:pres-norm}, again by induction over $e$.
 
@@
 
-From Theorem~\ref{thm:pres-handle} and Theorem~\ref{thm:pres-norm} we directly
+From \cref{thm:pres-handle} and \cref{thm:pres-norm} we directly
 obtain that the driving semantics also preserves types.
@@
 
-We define what we mean with progress in Theorem~\ref{thm:prog-norm}.
+We define what we mean with progress in \cref{thm:prog-norm}.
 \begin{theorem}[Progress under handling]
@@
 
-We now have the ingredients to prove Theorem~\ref{thm:prog-norm}.
+We now have the ingredients to prove \cref{thm:prog-norm}.
 
@@
   There does not exist an $i$ such that $e',s'\handle{i} e'',s''$. In this case, we
-  know that $\Failing(e',s')$ must be true, by Theorem~\ref{thm:failing}.
+  know that $\Failing(e',s')$ must be true, by \cref{thm:failing}.
 \end{proof}
@@
 and by complete we mean that the set of possible inputs contains all inputs that can be handled by the handle semantics.
-Theorem~\ref{thm:safety-i} expresses exactly this property.
+\Cref{thm:safety-i} expresses exactly this property.
 

### 2019-02-27 sections/introduction.tex
push fixmes in margin for last proofread

@@
 
-\fixme{Alejandro: The introduction seems to be more about why Task-Oriented Programming
+\todo{Alejandro: The introduction seems to be more about why Task-Oriented Programming
 is nice and useful. Personally, I would like it to tell me another
@@
 does it pose? I think this is what people from ICFP would care about more.}
-
 Many applications these days are developed to support workflows in institutions and businesses.
@@
 \subsection{Task-oriented programming}
-\fixme{Alejandro: Not entirely convincing}%Nico: hij heeft het hier volgens mij over de titel van de sectie en het woord aims in de eerste zin. Misschien moeten we een hardere claim doen hier?
+\todo{Alejandro: Not entirely convincing}%Nico: hij heeft het hier volgens mij over de titel van de sectie en het woord aims in de eerste zin. Misschien moeten we een hardere claim doen hier?
 % Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
@@
 Finally, there is communication with the \emph{outside} world, where information is entered into the system via input events.
-%\fixme{Should this be mentioned earlier?}
+\todo{Should this be mentioned earlier?}
 The end points where the outside world interacts with \TOP applications are called editors.

### 2019-02-27 sections/intuition.tex
call it interleaving concurrency to make sure what we mean
make notes about internal discussions
use cleveref consistently
push fixmes in margin for last proofread

@@
 Instead, what is meant by parallel is that the order in which we do tasks and the smaller tasks that they are composed of, does not matter.
-\fixme{Alejandro has placed a line and exclamation under this sentence. I'm not sure what to do here.}
+\todo{Alejandro has placed a line and exclamation under this sentence. I'm not sure what to do here.}
 Then finally, only when every item we want to have for breakfast is ready, can we sit down and enjoy it.
@@
 This results in a typed empty editor $\Enter \tau$, which is annotated with the same type.
-\fixme{Alejandro: unclear} %Nico: He has marked the above paragraph as being unclear.
+\todo{Alejandro: unclear} %Nico: He has marked the above paragraph as being unclear.
 
@@
 
-\fixme{Alejandro: What is a value? What is a type?}
+\todo{Alejandro: What is a value? What is a type?}
 
@@
 One of those is determining the value of a task $\Value(t)$.
-% \fixme{Should we change the type of Value to Maybe v? Both johan and alejandro mention that it is not completely clear to them.}
-% I don't think that we should use Maybe. We should use the term partial function and bottom. -- mkl
+%NOTE: Should we change the type of Value to Maybe v? Both johan and alejandro mention that it is not completely clear to them. --NN
+%NOTE: I don't think that we should use Maybe. We should use the term partial function and bottom. --MKL
 Not all tasks have a value, for example the empty editor $\Enter \tau$, which makes $\Value$ a partial function.
@@
 The parallel combination $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
-Therefore $\And$ introduces concurrency. \fixme{Alejandro: But before you said it was interleaving!}%nico: en later doen we dit ook.
+This operator introduces interleaving concurrency.
 For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
@@
 %
-% The \TS{delay} task is specified in Example~\ref{exm:delay} in Section~\ref{sec:language}.
+% The \TS{delay} task is specified in \cref{exm:delay} in \cref{sec:language}.
 %

### 2019-02-27 sections/language.tex
use cleveref consistently
push fixmes in margin for last proofread

@@
 Pretasks are tasks that have unevaluated subexpressions with respect to the host language.
-Who expressions are evaluated will be discussed in Section~\ref{sec:evaluation}.
+Who expressions are evaluated will be discussed in \cref{sec:evaluation}.
 Each pretask will be discussed in more detail in the following subsections.
@@
 The typing rules for pretasks are given in \cref{fig:typing-rules}.
+\todo{Say someting about why we need beta, and use beta in the acutal typing rule}
+\todo{Alejandro: Is there anything surprising? Say what}
 
-\fixme{Say someting about why we need beta, and use beta in the acutal typing rule}
-\fixme{Alejandro: Is there anything surprising? Say what}
 \begin{figure}[h]
@@
 An external choice requires users to pick a branch before continuing with it.
-This means users cannot work on the branches of an external choice before picking one. \fixme{Alejandro: Is this actually implied?}
+This means users cannot work on the branches of an external choice before picking one. \todo{Alejandro: Is this actually implied?}
 
@@
 %
-% \fixme{write this subsection}
-
+%
 % It is essential for a workflow system to support multiple collaborators working
@@
 % All semantic rules in \cref{sec:semantics} simply continue with their operations as if the label was not there.
-% \todo{we still have to fix the type of $\Inputs$ and state the type of $u$}

### 2019-02-27 sections/semantics.tex
some fixmes
F and S in inputs
use correct font for F and S in semantics
push fixmes in margin for last proofread

--- 2019-02-21 sections/semantics.tex
@@
 
+\todo{Alejandro: In the formalization you show quite some figures. I think that for
+each figure you need to be more explicit in the text about what is
+unsurprising and what is new or should get the attention of the reader.}
 In this section we formalise the semantics of the language constructs described in \cref{sec:language}.
@@
 \end{figure}
-
+\todo{Alejandro over tasks: I was not expecting these to be values...}
 The rules to evaluate expressions $e$ do not differ from standard work, except for the task constructs.
 The evaluation rules for tasks can be deduced from the value grammar.
-They are given in the appendix.
+They are given in the appendix. \todo{Alejandro: What is interesting about them?}
 
@@
 
-\fixme{Add $\UserInterface$ to appendix and refer.}
+%FIXME: Add $\UserInterface$ to appendix and refer.
 \TOPHAT is designed such that a user interface can be generated from a task's syntax tree.
@@
 To overcome this problem, the \refrule{N-Done} and \refrule{N-Repeat} rules ensure that striding is applied until the state $s$ becomes stable and no further normalisation can take place.
+\todo{The example is not 100\% clear to me}
 
@@
 Inputs $i$ are formed according to the grammar in \cref{fig:input-grammar}.
+$\First$ and $\Second$ in an input encode the path to the task for which the input is designated.
 %
@@
 \end{figure}
-
 \begin{figure}[h]

### 2019-02-27 sections/introduction.tex
spelling

@@
 story: given that some people find TOP useful (as shown by the many
-examples), why should be care about formalizing it and which challenges
+examples), why should be care about formalising it and which challenges
 does it pose? I think this is what people from ICFP would care about more.}

### 2019-02-27 sections/abstract.tex
merge pars in abstract for better readability, also mention mTasks

@@
 Software that models how people work is omnipresent in today's society.
-Current languages and frameworks often focus on usability by non-programmers, sacrificing flexibility and high level of abstraction.
+Current languages and frameworks often focus on usability by non-programmers, sacrificing flexibility and high level abstraction.
 Task-oriented programming (TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
@@
 Such specifications can be turned into interactive applications for different platforms, supporting collaboration during execution.
-
 % Inquiry
-TOP has been around for more than a decade, in the form of iTasks, which is tailored for real-world usability.
+TOP has been around for more than a decade, in the forms of iTasks and mTasks, which is tailored for real-world usability.
 So far, it has not been given a formalisation which is suitable for formal reasoning.
@@
 This approach allows for comparison with other work in the field.
-
 % Grounding
-We have implemented the semantic rules of TopHat in Idris, and the task layer on top of iTasks.
+We have implemented the semantic rules of TopHat in Idris, and the task layer on top of iTasks combinators.
 This shows that our approach is feasible, and lets us demonstrate the concepts by means of illustrative case studies.
-
 % Importance

### 2019-02-27 sections/introduction.tex
rewrite `scope` into `challenges` stating why we do this

--- 2019-02-27 sections/introduction.tex
@@
 
-\fixme{Alejandro: The introduction seems to be more about why Task-Oriented Programming
-is nice and useful. Personally, I would like it to tell me another
-story: given that some people find TOP useful (as shown by the many
-examples), why should be care about formalizing it and which challenges
-does it pose? I think this is what people from ICFP would care about more.}
+%XXX: Alejandro:
+% The introduction seems to be more about why Task-Oriented Programming is nice and useful.
+% Personally, I would like it to tell me another story:
+% given that some people find TOP useful (as shown by the many examples),
+% why should be care about formalising it and which challenges does it pose?
+% I think this is what people from ICFP would care about more.
+%NOTE: Ik hoop dat de `Challenges` sectie dit wat opvangt. --TS
 
@@
 \subsection{Task-oriented programming}
-\fixme{Alejandro: Not entirely convincing}%Nico: hij heeft het hier volgens mij over de titel van de sectie en het woord aims in de eerste zin. Misschien moeten we een hardere claim doen hier?
+\label{sec:top}
+
+%XXX: Alejandro: Not entirely convincing
+% Hij heeft het hier volgens mij over de titel van de sectie en het woord aims in de eerste zin. Misschien moeten we een hardere claim doen hier? --NN
+% Ik heb `aims` vervangen door `targets`. Doet dat wat? --TS
+
 % Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
-Task-oriented programming (\TOP) is a programming paradigm which aims to be the sweet spot between faithful modelling workflows
+Task-oriented programming (\TOP) is a programming paradigm which targets the sweet spot between faithful modelling workflows
 and rapid prototyping of multi-user web applications supporting these workflows \cite{conf/ppdp/PlasmeijerLMAK12}.
@@
 % First, it should allow to specify tasks from real world scenarios.
-Firstly, it should provide primitive building blocks that are useful for high-level descriptions of how users collaborate with each other and with machines.
+Firstly, it should provide primitive building blocks that are useful for high-level descriptions of how users collaborate,
+with each other and with machines.
 These building blocks are: \emph{editors}, \emph{composition}, and \emph{shared data}.
@@
 Users can work together in a number of ways, and this is reflected in \TOP by task compositions.
-There is sequential composition, parallel composition, and choice.
+There is \emph{sequential} composition, \emph{parallel} composition, and \emph{choice}.
 Users need to communicate in order to engage in these forms of collaboration.
@@
 Finally, there is communication with the \emph{outside} world, where information is entered into the system via input events.
-%\fixme{Should this be mentioned earlier?}
 The end points where the outside world interacts with \TOP applications are called editors.
@@
 It has been used to control home thermostats and other home automation applications \cite{koopman2018task}.
-
+%
 Both implementations currently lack formal semantics which are suited to prove properties about tasks.
 
-\subsection{Scope}
+\subsection{Challenges}
 
-In this paper, we set out to develop a formal basis for \TOP, called \TOPHAT (TopHat).
-Our goal is to pave the way for formal treatment of \TOP programs.
+% \ITASKS is the de-facto reference implementation of \TOP.
+Both \ITASKS and \MTASKS have been designed for developing real-world applications.
+They are constantly being extended and improved with this goal in mind.
+The different variations of task combinators and the details that come with real-world requirements,
+make it hard to see what the essence of \TOP is.
+% One of the defining aspects of both \ITASKS and \MTASKS is their embedding in a functional programming language.
+% This creates a synergy where they profit from the expressivity of Clean, while enhancing the language with functionality for user interaction.
+Also, the tight integration of both frameworks with the Clean, makes it difficult to see where the boundaries are.
+This makes formal reasoning about \TOP programs impossible.
 
-\ITASKS is the de-facto reference implementation of \TOP.
-It has been designed for developing real-world applications, and is constantly being extended and improved with this goal in mind.
-All the bells and whistles of \ITASKS, the different variations of task combinators, and the details that come with real-world requirements, make it hard to see what the essence of \TOP is.
-It also makes formal reasoning about \TOP programs impossible.
-
-In this paper, we want to take a step back and look at the essence of \TOP.
+In this paper, we want to take a step back and look at the spirit of \TOP.
 We do this both formally and informally.
 Informally in the sense that we give an intuitive description of the features that define task-oriented programming.
-Formally in the sense that we develop a calculus which formalises these features as language constructs, and we give them semantics in the style that is common in programming language research.
-
-One of the defining aspects of \ITASKS is its embedding in a functional programming language.
-This creates a synergy where \ITASKS profits from the expressivity of Clean, while enhancing the language with functionality for user interaction.
-The tight integration of task DSL and a host language makes it sometimes difficult to see where the boundaries are.
-
-In \TOPHAT, we separate the task layer and the underlying host language, both syntactically and semantically.
-In this way we make explicit which of the synergetic effects of \TOP come from the task layer, and which come from functional programming.
+Formally in the sense that we develop a calculus which formalises these features as language constructs,
+and we give them semantics in the style that is common in programming language research.
+We separate the task layer and the underlying host language, both syntactically and semantically.
+Thus making explicit which properties of \TOP come from the task layer, and which come from functional programming.
+Our challenge, therefore, is to model the properties of \TOP into a calculus
+and, pave the way for formal treatment of \TOP programs.
+We give this formal calculus the name \TOPHAT (TopHat).
 
@@
   \item
-    We present a formal calculus for modelling declarative workflows, embedded in a simply typed lambda-calculus.
-    This calculus is based on the aforementioned essential \TOP concepts.
+    We present a formal calculus for modelling declarative workflows, embedded in a simply typed lambda-calculus, called \TOPHAT.
+    It is based on the aforementioned essential \TOP concepts.
 
   \item
-    We develop an operational semantics that is driven by user input.
+    We develop a layered operational semantics for \TOPHAT that is driven by user input.
     The semantics of the task language is clearly separated from the semantics of the underlying host language.
@@
   \item
-    Along with the semantics, we present the following semantic observations on tasks:
+    Along with this semantics, we present the following semantic observations on tasks:
     the current value, whether a term is stuck, the current user interface, and the accepted inputs.
@@
   \item
-    We prove progress and type preservation for our calculus.
+    We prove progress and type preservation for \TOPHAT.
 
   \item
-    Using both the essential concepts and the formal calculus, we compare \TOP with related work, in areas ranging from business process modelling to process algebras and reactive programming.
+    Using both the essential concepts and the formal calculus, we compare \TOP with related work in areas ranging from business process modelling, to process algebras and reactive programming.
 
@@
   \item
-    %To create executable applications, we implemented \TOPHAT in iTasks.
-
     To create executable applications, we implemented the task layer of \TOPHAT in iTasks.
@@
 In \cref{sec:example} we demonstrate the functionality of \TOPHAT by means of an example,
-\Cref{sec:intuition} gives an overview of the essential concepts of \TOP and \TOPHAT.
+\Cref{sec:intuition} gives an overview of the essential concepts of \TOP.
 \Cref{sec:language} introduces the \TOPHAT calculus syntax

### 2019-02-27 sections/example.tex
small typos in example

@@
 
-In this section we develop a small example program to demonstrate the capabilities of \TOPHAT.
+In this section we develop an example program to demonstrate the capabilities of \TOPHAT.
 The example is a small flight booking system.
@@
     The ages are not filled in, therefore the \TS{Continue} button is disabled.
-    The third user finished a booking.
-    The first user can not pick seats \smallcaps{1b} and \smallcaps{1c} any more.
     The message bubble shows that the \TS{age} field only accepts integer values.
+    The third user finished a booking,
+    therefore, the first user can not pick seats \smallcaps{1b} and \smallcaps{1c} any more.
   }
@@
 Now we develop our workflow in a top down way.
-Our flight booking starts with an interactive task \TS{enter (List Passenger)}.
+Our flight booking starts with an interactive task \TS{enter (List Passenger)}, where users can enter a list of passengers.
 A task $\Enter \tau$ is an empty editor that asks for a value of the given type $\tau$.
-where users can enter a list of passengers.
-A passenger is valid if the name is not empty and the age is at least 0.
-A list of passengers is valid if each passenger is valid, and at least one of the passengers is an adult.
+Passengers are valid if their name is not empty and their age is at least 0.
+Lists of passengers are valid if each passenger is valid, and at least one of the passengers is an adult.
 When the user has entered a valid list of passengers, the step after \TS{>>?} becomes enabled,
@@
 The function \TS{confirmBooking} removes the picked seats from the shared list of free seats,
-and displays the end result using an editor \TS{edit}.
+and displays the end result using an editor, depicted by \TS{edit}.
 \begin{TASK}
@@
 The main task starts three \TS{bookFlight} tasks,
-which could be performed by three different users in parallel
+which could be performed by three different users in parallel.
 \begin{TASK}

### 2019-02-27 sections/intuition.tex
spelling
use more concrete examples in the intuition section
concentrate on the intuition: introduce symbols for tasks only, 
everthing else is introduced in the syntax section, less meta variables, 
no refereces to semantic functions

@@
 This section gives an overview over the core concepts of task-oriented programming.
+
+\subsection{Tasks model collaboration}
+%NOTE: Heb toch maar deze kop toegevoegd, onderstaande gaat er duidelijk over, en de sprong van het voorbeeld naar "Tasks are reusable" is anders abrupt!
 
@@
 
-The following program shows the different collaboration operators in action.
+The following program shows the different collaboration operators in the setting of making a breakfast.
 Users have a choice ($\Xor$) whether they want tea or coffee.
@@
 For example, it is impossible to scramble eggs, and put on the kettle for tea simultaneously.
-Instead, what is meant by parallel is that the order in which we do tasks and the smaller tasks that they are composed of, does not matter.
-\todo{Alejandro has placed a line and exclamation under this sentence. I'm not sure what to do here.}
+Instead, what is meant by parallel is that \emph{the order in which we do tasks and the smaller tasks that they are composed of, does not matter}.
+%XXX: Alejandro has placed a line and exclamation under this sentence. I'm not sure what to do here. --NN
+% I think he thinks it is an essential part of understanding tasks. I emphasised it. --TS
 Then finally, only when every item we want to have for breakfast is ready, can we sit down and enjoy it.
@@
 
-Tasks are modular in the following ways.
+There are three ways in which tasks are modular.
 First, larger tasks are composed from smaller ones.
@@
 
-In \TOPHAT, editors are the basic method of communication with the environment.
+In \TOPHAT, \emph{editors} are the basic method of communication with the environment.
+Editors are modelled after input widgets from graphical user interfaces.
 There are different editors, denoted by different box symbols.
-An editor of the form $\Edit v$, where $v$ is of type $\tau$, reacts to two kinds of events:
-Change events $v$ with a value of type $\tau$ and clear events $\Empty$ which empty the editor.
-This results in a typed empty editor $\Enter \tau$, which is annotated with the same type.
-\todo{Alejandro: unclear} %Nico: He has marked the above paragraph as being unclear.
+Take for example an editor holding the integer seven: $\Edit 7$.
+Such an editor reacts to two kinds of events.
+The first are change events, for example the values $42$ or $37$, which are of the same type.
+The second is a clear event $\Empty$, which empties the editor.
+This results is an empty editor $\Enter \Int$, annotated to only hold integers.
+
+%XXX: Alejandro: unclear
+%XXX: Alejandro: What is a value? What is a type?
+% He has marked the above paragraph as being unclear. --NN
+% Vraag is alleen wat dan precies... :-( Ik heb er een concreet voorbeeld van gemaakt met integers. --TS
 
@@
 As values of editors can be observed, for example by a user interface, editors serve as facilities for both input and output.
-An empty editor $\Enter \tau$ stands for a prompt to input data, while a filled editor can be seen either as outputting a value, or as an input that comes with a default value.
+An empty editor ($\Enter$) stands for a prompt to input data,
+while a filled editor ($\Edit$) can be seen either as outputting a value, or as an input that comes with a default value.
 
@@
 
-\todo{Alejandro: What is a value? What is a type?}
-
 \subsection{Tasks can be observed}
 
-Several observations can be made about tasks.
-One of those is determining the value of a task $\Value(t)$.
-%NOTE: Should we change the type of Value to Maybe v? Both johan and alejandro mention that it is not completely clear to them. --NN
-%NOTE: I don't think that we should use Maybe. We should use the term partial function and bottom. --MKL
-Not all tasks have a value, for example the empty editor $\Enter \tau$, which makes $\Value$ a partial function.
-For example, $\Value(\Edit 5) = 5$, $\Value(\Enter \Int) = \bot$.
+Several observations can be made on tasks.
+One of those is determining the value of a task.
+%NOTE: Should we change the type of Value to Maybe v? Both Johan and Alejandro mention that it is not completely clear to them. --NN
+% I don't think that we should use Maybe. We should use the term partial function and bottom. --MKL
+% I agree! I made the examples here also concrete instead of introducing the value function. --TS
+Not all tasks have a value, the empty editor for instance, which makes value observation partial.
+For example, the value of $\Edit 7$ is $7$, but the value of $\Enter \Int$ is $\bot$.
 
-Another observation is the set of events a task can react to.
-This is important for generating user interfaces.
-For example, the task $\Edit 5$ can react to value events $v$ and empty events $\Empty$.
+Another observation is the set of input events a task can react to.
+% This is important for generating user interfaces.
+For example, the task $\Edit 7$ can react to value events and empty events, as discussed before.
 
-In order to render a user interface, the system needs to observe a task's user interface.
+In order to render a task, we need to observe a task's user interface.
 This is done compositionally.
 User interfaces of combined tasks are composed of the user interfaces of the components.
-For example of two tasks combined with a step combinator, only the left hand side is rendered.
+For example, of two tasks combined with a step combinator, only the left hand side is rendered.
 Two parallel tasks are rendered next to each other.
+Combining this information with the task's value and possible inputs,
+we can display the current state of the task,
+together with buttons that show the actions a user can engage in.
 
-In order for the system to actually render a task, it needs to observe the task's value, possible inputs, and user interface.
-With this information, the system can then display the current state of the task, together with buttons that show the actions a user can engage in.
-
-The final observation is to determine whether a task results in $\Fail$ (failure).
+The final observation is to determine whether a task results in a failure, denoted by $\Fail$.
 The step combinator $\Next$ and the choice combinator $\Xor$ use this to prevent users from picking a failing task.
@@
 
-In a step $t \Next e$, the decision to move on is taken by $\Next$, not by $t$.
+In a step $t \Next e$, the decision to move on from a task $t$ to its continuation $e$ is taken by $\Next$, not by~$t$.
 The decision is based on a speculative evaluation of $e$.
-The step combinators in $t \Then e$ and $t \Next e$ pass the value $v$ of $t$ to the continuation $e$.
-Both steps act like $t$ as long as the step is guarded.
-A step is guarded if either the left task has no value, or the speculative evaluation $e\ v$ yields $\Fail$.
-Once it becomes unguarded, the step continues as $e\ v$.
+The step combinator in $t \Next e$ passes the value $v$ of $t$ to the continuation $e$.
+Steps act like $t$ as long as the step is guarded.
+A step is guarded if either the left task has no value, or the speculative evaluation of $e$ applied to $v$ yields the failure task $\Fail$.
+Once it becomes unguarded, the step continues as the result of $e\ v$.
+Speculative evaluation is designed so that possible side effects are undone.
 The task $t \Next e$ additionally requires a continue event $\Continue$ to proceed.
-Speculative evaluation is designed so that possible side effects are undone.
+% Its cousin $\Then$ makes a step without user interaction.
 
-The step combinators give rise to a form of internal communication:
+Step combinators give rise to a form of internal communication.
 They represent data flow that \emph{follows} control flow.
@@
 Users can not directly interact with shared data, a shared editor is required for that.
-If $r$ is a reference of type $\tau$, then $\Update r$ is an editor whose value is that of $r$.
+If $x$ is a reference of type $\tau$, then $\Update x$ is an editor whose value is that of $x$.
 
@@
 The tasks are truly independent of each other, all interleavings are permitted.
-The environment must prefix events to $t_1$ and $t_2$ by $\First$ (first) and $\Second$ (second) respectively.
+The environment must prefix events to $t_1$ and $t_2$ respectively by $\First$ (first) and $\Second$ (second).
 This unambiguously renames the inputs, removing any source of nondeterminism.
@@
 \TOPHAT is meant solely for implementation and does not have any form of nondeterminism.
-Input events for parallel tasks are disambiguated, internal steps have a well-defined evaluation order, and internal choice is left-biased.
+Input events for parallel tasks are disambiguated, internal steps ($\Then$) have a well-defined evaluation order, and internal choice ($\Or$) is left-biased.
 

### 2019-02-27 sections/example.tex
typo in example

@@
 The function \TS{confirmBooking} removes the picked seats from the shared list of free seats,
-and displays the end result using an editor, depicted by \TS{edit}.
+and displays the end result using an editor, denoted by \TS{edit}.
 \begin{TASK}

### 2019-02-27 sections/language.tex
add text about typing rules and use beta for references

@@
 The typing rules for pretasks are given in \cref{fig:typing-rules}.
-\todo{Say someting about why we need beta, and use beta in the acutal typing rule}
-\todo{Alejandro: Is there anything surprising? Say what}
+Most typing rules lift the type of their subexpressions into the $\Task$-type.
+The typing rules for steps make sure the continuations $e_2$ are functions which accept a well-typed value from the left hand side (\refrule{T-Then}, \refrule{T-Next}).
+References, and therefore shared editors, can only be of a basic type so they do not introduce implicit recursion (\refrule{T-Update}).
 

### 2019-02-27 sections/intuition.tex
fit in more

@@
 Input events drive evaluation of tasks.
-When the system receives a valid event, it applies the event to the current task, which results in a new task.
+When the system receives a valid event, it applies this event to the current task, which results in a new task.
 In this way the system communicates with the environment.
-Input events are synchronous.
+Inputs are synchronous.
 We assume systems react instantaneous to each such event.
@@
 There are three smokers, each having one of the ingredients and requiring the other two.
-There is an agent that randomly provides two of the ingredients.
+There is an agent that randomly provides two of those.
 The difficulty lies in the requirement that only the smoker may proceed whose missing ingredients are present.
@@
 Downey models availability of the ingredients with a semaphore for each ingredient.
-The agent randomly signals two of the three semaphores.
+The agent randomly signals two of the three.
 The solution proposed by Downey involves an additional mutex, three additional semaphores, three additional threads called \emph{pushers}, and three regular Boolean variables.
@@
 \TOPHAT allows a simple solution to this problem, using guarded steps.
-Steps can be guarded with arbitrary expressions, and the parallel combinator can be used to watch two shared editors at the same time.
+Steps can be guarded with arbitrary expressions.
+The parallel combinator can be used to watch two shared editors at the same time.
 Let \TS{match}, \TS{paper}, and \TS{tobacco} be references to Booleans.
 The smokers are defined as follows.
+% The problem is then specified as follows.
 \begin{TASK}

### 2019-02-27 sections/language.tex
fit in more

@@
 \end{TASK}
-Initially the step is guarded because the editor doesn't have a value.
-When the user enters an integer, the program continues immediately with either \TS|edit "Good"| or \TS|edit "Bad"|, depending on the input.
+Initially, the step is guarded because the editor does not have a value.
+When users enter an integer, the program continues immediately with either \TS{edit "Good"} or \TS{edit "Bad"}, depending on the input.
 
@@
 An external choice requires users to pick a branch before continuing with it.
-This means users cannot work on the branches of an external choice before picking one. \todo{Alejandro: Is this actually implied?}
+This means users cannot work on the branches of an external choice before picking one.
+%XXX: Alejandro: Is this actually implied?
+% Ik snap niet zo goed wat Alejandro hier mee bedoelt... --TS
 

### 2019-02-27 sections/semantics.tex
add some text to describe rules as tipped by Alejandro

--- 2019-02-21 sections/semantics.tex
@@
 
+%XXX: Alejandro: In the formalisation you show quite some figures.
+I think that for each figure you need to be more explicit in the text about what is unsurprising and what is new or should get the attention of the reader.
+% Geen tijd meer voor denk ik. --TS
+
 In this section we formalise the semantics of the language constructs described in \cref{sec:language}.
 We organise this by following the structure of the language.
+Firstly, the task language is embedded in a simply typed lambda calculus.
+This requires a specification of the \emph{evaluation} of terms in the host language, and how it handles the task language.
+Secondly, there are two ways to drive evaluation of task expressions, internally by the system itself, and externally by the user.
+This is done in two additional semantics, one for the internal \emph{normalisation} of tasks, and another for the \emph{interaction} with the end user.
 
-Firstly, the task language is embedded in a simply typed lambda calculus.
-This requires a specification of the evaluation of terms in the host language, and how it handles the task language.
-
-Secondly, there are two ways to drive evaluation of task expressions, internally by the system itself, and externally by the user.
-This is done in two additional semantics, one for the internal normalisation of tasks, and another for the interaction with the end user.
-
-One of our explicit goals is to keep the semantics for evaluation and normalisation separate,
-to not mix general purpose programming notions with workflow specific semantics.
-This is achieved by letting tasks be values in the host language.
-
-The three main layers of semantics are thus \emph{evaluation}, \emph{normalisation}, and \emph{interaction}.
+The three main layers of semantics are thus evaluation, normalisation, and interaction.
 The semantics, together with \emph{observations}, will be discussed in the following subsections.
-\Cref{fig:semantic-functions} shows the relation between the semantics.
+\Cref{fig:semantic-functions} shows the relation between all semantics arrows.
 It also shows that there are two helper semantics, \emph{handle} and \emph{stride}.
+We use the convention that downward arrows are big-step semantics, and rightward arrows are small-step semantics.
 
@@
 
-We use the convention that downward arrows are big-step semantics, and rightward arrows are small-step semantics.
+One of our explicit goals is to keep the semantics for evaluation and normalisation separate,
+to not mix general purpose programming notions with workflow specific semantics.
+This is achieved by letting tasks be values in the host language.
 
@@
 
+%XXX: Alejandro over tasks: I was not expecting these to be values...
+% Ik heb een verwijzing gemaakt hier naartoe, hoop dat het geen groot probleem is... --TS
+
 The rules to evaluate expressions $e$ do not differ from standard work, except for the task constructs.
@@
 They are given in the appendix.
+Most task constructors are strict in their arguments.
+Only steps keep their left hand side unevaluated to delay side effects till the moment the step is taken.
+The same holds for both branches of the external choice.
 
@@
 This expression is equivalent to $\Fail$, because it can not handle input and can not be further normalised.
-
+%
 This intuition is formalised by the function $\Failing$ in \cref{fig:observation-failing}.
@@
 \end{TASK}
+\pagebreak[2]
 \refrule{S-And} reduces this expression in one step to
@@
 To overcome this problem, the \refrule{N-Done} and \refrule{N-Repeat} rules ensure that striding is applied until the state $s$ becomes stable and no further normalisation can take place.
+%XXX: Alejandro: The example is not 100\% clear to me
 
@@
 On top of that, $t_2$ must not be failing: $\neg\Failing(t_2)$.
-
+%
 These principles lead to the stepping rules in \cref{fig:striding-semantics}.

### 2019-02-28 sections/abstract.tex
small typos in example
Minor nits fixed in abstract

@@
 % Inquiry
-TOP has been around for more than a decade, in the forms of iTasks and mTasks, which is tailored for real-world usability.
+TOP has been around for more than a decade, in the forms of iTasks and mTasks, which are tailored for real-world usability.
 So far, it has not been given a formalisation which is suitable for formal reasoning.
@@
 We describe TopHat by means of a layered semantics.
-These layers consist of multiple big step evaluations on expressions, and two labelled transition systems, handling user inputs.
+These layers consist of multiple big-step evaluations on expressions, and two labelled transition systems, handling user inputs.
 
@@
 % Grounding
-We have implemented the semantic rules of TopHat in Idris, and the task layer on top of iTasks combinators.
+We have implemented the semantic rules of TopHat in Idris, and the task layer on top of the iTasks framework.
 This shows that our approach is feasible, and lets us demonstrate the concepts by means of illustrative case studies.
 % Importance
-TOP has been employed in projects with the Dutch coast guard, tax office, and navy.
+TOP has been appplied in projects with the Dutch coast guard, tax office, and navy.
 Our work matters because formal program verification is important for mission-critical software, especially for systems with concurrency.

### 2019-02-28 sections/introduction.tex
Fixed nits in introduction

@@
 \TOP focusses on modelling collaboration patterns.
-This gives rise to user's need to interact and share information.
-Next to that, \TOP automatically provides solutions to common development jobs like designing \GUI\ s, connecting to databases, and communicating between servers and clients.
+This gives rise to a user's need to interact and share information.
+Next to that, \TOP automatically provides solutions to common development jobs like designing \GUI{s}, connecting to databases, and communicating between servers and clients.
 
@@
 % This creates a synergy where they profit from the expressivity of Clean, while enhancing the language with functionality for user interaction.
-Also, the tight integration of both frameworks with the Clean, makes it difficult to see where the boundaries are.
+Also, the tight integration of both frameworks with Clean, makes it difficult to see where the boundaries are.
 This makes formal reasoning about \TOP programs impossible.
@@
 Formally in the sense that we develop a calculus which formalises these features as language constructs,
-and we give them semantics in the style that is common in programming language research.
+and we give them a semantics in the style that is common in programming language research.
 We separate the task layer and the underlying host language, both syntactically and semantically.
@@
 Our challenge, therefore, is to model the properties of \TOP into a calculus
-and, pave the way for formal treatment of \TOP programs.
+and pave the way for formal treatment of \TOP programs.
 We give this formal calculus the name \TOPHAT (TopHat).

### 2019-03-01 sections/example.tex
fixed minor nits in example section

@@
 
-Now we develop our workflow in a top down way.
+Now we develop our workflow in a top-down manner.
 Our flight booking starts with an interactive task \TS{enter (List Passenger)}, where users can enter a list of passengers.
@@
 
-The function \TS{confirmBooking} removes the picked seats from the shared list of free seats,
+The function \TS{confirmBooking} removes the selected seats from the shared list of free seats,
 and displays the end result using an editor, denoted by \TS{edit}.

### 2019-03-01 sections/intuition.tex
Various fixes in intuition section

@@
 
-This section gives an overview over the core concepts of task-oriented programming.
+This section gives an overview of the core concepts of task-oriented programming.
 
@@
 
-The following program shows the different collaboration operators in the setting of making a breakfast.
+The following program shows the different collaboration operators in the setting of making breakfast.
 Users have a choice ($\Xor$) whether they want tea or coffee.
@@
 The drink and the food are prepared in parallel ($\And$).
-When drink and food are prepared, users can step ($\Next$) to eating the results.
+When both the drink and the food are prepared, users can step ($\Next$) to eating the result.
 
@@
 There are three ways in which tasks are modular.
-First, larger tasks are composed from smaller ones.
+First, larger tasks are composed of smaller ones.
 Second, tasks are first-class, they can be arguments and results of functions.
@@
 Inputs are synchronous.
-We assume systems react instantaneous to each such event.
+We assume systems react instantaneously to each if those events.
 
@@
 There are no output events.
-As values of editors can be observed, for example by a user interface, editors serve as facilities for both input and output.
+As values of editors can be observed, for example by a user interface, editors facilitate both input and output.
 An empty editor ($\Enter$) stands for a prompt to input data,
@@
   let paperSmoker = (update tobacco <&> update match) >>? continue in
-  let matchSmoker = (update tobacco <&> update paper) >>? continue
+  let matchSmoker = (update tobacco <&> update paper) >>? continue in
+  tobaccoSmoker <&> paperSmoker <&> matchSmoker
 \end{TASK}

### 2019-03-01 sections/language.tex
fixed minor nits in calculus section

@@
 Pretasks are tasks that have unevaluated subexpressions with respect to the host language.
-Who expressions are evaluated will be discussed in \cref{sec:evaluation}.
+How expressions are evaluated will be discussed in \cref{sec:evaluation}.
 Each pretask will be discussed in more detail in the following subsections.
@@
 Empty editors are annotated with a type in order to ensure type safety.
-Without that it would be possible to change the type of an editor, by clearing it and filling it with a value of different type.
+Without type annotations, it would be possible to change the type of an editor, by clearing it and filling it with a value of different type.
 
@@
 Changes to a shared editor are immediately visible to all shared editors watching the same reference.
-Imagine two users, Marco and Christopher, both watching shared editors of the same coordinate.
+Imagine two users, Marco and Christopher, both watching shared editors of the same coordinates.
 The editors are visualised as a pin on a map.

### 2019-03-25 sections/properties.tex
push fixmes in margin for last proofread
remove last note by Alejandro
rename state s to sigma everywhere (to be compatible with next paper)

@@
 
-Additionally, we show that the normalisation semantics is a big-step semantics. \fixme{Alejandro: ?}
+Additionally, we show that the normalisation semantics is a big-step semantics.
+%XXX: Alejandro: ?
+% Niks meer mee doen, was voor PLDI reviewers duidelijk genoeg. --TS
 While at first sight, this might seem obvious, the fact that we are dealing with
@@
 \begin{theorem}[Type preservation under evaluation]
-  % For all well typed expressions $e$ and states $s$,
-  For all expressions $e$ and states $s$
-  such that $\Gamma,\Sigma \infers e:\tau$ and $\Gamma,\Sigma \infers s$,
-  if $e,s \evaluate e',s'$,
-  then $\Gamma,\Sigma \infers e':\tau$ and $\Sigma \infers s'$.
+  % For all well typed expressions $e$ and states $\sigma$,
+  For all expressions $e$ and states $\sigma$
+  such that $\Gamma,\Sigma \infers e:\tau$ and $\Gamma,\Sigma \infers \sigma$,
+  if $e,\sigma \evaluate e',\sigma'$,
+  then $\Gamma,\Sigma \infers e':\tau$ and $\Sigma \infers \sigma'$.
   \label{thm:pres-eval}
@@
 
-Where $\Sigma \infers s$ means that for all $l\in s$, it holds that
-$\Sigma(l)=\Gamma,\Sigma\infers s(l)$.
+Where $\Sigma \infers \sigma$ means that for all $l\in \sigma$, it holds that
+$\Sigma(l)=\Gamma,\Sigma\infers \sigma(l)$.
 
@@
 \begin{theorem}[Type preservation under normalisation]
-  % For all well typed expressions $e$ and states $s$,
-  For all expressions $e$ and states $s$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
-  if $e,s \normalise e',s'$,
-  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers s'$.
+  % For all well typed expressions $e$ and states $\sigma$,
+  For all expressions $e$ and states $\sigma$
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  if $e,\sigma \normalise e',\sigma'$,
+  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers \sigma'$.
   \label{thm:pres-norm}
@@
 \begin{lemma}[Task value preserves types]
-  % For all well typed expressions $e$ and states $s$,
-  For all expressions $e$ and states $s$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
-  if $\Value{(e,s)}=v$,
+  % For all well typed expressions $e$ and states $\sigma$,
+  For all expressions $e$ and states $\sigma$
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  if $\Value{(e,\sigma)}=v$,
   then $v:\tau$.
@@
 \begin{lemma}[Striding preserves types]
-  For all expressions $e$ and states $s$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
-  if $e,s \stride e',s'$,
-  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers s'$.
+  For all expressions $e$ and states $\sigma$
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  if $e,\sigma \stride e',\sigma'$,
+  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers \sigma'$.
   \label{thm:pres-stride}
@@
 \begin{theorem}[Type preservation under handling]
-  % For all well typed expressions $e$, states $s$, and inputs $i$,
-  For all expressions $e$, states $s$ and inputs $i$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
-  if $ e,s \handle{i} e',s'$,
-  then $\Gamma,\Sigma\infers e':\Task\tau$ and $\Sigma\infers s'$.
+  % For all well typed expressions $e$, states $\sigma$, and inputs $i$,
+  For all expressions $e$, states $\sigma$ and inputs $i$
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  if $ e,\sigma \handle{i} e',\sigma'$,
+  then $\Gamma,\Sigma\infers e':\Task\tau$ and $\Sigma\infers \sigma'$.
   \label{thm:pres-handle}
@@
 \begin{theorem}[Progress under handling]
-  For all well typed expressions $e$ and states $s$,
-  % For all expressions $e$ and states $s$
-  % such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
-  if $e,s \normalise e',s'$,
-  then either $\Failing(e', s')$
-  or there exist $e''$, $s''$, and $i$ such that $e',s'\handle{i} e'',s''$.
+  For all well typed expressions $e$ and states $\sigma$,
+  % For all expressions $e$ and states $\sigma$
+  % such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  if $e,\sigma \normalise e',\sigma'$,
+  then either $\Failing(e', \sigma')$
+  or there exist $e''$, $\sigma''$, and $i$ such that $e',\sigma'\handle{i} e'',\sigma''$.
   \label{thm:prog-norm}
@@
 Where a well typed expression $e$ means that $\Gamma,\Sigma\infers e:\tau$ for
-some type $\tau$, and a well typed state means that $\Sigma\infers s$.
+some type $\tau$, and a well typed state means that $\Sigma\infers \sigma$.
 
-If an expression $e$ and state $s$ are well-typed, then after normalisation, the pair $e',s'$
+If an expression $e$ and state $\sigma$ are well-typed, then after normalisation, the pair $e',\sigma'$
 either fails, or there exists some input $i$ that can be handled by it under the handling semantics.
@@
 \begin{theorem}[Failing means no interaction possible]
-  For all well typed expressions $e$ and states $s$,
-  % For all expressions $e$ and states $s$
-  % such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
-  and $e,s \normalise e',s'$,
-  we have that $\Failing(e',s') = \True$,
+  For all well typed expressions $e$ and states $\sigma$,
+  % For all expressions $e$ and states $\sigma$
+  % such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  and $e,\sigma \normalise e',\sigma'$,
+  we have that $\Failing(e',\sigma') = \True$,
   if and only if there is no input $i$
-  such that $e',s'\handle{i} e'',s''$ for some $e''$ and $s''$.
+  such that $e',\sigma'\handle{i} e'',\sigma''$ for some $e''$ and $\sigma''$.
   \label{thm:failing}
@@
 
-The Theorem above states that an expression $e$ and state $s$ are failing, if,
+The Theorem above states that an expression $e$ and state $\sigma$ are failing, if,
 after normalisation, there exists no input that can be handled by it.
@@
 \begin{proof}
-  Given $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$ and after
-  normalisation $e,s \normalise e',s'$, we find ourselves in either one of the
+  Given $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers \sigma$ and after
+  normalisation $e,\sigma \normalise e',\sigma'$, we find ourselves in either one of the
   following situations:
 
-  There exists an $i$ such that $e',s'\handle{i} e'',s''$.
+  There exists an $i$ such that $e',\sigma'\handle{i} e'',\sigma''$.
 
-  There does not exist an $i$ such that $e',s'\handle{i} e'',s''$. In this case, we
-  know that $\Failing(e',s')$ must be true, by \cref{thm:failing}.
+  There does not exist an $i$ such that $e',\sigma'\handle{i} e'',\sigma''$. In this case, we
+  know that $\Failing(e',\sigma')$ must be true, by \cref{thm:failing}.
 \end{proof}
@@
 \begin{theorem}[Inputs function is sound and complete]
-  For all well typed expressions $e$, states $s$, and inputs $i$,
-  % For all expressions $e$, states $s$, and inputs $i$
-  % such that $\Gamma,\Sigma \infers e:\tau$ and $\Sigma \infers s$,
-  we have that $i \in \Inputs{(e,s)}$ if and only if $e,s \handle{i} e',s'$.
+  For all well typed expressions $e$, states $\sigma$, and inputs $i$,
+  % For all expressions $e$, states $\sigma$, and inputs $i$
+  % such that $\Gamma,\Sigma \infers e:\tau$ and $\Sigma \infers \sigma$,
+  we have that $i \in \Inputs{(e,\sigma)}$ if and only if $e,\sigma \handle{i} e',\sigma'$.
   \label{thm:safety-i}

### 2019-03-25 sections/semantics.tex
rename state s to sigma everywhere (to be compatible with next paper)

@@
 The host language evaluates expressions using a big-step semantics.
-Evaluating an expression $e$ in state $s$ into a value $v$ in state $s'$ is denoted by $\RelationE$.
+Evaluating an expression $e$ in state $\sigma$ into a value $v$ in state $\sigma'$ is denoted by $\RelationE$.
 To ease reasoning about references, we choose a call-by-value evaluation strategy.
@@
 We write $\RelationN$ to describe that
-an expression $e$ in state $s$ normalises to task $t$ in state $s'$.
+an expression $e$ in state $\sigma$ normalises to task $t$ in state $\sigma'$.
 
@@
 The split between striding and normalisation is due to mutable references.
-Consider the following example, where $s=\set{l\mapsto\False}$.
+Consider the following example, where $\sigma=\set{l\mapsto\False}$.
 \begin{TASK}
@@
 \end{TASK}
-with $s'=\set{l\mapsto\True}$.
+with $\sigma'=\set{l\mapsto\True}$.
 This expression is not normalised, because the left task can take a step.
 The issue here lies in the fact that the right task updates $l$.
-To overcome this problem, the \refrule{N-Done} and \refrule{N-Repeat} rules ensure that striding is applied until the state $s$ becomes stable and no further normalisation can take place.
+To overcome this problem, the \refrule{N-Done} and \refrule{N-Repeat} rules ensure that striding is applied until the state $\sigma$ becomes stable and no further normalisation can take place.
 %XXX: Alejandro: The example is not 100\% clear to me
@@
 The striding semantics is responsible for reducing internal steps and internal choices.
-A stride from task $t$ in state $s$ to $t'$ in state $s'$ is denoted by $\RelationS$.
+A stride from task $t$ in state $\sigma$ to $t'$ in state $\sigma'$ is denoted by $\RelationS$.
 The rules for striding are given in \cref{fig:striding-semantics}.
@@
 It is a small step semantics with labelled transitions.
-It takes a task $t$ in a state $s$ and an input $i$, and yields a new task $t'$ in a new state $s'$.
+It takes a task $t$ in a state $\sigma$ and an input $i$, and yields a new task $t'$ in a new state $\sigma'$.
 

### 2019-04-18 reviews/icfp-before-rebuttal.txt
add icfp reviews

ICFP 2019 Paper #111 Reviews and Comments
===========================================================================
Paper #111 TopHat: A formal foundation for task-oriented programming

Review #111A
===========================================================================

Overall merit
-------------
B. OK paper, but I will not champion it.

Reviewer expertise
------------------
Y. I am knowledgeable in the area, though not an expert.

Paper summary
-------------
Interactive workflow "tasks" are a common use of computing
resources. This paper provides a middleweight calculus modeling key
ideas of such interactions: a set of combinators specifies the inputs,
computations, steps, and orderings in such tasks.

The model is a subset of features of existing implementations iTasks
and mTasks, though there's no formal relationship.

The semantics are proven type safe and a small model is implemented in
Idris (though no detail is given on this mechanized model).

The paper concludes with a substantial comparison to related work.

Comments for author
-------------------
The work here is a nice semantics on an important topic that has
received too little attention in the PL community. My score is a B for
several reasons:

- Nothing is _done_ with the semantics beyond sanity checking. Such
  type safety proofs show only that you've built a sensible
  system. What is the semantics for?

- The papers aims and claims don't match up with its actual
  contributions.

  While "formal program verification is important for mission-critical
  software", no actual formal verification happens here---and there's
  no evidence that the workflows described here are mission critical.

  The "_____-oriented programming" framing leads one to compare the
  system to other orientations, which may be unfavorable either
  because other orientations are total views of what programming is
  (and so more substantial ideas than what any single semantics might
  offer) or because other orientations are seen as software
  engineering fads.

- The semantics is entirely operational, which offers the weakest
  reasoning principles as compared with denotational or axiomatic
  models. The semantics and combinators also have gaps, allowing one
  to express senseless workflows.

Despite these flaws, I think the combinators given here will be useful
elsewhere, so I lean towards acceptance.

# What might improve the paper?

- Moderate the claims. The key novelty here is an interesting library
  of combinators in an important domain and a nice semantics for a
  non-recursive fragment. Your properties don't quite "validate" your
  semantics (line 838), but rather serve as a sanity check.

- Use the semantics for something. An Iris-like logic would more than
  suffice for proofs of equivalence of programs. Why not prove that
  one can always eventually eat breakfast? (Such a property is highly
  desirable and sure to be popular.)

  In the related work, you frequently remark that other systems don't
  support formal reasoning---such statements would be more sensible if
  you used your semantics for formal reasoning.

- Summarize the interaction trees of tasks. Throughout the paper, it's
  stressed that TopHat merely specifies collaboration rather than
  explicit communication, as in, e.g., process calculi or session
  types. While such a system is perhaps nice to program in, having a
  term `e : Task t` doesn't tell me much of anything about the set of
  input events `e` can receive. Regular expressions (with embedded
  functional programs to enable transitions) should suffice here, and
  will make for a nice, abstract denotational semantics of tasks.

- Prove that your combinators are orthogonal. Pierce's spider calculus
  (2010) is a nice example: each primitive is proved 'independent' of
  the others.

- Clarify which kinds of tasks are realizable. There's no monadic
  "join" operation that allows us to work with `Task (Task t)`, yet
  the type is expressible. The set of editors is also surely limited:
  what can one do with a $\square (\lambda x:\textsc{Int}. x)$? What
  about $\blacksquare (\textsc{Int}\rightarrow\textsc{Int})$? What
  about $\square (\square 5)$? What about $\blacksquare (\textsc{Task}
  ~ \textsc{Int})$?

  Relatedly, tasks are values in the host language (line 591). What
  eliminators on tasks do you offer?

- You are able to write the smoker program simply because your
  semantics eats all the complexity. What would dining philosophers
  look like in your system? It seems you cannot write a version of it
  that dynamically determines the number of philosophers. What might
  you need to add to the language to support that?

  Later on (line 472) you talk about "immediately reflected" changes,
  but consistent, available (i.e., immediate), and persistent
  propagation is impossible in a distributed system. How should we
  think about the possibility of network delay or failure?
  Communication-based systems force programmers to think about these
  issues, for better or worse.

  Similarly, your comparison to HipHop misses a crucial fact: HipHop's
  Esterel-like semantics has no data races, while TopHat is full of
  data races. While not themselves a problem, there's strong anecdotal
  evidence that racy models are harder to program in.

- Clarify your implementations. What does the Idris implementation do?
  Is it actually executable, or is it a relational model? How does
  your Idris implementation compare to your embedding in iTasks? Have
  you tried testing whether the two semantics behave the same?

- Clarify your relation to FRP. You say that TopHat has no notion of
  time, but $\blacksquare \mathsf{time}$ seems to exactly be a
  time-varying value. It seems that one could use $\blacksquare$ and
  $\RHD$ to write FRP-ish programs, e.g., your derived `delay`
  combinator.

# Questions

You say you "show that the normalisation semantics is a big-step
semantics". Which property is that, and why is it important? Why does
having state complicate a big-step semantics?

Could you encode BPEL in TopHat? If not, which features does BPEL
support that you do not? (E.g., cycles?)

# Minor stuff

- Line 235: What is an invalid event?

- Line 269: What about explicit option types?

- Line 349: We haven't seen these internal operators yet. It might be
  good to at least name them all up front. Could you extend the
  airline example to include them?

- Line 403: You call $\bowtie$ "combination" but perhaps 'parallel
  composition' or 'interleaving' would be more helpful.

- Line 657: It may be worth clarifying that $\rhd$ doesn't need to
  look at $e_2$ because a `C` event is needed. Why doesn't
  $\mathcal{F}$ look at $e_2$ for $\RHD$, though, which needs no
  continue event? Isn't $\blacksquare 5 \RHD \Lightning$ failing?

- Line 793: The typing side conditions on operational are a strange
  violation of the phase distinction.

- Line 949: If iTasks "implicitly has a formal semantics", are there
  any languages that _don't_?

- A few typos: focussing, 'mad an implementation', moddeling

Review #111B
===========================================================================

Overall merit
-------------
B. OK paper, but I will not champion it.

Reviewer expertise
------------------
Z. I am not an expert. My evaluation is that of an informed outsider.

Paper summary
-------------
The paper presents a formalization of the essence of an existing "real-world task" workflow modelling language and methodology called TOP

Comments for author
-------------------
I am an outsider for this particular area of programming languages and was not familiar with TOP.

The paper is well written and interesting, especially for those who know and use TOP and need a detailed description of its formal semantics and some of its properties.

If I have misgivings, I think it is about TOP itself - the question is not its soundness, or its semantics, but its utility for the business domain and the "logical encoding" approach used to formalizing real-world business systems.  My own experience has been with systems such as Oracle Policy Automation which approaches modelling of real-world processes in a vastly different way.  So the sea of symbols as an "encoding" of the real world is imply daunting and unfamiliar to me, and I am skeptical of its utility.  However, since the focus of the paper is on formalization and not on TOP itself, this is me just admitting a (perhaps unfair) bias.

I've given my review a low confidence rating as a result.

Review #111C
===========================================================================

Overall merit
-------------
C. Weak paper, though I will not fight strongly against it.

Reviewer expertise
------------------
Y. I am knowledgeable in the area, though not an expert.

Paper summary
-------------
The paper presents TopHat, an imperative higher-order language with facilities
for interacting with an external environment.

In typical use cases, the environment can be thought of as a user interacting
with the program through a graphical interface. TopHat allows a program to
communicate with the environment via "tasks", which are logical descriptions of
input/output interfaces. Compound tasks can be formed via dedicated combinators:
pairing, monadic-bind-like sequential composition, choice, etc.

Technically, the paper contains an introduction to "task-oriented" programming
in TopHat, a formal semantics for TopHat, a type safety result, and a discussion
of related work.

Comments for author
-------------------

I welcome every effort to make UI programming easier, given the sorry state of
current programming practice. As such, I am largely sympathetic towards the aims
of "task-oriented programming", and find this paper intriguing. Unfortunately,
I'm not convinced that it would be a good fit for ICFP in its current form.

First, I don't think the paper contains enough technical contributions. I
understand that its goal is to give a standalone formalization of task-oriented
programming, which has been around for some time but always in the form of
libraries. But as is, this formalization does not pay off: all applications are
left for future work, which I find highly unsatisfactory. Apparently, we did not
learn anything new about the equivalence of task-oriented programs, nor gain
insight into their implementation, etc. Such points are left for future work,
but would greatly improve the relevance of the work.

Second, I am doubtful of certain design aspects of TopHat.

For example, the language mixes a form of shared memory with bespoke concurrency
constructs, which has several important consequences. In particular, sequential
composition performs speculative execution: if t is a task with a set value, the
sequential composition combinator t ▷ e tries to execute e with this value, and
backtracks if e fails ("failing" is an explicit concept in TopHat),
reestablishing memory as if e had never been executed. This makes composition
very expressive (see e.g. the example given in §3.6, p. 7), but also very
costly, as, in the worst case, the entire store must be duplicated. But I am
even more worried about the impact of this semantics on reasoning: can one
really understand what's going on in a TopHat program which is not very small?
Doesn't this make it hard to understand the complexity of TopHat programs?

Also, while TopHat seems to be a language dedicated to reactive programming,
i.e., continuous interaction with an environment, it does not help with managing
this interaction in a modular way. As far as I see, there is no way for a task A
to interpose itself between the environment and another task B, i.e., for A to
encapsulate B in a way that is transparent from B's perspective. This is
possible in the reactive languages I know of (e.g., Esterel).

Third, in my opinion the paper is poorly written and hard to understand. It
contains too many forward references and makes use of symbols or notions before
they are defined (see below). I also found several statements hard to interpret,
e.g. the task pairing combinator supposedly "introduces interleaving
concurrency" (p. 7), but as far as I see the semantics of TopHat is perfectly
deterministic (given a fixed input sequence) and does not involve interleaved
access to the store. My confusion was exacerbated by the related work section,
where TopHat is compared in vague terms to works whose relevance is not entirely
obvious to me (e.g., process algebra) but where closely-related work from the
ICFP community, such as Carlsson and Hallgren's Fudgets or Conal Elliott's
Tangible Values, is not mentioned at all. Similarly, the comparison to reactive
programming works are not technical enough in my opinion.

## Detailed comments

p. 4

- The "lightning strike" symbol used in the examples should be introduced in the
  text as soon as it is used.

- The "bookFlight ⋈ bookFlight ⋈ bookFlight" example is very problematic. At
  this point, I had zero understanding of TopHat's concurrency model, and thus
  this program looked like it had a data race, making it plainly incorrect. As a
  consequence, I found the paragraphs extolling the virtues of shared-memory
  programming flabbergasting! This should be addressed immediately in the text,
  explaining what is atomic and what isn't.

p. 5

- While discussing the semantics of parallel composition (⋈), it is stated that
  "what is meant by parallel is that the order in which we do tasks and the
  smaller tasks that they are composed of, does not matter". But I cannot
  reconcile this statement with the formal semantics: ⋈ is not commutative at
  all, since for instance "l := True ⋈ l := False" will always result in a store
  containing l ↦ True, and "l := False ⋈ l := True" in one containing l ↦ False.

- "Inputs are synchronous."

  I believe that only a single task may evolve per time step, judging from the
  grammar of inputs presented in Fig 14. So what does this mean, exaclty? That
  program and environment alternate?

p. 6

- I would have liked an intuitive description of the graphical representation of
  choice combinators.

p. 7

- "The details of the solution [Downey's] are not important here."

  Then perhaps one could simplify the preceding paragraph?

p. 8

- Internal choice is mentioned here, without ever being introduced before!

p. 9

- The grammar allows function-valued editors. What does ⊠(Int → Int) look like?

- The grammar allows higher-order tasks and editors. What does ⊠⊠Int look like?

p. 10

- "If e is a reference of type Ref τ, then ▪e is of type Task τ"

  Here τ should be changed into β, or T-Update should be modified.

- The rationale for the external vs. internal distinction is never given. I
  would expect that in an expressive language, external combinators would be
  definable in terms of their internal counterparts.

p. 11

- I liked example 4.4.

p. 13

- "When both branches have a value, it takes the value of the left branch"

  Why not give some combination function?

p. 14

- In Figure 10, F is a function only if ⇓ is deterministic. I believe it is, but
  perhaps this should be stated explicitly somewhere.

p. 15

- Isn't rule S-Xor lacking some premises?

p. 17

- l. 832: obfuscating the link to the public GitHub repository is worse than
  useless. Why not attach an anonymized copy of the source code to the paper?

p. 18

- "we mad": we made.

- "Also, we mad[e] an implementation of TopHat combinators into iTasks. This
  shows that TopHat is a subset of iTasks."

  I'm nitpicking, but I don't like that turn of phrase. Without additional
  information about the simplicity of this implementation, this conclusion is
  incorrect. Otherwise, the fact that you can implement, say, Clean on top of
  x86 assembly would imply that Clean is a subset of x86 assembly.

- l. 926: "understand" -> understanding.

- "We do not know yet if the more complex combinators of iTasks are expressible
  in the basic combinators of TopHat."

  I would have liked to know what these combinators look like.

p. 20

- "However, on IOT devices it is useful to continue running tasks endlessly,
  which is done in mTasks using a "forever" combinator. This is currently not
  possible in TopHat."

  So in TopHat tasks do not "run endlessly", yet they "never terminate" and
  "always keep reacting to events" (p. 6)? This is very confusing.

p. 21

- As I said above, I'm not sure TopHat, a deterministic functional language,
  compares to process algebra.

- l. 996-999: this paragraph is extremely vague ("model collaboration patterns",
  "declarative specification"). Moreover, I don't think comparing "goals" of
  formalisms leads to useful distinctions, otherwise we would conclude that
  functional languages (whose goals are to program computers) and Alonzo
  Church's λ-calculus (whose original goal was to give a foundation for
  mathematics) have nothing to do with each other.

- l. 1015: "syntonisation" -> synchronisation.

- l. 1025: "programatically" -> "programmatically".

- In my opinion, comparing TopHat to ReactiveML would be more relevant, since a
  lot of the HipHop work is about coping with the vagaries of the web
  programming environment.

p. 22

- "FRP and TOP are different systems and have different goals in mind."

  Same remarks as above.

- "TOP has no notion of time" Really? If so, why talk about synchrony and the
  instantaneous character of reactions earlier?

p. 23

- l. 1110: "moddeling" -> "modeling".

Review #111D
===========================================================================

Overall merit
-------------
C. Weak paper, though I will not fight strongly against it.

Reviewer expertise
------------------
Y. I am knowledgeable in the area, though not an expert.

Paper summary
-------------
The paper a formal operational model for a core task-oriented language
inspired by iTasks. The model has a stratified structure. At the bottom
layer there is a base functional typed language with first-class
cells. On top of that there is a language of typed task-related effects.
The semantics of the model is also stratified. At the bottom layer there
is a small-step operational semantics for the base language and on top of
that there are two layers for the effect layer. The bottom of these two
layers deals with task reductions that do not require user input. The
top deals with task reductions that are triggered by user input.
The model is proven type sound and has been implemented in Idris and the
task effect language on top of iTasks.

Comments for author
-------------------
+++++ Comments +++++

++ I think the goal of formalizing the core of iTasks is great. TOP is an
interesting programming paradigm and it would be good to be able to
understand its essence and furthermore to have techniques to prove the
properties of TOP programs.

++ The development of the model seems sensible and the provided examples
are quite interesting.

-- However I think this current paper is far from this goal. As is it
feels more like an exercise in formalisms rather than a research result.
In particular, the approach to describing TOP is a straightforward
adaptation of ideas from process calculi (see work on languages with
session types).

-- Moreover, the task layer and its type system are straightforward
adaptation of monadic effect descriptions. This is a missed opportunity as
a richer type system could be used to prove interesting properties such as
for instance that certain tasks never fail.

-- Finally, there is nothing surprising or challenging in proving type
soundness for the model.

-- This model doesn't meet the requirements to be called a calculus.
Plotkin's seminal paper on the call-by-name and call-by-value lambda
calculus provides a blueprint for how one constructs a calculus for an
existing language. For a more recent example see Florence et al. 's POPL
paper on a calculus for Esterel with the same motivation as this paper but
for Esterel. This effort builds on equatinal theories for effectual programs developed by Felleisen in the 80s. In general, a calculus should allow sound  and consistent
equational reasoning with respect to the existing semantics  of the language
(in this case the graph rewriting one of Clean) and this is possible even in the
presence of effects at the price of weakening reasoning power.

-- Again related to Esterel and the above, the paper ``An axiomatic semantics for
Esterel'' by Simone Tini presents an alternative approach to formalizing
Esterel with the goal of reasoning about Esterel programs. That paper opts
for bisimulation techniques instead of algebraic manipulations of syntax.
Thus similar to TOPHAT, the paper opts for an LTS approach. Again though,
and unlike TOPHAT, the design of the model is directly connected with certain
properties and techniques.

+++++ Small points/typos +++++

page 13: I is mentioned here but it is defined in page 16.

page 18, first line: "we mad" ---> "we made"

page 18, first line: "combinators into" ---> "combinators on top of"

page 21, section 7.3, last line: "syntonisation" ---> "synchronization"

++++++ QUESTIONS THAT THE AUTHORS ARE EXPECTED TO ANSWER ++++++

1) What is challenging/surprising about constructing TOPHAT?

2) Could you please explain what kind of properties of iTasks
programs you intend to intend to try to prove with TOPHAT and
how its design helps this process?

### 2019-04-18 reviews/icfp-rebuttal.md
add first stab on rebuttal

<!--
Ik probeer zo weinig mogelijk "our" te gebruiken om het wij <-> zij verschil kleiner te maken.
In plaats van "our" gebruik ik "the".

Verder ga ik er bij verschillende dingen wellicht te hard in op de details.
Kunst is om zoveel als mogelijk terug te grijpen naar de dingen die de reviewers positief vinden aan ons paper.
Dat is belangrijk voor jullie om te checken!

Daarbij zit ik nu 26 woorden over de limiet in het algemene deel (dus tot aan het kopje "specific questions").
Iets om rekening mee te houden!

Tekst [tussen blokhaken] ben ik niet tevreden over en zou ik graag anders hebben.

-- TS
-->

# Rebuttal paper #111

We thank the reviewers for their efforts and constructive comments.
We are pleased to read that reviewers consider the paper is "intriguing" and "well written and interesting".
We are glad to hear "the model seems sensible and the provided examples are quite interesting",
and that we are addressing "an important topic that has received too little attention in the PL community".
Also, we will make sure to incorporate all detailed comments in the final version of the paper.

First and foremost, we'd like to apologise naming our model a calculus.
In light of Plotkin's paper from 1973, our model does not meet the requirements to be called a calculus.
Therefore, we will rename our model to a language.
We are grateful for reviewer D to bring this up.

Next to this, reviewers raised three main points:

* The semantics presented is not challenging or special in any way.
* The work does not contain an application of the given semantics
  or a description why it will be usable for formal reasoning.
* Some undesired tasks can be constructed or unwanted effects could happen.

Below we would like to address these three concerns first, after which we answer specific questions of each reviewer.

## Challenge

The challenge in creating this model is indeed not the used techniques,
which are well established and proved their usefulness in computer science.
[The challenge is going back to the roots of workflow and business process modelling,
while adding facilities to specify higher order workflows and generate applications.]
The resulting applications should have automatically generated user interfaces
and no data races despite a multi user environment.

Although the iTasks framework already promises this,
it is a large library [with lots of bells and whistles].
So, in particular, the challenge is to reformulate the iTasks framework and its ideas in a handful of elementary and well-behaved combinators.
The iTasks framework is totally build on top of two "super combinators" called `step` and `parallel`.
Both were created to express every possible collaboration pattern the creators encountered during case studies.
All other simpler combinators are derivations of these two.

For example, `parallel` has the type (in Haskell syntax):
```
parallel
  :: ITask a
  => [ ( ParallelTaskType, ParallelTask a ) ]
  -> [ TaskCont [ ( Int, TaskValue a ) ] ( ParallelTaskType, ParallelTask a ) ]
  -> Task [ ( Int, TaskValue a ) ]
```
Where `ITask` is a type class facilitating (de)serialisability, datatype generic representations,
type information at runtime, and a way to produce user facing editors.
Trimming this combinator down to two elementary combinators (⋈ and ⧫) which,
together with shared memory (■), can model most of its use cases,
is the main challenge solved in this work.

## Application

The analysis that the current work does not apply the semantics is a valid one.
The exclusion of any [proofs] is an intentional one.
In this work we focus on properly introducing the core concepts of TOP,
formally describe them,
and placing this model into the right context by comparing it a large set of other frameworks and algebras.

In an upcoming article, we are using our system to symbolically execute tasks:
"Proving task properties using symbolic TopHat", which is under consideration for IFL'19.
Symbolic execution allows us to prove properties like "breakfast will always be served" (ex 3.1),
and "passengers always get a seat" (ex 2.1).
These kind of properties would not be expressible and provable using a library like the iTasks framework
and need a well described core set of functionality as presented in this work.
Also, symbolic execution allows us to relate traces of inputs to tasks and observations on them,
a suggestion made by reviewer A.

Also, this gives us space to discuss and analyse a bigger case studies for the Dutch tax office and the Dutch coast guard,
which is more mission critical than the examples given in the current work.

## Undesired tasks and effects

### Higher order tasks

As reviewers A and C point out, it is possible to express tasks of higher order types,
which we did not discuss in the paper.
This is an omission which we should fix for the final version of the paper.
Examples given are:

* ⊠(Int -> Int) or □(λx:τ. x)
* ⊠(Task Int) or ⊠(⊠Int)

It is indeed possible to have tasks of functions and tasks of tasks.
On the user interaction part, one could render such editors as a selection menu of functions or tasks already available in the system.
Users could pick one of them, and continue with execution,
such as picking a function to operate on an integer and applying it to a entered integer like such:

t1 := ⊠(Int -> Int) ⋈ ⊠Int ▶︎ λ⟨f,x⟩. □(f x)

Something similar holds for executing a task entered by a user or, more generally, saved by another task.
Here t1 is selected and will be executed in the next normalisation round:

t2 := □(t1) ▶︎ λt. t

This also gives an answer to the missing join operation in the language and is the only way programmers can eliminate tasks.
A top level task can only be eliminated by running the semantics,
as it needs user input.

Saving something else than an basic type in a reference is, however, not allowed by the grammar.
This is a deliberate choice to exclude recursion by using references.
This means the example below are not possible:

* ■l where l : Int -> Int
* ■l where l : Task Int

### Race conditions and distributed systems

Line 363, which states that our model model supports distributed applications,
should never have reached the submitted version.
It is simply not true.
The system can only "generate _multi-user applications_ to support [...] workflows".
Multiple users can interact with an application running on a single server.

There is an important consequence of this choice that we should state more clearly in the paper.
If one takes as a definition for a _data race_ that
_the same program with the same inputs sometimes gives different results_
then this is not possible in TopHat.
The order of execution is completely determined by the order of the inputs,
which we assume to be a serial stream.
Therefore, it is indeed the case that program and environment alternate:
a user gives an input, the system makes some calculations and the user interface will change.
Only then the user can give her next input, etc.

I.e., in the booking example (ex 2.1):
the user who books the seat first gets the seat.
This is exactly the desired behaviour.

## Specific questions

### Speculative evaluation

The speculative evaluation of task continuations in the semantics is needed to express guards.
We acknowledge that taking state into account needs to reestablish memory if the evaluation leads to failure.
A different approach would be move state from the expression layer to the task layer.
We choose not to do this to keep task modelling features and default language features as branching and references completely separate.
We should have mentioned this reasoning in section 5.3.

### Dining philosophers

The dining philosophers are very easy to express using TopHat.
<!-- Definition should be here.  -- TS -->
To express a dynamic version of this problem,
we would need a (restricted) version of recursion.
One could think of built in recursion principle on natural numbers
or a replicate function on tasks.

### 2019-04-18 reviews/icfp-rebuttal.md
extend rebuttal

@@
 
-Daarbij zit ik nu 26 woorden over de limiet in het algemene deel (dus tot aan het kopje "specific questions").
+Daarbij zit ik nu zo'n 100 woorden over de limiet van 1000 in het algemene deel (dus tot aan het kopje "specific questions").
 Iets om rekening mee te houden!
 
-Tekst [tussen blokhaken] ben ik niet tevreden over en zou ik graag anders hebben.
+Ik vraag me even af op welk punt we nog het beste de "de positieve opmerkingen van de reviewers [kunnen] meenemen",
+zoals Rinus aangaf:
+
+> TOP is, zoals de reviewers aangeven, een belangrijk concept die een oplossing biedt voor interactieve applicaties (Achilles hiel van FP),
+> met iTasks worden complexe commerciele applicaties gemaakt (dus het is praktisch bruikbaar),
+> een semantiek die het redeneren over programma's mogelijk maakt is belangrijk, maar niet eenvoudig door de vele mogelijkheden,
+> en daarom is TopHat waarbij jullie je bewust beperken een eerste noodzakelijke stap.
+
+Verder zijn er gaten voor een implementatie van de dining philosophers.
+Kun jij die toevoegen Markus?
+En heb ik de laatste vraag van reviewer D nog niet beantwoord.
 
@@
 and that we are addressing "an important topic that has received too little attention in the PL community".
-Also, we will make sure to incorporate all detailed comments in the final version of the paper.
 
@@
 
-Next to this, reviewers raised three main points:
+Next to this, reviewers raised three main concerns:
 
 * The semantics presented is not challenging or special in any way.
-* The work does not contain an application of the given semantics
-  or a description why it will be usable for formal reasoning.
+* The work does not contain an application of the given semantics.
 * Some undesired tasks can be constructed or unwanted effects could happen.
@@
 Below we would like to address these three concerns first, after which we answer specific questions of each reviewer.
+Naturally, will make sure to incorporate all detailed comments in the final version of the paper.
 
-## Challenge
+## General concerns
 
-The challenge in creating this model is indeed not the used techniques,
-which are well established and proved their usefulness in computer science.
-[The challenge is going back to the roots of workflow and business process modelling,
-while adding facilities to specify higher order workflows and generate applications.]
+### Challenge
+
+As reviewers point out,
+the challenge in creating this model is indeed not the used techniques.
+It is indeed the case that proving type soundness is just a sanity check and not a validation.
+This should be reflected in the paper.
+
+<!-- Deze sectie kan iemand vast veeeeel korter herformuleren... ;-)  --TS -->
+
+The main challenge we faced was creating a language which models workflows and business processes,
+while being able to specify higher order workflows and generate applications out of such specifications.
 The resulting applications should have automatically generated user interfaces
@@
 Although the iTasks framework already promises this,
-it is a large library [with lots of bells and whistles].
-So, in particular, the challenge is to reformulate the iTasks framework and its ideas in a handful of elementary and well-behaved combinators.
+it is a large library embedded in an general purpose functional programming language.
+So, our challenge was to reformulate the iTasks framework and its ideas in a handful of elementary and well-behaved combinators.
+
 The iTasks framework is totally build on top of two "super combinators" called `step` and `parallel`.
@@
 Where `ITask` is a type class facilitating (de)serialisability, datatype generic representations,
-type information at runtime, and a way to produce user facing editors.
+type information at runtime, and a way to produce graphical editors.
 Trimming this combinator down to two elementary combinators (⋈ and ⧫) which,
@@
 is the main challenge solved in this work.
+We did not give this example in section 1.4, where it belongs.
 
-## Application
+### Application
 
 The analysis that the current work does not apply the semantics is a valid one.
-The exclusion of any [proofs] is an intentional one.
-In this work we focus on properly introducing the core concepts of TOP,
-formally describe them,
-and placing this model into the right context by comparing it a large set of other frameworks and algebras.
+The exclusion of any application is an intentional one.
+In this work, we focus on properly introducing the core concepts of TOP,
+linking them to a formal description,
+showing this description is sane,
+and placing the resulting model into the right context by comparing it a fair set of other frameworks and algebras.
+These four things lead to a holistic description of TOP which can be used for further research as well as implementation.
+Summarising: the presented system is "useful elsewhere", as indicated by reviewer A.
 
-In an upcoming article, we are using our system to symbolically execute tasks:
-"Proving task properties using symbolic TopHat", which is under consideration for IFL'19.
+In an upcoming article, we are using our system to symbolically execute tasks.
+<!-- Moet zo'n verwijzing er in?  --TS -->
+["Proving task properties using symbolic TopHat", under consideration for IFL'19.]
 Symbolic execution allows us to prove properties like "breakfast will always be served" (ex 3.1),
 and "passengers always get a seat" (ex 2.1).
-These kind of properties would not be expressible and provable using a library like the iTasks framework
-and need a well described core set of functionality as presented in this work.
-Also, symbolic execution allows us to relate traces of inputs to tasks and observations on them,
+It will be possible to relate traces of inputs to tasks and observations on them,
 a suggestion made by reviewer A.
+These kind of properties would not be expressible and provable using a library like the iTasks framework.
+They need a well described core set of functionality as presented in this work.
 
-Also, this gives us space to discuss and analyse a bigger case studies for the Dutch tax office and the Dutch coast guard,
-which is more mission critical than the examples given in the current work.
+A split also gives us the opportunity to discuss and analyse bigger case studies from the Dutch tax office and the Dutch coast guard.
+They are certainly more mission critical than the small examples given in the current work.
 
-## Undesired tasks and effects
+### Undesired tasks and effects
 
-### Higher order tasks
+#### Higher order tasks
 
@@
 which we did not discuss in the paper.
-This is an omission which we should fix for the final version of the paper.
+This is an omission which we should fix for the final version.
 Examples given are:
@@
 It is indeed possible to have tasks of functions and tasks of tasks.
-On the user interaction part, one could render such editors as a selection menu of functions or tasks already available in the system.
+On the user interaction part, one could render above editors as a selection menu of functions or tasks already available in the system.
 Users could pick one of them, and continue with execution,
@@
 
-### Race conditions and distributed systems
+#### Race conditions and distributed systems
 
-Line 363, which states that our model model supports distributed applications,
+Line 363, which states that our model supports distributed applications,
 should never have reached the submitted version.
 It is simply not true.
-The system can only "generate _multi-user applications_ to support [...] workflows".
-Multiple users can interact with an application running on a single server.
+The system can only "generate _multi-user applications_ to support [...] workflows":
+multiple users can interact with an application running on a single server.
 
 There is an important consequence of this choice that we should state more clearly in the paper.
-If one takes as a definition for a _data race_ that
+If one takes as a definition for a data race that
 _the same program with the same inputs sometimes gives different results_
@@
 which we assume to be a serial stream.
-Therefore, it is indeed the case that program and environment alternate:
+Therefore, it is indeed the case that program and environment alternate as mentioned by reviewer C:
 a user gives an input, the system makes some calculations and the user interface will change.
 Only then the user can give her next input, etc.
-
 I.e., in the booking example (ex 2.1):
@@
 
-### Speculative evaluation
+### Reviewer A
+
+> You say you "show that the normalisation semantics is a big-step semantics".
+> Which property is that, and why is it important?
+> Why does having state complicate a big-step semantics?
+
+This is an leftover and should not have made it into this version.
+We will remove it.
+
+> Could you encode BPEL in TopHat? If not, which features does BPEL
+> support that you do not? (E.g., cycles?)
+
+It should be possible to express most BPEL programs in TopHat
+as BPEL supports the same basic compositions as TopHat (sequence and parallel)
+and the same structured programming constructs.
+Only loop constructs are not supported, as TopHat does not have recursion.
+
+An embedding in the other direction, however, would not be possible.
+BPEL lacks the higher order constructs which TopHat has.
+
+> You are able to write the smoker program simply because your semantics eats all the complexity.
+> What would dining philosophers look like in your system?
+> It seems you cannot write a version of it that dynamically determines the number of philosophers.
+> What might you need to add to the language to support that?
+
+The dining philosophers are very easy to express using TopHat.
+<!-- Definition should be here.  -- TS -->
+To express a dynamic version of this problem,
+we would need a (restricted) version of recursion.
+One could think of built in recursion principle on natural numbers
+or a replicate function on tasks.
+
+### Reviewer C
+
+#### On speculative evaluation
 
@@
 
-### Dining philosophers
+#### On the notion of time in comparison to FRP
 
-The dining philosophers are very easy to express using TopHat.
-<!-- Definition should be here.  -- TS -->
-To express a dynamic version of this problem,
-we would need a (restricted) version of recursion.
-One could think of built in recursion principle on natural numbers
-or a replicate function on tasks.
+The main difference between time in TopHat and in FRP is that
+behaviours in FRP are explicitly functions from time to values.
+In TopHat, tasks have a notion of order, but do not explicitly depend on time.
+
+This doesn't mean, however, that task cannot react on a time value, as is done in example 4.5.
+Takeaway is that _a clock in TopHat is just a user which updates a shared time value on a regular base_.
+We will explicitly mention this in the paper in example 4.5 and in section 7.4.
+
+### Reviewer D
+
+> 1) What is challenging/surprising about constructing TOPHAT?
+
+We hope to have answered this question in the general section,
+in the paragraphs about "Challenge".
+
+> 2) Could you please explain what kind of properties of iTasks programs you intend to prove with TOPHAT
+>    and how its design helps this process?
+
+<!-- How does the design help??  --TS -->

### 2019-04-18 reviews/icfp-rebuttal.md
add notes on fail and editors

@@
 We did not give this example in section 1.4, where it belongs.
+
+The insight that a fail task (☇) is a basic task which will never have an observable value
+and also can be used for guards is a total change from the iTasks implementation.
+Also, our system radically simplifies the notion of editors and connects them to the already existing notion of widgets.
+All these points belong to the process of creating TopHat and are not apparent in the paper.
 

### 2019-04-19 reviews/icfp-rebuttal.md
rebuttal: general, challenges

--- 2019-04-18 reviews/icfp-rebuttal.md
@@
 We thank the reviewers for their efforts and constructive comments.
-We are pleased to read that reviewers consider the paper is "intriguing" and "well written and interesting".
-We are glad to hear "the model seems sensible and the provided examples are quite interesting",
+We are glad to hear that the reviewers consider the paper "intriguing" and "well written and interesting",
+that "the model seems sensible and the provided examples are quite interesting",
 and that we are addressing "an important topic that has received too little attention in the PL community".
 
-First and foremost, we'd like to apologise naming our model a calculus.
-In light of Plotkin's paper from 1973, our model does not meet the requirements to be called a calculus.
-Therefore, we will rename our model to a language.
-We are grateful for reviewer D to bring this up.
+First and foremost we are grateful for reviewer D for bringig up that in light
+of Plotkin's paper from 1973, our model does not meet the requirements to be
+called a calculus.
+We will call it a domain specific language, or just language.
 
-Next to this, reviewers raised three main concerns:
+The reviewers concerns fall into three main categories:
 
-* The semantics presented is not challenging or special in any way.
+* The semantics presented is not challenging or special.
 * The work does not contain an application of the given semantics.
@@
 
-Below we would like to address these three concerns first, after which we answer specific questions of each reviewer.
-Naturally, will make sure to incorporate all detailed comments in the final version of the paper.
+We will address these concerns, after which we will answer specific questions of each reviewer.
+Naturally, all comments will be addressed in the final version of the paper.
 
@@
 
-As reviewers point out,
-the challenge in creating this model is indeed not the used techniques.
-It is indeed the case that proving type soundness is just a sanity check and not a validation.
+As the reviewers point out,
+the challenge in creating our language is neither the used techniques,
+nor proving soundness.
 This should be reflected in the paper.
 
-<!-- Deze sectie kan iemand vast veeeeel korter herformuleren... ;-)  --TS -->
+The main challenge we faced was stripping down iTasks into a handful of
+elementary and well-behaved combinators while retaining the essential features
+of TOP.
 
-The main challenge we faced was creating a language which models workflows and business processes,
-while being able to specify higher order workflows and generate applications out of such specifications.
-The resulting applications should have automatically generated user interfaces
-and no data races despite a multi user environment.
+iTasks is built on the two feature-rich combinators `step` and `parallel`.
+Both were designed to express every possible collaboration pattern the creators encountered during case studies.
+All other combinators are derived from these two.
 
-Although the iTasks framework already promises this,
-it is a large library embedded in an general purpose functional programming language.
-So, our challenge was to reformulate the iTasks framework and its ideas in a handful of elementary and well-behaved combinators.
-
-The iTasks framework is totally build on top of two "super combinators" called `step` and `parallel`.
-Both were created to express every possible collaboration pattern the creators encountered during case studies.
-All other simpler combinators are derivations of these two.
-
-For example, `parallel` has the type (in Haskell syntax):
+For example, `parallel` has the following type, given in Haskell syntax:
 ```
@@
 ```
-Where `ITask` is a type class facilitating (de)serialisability, datatype generic representations,
-type information at runtime, and a way to produce graphical editors.
+Where `ITask` is a type class that provides serialisation, datatype generic representations,
+run-time type information, and generation of graphical editors.
 Trimming this combinator down to two elementary combinators (⋈ and ⧫) which,
@@
 We did not give this example in section 1.4, where it belongs.
+
+The insight that the fail task (☇) is a basic task which will never have an observable value
+and can be used for guards is a deviation from iTasks that lets us simplify the step combinator significantly.
+Also, our system simplifies the notion of editors and connects them to the already existing notion of widgets.
+All these points are not sufficiently emphasized in the paper.
 

### 2019-04-19 reviews/icfp-rebuttal.md
rebuttal: challenges

@@
 run-time type information, and generation of graphical editors.
-Trimming this combinator down to two elementary combinators (⋈ and ⧫) which,
-together with shared memory (■), can model most of its use cases,
-is the main challenge solved in this work.
+The main challenge solved in this work is
+trimming this combinator down to two elementary combinators (⋈ and ⧫) which,
+together with shared memory (■), can model most of its use cases.
 We did not give this example in section 1.4, where it belongs.
 
-The insight that the fail task (☇) is a basic task which will never have an observable value
-and can be used for guards is a deviation from iTasks that lets us simplify the step combinator significantly.
+Furthermore, we were able to significantly simplify the step combinator using the
+insight that the fail task (☇) is a basic task which will never have an observable value
+and can be used for guards.
+
 Also, our system simplifies the notion of editors and connects them to the already existing notion of widgets.
+
 All these points are not sufficiently emphasized in the paper.

### 2019-04-19 reviews/icfp-rebuttal.md
rebuttal: application

@@
 
-The analysis that the current work does not apply the semantics is a valid one.
-The exclusion of any application is an intentional one.
-In this work, we focus on properly introducing the core concepts of TOP,
+It is true that the paper does not use the semantics to prove properties about tasks.
+This is intentional.
+In this work, we focus on introducing the core concepts of TOP,
 linking them to a formal description,
-showing this description is sane,
-and placing the resulting model into the right context by comparing it a fair set of other frameworks and algebras.
+showing that this description is sane,
+and puting the resulting language into context by comparing it to other similar systems.
 These four things lead to a holistic description of TOP which can be used for further research as well as implementation.
-Summarising: the presented system is "useful elsewhere", as indicated by reviewer A.
 
@@
 Symbolic execution allows us to prove properties like "breakfast will always be served" (ex 3.1),
-and "passengers always get a seat" (ex 2.1).
+"passengers always get a seat", and "no two passengers can book the same seat" (ex 2.1).
 It will be possible to relate traces of inputs to tasks and observations on them,
 a suggestion made by reviewer A.
-These kind of properties would not be expressible and provable using a library like the iTasks framework.
-They need a well described core set of functionality as presented in this work.
 
-A split also gives us the opportunity to discuss and analyse bigger case studies from the Dutch tax office and the Dutch coast guard.
-They are certainly more mission critical than the small examples given in the current work.
+We plan to use this technique to discuss and analyse bigger case studies from the Dutch tax office and the Dutch coast guard.
+There is an implementation of a part of the Dutch law for solar panel subsidies, and the command and control process for search and rescue operations.
+These are certainly more mission critical than the small examples given in the current work.
 

### 2019-04-19 reviews/icfp-rebuttal.md
data races

@@
 <!-- Moet zo'n verwijzing er in?  --TS -->
+<!-- Ja!  --mkl -->
 ["Proving task properties using symbolic TopHat", under consideration for IFL'19.]
@@
 Line 363, which states that our model supports distributed applications,
-should never have reached the submitted version.
-It is simply not true.
+should not have reached the submitted version.
 The system can only "generate _multi-user applications_ to support [...] workflows":
 multiple users can interact with an application running on a single server.
+It is the users who are distributed, but their events are serialized by the server before being fed to the program.
 
 There is an important consequence of this choice that we should state more clearly in the paper.
-If one takes as a definition for a data race that
-_the same program with the same inputs sometimes gives different results_
+If one takes the following definition for a data race:
+
+  the same program with the same inputs sometimes gives different results
+
 then this is not possible in TopHat.
-The order of execution is completely determined by the order of the inputs,
-which we assume to be a serial stream.
-Therefore, it is indeed the case that program and environment alternate as mentioned by reviewer C:
-a user gives an input, the system makes some calculations and the user interface will change.
-Only then the user can give her next input, etc.
-I.e., in the booking example (ex 2.1):
+The order of execution is completely determined by the order of the inputs.
+Therefore, it is the case that program and environment alternate as mentioned by reviewer C.
+A user gives an input, the system makes some calculations and the user interface will change.
+Only then can the user give the next input.
+For the booking example (ex 2.1) this means that
 the user who books the seat first gets the seat.
-This is exactly the desired behaviour.
+This is the desired behaviour.
 

### 2019-04-19 reviews/icfp-rebuttal.md
rebuttal: task inception

@@
 It is indeed possible to have tasks of functions and tasks of tasks.
-On the user interaction part, one could render above editors as a selection menu of functions or tasks already available in the system.
+One could render editors like this as a selection menu of functions or tasks already available in the system.
 Users could pick one of them, and continue with execution,
@@
 
-Something similar holds for executing a task entered by a user or, more generally, saved by another task.
+Something similar holds for executing a task inside a task.
 Here t1 is selected and will be executed in the next normalisation round:
@@
 
-This also gives an answer to the missing join operation in the language and is the only way programmers can eliminate tasks.
-A top level task can only be eliminated by running the semantics,
-as it needs user input.
+This also gives an answer to the missing join operation.
+A task can only be eliminated by the semantics, and only if it is the top-level expression.
 
-Saving something else than an basic type in a reference is, however, not allowed by the grammar.
+Our grammar only allows putting values of basic types in references.
 This is a deliberate choice to exclude recursion by using references.
-This means the example below are not possible:
+This means the following are not possible:
 

### 2019-04-19 reviews/icfp-rebuttal.md
dining philosophers, individual questions

@@
 It should be possible to express most BPEL programs in TopHat
-as BPEL supports the same basic compositions as TopHat (sequence and parallel)
-and the same structured programming constructs.
-Only loop constructs are not supported, as TopHat does not have recursion.
+as BPEL supports the same basic constructs as TopHat (sequence and parallel)
+and the same structured programming features.
+Only loops are not supported, as TopHat does not have recursion.
 
-An embedding in the other direction, however, would not be possible.
+An embedding in the other direction would not be possible.
 BPEL lacks the higher order constructs which TopHat has.
@@
 
-The dining philosophers are very easy to express using TopHat.
-<!-- Definition should be here.  -- TS -->
-To express a dynamic version of this problem,
+The dining philosophers can be expressed in iTasks as follows.
+<!-- Ik heb geen tijd meer voor een vertaling naar TopHat  --TS -->
+
+```
+module Main
+
+import Data.Func
+import Data.Functor
+import iTasks
+
+Start world = startEngine main world
+
+fork0 = sharedStore "fork0" True
+fork1 = sharedStore "fork1" True
+fork2 = sharedStore "fork2" True
+numSitting = sharedStore "numSitting" 0
+
+takeFork side fork =
+  watch fork -|| (viewInformation "Philosopher" [] ("waiting for "+++side+++" fork")) >>*
+  [ OnAction (Action "Take fork") (ifValue id (const continue))
+  ]
+where
+  continue :: Task ()
+  continue = void $ set False fork
+
+putFork fork =
+  void $ set True fork
+
+sitDown =
+  watch numSitting -|| viewInformation "Philosopher" [] "hungry" >>*
+  [ OnAction (Action "Sit down") (ifValue (\x -> x < 2) (const $ void $ upd inc numSitting))
+  ]
+
+getUp =
+  viewInformation "Philosopher" [] "done eating" >>*
+  [ OnAction (Action "Get up") (always $ void $ upd dec numSitting)
+  ]
+
+philosopher leftFork rightFork =
+  sitDown >>- \_ ->
+  takeFork "left" leftFork >>- \_ ->
+  takeFork "right" rightFork >>- \_ ->
+  viewInformation "Philosopher" [] "eating..." >>*
+  [ OnAction (Action "Eat") (always $ return ())
+  ] >>- \_ ->
+  putFork rightFork >>- \_ ->
+  putFork leftFork >>- \_ ->
+  getUp
+
+main =
+  allTasks
+    [ forever $ philosopher fork0 fork1
+    , forever $ philosopher fork1 fork2
+    , forever $ philosopher fork2 fork0
+    , void $ viewSharedInformation "fork 0" [] fork0
+    , void $ viewSharedInformation "fork 1" [] fork1
+    , void $ viewSharedInformation "fork 2" [] fork2
+    , void $ viewSharedInformation "num sitting" [] numSitting
+    ]
+```
+
+To express version with dynamic number of philosophers
 we would need a (restricted) version of recursion.
@@
 We acknowledge that taking state into account needs to reestablish memory if the evaluation leads to failure.
-A different approach would be move state from the expression layer to the task layer.
+A different approach would be move state from the expression layer to the task layer, which would limit guards to pure expressions.
 We choose not to do this to keep task modelling features and default language features as branching and references completely separate.
@@
 
-<!-- How does the design help??  --TS -->
+The design of TopHat makes it possible to apply different kinds of program verification techniques to it.
+So far we have been considering the Dijkstra monad and symbolic execution.
+We are currently working on implementing a symbolic execution engine for TopHat.
+This would allow us to prove properties like the ones mentioned above:
+- "breakfast will always be served" (ex 3.1),
+- "passengers always get a seat"
+- "no two passengers can book the same seat" (ex 2.1).

### 2019-04-24 sections/relatedwork.tex
push fixmes in margin for last proofread
last and only todo
Fixed chorus related work
convert back to two column layout
Hide related work section: related web services

@@
 
-\subsection{Task-oriented programming implementations}
+\subsection{TOP implementations}
 
@@
 
-\subsection{Related web services}
-
-\paragraph{IFTTT}
-
-If This Then That (\IFTTT)~\cite{IFTTT} is a web service that allows users to write and
-configure simple conditional statements called applets, that can interact with
-web services and IoT devices. These applets can be seen as tiny workflows, that
-facilitate collaboration between different services and platforms. The
-complexity of the flow is very limited however, and is not geared towards humans
-collaborating.
-
-\paragraph{Google forms \& SurveyMonkey}
-
-Google Forms~\cite{googleforms} and SurveyMonkey~\cite{wufoo} (and their various other tools like Wufoo, Precise Polling and Zoomerang) are web based survey administration apps that allows
-users to compose forms and gather data from them. They are comparable to \TOP and
-iTasks in that they provide an easy way for users to construct forms on the web.
-However, Google Forms and Wufoo do not allow users to define what the system
-then should do with this information, and can therefore not be regarded as a
-workflow or business process modelling systems.
-
-\paragraph{Chorus} % (http://www.chorus-home.org/)
-
-Chorus~\cite{chen2017chorus} is a visual programming environment for online
-collaboration apps. The goal of Chorus is to allow users to program their own
-application, though which a collective task can be coordinated. Users work on a
-preset data-type that drives the behaviour of the user defined applications.
-\fixme{compare with tophat}
+% \subsection{Related web services}
+% 
+%
+% \paragraph{IFTTT}
+%
+% If This Then That (\IFTTT)~\cite{IFTTT} is a web service that allows users to write and
+% configure simple conditional statements called applets, that can interact with
+% web services and IoT devices. These applets can be seen as tiny workflows, that
+% facilitate collaboration between different services and platforms. The
+% complexity of the flow is very limited however, and is not geared towards humans
+% collaborating.
+%
+% \paragraph{Google forms \& SurveyMonkey}
+%
+% Google Forms~\cite{googleforms} and SurveyMonkey~\cite{wufoo} (and their various other tools like Wufoo, Precise Polling and Zoomerang) are web based survey administration apps that allows
+% users to compose forms and gather data from them. They are comparable to \TOP and
+% iTasks in that they provide an easy way for users to construct forms on the web.
+% However, Google Forms and Wufoo do not allow users to define what the system
+% then should do with this information, and can therefore not be regarded as a
+% workflow or business process modelling systems.
+%
+% \paragraph{Chorus} % (http://www.chorus-home.org/)
+%
+% Chorus~\cite{chen2017chorus} is a visual programming environment for online
+% collaboration apps. The goal of Chorus is to allow users to program their own
+% application, though which a collective task can be coordinated. Users work on a
+% preset data-type that drives the behaviour of the user defined applications.
+% Chorus can not be regarded as a fully fledged workflow system, nor does it aim
+% to be. The data-type that drives the behaviour of the app does not change during
+% execution, and therefore heavily limits what flows can be encoded.

### 2019-04-24 sections/intuition.tex
correct spelling in intuition
convert back to two column layout
rename main tex file to main
use paragraph instead of subsubsection*

@@
-% !TEX root=../icfp2019.tex
+% !TEX root=../main.tex
 
@@
 
-\subsection{Tasks model collaboration}
+\paragraph{Tasks model collaboration}
 %NOTE: Heb toch maar deze kop toegevoegd, onderstaande gaat er duidelijk over, en de sprong van het voorbeeld naar "Tasks are reusable" is anders abrupt!
@@
 
-\subsection{Tasks are reusable}
+\paragraph{Tasks are reusable}
 
@@
 
-\subsection{Tasks are driven by user input}
+\paragraph{Tasks are driven by user input}
 
@@
 Inputs are synchronous.
-We assume systems react instantaneously to each if those events.
+We assume systems react instantaneously to each of those events.
 
@@
 
-\subsection{Tasks can be observed}
+\paragraph{Tasks can be observed}
 
@@
 
-\subsection{Tasks are never done}
+\paragraph{Tasks are never done}
 
@@
 
-\subsection{Tasks can share information}
+\paragraph{Tasks can share information}
 
@@
 
-\subsection{Tasks are predictable}
+\paragraph{Tasks are predictable}
 
@@
 
-\subsection{Recap}
+\paragraph{Recap}
 
@@
 For example, the cigarette smokers communicate with each other, but the programs do not explicitly mention sending or receiving events.
-
-By focussing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate distributed, multi-user applications to support such workflows.
+By focussing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate multi-user applications to support such workflows.

### 2019-04-24 sections/semantics.tex
convert back to two column layout
rename main tex file to main
fixed typo

@@
   \centering
-  \includegraphics[width=0.6\columnwidth,page=5]{figures/drawings-crop.pdf}
+  \includegraphics[width=\columnwidth,page=5]{figures/drawings-crop.pdf}
   \caption{
@@
 
-Also, we mad an implementation of \TOPHAT combinators into \ITASKS,
+Also, we made an implementation of \TOPHAT combinators into \ITASKS,
 so that \TOPHAT specifications can be compiled to runnable applications.

### 2019-04-24 sections/intuition.tex
remove reference to distributed apps

--- 2019-03-01 sections/intuition.tex
@@
 
-By focussing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate distributed, multi-user applications to support such workflows.
+By focussing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate multi-user applications to support such workflows.

### 2019-04-24 reviews/icfp-rebuttal.md

@@
 
-All these points are not sufficiently emphasized in the paper.
+All these points are not sufficiently emphasised in the paper.
 
@@
 showing that this description is sane,
-and puting the resulting language into context by comparing it to other similar systems.
+and putting the resulting language into context by comparing it to other similar systems.
 These four things lead to a holistic description of TOP which can be used for further research as well as implementation.
@@
 <!-- Ja!  --mkl -->
-["Proving task properties using symbolic TopHat", under consideration for IFL'19.]
+["Proving task properties using symbolic TopHat", to be submitted to IFL'19.]
 Symbolic execution allows us to prove properties like "breakfast will always be served" (ex 3.1),
@@
 multiple users can interact with an application running on a single server.
-It is the users who are distributed, but their events are serialized by the server before being fed to the program.
+It is the users who are distributed, but their events are serialised by the server before being fed to the program.
 

### 2019-04-24 sections/intuition.tex
shorten code examples

--- 2019-04-24 sections/intuition.tex
@@
 
-\lstset{emph={makeDrink,makeFood}}
+\lstset{emph={mkDrink,mkFood}}
 
@@
 \begin{TASK}
-  let makeBreakfast : Task Drink -> Task Food -> Task <<Drink,Food>> =
-    \makeDrink. \makeFood. makeDrink <&> makeFood in
-  makeBreakfast (makeTea <?> makeCoffee) makeEgg >>? enjoyBreakfast
+  let mkBrkfst : Task Drink -> Task Food -> Task <<Drink,Food>>
+    = \mkDrink. \mkFood. mkDrink <&> mkFood in
+  mkBrkfst (mkTea <?> mkCoffee) mkEgg >>? enjoyMorning
 \end{TASK}
@@
   let vend : Task Snack = enter Int >>? \n. if n == 1 then edit Biscuit
-    else if n == 2 then edit ChocolateBar
-    else fail
+    else if n == 2 then edit ChocolateBar else fail
 \end{TASK}

### 2019-04-24 sections/language.tex
convert back to two column layout
rename main tex file to main
shorten code examples

@@
   \centering
-  \includegraphics[width=0.6\columnwidth,page=3]{figures/drawings-crop.pdf}
+  \includegraphics[width=\columnwidth,page=3]{figures/drawings-crop.pdf}
   \caption{
@@
   let cancel : Task Unit = edit <<>> in
-  let delay : Int -> Task Unit -> Task Unit = \n : Int. \proceed : Task Unit.
+  let delay : Int -> Task Unit -> Task Unit = \n. \proceed.
     (proceed <?> cancel) <|> (wait n >>= \u : Unit. proceed)

### 2019-04-24 appendices/proofs.tex
rename main tex file to main
shorten proofs section

@@
 
+\begin{lemma}[Task value preserves types]
+  % For all well typed expressions $e$ and states $\sigma$,
+  For all expressions $e$ and states $\sigma$
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  if $\Value{(e,\sigma)}=v$,
+  then $v:\tau$.
+  \label{lem:presvalue}
+\end{lemma}
+
 \begin{proof}
@@
 \subsection{\cref{thm:pres-stride}}
+
+\begin{lemma}[Striding preserves types]
+  For all expressions $e$ and states $\sigma$
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  if $e,\sigma \stride e',\sigma'$,
+  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers \sigma'$.
+  \label{thm:pres-stride}
+\end{lemma}
 

### 2019-04-24 sections/properties.tex
rename main tex file to main
remove reference to proofs about big step semantics
shorten properties section

@@
 theorem for our small-step handling semantics.
-
-Additionally, we show that the normalisation semantics is a big-step semantics.
-%XXX: Alejandro: ?
-% Niks meer mee doen, was voor PLDI reviewers duidelijk genoeg. --TS
-While at first sight, this might seem obvious, the fact that we are dealing with
-state complicates matters.
-
 We show that our failing function $\Failing$ indeed only indicates expressions
 that can not be normalised and that allow no further interaction.
-
 Finally, we prove that the function to compute all possible inputs $\Inputs$ is sound and complete.
@@
 
-\Cref{thm:pres-eval} is shown to be correct by induction over $e$. The full
-proof can be found in the appendix.
-
-Moving on, we show that normalisation also preserves types, by showing that the following Theorem holds.
-
 \begin{theorem}[Type preservation under normalisation]
@@
 \end{theorem}
-
-Since this semantics makes use of the value function $\Value$ and the striding semantics $\stride$,
-we first need to show that preservation also holds for these two semantics.
-
-\begin{lemma}[Task value preserves types]
-  % For all well typed expressions $e$ and states $\sigma$,
-  For all expressions $e$ and states $\sigma$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
-  if $\Value{(e,\sigma)}=v$,
-  then $v:\tau$.
-  \label{lem:presvalue}
-\end{lemma}
-
-\begin{lemma}[Striding preserves types]
-  For all expressions $e$ and states $\sigma$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
-  if $e,\sigma \stride e',\sigma'$,
-  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers \sigma'$.
-  \label{thm:pres-stride}
-\end{lemma}
-
-\Cref{lem:presvalue} and \cref{thm:pres-stride} are proven in the
-appendix by induction over $e$. This subsequently allows us to prove
-\Cref{thm:pres-norm}, again by induction over $e$.
-
-This brings us finally to the type preservation property of the handling semantics.
 
@@
 
-And again, this is proven by induction over $e$.
-
+All three Theorems are proven to be correct by induction over $e$.
+The full proofs are listed in the appendix.
 From \cref{thm:pres-handle} and \cref{thm:pres-norm} we directly
@@
 either fails, or there exists some input $i$ that can be handled by it under the handling semantics.
-In order to prove this Theorem, we require an additional theorem.
-
-% \paragraph{Failing}
-
-We need to show that the failing function $\Failing$ behaves as desired.
+In order to prove this Theorem, we need to show that the failing function $\Failing$ behaves as desired.
 

### 2019-04-24 sections/introduction.tex
convert back to two column layout
rename main tex file to main
micro-optimise sentence in introduction

@@
 
-\subsection{Tasks}
+\paragraph{Tasks}
 
@@
 
-\subsection{Task-oriented programming}
+\paragraph{Task-oriented programming}
 \label{sec:top}
@@
 This gives rise to a user's need to interact and share information.
-Next to that, \TOP automatically provides solutions to common development jobs like designing \GUI{s}, connecting to databases, and communicating between servers and clients.
+Next to that, \TOP automatically provides solutions to common development jobs like designing \GUI{s}, connecting to databases, and servers-client communication.
 
@@
 
-\subsection{Utilisation}
+\paragraph{Utilisation}
 
 Currently, we know of two frameworks implementing \TOP: \ITASKS and \MTASKS.
-
+%
 \ITASKS is an implementation of \TOP, in the form of a shallowly embedded domain-specific language in the lazy functional programming language Clean.
@@
 
-\subsection{Challenges}
+\paragraph{Challenges}
 
@@
 
-\subsection{Contributions}
+\paragraph{Contributions}
 
@@
 
-\begin{itemize}
+\begin{enumerate*}
   \item
     We describe the essential concepts of task-oriented programming.
-
   \item
@@
     It is based on the aforementioned essential \TOP concepts.
-
   \item
@@
     The semantics of the task language is clearly separated from the semantics of the underlying host language.
-
   \item
@@
     the current value, whether a term is stuck, the current user interface, and the accepted inputs.
-
   \item
     We prove progress and type preservation for \TOPHAT.
-
   \item
     Using both the essential concepts and the formal calculus, we compare \TOP with related work in areas ranging from business process modelling, to process algebras and reactive programming.
-
   \item
     We implemented the whole semantic system in the dependently typed programming language Idris \cite{journals/jfp/Brady13}.
-
   \item
@@
     This also demonstrates that the former is a subset of the latter.
+\end{enumerate*}
 
-\end{itemize}
-
-\subsection{Structure}
+\paragraph{Structure}
 

### 2019-04-24 sections/intuition.tex
rename main file to main
micro-optimise sentence in intuition

@@
 % \begin{TASK}
-%   let makeBreakfast : Task Drink -> Task Food -> Task Unit =
+%   let mkBrkfst : Task Drink -> Task Food -> Task Unit =
 %     \makeDrink. \makeFood. makeDrink <&> makeFood >>? \<<d, f>>. eat d f
 %   let main : Task Unit =
-%     makeBreakfast (makeTea <?> makeCoffee) makeEgg
+%     mkBrkfst (makeTea <?> makeCoffee) makeEgg
 % \end{TASK}
@@
 % \begin{TASK}
-%   let makeBreakfast : Task Drink -> Task Food -> Task Breakfast =
+%   let mkBrkfst : Task Drink -> Task Food -> Task Breakfast =
 %     \makeDrink. \makeFood. makeDrink <&> makeFood
 %   let main : Task Unit =
-%     makeBreakfast (makeTea <?> makeCoffee) makeEgg >>? \breakfast. eat breakfast
+%     mkBrkfst (makeTea <?> makeCoffee) makeEgg >>? \breakfast. eat breakfast
 % \end{TASK}
@@
 These aspects make it possible for programmers to model custom collaboration patterns.
-
-\Cref{exm:breakfast} demonstrates how tasks can be parameterised by other tasks: \TS{makeBreakfast} is a collaboration pattern that always works the same way, regardless of which food and drink are being prepared.
+%
+\Cref{exm:breakfast} demonstrates how tasks can be parameterised by other tasks: \TS{mkBrkfst} is a collaboration pattern that always works the same way, regardless of which food and drink are being prepared.
 
@@
 When the user presses the continue button, the machine dispenses either a biscuit or a chocolate bar, depending on the amount of money.
-Snacks are modelled using a custom data type.
+Snacks are modelled using a custom type.
 
@@
 Not all tasks have a value, the empty editor for instance, which makes value observation partial.
-For example, the value of $\Edit 7$ is $7$, but the value of $\Enter \Int$ is $\bot$.
+I.e., the value of $\Edit 7$ is $7$, but the value of $\Enter \Int$ is $\bot$.
 
@@
 
-The details of the solution are not important here.
+% The details of the solution are not important here.
 What is important is that the implementation of what is essentially deadlock-free waiting for two semaphores requires a substantial amount of additional synchronisation, together with non-trivial conditional statements.
-
+%
 \TOPHAT allows a simple solution to this problem, using guarded steps.

### 2019-04-24 sections/abstract.tex
rename main tex file to main
rename calculus to language

@@
 In this paper we give a description of the TOP paradigm and then decompose its rich features into elementary language elements, which makes them suitable for formal treatment.
-We use the simply typed lambda calculus, extended with pairs and references, as a base language.
-On top of this language, we develop TopHat, a calculus for modular interactive workflows.
+We use the simply typed $lambda$-calculus, extended with pairs and references, as a base language.
+On top of this language, we develop TopHat, a language for modular interactive workflows.
 We describe TopHat by means of a layered semantics.

### 2019-04-24 sections/conclusions.tex
Updated reference to TFP paper naus/jeuring
convert back to two column layout
rename main tex file to main
rename main file to main
rename calculus to language

@@
 In this paper we have identified and intuitively described the essence of task-oriented programming.
-We then formalized this essence by developing a domain-specific calculus for declarative interactive workflows, called \TOPHAT.
-The task calculus and the host language are clearly separated, to make explicit where the boundaries are.
+We then formalized this essence by developing a domain-specific language for declarative interactive workflows, called \TOPHAT.
+The task language and the host language are clearly separated, to make explicit where the boundaries are.
 The semantics of the task layer is driven by user input.
@@
 To this end, we are very interested to see if it is possible to develop an axiomatic semantics for \TOPHAT that allows us to do so.
-There are certain properties of our calculus that make this particularly complex:
+There are certain properties of our language that make this particularly complex:
 We have to deal with parallelism, user interaction, and references.
@@
 \citet{conf/ifl/KlinikJP17} have developed a cost analysis for tasks that require resources in order to be executed.
-This analysis was developed for a simpler task calculus, and could be brought over to the one developed here.
+This analysis was developed for a simpler task language, and could be brought over to the one developed here.
 
-\citet{UUCS2017013} have looked at building a generic feedback system for rule-based problems.
+\citet{DBLP:conf/sfp/NausJ16} have looked at building a generic feedback system for rule-based problems.
 A workflow system typically is rule based, as outlined in their work.
-It would be interesting to fit the generic feedback system to \TOPHAT in order to support end-users working in applications developed in this calculus.
+It would be interesting to fit the generic feedback system to \TOPHAT in order to support end-users working in applications developed in this language.
 

### 2019-04-24 sections/example.tex
rename main tex file to main
shorten code examples
introduce fail symbol in example
rename calculus to language

@@
 and the user can proceed to picking seats.
+In case of an invalid list of passengers,
+the step is guarded by the failing task \TS{fail}.
 \begin{TASK}
@@
 All instances of the \TS{bookFlight} task have access to the shared list of free seats.
-Rewriting the example in a calculus without side effects would not only be cumbersome,
+Rewriting the example in a language without side effects would not only be cumbersome,
 obfuscating the code with explicit threading of state,

### 2019-04-24 sections/introduction.tex
rename main file to main
rename calculus to language

--- 2019-02-28 sections/introduction.tex
@@
 Informally in the sense that we give an intuitive description of the features that define task-oriented programming.
-Formally in the sense that we develop a calculus which formalises these features as language constructs,
+Formally in the sense that we develop a language which formalises these features as language constructs,
 and we give them a semantics in the style that is common in programming language research.
@@
 Thus making explicit which properties of \TOP come from the task layer, and which come from functional programming.
-Our challenge, therefore, is to model the properties of \TOP into a calculus
+Our challenge, therefore, is to model the properties of \TOP into a language
 and pave the way for formal treatment of \TOP programs.
-We give this formal calculus the name \TOPHAT (TopHat).
+We give this formal language the name \TOPHAT (TopHat).
 
@@
   \item
-    We present a formal calculus for modelling declarative workflows, embedded in a simply typed lambda-calculus, called \TOPHAT.
+    We present a formal language for modelling declarative workflows, embedded in a simply typed $lambda$-calculus, called \TOPHAT.
     It is based on the aforementioned essential \TOP concepts.
@@
   \item
-    Using both the essential concepts and the formal calculus, we compare \TOP with related work in areas ranging from business process modelling, to process algebras and reactive programming.
+    Using both the essential concepts and the formal language, we compare \TOP with related work in areas ranging from business process modelling, to process algebras and reactive programming.
 
@@
 \Cref{sec:intuition} gives an overview of the essential concepts of \TOP.
-\Cref{sec:language} introduces the \TOPHAT calculus syntax
+\Cref{sec:language} introduces the \TOPHAT language syntax
 and \cref{sec:semantics} the semantics.
-Then in \cref{sec:properties} we show that certain properties hold for the calculus.
+Then in \cref{sec:properties} we show that certain properties hold for the language.
 We take a look at related work in \cref{sec:relatedwork}

### 2019-04-24 sections/language.tex
make grammars fit columns
make rules fit columns
exclude editor figure
rename main file to main
rename calculus to language

--- 2019-03-01 sections/language.tex
@@
 
-In this section, we present the constructs of \TOPHAT, our modular interactive workflow calculus.
+In this section, we present the constructs of \TOPHAT, our modular interactive workflow language.
 We define the host and task language, the types, and the static semantics.
@@
 \label{sub:expressions}
-The host language is a simply typed lambda calculus, extended with some basic types and \ML-style references.
+The host language is a simply typed $lambda$-calculus, extended with some basic types and \ML-style references.
 We use references to represent shared data sources.
@@
 % It is essential for a workflow system to support multiple collaborators working
-% together. Adding this funcitonaltiy to the \TOPHAT calculus is very
+% together. Adding this funcitonaltiy to the \TOPHAT language is very
 % straightforward. We identify users by means of a name $u$ which we can then use

### 2019-04-24 sections/relatedwork.tex
rename main tex file to main
rename main file to main
rename calculus to language

--- 2019-02-26 sections/relatedwork.tex
@@
 
-\BPEL~\cite{bpel} is another popular business process language. The standardised
+\BPEL~\cite{bpel} is another popular business process calculus. The standardised
 language allows for the specification of actions within business processes,
@@
 As with \TOPHAT, HipHop is a \DSL embedded in a general purpose programming language.
-% the same way \TOPHAT is embedded into the simply typed $\lambda$-calculus.
+% the same way \TOPHAT is embedded into the simply typed $\lambda$-language.
 Another similarity is that both specifications lead to executable server and client applications from a single source.
@@
 preset data-type that drives the behaviour of the user defined applications.
-\fixme{compare with tophat}
+Chorus can not be regarded as a fully fledged workflow system, nor does it aim
+to be. The data-type that drives the behaviour of the app does not change during
+execution, and therefore heavily limits what flows can be encoded.

### 2019-04-24 sections/introduction.tex

@@
     the current value, whether a term is stuck, the current user interface, and the accepted inputs.
-
   \item
     We prove progress and type preservation for \TOPHAT.
-
   \item
@@
     We implemented the whole semantic system in the dependently typed programming language Idris \cite{journals/jfp/Brady13}.
-
   \item

### 2019-04-24 sections/language.tex

@@
   \small
-  \usemacro{G-Language-Compact}
+  \usemacro{G-Language-Minimal}
   \caption{Language grammar} \label{fig:language-grammar}
 \end{figure}
+
 \label{sub:notation}
@@
   \small
-  \usemacro{G-Pretasks-Compact}
+  \usemacro{G-Pretasks-Minimal}
   \caption{Task grammar} \label{fig:task-grammar}
@@
   \small
-  \usemacro{G-Types-Compact}
+  \usemacro{G-Types-Minimal}
   \caption{Type grammar} \label{fig:type-grammar}
@@
     \userule{T-Enter} \Quad
-    \userule{T-Update} \Quad
-    \userule{T-Fail} \Break
+    \userule{T-Update} \Break
     \userule{T-Then} \Quad
     \userule{T-Next} \Break
-    \userule{T-And} \Quad
+    \userule{T-Fail} \Quad
+    \userule{T-And} \Break
     \userule{T-Or} \Quad
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.6\columnwidth,page=3]{figures/drawings-crop.pdf}
-  \caption{
-    Possible states of an editor and its transitions.
-    Shared editors cannot be cleared.
-  }
-  \label{fig:editor-state}
-\end{figure}
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=\columnwidth,page=3]{figures/drawings-crop.pdf}
+%   \caption{
+%     Possible states of an editor and its transitions.
+%     Shared editors cannot be cleared.
+%   }
+%   \label{fig:editor-state}
+% \end{figure}
 

### 2019-04-24 sections/language.tex
micro-optimise sentences in language section

@@
 
-Consider the following program.
+Consider the following:
 \begin{TASK}
@@
 
-Consider this slight variation on \cref{exm:conditions}.
+Consider this slight variation on \cref{exm:conditions}:
 \begin{TASK}
@@
     update time >>= \now : Int.
-      if now >= start + amount then edit <<>> else fail
+      if now > start + amount then edit <<>> else fail
 \end{TASK}
 The first step is immediately taken, resulting in \TS{start} to be the time at the moment \TS{wait} is executed.
-The second step is guarded until the current time is greater or equal to the start time plus the requested amount.
+The second step is guarded until the current time is greater to the start time plus the requested \TS{amount}.
 
@@
 In \TOPHAT, all parallel branches can progress independently, driven by input events.
-This requires input events to be tagged in order to reach the intended task.
+This requires inputs to be tagged in order to reach the intended~task. %%XXX: Hack to force `task` on the same line.
 
@@
 can only step when both editors have values.
-When it steps, the continuation uses the pair of integer and string to calculate the result.
+When it steps, the continuation uses the pair to calculate the result.
 

### 2019-04-25 sections/conclusions.tex
fix typos suggested by the reviewers

@@
 In this paper we have identified and intuitively described the essence of task-oriented programming.
-We then formalized this essence by developing a domain-specific language for declarative interactive workflows, called \TOPHAT.
+We then formalised this essence by developing a domain-specific language for declarative interactive workflows, called \TOPHAT.
 The task language and the host language are clearly separated, to make explicit where the boundaries are.
 The semantics of the task layer is driven by user input.
-We have compared \TOPHAT with workflow moddeling languages, process algebras, functional reactive programming and session types to point out differences and similarities.
+We have compared \TOPHAT with workflow modelling languages, process algebras, functional reactive programming and session types to point out differences and similarities.
 Finally, we have proven type safety and progress for our language.
@@
 
-One of the main motivations to formalize task-oriented programming is to be able to reason about programs.
+One of the main motivations to formalise task-oriented programming is to be able to reason about programs.
 In this paper we reason about the language itself, but it would be nice to prove properties about individual programs.
@@
 
-Additionally, we would like to develop visualizations for \TOPHAT language constructs.
-An assistive development environment integrating these visualizations and the presented textual language
+Additionally, we would like to develop visualisations for \TOPHAT language constructs.
+An assistive development environment integrating these visualisations and the presented textual language
 would aid domain experts to model workflows in a more accessible manner.
-A system that visualizes iTask programs has been developed in the past~\cite{DBLP:conf/sfp/StutterheimPA14}.
+A system that visualises iTask programs has been developed in the past~\cite{DBLP:conf/sfp/StutterheimPA14}.

### 2019-04-25 sections/intuition.tex
fix typos suggested by the reviewers

--- 2019-04-24 sections/intuition.tex
@@
 
-By focussing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate multi-user applications to support such workflows.
+By focusing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate multi-user applications to support such workflows.

### 2019-04-25 sections/properties.tex
micro-optimise sentences in properties section
fix typos suggested by the reviewers

--- 2019-03-25 sections/properties.tex
@@
 
-In order to validate our semantics, we show that our evaluation, normalisation
+In order to show our semantics is sane, we show that our evaluation, normalisation
 and handling semantics is type preserving. We additionally prove a progress
 theorem for our small-step handling semantics.
-
-Additionally, we show that the normalisation semantics is a big-step semantics.
-%XXX: Alejandro: ?
-% Niks meer mee doen, was voor PLDI reviewers duidelijk genoeg. --TS
-While at first sight, this might seem obvious, the fact that we are dealing with
-state complicates matters.
-
 We show that our failing function $\Failing$ indeed only indicates expressions
 that can not be normalised and that allow no further interaction.
-
 Finally, we prove that the function to compute all possible inputs $\Inputs$ is sound and complete.

### 2019-04-25 sections/relatedwork.tex
remove acknowledgements for more space
remove remark that itasks has an "implicit semantics through clean"
fix typos suggested by the reviewers

@@
 These semantics however do not make an explicit distinction between the host language and task language and they do not provide an explicit formal semantics.
-\ITASKS implicitly has a formal semantics, inherited from the graph-rewriting semantics of Clean.
-This semantics however does not lend itself well for formal reasoning.
+Therefore, the do not lend itself well for formal reasoning.
 
@@
 Synchronisation means that only some, but not all, of the possible interleavings are desirable.
-The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise and intuitive syntonisation code.
+The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise and intuitive synchronisation code.
 
@@
 
-Input events are sent to such a machine programatically using Hop, or are explicitly wired to events from the client.
+Input events are sent to such a machine programmatically using Hop, or are explicitly wired to events from the client.
 They are optionally associated with a Hop value.

### 2019-04-25 sections/semantics.tex
make grammars fit columns
make maps fit columns
make rules fit columns
rename main file to main
rename calculus to language
fix typos suggested by the reviewers

--- 2019-03-25 sections/semantics.tex
@@
 We organise this by following the structure of the language.
-Firstly, the task language is embedded in a simply typed lambda calculus.
+Firstly, the task language is embedded in a simply typed $lambda$-calculus.
 This requires a specification of the \emph{evaluation} of terms in the host language, and how it handles the task language.
@@
 
-Also, we mad an implementation of \TOPHAT combinators into \ITASKS,
+Also, we made an implementation of \TOPHAT combinators on top of \ITASKS,
 so that \TOPHAT specifications can be compiled to runnable applications.

### 2019-04-25 sections/abstract.tex
remove from abstract that we prove interesting properties of tophat

@@
 In this paper we give a description of the TOP paradigm and then decompose its rich features into elementary language elements, which makes them suitable for formal treatment.
-We use the simply typed $lambda$-calculus, extended with pairs and references, as a base language.
+We use the simply typed lambda-calculus, extended with pairs and references, as a base language.
 On top of this language, we develop TopHat, a language for modular interactive workflows.
@@
 With TopHat we prepare a way to formally reason about TOP languages and programs.
-We show some interesting properties of the language.
 This approach allows for comparison with other work in the field.
@@
 % Importance
-TOP has been appplied in projects with the Dutch coast guard, tax office, and navy.
+TOP has been applied in projects with the Dutch coast guard, tax office, and navy.
 Our work matters because formal program verification is important for mission-critical software, especially for systems with concurrency.
@@
 % Knowledge: What new facts were uncovered? If the research was not results oriented, what new capabilities are enabled by the work?
-% Grounding: What argument, feasibility proof, artifacts, or results and evaluation support this work?
+% Grounding: What argument, feasibility proof, artefacts, or results and evaluation support this work?
 % Importance: Why does this work matter?

### 2019-04-25 sections/intuition.tex

--- 2019-04-24 sections/intuition.tex
@@
 For example, the cigarette smokers communicate with each other, but the programs do not explicitly mention sending or receiving events.
-By focussing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate multi-user applications to support such workflows.
+
+By focusing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate multi-user applications to support such workflows.

### 2019-04-25 sections/properties.tex

--- 2019-04-24 sections/properties.tex
@@
 
-In order to validate our semantics, we show that our evaluation, normalisation
+In order to show our semantics is sane, we show that our evaluation, normalisation
 and handling semantics is type preserving. We additionally prove a progress
@@
 
+\noindent
 Where $\Sigma \infers \sigma$ means that for all $l\in \sigma$, it holds that

### 2019-05-02 sections/introduction.tex
fix new par in introduction
fix typos suggested by the reviewers
clarify sentence structure

@@
 \MTASKS is a subset of \ITASKS,
-focussing on \IOT devices and deployment on micro controllers.
+focusing on \IOT devices and deployment on micro controllers.
 It has been used to control home thermostats and other home automation applications \cite{koopman2018task}.
@@
   \item
-    We present a formal language for modelling declarative workflows, embedded in a simply typed $lambda$-calculus, called \TOPHAT.
+    We present a formal language for modelling declarative workflows, called \TOPHAT, embedded in a simply typed $\lambda$-calculus.
     It is based on the aforementioned essential \TOP concepts.
@@
     Using both the essential concepts and the formal language, we compare \TOP with related work in areas ranging from business process modelling, to process algebras and reactive programming.
-
   \item

### 2019-05-02 sections/language.tex
rename Calculus -> Language
lambda -> \lambda
remove reference to deleted figure

@@
 
-\section{Calculus}
+\section{Language}
 \label{sec:language}
@@
 \label{sub:expressions}
-The host language is a simply typed $lambda$-calculus, extended with some basic types and \ML-style references.
+The host language is a simply typed $\lambda$-calculus, extended with some basic types and \ML-style references.
 We use references to represent shared data sources.
@@
 Empty editors can be \emph{filled}.
-This is depicted as a state diagram in \cref{fig:editor-state} below.
 

### 2019-05-02 sections/semantics.tex
lambda -> \lambda
steps keep their right hand side unevaluated

@@
 We organise this by following the structure of the language.
-Firstly, the task language is embedded in a simply typed $lambda$-calculus.
+Firstly, the task language is embedded in a simply typed $\lambda$-calculus.
 This requires a specification of the \emph{evaluation} of terms in the host language, and how it handles the task language.
@@
   \small
-  \usemacro{G-Values-Compact}
+  \usemacro{G-Values-Minimal}
   \caption{Value grammar} \label{fig:value-grammar}
@@
 Most task constructors are strict in their arguments.
-Only steps keep their left hand side unevaluated to delay side effects till the moment the step is taken.
+Only steps keep their right hand side unevaluated to delay side effects till the moment the step is taken.
 The same holds for both branches of the external choice.
@@
 %   \small
-%   \usemacro{O-Value-Compact} \hfill \usemacro{O-Failing-Compact} \hfill \usemacro{O-Inputs-Compact}
+%   \usemacro{O-Value-Minimal} \hfill \usemacro{O-Failing-Minimal} \hfill \usemacro{O-Inputs-Minimal}
 %   \caption{Values} \label{fig:observation-value}
@@
   \small
-  \usemacro{O-Failing}
+  \usemacro{O-Failing-Compact}
   \caption{Failing} \label{fig:observation-failing}
@@
   \paragraph{Step}
-  \begin{align*}
-    \userule{S-ThenStay} \Quad
-    & \userule{S-ThenFail} \Break
-    & \userule{S-ThenCont}
-  \end{align*}
+  \begin{gather*}
+    \userule{S-ThenStay} \Break
+    \userule{S-ThenFail} \Break
+    \userule{S-ThenCont}
+  \end{gather*}
 
   \paragraph{Choose}
-  \begin{align*}
-    \userule{S-OrLeft} \Quad
-    & \userule{S-OrRight} \Break
-    & \userule{S-OrNone}
-  \end{align*}
+  \begin{gather*}
+    \userule{S-OrLeft} \Break
+    \userule{S-OrRight} \Break
+    \userule{S-OrNone}
+  \end{gather*}
 
@@
   \begin{gather*}
-    \userule{S-Edit} \Quad \userule{S-Fill} \Quad \userule{S-Update} \Quad
+    \userule{S-Edit} \Quad \userule{S-Fill} \Quad \userule{S-Update} \Break
     \userule{S-Fail} \Quad \userule{S-Xor}
@@
   \small
-  \usemacro{G-Inputs-Compact}
+  \usemacro{G-Inputs-Minimal}
   \caption{Input grammar} \label{fig:input-grammar}
@@
   \small
-  \usemacro{O-Inputs}
+  \usemacro{O-Inputs-Compact}
   \caption{Inputs} \label{fig:observation-input}
@@
   \begin{gather*}
-    \userule{H-PassThen} \Quad \userule{H-FirstAnd} \Quad \userule{H-FirstOr} \Break
-    \userule{H-PassNext} \Quad \userule{H-SecondAnd} \Quad \userule{H-SecondOr}
+    \userule{H-PassThen} \Quad \userule{H-PassNext} \Break
+    \userule{H-FirstAnd} \Quad \userule{H-SecondAnd} \Break
+    \userule{H-FirstOr}  \Quad \userule{H-SecondOr}
   %  \userule{H-Appoint}

### 2019-05-07 sections/language.tex
refer to Haskell impl instead of Idris

@@
 Annotations are not covered in this paper.
-Our Idris implementation of \TOPHAT supports annotating tasks with user \ID{}s, so that individual tasks in a large workflow can be assigned to different users.
+Our Haskell implementation of \TOPHAT supports annotating tasks with user \ID{}s, so that individual tasks in a large workflow can be assigned to different users.
 These annotations are used to filter the user interfaces for each user so that they can only see their part of the workflow.

### 2019-05-07 sections/semantics.tex
add github link

@@
 
-The semantics have been implemented in the Idris programming language \cite{journals/jfp/Brady13}.
+The semantics have been implemented in the Haskell programming language \cite{marlow2010haskell}.
 We use techniques presented by \textcite{journals/entcs/JaskelioffGH11}, \textcite{journals/jfp/Swierstra08}, and \textcite{school/maktoberdorf/PeytonJones01}.
 % Such an implementation increases the confidence in the correctness and completeness of the semantics.
-The source code can be found on GitHub.\footnote{\url{http://github.com/XXX/XXX}}
+The source code can be found on GitHub.\footnote{\url{https://github.com/timjs/tophat-haskell}}
 % A proof of concept implementation in Haskell is also available, showing that \TOPHAT can be implemented without dependent types.

### 2019-05-24 appendices/proofs.tex
Removed all mention of clearing editors

@@
   \case
-    {$\userule{H-Empty}$}
-    {Given that $\Gamma,\Sigma\infers\Edit v:\Task\beta$ and $\Sigma\infers s$, the \refrule{H-Empty} rule additionally gives us that $v:\tau$.
-    Then by \refrule{T-Enter} we have $\Gamma,\Sigma\infers\Enter \tau:\Task\tau$.}
-
-  \case
     {$\userule{H-Fill}$}
@@
 
-  \case{$e=\Edit v:\Task\tau, i= \Empty$}
-       {Given that $\userule{H-Empty}$, we have by definition that $\Inputs(\Edit v:\Task\tau,s) = \set{v':\tau,\Empty}$, which includes $\Empty$.}
-
   \case{$e=\Enter \tau, i= v':\tau$}

### 2019-05-24 sections/intuition.tex
explain synchronous events
Removed all mention of clearing editors

@@
 In this way the system communicates with the environment.
-Inputs are synchronous.
-We assume systems react instantaneously to each of those events.
+Inputs are synchronous, which means the order of execution is completely determined by the order of the inputs.
 
@@
 Take for example an editor holding the integer seven: $\Edit 7$.
-Such an editor reacts to two kinds of events.
-The first are change events, for example the values $42$ or $37$, which are of the same type.
-The second is a clear event $\Empty$, which empties the editor.
-This results is an empty editor $\Enter \Int$, annotated to only hold integers.
+Such an editor reacts to change events, for example the values $42$ or $37$, which are of the same type.
 
@@
 % This is important for generating user interfaces.
-For example, the task $\Edit 7$ can react to value events and empty events, as discussed before.
+For example, the task $\Edit 7$ can react to value events, as discussed before.
 

### 2019-06-25 sections/language.tex
Removed all mention of clearing editors
Todo's from reviews

@@
 
+\todo{Speaking about references you say "so they do not introduce implicit recursion (T-Update)".
+What do you want to mean by "implicit recursion".
+Is it because in the other cases $\tau$ can be again a Task and here $\beta$ can only be a basic type?}
+
 \begin{figure}[h]
@@
 Editors are typed containers that either hold a value or are empty.
-Editors that have a value can be \emph{changed} or \emph{emptied}.
+Editors that have a value can be \emph{changed}.
 Empty editors can be \emph{filled}.
@@
 Empty editors are annotated with a type in order to ensure type safety.
-Without type annotations, it would be possible to change the type of an editor, by clearing it and filling it with a value of different type.
+
+\todo{"Without type annotations, it would be possible to change the type of an editor, by clearing it and filling it with a value of different type." Is this a hypothetical affirmation (for the case the language wasn't type-safe) or a current feature of the language?}
 
@@
 \end{example}
+\todo{The starting double-quotes around 'Batman' and 'Na' are backwards.}
 

### 2019-06-25 sections/semantics.tex
Removed all mention of clearing editors
Todo's from reviews

@@
 with $\sigma'=\set{l\mapsto\True}$.
+\todo{you mention S-AND before having introduced the striding semantics (perhaps add a reference to the figure where this rule appears).}
 This expression is not normalised, because the left task can take a step.
@@
 \end{figure}
-
+\todo{when you write $v' : \tau$ (in the set comprehensions) I assume the intention is to say any input value of type $\tau$. I would write it as ${v' | v' : \tau}$.}
 Handling input is done by the \emph{handling} semantics shown in \cref{fig:handling-semantics}.
@@
     \userule{H-Change} \Quad
-    \userule{H-Empty} \Break
-    \userule{H-Fill} \Quad
+    \userule{H-Fill} \Break
     \userule{H-Update}
@@
 \refrule{H-Change},
-\refrule{H-Empty},
 \refrule{H-Fill},
 \refrule{H-Update}:
-Input events $v$ and $\Empty$ are used to change the value of editors and empty them.
+Input events $v$ are used to change the value of editors.
 Editors only accept values of the correct type.

### 2019-07-02 sections/language.tex
Fixed reviewer comments. One TODO left; fix proofs

--- 2019-05-07 sections/language.tex
@@
 Editors that hold an expression $e : \tau$ have type $\Task \tau$.
-Empty editors are annotated with a type in order to ensure type safety.
-Without type annotations, it would be possible to change the type of an editor, by clearing it and filling it with a value of different type.
+Empty editors are annotated with a type in order to ensure type safety and type preservation during evaluation.
 

### 2019-07-02 sections/properties.tex
Todo's from reviews
Fixed reviewer comments. One TODO left; fix proofs

@@
   if $e,\sigma \evaluate e',\sigma'$,
-  then $\Gamma,\Sigma \infers e':\tau$ and $\Sigma \infers \sigma'$.
+  then $\Gamma,\Sigma \infers e':\tau$ and $\Gamma,\Sigma \infers \sigma'$.
   \label{thm:pres-eval}
@@
 \noindent
-Where $\Sigma \infers \sigma$ means that for all $l\in \sigma$, it holds that
-$\Sigma(l)=\Gamma,\Sigma\infers \sigma(l)$.
+Where $\Gamma,\Sigma \infers \sigma$ means that for all $l\in \sigma$, it holds that
+$\Gamma,\Sigma\infers \sigma(l):\Sigma(l)$.
 
@@
   For all expressions $e$ and states $\sigma$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Gamma,\Sigma \infers \sigma$,
   if $e,\sigma \normalise e',\sigma'$,
-  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers \sigma'$.
+  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Gamma,\Sigma \infers \sigma'$.
   \label{thm:pres-norm}
@@
   For all expressions $e$, states $\sigma$ and inputs $i$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Gamma,\Sigma \infers \sigma$,
   if $ e,\sigma \handle{i} e',\sigma'$,
-  then $\Gamma,\Sigma\infers e':\Task\tau$ and $\Sigma\infers \sigma'$.
+  then $\Gamma,\Sigma\infers e':\Task\tau$ and $\Gamma,\Sigma\infers \sigma'$.
   \label{thm:pres-handle}

### 2019-07-02 appendices/proofs.tex
Last corrections for camera ready

@@
     {$\userule{E-Edit}$}
-    {Given that $\Gamma,\Sigma\infers\Edit e:\Task \tau$ and $\Sigma\infers s$,\refrule{T-Edit} gives us that $\Gamma,\Sigma\infers e:\tau$.
-    The induction hypothesis gives us that $e,s\evaluate v,s'$ also preserves, and thus $\Gamma,\Sigma\infers v:\tau$ and $\Sigma\infers s'$.
+    {Given that $\Gamma,\Sigma\infers\Edit e:\Task \tau$ and $\Gamma,\Sigma\infers s$,\refrule{T-Edit} gives us that $\Gamma,\Sigma\infers e:\tau$.
+    The induction hypothesis gives us that $e,s\evaluate v,s'$ also preserves, and thus $\Gamma,\Sigma\infers v:\tau$ and $\Gamma,\Sigma\infers s'$.
     Therefore $\Gamma,\Sigma\infers\Edit v:\Task\tau$.}
@@
     {$\userule{E-Update}$}
-    {Given that $\Gamma,\Sigma\infers \Edit e:\Task \tau$ and $\Sigma\infers s$, \refrule{T-Update} gives us that $\Gamma,\Sigma\infers e:\Ref \tau$.
-    The induction hypothesis gives us that $e,s\evaluate l,s'$ also preserves, and thus $\Gamma,\Sigma\infers l:\Ref\tau$ and $\Sigma\infers s'$.
+    {Given that $\Gamma,\Sigma\infers \Edit e:\Task \tau$ and $\Gamma,\Sigma\infers s$, \refrule{T-Update} gives us that $\Gamma,\Sigma\infers e:\Ref \tau$.
+    The induction hypothesis gives us that $e,s\evaluate l,s'$ also preserves, and thus $\Gamma,\Sigma\infers l:\Ref\tau$ and $\Gamma,\Sigma\infers s'$.
     Therefore $\Gamma,\Sigma\infers\Update l:\Task\tau$.}
@@
     {$\userule{E-Then}$}
-    {Given that $\Gamma,\Sigma\infers e_1\Then e_2:\Task \tau$ and $\Sigma\infers s$, \refrule{T-Then} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$.
-    By the induction hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Sigma\infers s'$.
+    {Given that $\Gamma,\Sigma\infers e_1\Then e_2:\Task \tau$ and $\Gamma,\Sigma\infers s$, \refrule{T-Then} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$.
+    By the induction hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers s'$.
     Therefore $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$.}
@@
     {$\userule{E-Next}$}
-    {Given that $\Gamma,\Sigma\infers e_1\Next e_2:\Task \tau$ and $\Sigma\infers s$, \refrule{T-Next} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$.
-    By the induction hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Sigma\infers s'$.
+    {Given that $\Gamma,\Sigma\infers e_1\Next e_2:\Task \tau$ and $\Gamma,\Sigma\infers s$, \refrule{T-Next} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1 \to \Task \tau$.
+    By the induction hypothesis, we know that $e_1,s\evaluate t_1,s'$ preserves and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers s'$.
     Therefore $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$.}
@@
     {$\userule{E-And}$}
-    {Given that $\Gamma,\Sigma\infers e_1\And e_2:\Task(\tau_1\times\tau_2)$ and $\Sigma\infers s$, \refrule{T-And} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\Task\tau_2$.
-    By the induction hypothesis, we know that both $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$ preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$, $\Sigma\infers s'$, $\Gamma,\Sigma\infers t_2:\Task\tau_2$ and $\Sigma\infers s''$.
+    {Given that $\Gamma,\Sigma\infers e_1\And e_2:\Task(\tau_1\times\tau_2)$ and $\Gamma,\Sigma\infers s$, \refrule{T-And} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\Task\tau_2$.
+    By the induction hypothesis, we know that both $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$ preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau_1$, $\Gamma,\Sigma\infers s'$, $\Gamma,\Sigma\infers t_2:\Task\tau_2$ and $\Gamma,\Sigma\infers s''$.
     Therefore $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$.}
@@
     {$\userule{E-Or}$}
-    {Given that $\Gamma,\Sigma\infers e_1\Or e_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau$ and $\Gamma,\Sigma\infers e_2:\Task\tau$.
-    By the induction hypothesis, we have that both $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$ preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau$, $\Sigma\infers s'$, $\Gamma,\Sigma\infers t_2:\Task\tau$ and $\Sigma\infers s''$.
+    {Given that $\Gamma,\Sigma\infers e_1\Or e_2:\Task\tau$ and $\Gamma,\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers e_1:\Task\tau$ and $\Gamma,\Sigma\infers e_2:\Task\tau$.
+    By the induction hypothesis, we have that both $e_1,s\evaluate t_1,s'$ and $e_2,s'\evaluate t_2,s''$ preserve and thus $\Gamma,\Sigma\infers t_1:\Task\tau$, $\Gamma,\Sigma\infers s'$, $\Gamma,\Sigma\infers t_2:\Task\tau$ and $\Gamma,\Sigma\infers s''$.
     Therefore $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$.}
@@
   For all expressions $e$ and states $\sigma$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Gamma,\Sigma \infers \sigma$,
   if $\Value{(e,\sigma)}=v$,
@@
 \case {$\Value{(\Update l,s)}=s(l)$ }
-      {Given that $\Gamma,\Sigma\infers\Update l:\Task \tau$ and $\Sigma\infers s$,
+      {Given that $\Gamma,\Sigma\infers\Update l:\Task \tau$ and $\Gamma,\Sigma\infers s$,
       we know that $\Gamma,\Sigma\infers s(l):\tau$ by definiton.}
@@
   For all expressions $e$ and states $\sigma$
-  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers \sigma$,
+  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Gamma,\Sigma \infers \sigma$,
   if $e,\sigma \stride e',\sigma'$,
-  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers \sigma'$.
+  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Gamma,\Sigma \infers \sigma'$.
   \label{thm:pres-stride}
@@
     {Given that
-    $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that $\Gamma,\Sigma\infers t:\Task\tau$ and $\Sigma\infers s'$ by
+    $\Gamma,\Sigma\infers e:\Task\tau$ and $\Gamma,\Sigma\infers s$, we know that $\Gamma,\Sigma\infers t:\Task\tau$ and $\Gamma,\Sigma\infers s'$ by
     \Cref{thm:pres-eval}.
-    Then by \cref{thm:pres-stride}, we have $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$.}
+    Then by \cref{thm:pres-stride}, we have $\Gamma,\Sigma\infers t':\Task\tau$ and $\Gamma,\Sigma\infers s''$.}
 
@@
     {$\userule{N-Repeat}$}
-    { Given that $\Gamma,\Sigma\infers e:\Task\tau$ and $\Sigma\infers s$, we know that $\Gamma,\Sigma\infers t:\Task\tau$ and $\Sigma\infers s'$ by \cref{thm:pres-eval}.
-    Then by \cref{thm:pres-stride}, we have $\Gamma,\Sigma\infers t':\Task\tau$ and $\Sigma\infers s''$.
-    Then by the induction hypothesis, we finally obtain that $\Gamma,\Sigma\infers t'':\Task\tau$ and $\Sigma\infers s'''$.}
+    { Given that $\Gamma,\Sigma\infers e:\Task\tau$ and $\Gamma,\Sigma\infers s$, we know that $\Gamma,\Sigma\infers t:\Task\tau$ and $\Gamma,\Sigma\infers s'$ by \cref{thm:pres-eval}.
+    Then by \cref{thm:pres-stride}, we have $\Gamma,\Sigma\infers t':\Task\tau$ and $\Gamma,\Sigma\infers s''$.
+    Then by the induction hypothesis, we finally obtain that $\Gamma,\Sigma\infers t'':\Task\tau$ and $\Gamma,\Sigma\infers s'''$.}
 
@@
 \begin{lemma}
-  Given that $\Sigma\infers s$, $\Sigma(l)=\tau$ and $\Gamma,\Sigma\infers v:\tau$, it holds that $\Sigma\infers s[l\mapsto v]$
+  Given that $\Gamma,\Sigma\infers s$, $\Sigma(l)=\tau$ and $\Gamma,\Sigma\infers v:\tau$, it holds that $\Gamma,\Sigma\infers s[l\mapsto v]$
   \label{lemmasigmaconsistent}
@@
     {$\userule{H-Change}$}
-    {Given that $\Gamma,\Sigma\infers\Edit v:\Task\tau$ and $\Sigma\infers s$, the \refrule{H-Change} rule additionally gives us that $v,v':\tau$.
+    {Given that $\Gamma,\Sigma\infers\Edit v:\Task\tau$ and $\Gamma,\Sigma\infers s$, the \refrule{H-Change} rule additionally gives us that $v,v':\tau$.
 Therefore by \refrule{T-Edit} we have that $\Gamma,\Sigma\infers\Edit v':\Task\tau$.}
@@
     {$\userule{H-Fill}$}
-    {Given that $\Gamma,\Sigma\infers\Enter\tau$ and $\Sigma\infers s$, the \refrule{H-Fill} rule additionally gives us that $v':\tau$.
+    {Given that $\Gamma,\Sigma\infers\Enter\tau$ and $\Gamma,\Sigma\infers s$, the \refrule{H-Fill} rule additionally gives us that $v':\tau$.
     Then by \refrule{T-Enter} we have $\Gamma,\Sigma\infers \Edit v':\Task\tau$.}
@@
     {$\userule{H-Update}$}
-    {Given that $\Gamma,\Sigma\infers\Update l:\Task\tau$ and $\Sigma\infers s$.
+    {Given that $\Gamma,\Sigma\infers\Update l:\Task\tau$ and $\Gamma,\Sigma\infers s$.
     This gives us that $\Sigma(l)=\tau$, and we additionally obtain $s(l),v':\tau$ by \refrule{H-Update}.
@@
     {$\userule{H-PickLeft}$}
-    {Given that $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Sigma\infers s$,
+    {Given that $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Gamma,\Sigma\infers s$,
     then by \refrule{T-Xor} we have $\Gamma,\Sigma\infers t_1:\Task \tau$.}
@@
     {$\userule{H-PickRight}$}
-    {Given that $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Sigma\infers s$,
+    {Given that $\Gamma,\Sigma\infers t_1\Xor t_2:\Task\tau$ and $\Gamma,\Sigma\infers s$,
     then by \refrule{T-Xor} we have $\Gamma,\Sigma\infers t_2:\Task \tau$.}
@@
     {$\userule{H-Next}$}
-    {Given that $\Gamma,\Sigma\infers t_1\Next e_2 :\Task\tau$ and $\Sigma\infers s$.
+    {Given that $\Gamma,\Sigma\infers t_1\Next e_2 :\Task\tau$ and $\Gamma,\Sigma\infers s$.
     Then by \refrule{T-Next}, we have $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$.
@@
     {$\userule{H-PassThen}$}
-    {Given that $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$ and $\Sigma\infers s$,
+    {Given that $\Gamma,\Sigma\infers t_1\Then e_2:\Task\tau$ and $\Gamma,\Sigma\infers s$,
     \refrule{T-Then} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$.
@@
 \case{$\userule{H-PassNext}$}
-     {Given that $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Next} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$.
+     {Given that $\Gamma,\Sigma\infers t_1\Next e_2:\Task\tau$ and $\Gamma,\Sigma\infers s$, \refrule{T-Next} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers e_2:\tau_1\to\Task\tau$.
      By the induction hypothesis, we know that $t_1,s\handle{i}t_1',s'$ also preserves and thus $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Gamma,Sigma\infers s'$.
@@
 \case{$\userule{H-FirstAnd}$}
-     {Given that $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and $\Sigma\infers s$, \refrule{T-And} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers t_2:\Task\tau_2$.
-     By the induction hypothesis, we know that $t_1,s\handle{i}t_1',s'$ also preserves and thus $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Sigma\infers s'$.
+     {Given that $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and $\Gamma,\Sigma\infers s$, \refrule{T-And} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers t_2:\Task\tau_2$.
+     By the induction hypothesis, we know that $t_1,s\handle{i}t_1',s'$ also preserves and thus $\Gamma,\Sigma\infers t_1':\Task\tau_1$ and $\Gamma,\Sigma\infers s'$.
      Therefore by \refrule{T-Next} we obtain $\Gamma,\Sigma\infers t_1'\And t_2:\Task(\tau_1\times\tau_2)$.}
@@
 \case{$\userule{H-SecondAnd}$}
-     {Given that $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and $\Sigma\infers s$, \refrule{T-And} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers t_2:\Task\tau_2$.
-     By the induction hypothesis, we know that $t_2,s\handle{i}t_2',s'$ also preserves and thus $\Gamma,\Sigma\infers t_2':\Task\tau_2$ and $\Sigma\infers s'$.
+     {Given that $\Gamma,\Sigma\infers t_1\And t_2:\Task(\tau_1\times\tau_2)$ and $\Gamma,\Sigma\infers s$, \refrule{T-And} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau_1$ and $\Gamma,\Sigma\infers t_2:\Task\tau_2$.
+     By the induction hypothesis, we know that $t_2,s\handle{i}t_2',s'$ also preserves and thus $\Gamma,\Sigma\infers t_2':\Task\tau_2$ and $\Gamma,\Sigma\infers s'$.
      Therefore by \refrule{T-Next} we obtain $\Gamma,\Sigma\infers t_1\And t_2':\Task(\tau_1\times\tau_2)$.}
@@
 \case{$\userule{H-FirstOr}$}
-     {Given that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
-     By the induction hypothesis we know that $t_1,s\handle{i}t_1',s'$ also preserves, and therefore $\Gamma,\Sigma\infers t_1':\Task\tau$ and $\Sigma\infers s'$.
+     {Given that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Gamma,\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
+     By the induction hypothesis we know that $t_1,s\handle{i}t_1',s'$ also preserves, and therefore $\Gamma,\Sigma\infers t_1':\Task\tau$ and $\Gamma,\Sigma\infers s'$.
      By \refrule{T-Or} we now obtain $\Gamma,\Sigma\infers t_1'\Or t_2:\Task\tau$.}
@@
 \case{$\userule{H-SecondOr}$}
-     {Given that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
-     By the induction hypothesis we know that $t_2,s\handle{i}t_2',s'$ also preserves, and therefore $\Gamma,\Sigma\infers t_2':\Task\tau$ and $\Sigma\infers s'$.
+     {Given that $\Gamma,\Sigma\infers t_1\Or t_2:\Task\tau$ and $\Gamma,\Sigma\infers s$, \refrule{T-Or} gives us that $\Gamma,\Sigma\infers t_1:\Task\tau$ and $\Gamma,\Sigma\infers t_2:\Task\tau$.
+     By the induction hypothesis we know that $t_2,s\handle{i}t_2',s'$ also preserves, and therefore $\Gamma,\Sigma\infers t_2':\Task\tau$ and $\Gamma,\Sigma\infers s'$.
      By T-Or we now obtain $\Gamma,\Sigma\infers t_1\Or t_2':\Task\tau$.}

### 2019-09-16 sections/language.tex
Last corrections for camera ready
merge remote-tracking branch 'origin/ppdp2019' into ppdp2019

@@
 Typing rules are of the form $\RelationT$, which should be read as \enquote{in environment $\Gamma$ and store typing $\Sigma$, expression $e$ has type $\tau$}.
-Typing rules for expressions in the host language are presented in the appendix.
+Typing rules for expressions in the host language are presented in the appendix\footnote{\url{https://github.com/timjs/tophat/appendix.pdf}}.
 The typing rules for pretasks are given in \cref{fig:typing-rules}.
@@
 Empty editors can be \emph{filled}.
+This is depicted as a state diagram in \cref{fig:editor-state} below.
 
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=\columnwidth,page=3]{figures/drawings-crop.pdf}
-%   \caption{
-%     Possible states of an editor and its transitions.
-%     Shared editors cannot be cleared.
-%   }
-%   \label{fig:editor-state}
-% \end{figure}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth,page=3]{figures/drawings-crop.pdf}
+  \caption{
+    Possible states of an editor and its transitions.
+    Shared editors cannot be cleared.
+  }
+  \label{fig:editor-state}
+\end{figure}
 

### 2019-09-16 sections/semantics.tex
Fixed reviewer comments. One TODO left; fix proofs
merge remote-tracking branch 'origin/ppdp2019' into ppdp2019

--- 2019-05-07 sections/semantics.tex
@@
 The evaluation rules for tasks can be deduced from the value grammar.
-They are given in the appendix.
+They are given in the appendix\footnote{\url{https://github.com/timjs/tophat/appendix.pdf}}.
 Most task constructors are strict in their arguments.
@@
 
+The striding semantics is responsible for reducing internal steps and internal choices.
+A stride from task $t$ in state $\sigma$ to $t'$ in state $\sigma'$ is denoted by $\RelationS$.
+The rules for striding are given in \cref{fig:striding-semantics}.
+Tasks like editors, fail and external choice are not further reduced.
+For external choice and parallel there are congruence rules.
+
 The split between striding and normalisation is due to mutable references.
@@
 %XXX: Alejandro: The example is not 100\% clear to me
-
-The striding semantics is responsible for reducing internal steps and internal choices.
-A stride from task $t$ in state $\sigma$ to $t'$ in state $\sigma'$ is denoted by $\RelationS$.
-The rules for striding are given in \cref{fig:striding-semantics}.
-Tasks like editors, fail and external choice are not further reduced.
-For external choice and parallel there are congruence rules.
 

### 2021-01-08 make-with-docker.sh
add make-with-docker script

@@ -1,0 +1,5 @@
+#!/usr/bin/env bash
+
+# for more information see https://hub.docker.com/r/blang/latex
+
+docker run --rm -i --user="$(id -u):$(id -g)" -v "$PWD":/data blang/latex:ubuntu latexmk -pdf main.tex
